### 2020-12-10 Notes/Latent_field_and_commercial_data.rmd
Ajout de notre script de simulation et du Rmd explicatif

---
title: "Simualation du modele"
author: "Oceane GUITTON, Chloe TELLIER, Juliette THEOLEYRE"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

Importations des packages :

```{r, message=FALSE}
library(spatstat)
library(tidyverse)
library(ggplot2)
```

On considere un modele integre, qui prend en compte l'echantillonage preferentiel (des donnees commerciales), et qui deduit la distribution spatiale de densites de poissons a partir des donnees scientifiques et/ou des donnees commerciales.

Ce domument presente une situation dans le cadre de la generation de donnees commerciales.

On considere une grille de taille 10*10 qui va definir notre zone d'etude.

# Champ Latent S(x)

Dans un premier premier temps, on s'interesse a la generation du *champ latent S(x)*.

Le champ latent represente une densite de poisson au point x considere.

Ce champ latent est modelise selon champ spatial gaussien (plus precisément "latent log gaussian") sur un domaine discret.

Ce champ latent depend de differentes covariables (comme par exemple le type de sol et la profondeur) et d'un effet spatial aleatoire.

Le champ latent est exprime en fonction dun intercept (alpha_s ou $\alpha_{s}$), des covariables environnementales et des parametres (betas_s ou $\beta_{s}$) qui leur sont associes (ces parametres traduisent la relation entre les especes et leur habitat (effets fixes)), et enfin de l'effet spatial aleatoire qui traduit la correlation spatiale des captures ($\delta(x)$).

Voici ci dessous les expressions du champ latent et de l'effet spatial aleatoire :

\begin{eqnarray*}
\text{log}(S(x)) & = & \alpha_{s} + \Gamma_{s}(x)^{T} \cdot \beta_{s} + \delta(x)
\end{eqnarray*}

\begin{eqnarray*}
\delta(x) & = & \text{GRF} (0 , M (x , x':\rho) )
\end{eqnarray*}

Par simplification, on ne considere ici qu'une seule covariable (qui est le type de sol) et on ne prend pas en compte l'effet spatial aleatoire. On considere quatre modalites pour la variable type de sol numerotees de 1 a 4.

La fonction zonage prend pour argument les coordonnees d'un point de la zone d'etude (coodonnees x et y) et renvoie pour ce point la valeur associee au type de sol du point considere.  Par exemple, le point de coordonnees (7,7) presente la modalite 4 pour le type de sol.

```{r}

zonage <- function(x,y, xmed = 5, ymed = 5){
  ifelse(x<xmed, 1, 2) + ifelse(y<ymed, 0, 2) 
}

zonage(7,7)
```

On initialise ensuite les valeurs des coefficients $\alpha_{s}$ et $\beta_{s}$.
Le coefficient $\alpha_{s}$ est initialise a -0.5.
La covariable type de sol etant discrete a quatre modalites, on initialise quatre valeurs de $\beta_{s}$, une pour chaque modalite. 
Les $\beta_{s}$ sont initialises a 0, 1, -1 et 2.

```{r}
alpha_s <- -0.5
betas_s <- c(0, 1, -1, 2)
```

On cree ensuite une fonction champ latent qui va calculer la densite de poisson pour un point donne (la valeur de ce champ latent depend de la modalite du type de sol).

La fonction champ latent  prend pour argument les coordonnees x et y du point considere, ainsi que les coefficients $\alpha_{s}$ et $\beta_{s}$. Elle renvoie la valeur du champ latent au point considere. Par exemple, la valeur du champ latent au point de coordonnees (7,7) vaut 4.48.

```{r}
latent_field_function <- function(x,y, alpha_s, betas_s){
  exp(alpha_s +  betas_s[zonage(x,y)])
}

latent_field_function(7,7,alpha_s, betas_s)
```

On propose une representation graphique de la repartition ce champ latent pour des points simules au hasard au sein de notre grille d'etude.

On voit donc que la zone en haut a droite de la grille d'etude est la zone de plus forte densite.

```{r}
dta <- data.frame(x = runif(n = 10000, 0,10 ), y = runif(10000, 0, 10)) %>% 
  mutate(z= latent_field_function(x, y, alpha_s, betas_s))
ggplot(data= dta) + geom_point(aes(x=x, y=y, col = z)) + labs(colour= "Champ latent")
```

# Generation des points de peche et processsus de Poisson non homogene

Desormais, nous nous interessons au **processus d'echantillonage**, c'est a dire a la **generation des points de peche** (et de leur distribution spatiale) selon un processus de poisson non homogene. L'intensite des points de peche dans une zone donnee dépend de l'abondance du champ latent a cet endroit.

La distribution spatiale des points de peche peut dependre d'un echantillonage aleatoire (c'est le cas de la recolte des donnees scientifiques) ou bien d'un echantillonage preferentiel (c'est le cas des donnees commerciales).

En effet, les pecheurs ciblent en general leurs zones de peche.

La localisation des points de peche peut dependre de la connaissance a priori de la distribution spatiale de la densite de poisson (champ latent), d'autres effets incarnes par des covariables (comme la connaissance de la presence d'especes non cibles (cf prises accessoires), les technologies disposees, les contraintes logistiques etc) et enfin d'un effet spatial aleatoire.

On note $X_{com f}$ les points spatiaux ou les bateaux de la flotte f ont peche (ou ont ete identifies comme etant en train de pecher). $X_{com f}$ est considere comme un "poisson point process" avec une variation spatiale d'intensite $\lambda_{f}(x)$.

Cette insensite $\lambda_{f}(x)$ est exprimee en fonction d'un intercept (alpha_Xf ou $\alpha_{Xf}$), d'un parametre $b_{f}$ (qui traduit un echantillonage preferentiel plus ou moins fort), du champ latent au point considere, de covariables et de leurs parametres associes (betas_Xf ou $\beta_{Xf}$), et enfin de l'effet spatial aleatoire ($\eta{f}(x)$).

Voici ci dessous les expressions de $X_{com f}$ et de $\lambda_{f}(x)$.

\begin{eqnarray*}
X_{com f} \sim {\cal IPP} (\lambda_{f}(x))
\end{eqnarray*}

\begin{eqnarray*}
\text{log}(\lambda_{f}(x)) & = & \alpha_{Xf} +  b_{f} \cdot \text{log}(S(x)) + \Gamma_{X}(x)^{T} \cdot \beta_{Xf} + \eta_{f}(x)
\end{eqnarray*}

On commence donc tout d'abord exprimer $\lambda_{f}(x)$.

On commence par initialiser les parametres necessaires a l'expression de $\lambda_{f}(x)$ a savoir $\alpha_{Xf}$ (que l'on fixe a 1) et $b_{f}$ (que l'on fixe a 1).

$b_{f}$ peut prendre pour valeurs 0, 1 et 3.

$b_{f}$ = 0 signifie qu'il n'y a pas d'echantillonage preferentiel.

Ensuite, plus $b_{f}$ augmente plus l'echantillonage preferentiel est important.

```{r}
alpha_xf <- 1
b_f <- 1
```

On cree ensuite une fonction qui calcule la valeur de l'intensite $\lambda_{f}(x)$ a un point donne de coordonnees (x,y).

Cette fonction prend en argument les coordonnees du point (x,y), ainsi que les parametres $\alpha_{Xf}$, $b_{f}$, $\alpha_{s}$ et $\beta_{s}$.

Par exemple, on calcule l'intensite $\lambda_{f}(x)$ au point de coordonnees (7,7).

```{r}
lambda_function <- function(x, y, alpha_xf, b_f, alpha_s, betas_s){
  exp(alpha_xf + b_f * log(latent_field_function(x,y, alpha_s, betas_s)))
}

lambda_function(7,7,alpha_xf, b_f, alpha_s, betas_s)
```

On genere ensuite des points de peche selon un processus de poisson non homogene grace a la fonction **rpoispp** du package *spatstat*.

Cette fonction prend pour arguments :
- lambda : la fonction de densite
- lmax : pour fixer un maximum pour la densite, eviter qu'elle n'explose
- win : la fenetre dans laquelle on simule nos points de densite, pour nous c'est 10*10
- les autres arguments utilises par lambda (sauf x et y qui du coup sont definis par la fenetre win)

```{r}
pp = rpoispp(lambda_function, lmax=4.5, win=owin(c(0,10),c(0,10)),
             alpha_xf=alpha_xf, b_f=b_f, alpha_s= alpha_s, betas_s = betas_s)
plot(pp)
```

On voit que la zone de droite (zone qui presente les densites de poisson (champ latent) les plus fortes) est la zone qui concentre le plus de points de peches. L'echantillonage preferentiel est donc bien pris en compte : les pecheurs visent les zones avec les plus fortes densites de poissons.

# Zero inflating

L'objectif est d'attribuer a chaque point de peche prealablement determine avec la fonction rpoispp la quantite de poissons peches, sachant qu'il est possible qu'en certains points de peche aucun poisson n'ai ete peche.

On determine alors dans un premier temps quels sont les points ou la peche a ete 'sans succes' et ceux ou elle s'est revelee etre fructueuse.

Ensuite on attribue a chaque point de peche 'fructueux' la quantite de poisson pechee.

Pour cela, on fait appel a un modele zero-inflated et on modelise la probabilite de presence/absence de capture par une loi de Bernoulli.

## Determination de la nature des peches : fructueuses ou non, loi de Bernoulli

On cree un vecteur de longueur notre nombre de point de peches qui ne contient pour l'instant que des 0.

Ce vecteur sera porteur de l'information de peche : abscence de capture (Y = 0) ou presence de capture (Y = 1).

```{r}
Y = rep(0, len = length(pp$x))
```

La fonction peche_attendue permet de calculer la quantite $\mu_f(x_i)$ caracterisee par l'expression $\mu_f(x_i) = q_f\cdot S(x_{i})$.

Elle prend pour arguments les coordonnees du point de peche (x,y), la capturabilite de la flotte q (specifique a chaque flotte) et retourne la peche attendue.

On prend l'exemple du point de peche de coordonnees (7,7) et on obtient sa peche attendue.

```{r}
# On fixe la capturabilité a 1
q = 1

# Pêche attendue
peche_attendue = function(x, y, q){
  q * latent_field_function(x, y, alpha_s, betas_s)
}

peche_attendue(7,7,q)
```

La fonction p0_bernoulli calcule, pour chaque point de peche, la probabilite de ne rien pecher :

si $y_i = 0$ alors $P(Y = y_i|x_i, S(x_i)) = exp(-e^{\xi_f}\cdot \mu_f(x_i))$

La fonction prend pour arguments les coordonnees du point (x,y), $\xi$i (parametre qui controle l'intensite de 0), et q (capturabilite de la flotte).

Par exemple, pour le point de coordonnees (7,7), la probabilite de ne capturer aucun poisson vaut 0,01 (il est coherent que cette probabilite soit faible puisque ce point est dans une zone de forte densite, il est donc plus probable que la peche soit fructueuse).

```{r}
#  On fixe l'intensité de 0 captures, ksi, a 0
ksi = 0

# Proba d'absence de capture
p0_bernoulli = function(x, y, ksi, q){
  exp(-exp(ksi) * peche_attendue(x, y, q))
}

p0_bernoulli(7,7,ksi,q)
```

Dans le cas d'un "succes" de Bernoulli, la flotte n'a peche aucun poisson en ce point de peche.

On fait alors (1 - p0_bernoulli(x,y,ksi,q)) afin que Y prenne la valeur 1 lorsque la peche a ete fructueuse et la valeur 0 si la peche a ete 'sans succes'.

La loi de Bernouilli de parametre p0_bernoulli permet de determiner le succes ou l'echec de la peche pour chacun des points.

```{r}
Y = 1 - rbinom(n=length(Y), size=1, prob=p0_bernoulli(pp$x, pp$y, ksi, q))
```

On repertorie nos donnees dans un dataframe qui contient les coordonnees de chaque point de peche ainsi que la nature de la peche : 'fructueuse' ou 'sans succes'.

```{r}
table = data.frame(x=pp$x, y=pp$y, peche=Y)
```

## Visualisation de la nature des peches

On affiche dans notre grille les points de peche 'sans succes' et 'fructueux'.

```{r}
ggplot(table) + geom_point(aes(x=x, y=y, col=as.factor(peche))) + labs(color= "Statut de la peche")
```

## Determination des quantites pechees en chaque point de peche fructueuse : loi logNormale

On cherche maintenant quel est le nombre de poissons potentiellement peches en chaque point de peche.

Pour cela on modelise une loi lognormale : ${\cal L} (y_i,\frac{\mu_f(x_i)}{1-exp(-e^{\xi_f}\cdot \mu_f(x_i)},\sigma^2)$.

Remarque : on fixe notre ecart-type $\sigma^2$ a 1.

```{r}
L = rlnorm(n=length(Y), 
           meanlog = peche_attendue(pp$x, pp$y, q)/(1 - p0_bernoulli(pp$x, pp$y, ksi, q)),
           sdlog=0.1)
```

On ajoute a notre table une colonne quantite qui materialise la quantite de poisson pechee en chaque point de peche :

Le nombre de poissons peches = le succes ou non de la peche (Y) * le nombre de poissons potentiellement peches (L).

Ce qui donne sous forme mathematique : $(1-exp(-e^{\xi_f}\cdot \mu_f(x_i))\cdot{\cal L} (y_i,\frac{\mu_f(x_i)}{1-exp(-e^{\xi_f}\cdot \mu_f(x_i)},\sigma^2$

```{r}
table <-table %>%
  mutate(quantite = L*Y)
```

## Visualisation des quantites de poissons peches en chaque point de peche

On represente graphiquement les quantites de poisson peches en differents points de notre grille d'etude.

```{r}
ggplot(table) + 
  geom_point(aes(x=x, y=y, col=quantite)) + labs(color= "Quantite pechee")
```

L'histogramme ci-dessous permet de visualiser la loi zero inflated :

```{r}
hist(table$quantite, main = "Frequence des quantites de poissons peches", xlab = "Quantite de poisson peches", ylab = "Frequence")
```

# Re allocation uniforme des captures

On va maintenant considerer que la capture totale est enregistree a l'echelle d'une zone.

```{r}
capturetotale = sum(table$peche)
```

Ensuite, au sein de cette zone, nous n'avons pas le detail de chaque capture a chaque point de peche : la capture totale est repartie de facon uniforme entre tous les points de peche.

## Division de notre grille d'etude en une seule zone

```{r}
table$L_unezone = mean(table$peche)
ggplot(table) + geom_point(aes(x=x, y=y, col=as.factor(L_unezone))) + labs(color= "Quantite pechee")
```

## Division de notre grille d'etude en deux zones

```{r}

# On coupe selon l'axe des x, a une valeur definie
xmed = 5
premierezone = table[which(table$x < xmed),]
deuxiemezone = table[which(table$x >= xmed),]
mean_premierezone = mean(premierezone$peche)
mean_deuxiemezone = mean(deuxiemezone$peche)
table$L_deuxzones = rep(0, nrow(table))
table[which(table$x < xmed), "L_deuxzones"] = mean_premierezone
table[which(table$x >= xmed), "L_deuxzones"] = mean_deuxiemezone
hist(table$L_deuxzones)
ggplot(table) + geom_point(aes(x=x, y=y, col=as.factor(L_deuxzones))) + labs(color= "Quantite pechee")
```

## Division de notre grille d'etude en quatre zones

```{r}
# On coupe selon l'axe des x et l'axe des y, a des valeurs definies
xmed = 5
ymed = 5
premierezone = table[which(table$x < xmed & table$y < ymed),]
deuxiemezone = table[which(table$x < xmed & table$y >= ymed),]
troisiemezone = table[which(table$x >= xmed & table$y < ymed),]
quatriemezone = table[which(table$x >= xmed & table$y >= ymed),]
mean_premierezone = mean(premierezone$peche)
mean_deuxiemezone = mean(deuxiemezone$peche)
mean_troisiemezone = mean(troisiemezone$peche)
mean_quatriemezone = mean(quatriemezone$peche)
table$L_quatrezones = rep(0, nrow(table))
table[which(table$x < xmed & table$y < ymed), "L_quatrezones"] = mean_premierezone
table[which(table$x < xmed & table$y >= ymed), "L_quatrezones"] = mean_deuxiemezone
table[which(table$x >= xmed & table$y < ymed), "L_quatrezones"] = mean_troisiemezone
table[which(table$x >= xmed & table$y >= ymed), "L_quatrezones"] = mean_quatriemezone
hist(table$L_quatrezones, breaks = 100)
ggplot(table) + geom_point(aes(x=x, y=y, col=as.factor(L_quatrezones))) + labs(color= "Quantite pechee")
```

### 2020-12-10 Scripts_groupe_M2/Latent_field_and_commercial_data.r

@@ -1,0 +1,173 @@
+###SIMULATIONS
+
+##IMPORTATION DES PACKAGES
+
+library(spatstat)
+library(tidyverse)
+library(ggplot2)
+
+## DEFINITION DES 4 SOUS ZONES
+
+#On d√©finit 4 sous zones dans la zone d'√©tude, chaque sous zone correspondant
+#√† un type de sol : par exemple sable, boue, zoche et gravier
+#on attribue une valeur fix√©e (1,2,3 ou 4) a chacune des 4 modalit√©s
+#la fonction zonage prend pour argument les coordonn√©es x et y d'un point de la grille
+zonage <- function(x,y, xmed = 5, ymed = 5){
+  ifelse(x<xmed, 1, 2) + ifelse(y<ymed, 0, 2) 
+}
+
+
+## CALCUL DU CHAMP LATENT
+
+#dans le champ latent, on ne prend pas en compte l'effet spatial al√©atoire,
+#et on considere qu'une seule covariable (type de sol)
+alpha_s <- -0.5 #on fixe l'intercept √† -0,5
+betas_s <- c(0, 1, -1, 2) #on fixe les beta correspondant a chacune des modalit√©s de la
+#covariable type de sol
+#la fonction du champ latent prend pour argument les coordonn√©es x et y d'un point
+#de la grille, les valeurs fix√©es de alpha_s et des beta_s et renvoie la valeur du 
+#champ latent en ce point
+latent_field_function = function(x,y, alpha_s, betas_s){
+  exp(alpha_s +  betas_s[zonage(x,y)])
+}
+
+## REPRESENTATION GRAPHIQUE DU CHAMP LATENT DANS LA GRILLE
+
+dta <- data.frame(x = runif(n = 100, 0,10 ), y = runif(100, 0, 10)) %>% 
+  mutate(z= latent_field_function(x, y, alpha_s, betas_s))
+ggplot(data= dta) + geom_point(aes(x=x, y=y, col = z))
+
+
+## CALCUL DE L'INTENSITE LAMBDA
+
+alpha_x <- 1
+b <- 1  # b peut aussi valoir 0 ou 3
+
+#fonction de calcul du lambda
+lambda_function <- function(x, y, alpha_x, b, alpha_s, betas_s){
+  lambda <- exp(alpha_x + b * log(latent_field_function(x,y, alpha_s, betas_s)))
+}
+
+## GENERATION DES POINTS DE PECHE
+
+# on utilise la fonction rpoispp pour g√©n√©rer les points de peche
[…]

### 2020-12-11 Scripts/function/commercial_scientific_14.R
RANGEMENT du dossier

@@ -1,357 +1,0 @@
-########################################################
-## Simulation loop function (simu_commercial_scientific)
-########################################################
-
-#' @title simu_commercial_scientific()
-#' 
-#' @param Results dataframe summarising simulation/estimation
-#' @param simu_file simulation file to save 'Results' and outputs
-#' 
-#' ## Simulations scenarios
-#' # Latent field
-#' @param grid_dim grid dimension
-#' @param n_cells cell number
-#' @param beta0 intercept of the abundance distribution equation
-#' @param beta  fixed parameters of the abundance distribution equation
-#' @param range range parameter. Must be provided for the "MaternCov" package.
-#' @param nu nu parameter. Must be provided for the "MaternCov" package.
-#' @param SD_x Standard deviation for the simulation of covariate related to the abundance distribution equation.
-#' 
-#' 
-#' # Observations
-#' # Scientific
-#' @param n_samp_sci number number of scientific samples
-#' @param logSigma_sci observation error for the scientific data (log scale)
-#' @param q1_sci parameter related to the zero-inflation for scientific data
-#' @param q2_sci relative catchability of the scientific data
-#' # Commercial
-#' @param n_samp_com number of commercial samples
-#' @param logSigma_com observation error for the scientific data (log scale)
-#' @param q1_com parameter related to the zero-inflation for commercial data
-#' @param q2_com relative catchability of the commercial data
-#' @param b_set preferential sampling parameter levels
-#' 
-## Model configuration
-#' @param Data_source Alternative for the estimation model ("scientific_commercial" : both data are fitted to the model, "scientific_only" : only scientific data is fitted to the model, "commercial_only" : only commercial data is fitted to the model)
-#' @param Samp_process If 1 : the sampling process contribute to the likelihood, else it doesn't
-#' @param EM if "est_b" : b is estimated, if "fix_b" : b is fixed to 0
-#' 
-#' # TMB model version = "com_data"
-#' @param Version estimation model version
-#' @param TmbFile file for the C++ template
-#' @param ignore.uncertainty if TRUE, ignore uncertainty in estimation
-#' 
-#' ## Loop indices
-#' @param counter index for the dataframe
-#' @param i simulation index
-#' @param n_sim ending iteration
-#' 
-#' @return Results : summary dataframe of simulation/estimation
-#' @return List_param : List of model outputs
-#' @return counter : index to browse Results and List_param
-#' 
[…]

### 2020-12-11 Version_Baptiste/Scripts/function/commercial_scientific_14.R

@@ -11,57 +11,30 @@
 #' # Latent field
 #' @param grid_dim grid dimension
 #' @param n_cells cell number
-#' @param latent_fields_simu if == "simulate" : simulate data, if == "load_from_RData" : load already simulated data, to have the same data between different simulation loops (here `latent_field` and `scientific_data`)
-#' @param latent_field already simulated data for the abundance distribution equation
-#' @param scientific_data already simulated data for the scientific data
 #' @param beta0 intercept of the abundance distribution equation
 #' @param beta  fixed parameters of the abundance distribution equation
-#' @param simu_tool simulation functions ("RandomFields" or "MaternCov")
-#' @param SpatialScale spatial scale parameter. Must be provided for the `RandomFields` package
 #' @param range range parameter. Must be provided for the "MaternCov" package.
 #' @param nu nu parameter. Must be provided for the "MaternCov" package.
 #' @param SD_x Standard deviation for the simulation of covariate related to the abundance distribution equation.
-#' @param SD_delta Standard deviation for the random effect delta (abundance distribution).
-#' @param SD_xfb Standard deviation for the simulation of covariate related to the sampling process.
-#' @param SD_eta Standard deviation for the random effect eta (sampling process).
 #' 
-#' 
-#' # Commercial sampling
-#' @param beta0_fb intercept of the sampling process
-#' @param beta_fb fixed parameters of the sampling process
 #' 
 #' # Observations
-#' @param DataObs Observation model (1 : zero-inflated gamma, 2 : zero-inflated lognormal, 3 : lognormal)
-#' @param zero.infl_model Zero-inflated model parametrisation
 #' # Scientific
 #' @param n_samp_sci number number of scientific samples
 #' @param logSigma_sci observation error for the scientific data (log scale)
 #' @param q1_sci parameter related to the zero-inflation for scientific data
 #' @param q2_sci relative catchability of the scientific data
-#' @param sci_sampling type of sampling plan for scientific data ("random_stratified";"fixed")
-#' @param n_strate parameter for the size of the strata (n=9)
 #' # Commercial
 #' @param n_samp_com number of commercial samples
 #' @param logSigma_com observation error for the scientific data (log scale)
 #' @param q1_com parameter related to the zero-inflation for commercial data
 #' @param q2_com relative catchability of the commercial data
 #' @param b_set preferential sampling parameter levels
[…]

### 2020-12-11 Version_Baptiste/Scripts/function/fit_IM.R

@@ -189,7 +189,8 @@ fit_IM <- function(Estimation_model_i = 1,
   #Upper = Inf
   Lower = -50  #getting a few cases where -Inf,Inf bounds result in nlminb failure (NaN gradient)
   Upper = 50
-  Opt = nlminb( start=Obj$par, objective=Obj$fn, gradient=Obj$gr, lower=Lower, upper=Upper, control=list(trace=1, maxit=1000))         #
+  Opt = nlminb( start=Obj$par, objective=Obj$fn, gradient=Obj$gr, lower=Lower, upper=Upper,
+                control=list(trace=1, maxit=1000))         #
   Opt[["diagnostics"]] = data.frame( "Param"=names(Obj$par), "Lower"=-Inf, "Est"=Opt$par, "Upper"=Inf, "gradient"=Obj$gr(Opt$par) )
   Report = Obj$report()
   

### 2020-12-11 Version_Baptiste/Scripts/simu_main_script.R

@@ -10,7 +10,7 @@ source("Scripts/function/load_packages.R")
 #-----------------------------------------------------------
 
 # Simulation name
-simu_name = "Lognormal_0Effect"
+simu_name = "test"
 
 #---------------------
 # Simulation scenarios
@@ -26,23 +26,17 @@ grid_side_y <- 25
 grid_dim = c("x"=grid_side_x, "y"=grid_side_y)
 n_cells = grid_dim[1]*grid_dim[2]
 
-# latent field simulation
-latent_fields_simu = "load_from_RData" # "load_from_RData" or "simulate"
-load("Results/Simu/simulated_data/scientific_data.RData")
-load("Results/Simu/simulated_data/latent_field.RData")
-
-# Intercepts and covariates
+# Intercepts and covariates of the latent field
 beta0 = 2 # intercept
 beta = c(0,0,0,0,0) # covariates
 # parameters covariates vector : 1st param : continuous variable, 4 nexts : strata_1, strata_2, strata_3, strata_4
 
-# Correlation structure of the latent fied
+# Correlation structure of the covariate
 simu_tool <- "MaternCov" # 2 way to simulate latent fields : "RandomFields", "MaternCov"
-SpatialScale = sqrt(prod(grid_dim))/5  # Range ~ 2*Scale  --> RandomField
 range = sqrt(prod(grid_dim))/(5)*2 # --> MaternCov
 nu = 1 
-SD_eta = SD_x = SD_xfb = SD_delta = 1
-
+SD_x = 0.5
+SD_delta = SD_eta= 0.5
 ###
 ## Commercial sampling process
 ###
@@ -50,37 +44,19 @@ SD_eta = SD_x = SD_xfb = SD_delta = 1
 # intercept
 beta0_fb = 2
 
-# covariate parameter (continuous)
-beta_fb <- 0
-
-# constraint on b 
-# (b = 1 : b is positive, b = 2 : b is free)
-b_constraint <- 2
-
 ###
 ## Observation process
 ###
 
-## Model for the observation process (it is the same for simulating scientific and commercial data and in estimations)
-DataObs <- 2 # 1 : zeroinflated gamma, 2 : zeroinflated lognormal, 3 : lognormal
-zero.infl_model <- 2
-# 0 : plogis( q1_com(0) + q1_com(1)*log(exp_catch))
[…]

### 2020-12-11 Version_M2/Scripts/In_progress/plot_simu_bis.R
Rangement du dossier

@@ -1,0 +1,166 @@
+
+
+
+# Results$n_samp_com
+# Results$nsci_ncom <- 0
+# Results$nsci_ncom[which(Results$n_samp_sci == 0 | Results$n_samp_com == 200)] <- "0_200"
+# Results$nsci_ncom[which(Results$n_samp_sci == 50 | Results$n_samp_com == 150)] <- "50_150"
+# Results$nsci_ncom[which(Results$n_samp_sci == 100 | Results$n_samp_com == 100)] <- "100_100"
+# Results$nsci_ncom[which(Results$n_samp_sci == 150 | Results$n_samp_com == 50)] <- "150_50"
+# Results$nsci_ncom[which(Results$n_samp_sci == 200 | Results$n_samp_com == 0)] <- "200_0"
+
+# Results$nsci_ncom <- factor(Results$nsci_ncom,levels = c("200_0","150_50","100_100","50_150","0_200"))
+Results$Data_source <- factor(Results$Data_source,levels = c("scientific_only","commercial_only","scientific_commercial"))
+
+Results[,"OnBoundary"]=rep(0,nrow(Results))
+Results[which(abs(Results[,"b_est"])>10),"OnBoundary"]=1
+Results[,"Converge"]=Results[,"OnBoundary"]+Results[,"Convergence"]
+#For figures, limit to results for which |b estimate| < 10
+Results[which(Results[,"Converge"]==2),"Converge"]=1
+
+Converge_table = summaryBy(Converge~sci_sampling+Data_source,data=Results,FUN=sum)
+
+# length(Results$Convergence[which(Results$Convergence == 1 & Results$Data_source == "scientific_only")])/length(Results$Convergence[which(Results$Data_source == "scientific_only")])
+
+# Filter simulation which converged
+WhichFailed <- which(Results[,"Convergence"]!=0)
+Results_cvg.failed <- Results[WhichFailed,]
+WhichDone = which(Results[,"Convergence"]==0)
+Results=Results[WhichDone,]
+
+# add relative bias for abundance
+Results[,"RelBias_N"]=(Results[,"N_est"]-Results[,"N_true"])/Results[,"N_true"]
+Results[,"Bias_b"]=Results[,"b_est"]-Results[,"b_true"]
+
+# Handle data to have pretty graph 
+Results %>%
+  filter(Data_source == "scientific_only") -> Results_sc_only
[…]

### 2020-12-11 Version_M2/Scripts/function/commercial_scientific_14.R

@@ -1,0 +1,359 @@
+########################################################
+## Simulation loop function (simu_commercial_scientific)
+########################################################
+
+#' @title simu_commercial_scientific()
+#' 
+#' @param Results dataframe summarising simulation/estimation
+#' @param simu_file simulation file to save 'Results' and outputs
+#' 
+#' ## Simulations scenarios
+#' # Latent field
+#' @param grid_dim grid dimension
+#' @param n_cells cell number
+#' @param beta0 intercept of the abundance distribution equation
+#' @param beta  fixed parameters of the abundance distribution equation
+#' @param range range parameter. Must be provided for the "MaternCov" package.
+#' @param nu nu parameter. Must be provided for the "MaternCov" package.
+#' @param SD_x Standard deviation for the simulation of covariate related to the abundance distribution equation.
+#' 
+#' 
+#' # Observations
+#' # Scientific
+#' @param n_samp_sci number number of scientific samples
+#' @param logSigma_sci observation error for the scientific data (log scale)
+#' @param q1_sci parameter related to the zero-inflation for scientific data
+#' @param q2_sci relative catchability of the scientific data
+#' # Commercial
+#' @param n_samp_com number of commercial samples
+#' @param logSigma_com observation error for the scientific data (log scale)
+#' @param q1_com parameter related to the zero-inflation for commercial data
+#' @param q2_com relative catchability of the commercial data
+#' @param b_set preferential sampling parameter levels
+#' 
+## Model configuration
+#' @param Data_source Alternative for the estimation model ("scientific_commercial" : both data are fitted to the model, "scientific_only" : only scientific data is fitted to the model, "commercial_only" : only commercial data is fitted to the model)
+#' @param Samp_process If 1 : the sampling process contribute to the likelihood, else it doesn't
+#' @param EM if "est_b" : b is estimated, if "fix_b" : b is fixed to 0
+#' 
+#' # TMB model version = "com_data"
+#' @param Version estimation model version
+#' @param TmbFile file for the C++ template
+#' @param ignore.uncertainty if TRUE, ignore uncertainty in estimation
+#' 
+#' ## Loop indices
+#' @param counter index for the dataframe
+#' @param i simulation index
+#' @param n_sim ending iteration
+#' 
+#' @return Results : summary dataframe of simulation/estimation
+#' @return List_param : List of model outputs
+#' @return counter : index to browse Results and List_param
+#' 
[…]

### 2020-12-11 Version_M2/Scripts/function/fit_IM.R

@@ -1,0 +1,256 @@
+############
+## Fit model
+############
+
+#' @title fit_IM()
+#' 
+## Model configuration
+#' @param Estimation_model_i index of the estimation model (scientific only, commercial only, integrated)
+#' @param Samp_process If 1 : the sampling process contribute to the likelihood, else it doesn't
+#' @param EM if "est_b" : b is estimated, if "fix_b" : b is fixed to 0
+#' 
+#' # TMB model version = "com_data"
+#' @param Version estimation model version
+#' @param TmbFile file for the C++ template
+#' @param ignore.uncertainty if TRUE, ignore uncertainty in estimation
+#' 
+#' # Data/model Inputs
+#' @param c_com_x number of sample in each cell (line) for each fleet (column)
+#' @param y_com_i catch data
+#' @param index_com_i sampled cell for commercial observation
+#' 
+#' @param y_sci_i scientific observation
+#' @param index_sci_i sampled cell for scientific observation
+#' 
+#' @param Cov_x covariate for species distribution
+#' 
+#' @return SD 
+#' @return Opt 
+#' @return Report
+#' @return Converge
+#' @return Data
+#' @return Params
+#' @return Map
+
+fit_IM <- function(Estimation_model_i = 1,
+                   Samp_process = 1,
+                   EM = "est_b",
+                   TmbFile,
+                   ignore.uncertainty,
+                   c_com_x,
+                   y_com_i,
+                   index_com_i,
+                   y_sci_i,
+                   index_sci_i,
+                   Cov_x){
+
+  #----------------------
+  ## Shape models outputs
+  #----------------------
+  # Option_vec : model configuration
+  # Data : list of data (catch data, covariates, Option_vec, spde objects, etc.)
+  # Params : list of parameters
+  # Map : parameters that are fixed to a specific value
+  
+  Options_vec = c( 'Prior'=0, # (DEPRECATED)
+                   'Alpha'=2, # (DEPRECATED)
+                   'IncludeDelta'=1, # (DEPRECATED)
+                   'IncludeEta'=1, # (DEPRECATED)
[…]

### 2020-12-11 Version_M2/Scripts/function/plot_simu.R

@@ -1,0 +1,295 @@
+###############
+## Plot_Results
+###############
+#' @title Plot_Results()
+#' 
+#' @param Results summary dataframe of the simulation/estimation loop
+#' @param b_set levels of preferential sampling
+#' 
+#' @return RelBias_N : Plot of the relative bias of abundance
+#' @return Bias_b : Plot of the relative bias of b
+#' @return MSPE_S : Plot of the mean squared prediction error
+#' @return Converge_table : convergence table
+#' 
+
+Plot_Results <- function(Results,b_set){
+  
+  # Identify simulations/estimations that have failed
+  Results[,"OnBoundary"]=rep(0,nrow(Results))
+  Results[which(abs(Results[,"b_est"])>10),"OnBoundary"]=1
+  Results[,"Converge"]=Results[,"OnBoundary"]+Results[,"Convergence"]
+  
+  # For figures, limit to results for which |b estimate| < 10
+  Results[which(Results[,"Converge"]==2),"Converge"]=1
+
+  # Convergence table
+  Converge_table = summaryBy(Converge~b_true+Data_source,data=Results,FUN=sum)
+
+  # Filter simulation which converged
+  WhichFailed <- which(Results[,"Convergence"]!=0)
+  Results_cvg.failed <- Results[WhichFailed,]
+  WhichDone = which(Results[,"Convergence"]==0)
+  Results=Results[WhichDone,]
+  
+  # add relative bias for abundance
+  Results[,"RelBias_N"]=(Results[,"N_est"]-Results[,"N_true"])/Results[,"N_true"]
+  Results[,"Bias_b"]=(Results[,"b_est"]-Results[,"b_true"]) / ifelse(Results[,"b_true"] != 0,Results[,"b_true"],1)
+  # plot estimates of 'b'
+  Which_plot = which(Results[,"type_b"] == "est_b")
+
+  # Handle data to have pretty graph (a bit dirty...)
+  Results %>%
+    filter(Data_source == "scientific_only") -> Results_sc_only
+  Results_sc_only <- Results_sc_only[which(is.na(Results_sc_only$b_true)),]
+  Results %>%
+    filter(Data_source != "scientific_only") -> Results_com
+  
+  Results_sc_only$b_true <- b_set[1]
[…]

### 2020-12-11 Version_M2/Scripts/function/simu_latent_field.R

@@ -1,0 +1,99 @@
+#####################################
+## Simulate or load latent field data 
+#####################################
+
+#' @title simu_latent_field()
+#' 
+#' @param loc_x dataframe with grid coordinate and strata
+#' @param latent_fields_simu if == "simulate" : simulate data, if == "load_from_RData" : load already simulated data, to have the same data between different simulation loops (here `latent_field` and `scientific_data`)
+#' @param latent_field already simulated data for the abundance distribution equation
+#' @param beta0 intercept of the abundance distribution equation
+#' @param beta  fixed parameters of the abundance distribution equation
+#' @param simu_tool simulation functions ("RandomFields" or "MaternCov")
+#' @param SpatialScale spatial scale parameter. Must be provided for the `RandomFields` package
+#' @param range range parameter. Must be provided for the "MaternCov" package.
+#' @param nu nu parameter. Must be provided for the "MaternCov" package.
+#' @param SD_x Standard deviation for the simulation of covariate related to the abundance distribution equation.
+#' @param SD_delta Standard deviation for the random effect delta (abundance distribution).
+#' @param SD_xfb Standard deviation for the simulation of covariate related to the sampling process.
+#' @param SD_eta Standard devioation for the random effect eta (sampling process).
+#' 
+#' @return Cov_x : design matrix (matrix for covariates levles / values)
+#' @return Strue_x Latent field values
+#' @return beta : fixed parameters for the abundance distribution equation
+#' @return delta_x : random effect for the abundance distribution equation
+#' @return xfb_x : fixed parameters for the sampling process equation
+#' @return eta_x : random effect for the sampling process equation
+#' 
+#' @author B. Alglave
+
+# for example : run main_script and beginning of commercial_scientific_14
+
+simu_latent_field <- function(loc_x,
+                              latent_fields_simu,
+                              latent_field,
+                              scientific_data,
+                              beta0,
+                              beta,
+                              simu_tool,
+                              range,
[…]

### 2020-12-11 Version_M2/Scripts/function/simu_scientific_data.R

@@ -1,0 +1,111 @@
+###################################
+## Simulate or load scientific data 
+###################################
+
+#' @title simu_scientific_data()
+#' 
+#' @param loc_x dataframe with grid coordinate and strata
+#' @param grid_dim grid dimension
+#' @param Strue_x spatial abundance data
+#' @param scientific_data already simulated data for the scientific data
+#' @param DataObs Observation model (1 : zero-inflated gamma, 2 : zero-inflated lognormal, 3 : lognormal)
+#' @param zero.infl_model Zero-inflated model parametrisation
+#' @param n_samp_sci number number of scientific samples
+#' @param logSigma_sci observation error for the scientific data (log scale)
+#' @param q1_sci parameter related to the zero-inflation for scientific data
+#' @param q2_sci relative catchability of the scientific data
+#' @param sci_sampling type of sampling plan for scientific data ("random_stratified";"fixed")
+#' @param n_strate parameter for the size of the strata (n=9)
+#' 
+#' 
+#' @return index_sci_i : cells sampled by the scientific survey
+#' @return c_sci_x : vector filled with 0/1. If 1 : the cell has been sampled. If 0 :  the cell has not been sampled. 
+#' @return y_sci_i : scientific observations 
+#' 
+#' @author B. Alglave
+
+
+simu_scientific_data <- function(loc_x,
+                                 grid_dim,
+                                 Strue_x,
+                                 zero.infl_model,
+                                 n_samp_sci,
+                                 logSigma_sci,
+                                 q1_sci,
+                                 q2_sci,
+                                 n_strate,
+                                 scientific_data){
+
+  index_sci_i <- c()
+  
+  # ## Uniform sampling
+  # index_sci_i <- sample(loc_x$cell,size=n_samp_sci)
+
+  ## Random stratified sampling
+  loc_x %>%
+    tidyr::pivot_longer(cols = starts_with("strata"),names_to = "strata") %>%
+    data.frame() %>% filter(value > 0) -> loc_x_2
+  # nb of samples in each strata
+
[…]

### 2020-12-11 Version_M2/Scripts/simu_main_script.R

@@ -1,0 +1,245 @@
+####################
+## Simulation script
+####################
+# B. Alglave based on Conn et al. (2017)
+
+source("Scripts/function/load_packages.R")
+
+#-----------------------------------------------------------
+#-------------- Simulation/estimation settings -------------
+#-----------------------------------------------------------
+
+# Simulation name
+simu_name = "test"
+
+#---------------------
+# Simulation scenarios
+#---------------------
+
+###
+## Latent field
+###
+
+# Grid dimension
+grid_side_x <- 25
+grid_side_y <- 25
+grid_dim = c("x"=grid_side_x, "y"=grid_side_y)
+n_cells = grid_dim[1]*grid_dim[2]
+
+# Intercepts and covariates of the latent field
+beta0 = 2 # intercept
+beta = c(0,0,0,0,0) # covariates
+# parameters covariates vector : 1st param : continuous variable, 4 nexts : strata_1, strata_2, strata_3, strata_4
+
+# Correlation structure of the covariate
+simu_tool <- "MaternCov" # 2 way to simulate latent fields : "RandomFields", "MaternCov"
+range = sqrt(prod(grid_dim))/(5)*2 # --> MaternCov
+nu = 1 
+SD_x = 0.5
+SD_delta = SD_eta= 0.5
+###
+## Commercial sampling process
+###
+
+# intercept
+beta0_fb = 2
+
+###
+## Observation process
+###
+
+## Scientific data
+# Number of scientific data
+n_samp_sci = 50
+# observation error 
+logSigma_sci = log(1)
+# zero-inflation parameter
+q1_sci <- 1
+# Relative catchability
+q2_sci <- 1
+# parameter controling size of the strata
+n_strate <- 9 
+
+## Commercial data
+# Number of commercial samples
+n_samp_com = 3000
+# observation error 
+logSigma_com = log(1)
+# zero-inflation parameter
+q1_com <- 1
+# Relative catchability
+q2_com <- 1
+# Levels of preferential sampling
+b_set = c(0,1,3) 
+
+
+
+#--------------------
+# Model configuration
+#--------------------
+
+# Data sources feeding the model
+Data_source = c("scientific_commercial","scientific_only","commercial_only")
+
+# b estimated or b fixed to 0 in estimation
+EM_set = c("fix_b","est_b")
+EM = EM_set[2]
+
+# if 1, sampling process is accounted for in estimation
+Samp_process = 1
+
+# TMB model version
[…]

### 2021-01-11 Notes/Rmarkdown-Synthese.rmd
Ajout RMD synthese

---
title: "Synthèse : Effet de l'allocation proportionnelle des captures de soles"
author: "Oceane GUITTON, Chloe TELLIER, Juliette THEOLEYRE"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***
# Contexte

Dans sa thèse Baptiste développe un modèle hiérarchique qui couple des données de captures de soles issues de protocole d’échantillonnage scientifique (protocole standardisé mais couteux et limités dans le temps) et des données de captures commerciales (peu chères, mais qui souffrent d’un biais d’échantillonnage) pour estimer l’abondance de la sole dans le golfe de Gascogne. Son modèle est nourri par des données de capture géoréférencées.  
A terme, l'objectif de la thèse de Baptiste est de décrire la dynamique spatio-temporel des poissons (ici cas appliqué aux soles) à petite échelle et identifier leurs zones fonctionnelles.

***
# Problématique

En pratique les données de captures commerciales ne sont pas géoréférencées. La capture débarquée au port est connue ainsi que la trajectoire des bateaux (données réglementaire de positionnement des navires de pêche). D’après la trajectoire des bateaux, des positions du bateau associés à la pêche sont identifiées et la capture totale est répartie équitablement sur chaque point de pêche. C'est ce qu'on appellera par la suite l'uniformisation des pêches.

Dans le modèle que Baptiste utilise, on connait les quantités exactes pêchées à chaque point de pêche. Or, dans la réalité on ne dispose pas de ces valeurs exactes mais seulement d'une valeur moyenne qui est attribuée à chaque point de pêche pour un même bateau.

Se pose alors la problématique suivante : *" Est-ce que l'uniformisation des captures (par carré statistique/ par bateau) amène un biais à la simulation et à l'estimation des paramètres ? "*  
On peut aussi la poser autrement : *" La prise en compte de l'échantillonnage préférentiel dans les données commerciales couplée aux informations standardisées des données scientifiques permettent ils de compenser le lissage qu'entraine l'uniformisation des points de pêche ? "*

***
# Modèle et hypotèses de Baptise

La démarche de Baptiste est la suivante :  
Il utilise un modèle intégré pour déduire la distribution spatiale des poissons en combinant à la fois des données scientifiques et commerciales tout en tenant compte de l'échantillonnage préférentiel induit par le comportement ciblé des pêcheurs.

(Modèle intégré : Les inférences sont dérivées dans un cadre classique de maximum de vraisemblance.)

(Afin d’explorer les bénéfices de ce modèle : nous évaluons la performance de la méthode d'estimation basée sur des expériences de simulation-estimation.)

Le modèle hiérarchique intégré de Baptiste est contitué de 3 composantes :  
  * Un champs latent S correspondant à la densité/biomasse de poisson.  
  * Un processus de Poisson non-homogène qui permet de prendre en compte le biais relatif à l'échantillonnage préférentiel.  
  * Des valeurs latentes qui conditionnent les processus d'observation des données scientifiques et commerciales. (A revoir niveau formulation)

## Champ Latent

La densité de poissons 𝑆 est modélisée par un **champ spatial latent gaussien logarithmique**. La valeur du champ latent est exprimée par une combinaison linéaire entre un intercept, des co-variables environnementales associées/pondérées? à des paramètres de relation espèce-habitat et un effet aléatoire spatial.

## Processus ponctuel de Poisson non-homogène : génération des points de pêche.

Hypothèses :   
* Les bateaux d'une même flotte présentent un comportement homogène et partagent donc les mêmes paramètres (notamment de capturabilité).    
* Le processus d’échantillonnage des données scientifiques est considéré indépendant du champ latent (non préférentiel).    
* L’échantillonnage des données commerciales est préférentiel, et les emplacements de pêche sont conditionnés à l’abondance S.

Les points de pêches des bateaux sont modélisés par un **processus ponctuel de Poisson non-homogène** avec une intensité qui dépend de l'abondance du champ latent a cet endroit (S). Elle est aussi associée à des co-variables et à un effet spatial résiduel.  
On définit de plus une métrique sans dimension permettant la quantification du degrès d'échantillonnage préférentiel (b).

## Le modèle Zero-Inflated : attribution des quantités poissons pêchés en chaque point de pêche.

Dans le cadre du modèle de Baptiste, chaque point de pêche se voit attribuer une quantité exact de poissons pêchés.

On determine  dans un premier temps quels sont les points ou la peche a ete 'sans succes' et ceux ou elle s'est revelee etre fructueuse. Pour cela, on fait appel a un **modele zero-inflated** et on modelise la probabilite de presence/absence de capture par une **loi de Bernoulli**.

Puis on attribue a chaque point de peche 'fructueux' la quantite de poisson pechee. Pour cela on modelise une **loi lognormale** et chaque point de pêche fructueux se voit attribuer une quantité de poisson pêchés.

On obtient alors les données sous la forme : quantité de poisson exactement capturés en chaque point de pêche.

## Estimation des paramètres.

Une fois ces données de captures simulées, il ne suffit plus qu'à les rentrer dans le modèle de Baptiste pour ensuite comparer les paramètres estimés à ceux utilisés pour générer les données.

RQ : est ce qu'on doit parler des données VMS et dire qu'on ne connait pas exactement les coordonnées des points de pêche mais on peut les estimer à partir des données VMS ?

***
# Notre objectif au sein de ce projet

Cependant, les données commerciales ne sont pas rendues sous cette forme. En effet, chaque bateau ne déclare qu'une quantité totale de poissons capturés au cours d'une même journée et non en chaque point de pêche.

De la même manière que Baptiste, on va chercher à simuler un jeu de données commerciales, mais cette fois la capture totale de chaque bateau est repartie de facon uniforme entre tous ses points de peche. Cela implique que les points de pêche visités par un même bateau ont une même valeur de quantité de poissons pêchés.

# Simplification

Afin d'aborder le problème sous un angle simplifié, on retire du modèle hiérarchique intégré tous les effets spatialisés. La ressource est uniquement guidée par des co-variables (sol, teneur en chlorophylle, etc.) et on conserve l’échantillonnage préférentiel.

L'étude se concentre dans un premier temps sur un seul carré statistique. En effet ce qui se passes à l’échelle d’un carré pourra être retranscrit à l’echelle de plusieurs.

On fixe notre flotte à 10 bateaux executant en une journée 3000 coups de pêche au total. Tous les bateaux de la flotte présentent les mêmes caractéristiques : le paramètre de capturabilité est le même pour toute la flotte.

# Notre modèle

## Génération du champ latent S

Par simplification par rapport au modèle complet de Baptiste, on ne considère pour la génération du champ latent que les covariables environnementales. La valeur du champ latent représente alors la densité de poissons au point x.

## Génération des points de pêche par le processus ponctuel de Poisson non-homogène

De premier abord, nous nous sommes intéressées au processus d'echantillonage, c'est a dire a la generation des points de peche (et de leur distribution spatiale) selon un **processus de poisson non homogene**. L'intensite des points de peche dans une zone donnee dépendant de l'abondance du champ latent a cet endroit.

Or, au fil de notre appréhension du modèle, on a pu relever un manque de cohérence spatial dans les pêches des différents bateaux. On cherche alors à palier se problème en imaginant une nouvelle stratégie de répartition des points de pêches et des bateaux qui auquels ils sont associés.

On tente alors de rétablir une cohérence en cherchant à tracer le trajet du bateau cible, ou du moins les zones dans lesquels celui-ci va exécuter ses coups de pêche. Ces zones sont délimitées en prenant en compte la force d'attraction de la biomasse (densité de soles) comme avec le premier processus d'échantillonnage.
Cette stratégie va alors contraindre les points de pêches à être localisés dans les zones où les bateaux sont passés.

**La stratégie se révèle être la suivante**

1. Pour chaque bateau, on commence par tirer un point au hasard au moyen d’un **processus ponctuel de Poisson non-homogène**. Ce point va alors constituer le centre de la zone de pêche du bateau i pour le jour j.

Remarque : Ce point tiré à l’aide d’un processus ponctuel a plus de chance de se retrouver dans une zone où la densité en soles est grande.

2. L’amplitude de la zone de pêche du bateau est ensuite délimitée par un rayon (reste à déterminer ce rayon).  
On définit alors un cercle de centre I et de rayon R correspondant à la zone de pêche du bateau i pour le jour j.

3. Une fois la zone de pêche délimitée, on va tirer les coups de pêche de bateau cible au sein même de cette zone.  
Les points de pêches, sont comme dans le modèle précédent tirés selon le **processus ponctuel de Poisson non homogène**.

Remarque : Pour l’instant, on considère que chacun des bateaux execute le même nombre de coups de pêche. Ainsi, sur nos 3000 points de pêche total souhaités, et avec une flotte de 10 bateaux, cela nous ramène à 300 points de pêche par bateau et par jour.

Dans un deuxième temps, on pourra réfléchir à ne pas forcer chaque bateau à avoir autant de point de pêche que son voisin. En effet, cela permettrait d’attribuer aux bateaux pêchant dans des zones moins denses en soles d’effectuer moins de coups de pêches, attestant de leur pêche non fructueuse. On gagnerai ainsi en authenticité.

## Le modèle Zero-Inflated : réallocation uniforme des quantités poissons pêchés en chaque point de pêche

Le **modèle zero-inflated** nous permet toujours de déterminer la nature des coups de pêche (raté ou fructueux) à l'aide d'une **loi de Bernoulli**.   
Comme dans le modèle de Baptiste, on commence par attribuer une quantité de poissons capturé par point de pêche à l'aide d'une loi lognormale.   
On somme ensuite les quantités de poissons en chaque points de pêche d'un même bateau que l'on divise par le nombre de point de pêche du bateau (pour l'instant 300 point de pêche par bateau). Puis, on réalloue à chacun des points de pêche du bateau cette moyenne. On simule ainsi les données commerciales telles quelles sont obtenues, i.e. une quantité totale de poissons capturés par bateau et par jour. Il s'agit de l'uniformisation de nos points de pêche.  
On se retrouve alors avec autant de quantité différentes qu'il y a de bateau (soit 10 selon notre initialisation actuelle).

## Limites du modèle : le rôle du nombre de bateaux

On a remarqué que le “niveau d’uniformisation” pouvait être influencé par le nombre de bateaux au sein de la flotte, soit le nombre de captures totales différentes déclarées par jour. En effet, chaque bateau déclare son nombre total de capture pour la journée, au plus il y a de bateau, au plus il y a de totaux différents et donc au plus l’uniformisation serait ‘smooth’.

On pourrait alors formuler la problématique suivante : *“Au maximum, combien de pings VMS peut on uniformiser pour un même bateau afin que les estimations des paramètres du modèles restent plausibles/efficaces ?”*

Pour répondre à cette problématique, il nous suffirait de faire varier le nombre de bateau dans notre modèle ci-dessus. De ce fait le nombre de pings associé à chaque bateau suivrait une évolution inversement proportionnelle au nombre de bateau.

## Estimation des paramètres

Une fois ces données de captures simulées, il ne suffit plus qu'à les rentrer dans le modèle simplifié pour ensuite comparer les paramètres estimés à ceux utilisés pour générer les données.

# Résultats : les graphiques obtenus ?

### 2021-01-11 Version_M2_10bateaux/Scripts/function/commercial_scientific_14.R

@@ -269,15 +269,15 @@ simu_commercial_scientific <- function(Results,
       labs(color= "Bateau") +
       theme_bw() +
       theme(axis.title = element_blank(),legend.position='none')
-    if (b==0){
+    if (b==b_set[1]){
       plot_pointsdepechecell <- plot_pointsdepechecell +
-        facet_wrap(~"b = 0")
-    }else if (b==1){
+        facet_wrap(~paste0("b = ", b_set[1]))
+    }else if (b==b_set[2]){
       plot_pointsdepechecell <- plot_pointsdepechecell +
-        facet_wrap(~"b = 1")
+        facet_wrap(~paste0("b = ", b_set[2]))
     } else {
       plot_pointsdepechecell <- plot_pointsdepechecell +
-        facet_wrap(~"b = 3")
+        facet_wrap(~paste0("b = ", b_set[3]))
     }
     
     if (k==0){
@@ -287,39 +287,39 @@ simu_commercial_scientific <- function(Results,
         scale_color_gradient2(midpoint = mean(pointsdepeche_com$y_com_i), low = "#E6F2FC", mid = "#62B4FC",
                               high = "#02182C", space = "Lab" ) + 
         theme(axis.title = element_blank(),legend.position='none')
-      if (b==0){
+      if (b==b_set[1]){
         plot_pointsdepecheqte <- plot_pointsdepecheqte +
-          facet_wrap(~"b = 0")
-      }else if (b==1){
+          facet_wrap(~paste0("b = ", b_set[1]))
+      }else if (b==b_set[2]){
         plot_pointsdepecheqte <- plot_pointsdepecheqte +
-          facet_wrap(~"b = 1")
+          facet_wrap(~paste0("b = ", b_set[1]))
       } else {
         plot_pointsdepecheqte <- plot_pointsdepecheqte +
-          facet_wrap(~"b = 3")
+          facet_wrap(~paste0("b = ", b_set[1]))
       }
         
     } else{
       plot_pointsdepecheqte = ggplot(pointsdepeche_com) + geom_point(aes(x=x, y=y, col=as.factor(round(y_com_i, 1))), size=4) +
         labs(color= "Quantite pechee") + theme_bw() +
         theme(axis.title = element_blank(),legend.position='none')
-      if (b==0){
+      if (b==b_set[1]){
         plot_pointsdepecheqte <- plot_pointsdepecheqte +
[…]

### 2021-01-11 Version_M2_10bateaux/Scripts/function/simu_commercial_data.R

@@ -81,20 +81,20 @@ simu_commercial_data <- function(loc_x,
   # X correspond donc aux coordonnées des 10 "points centraux" des zones de peche de chaque bateau
   
   # Visualisation des 10 centres des zones de peche de chaque bateau
-  centres = as.data.frame(cbind(x = X$x, y = X$y, boats = seq(1, 10)))
+  centres = as.data.frame(cbind(x = X$x, y = X$y, boats = seq(1, nboat)))
   plot_centres = ggplot(centres) + geom_point(aes(x=x, y=y, col=as.factor(boats)), size=4) +
     labs(color= "Bateau") +
     theme_bw() +
     theme(axis.title = element_blank(),legend.position='none')
-  if (b==0){
+  if (b==b_set[1]){
     plot_centres <- plot_centres +
-      facet_wrap(~"b = 0")
-  }else if (b==1){
+      facet_wrap(~paste0("b = ", b_set[1]))
+  }else if (b==b_set[2]){
     plot_centres <- plot_centres +
-      facet_wrap(~"b = 1")
+      facet_wrap(~paste0("b = ", b_set[2]))
   } else {
     plot_centres <- plot_centres +
-      facet_wrap(~"b = 3")
+      facet_wrap(~paste0("b = ", b_set[3]))
   }
 
   
@@ -137,15 +137,15 @@ simu_commercial_data <- function(loc_x,
     labs(color= "Bateau") +
     theme_bw() + 
     theme(axis.title = element_blank(),legend.position='none')
-  if (b==0){
+  if (b==b_set[1]){
     plot_pointsdepechecomperboat <- plot_pointsdepechecomperboat +
-      facet_wrap(~"b = 0")
-  }else if (b==1){
+      facet_wrap(~paste0("b = ", b_set[1]))
+  }else if (b==b_set[2]){
     plot_pointsdepechecomperboat <- plot_pointsdepechecomperboat +
-      facet_wrap(~"b = 1")
+      facet_wrap(~paste0("b = ", b_set[2]))
   } else {
     plot_pointsdepechecomperboat <- plot_pointsdepechecomperboat +
-      facet_wrap(~"b = 3")
+      facet_wrap(~paste0("b = ", b_set[3]))
   }
   
   # # Visualisation des points de peche du premier bateau dans sa zone

### 2021-01-11 Version_M2_10bateaux/Scripts/simu_main_script.R

@@ -10,7 +10,7 @@ source("Scripts/function/load_packages.R")
 #-----------------------------------------------------------
 
 # Simulation name
-simu_name = "Bateaux10_Repets100_b013_k01"
+simu_name = "Test2"
 
 #---------------------
 # Simulation scenarios
@@ -81,7 +81,7 @@ q1_com <- 1
 # Relative catchability
 q2_com <- 1
 # Levels of preferential sampling
-b_set = c(0,1,3) 
+b_set = c(0,1,3)
 
 
 
@@ -111,7 +111,7 @@ counter <- 1
 # index of the first iteration
 i0 <- 1
 # Number of simulation
-n_sim = 100
+n_sim = 1
 
 RandomSeed = 123456
 

### 2021-01-11 Version_M2_10bateaux/Scripts/function/commercial_reallocation_uniforme.R
Tout marche pour differentes sequences de peche, differentes zones de peche par sequence de peche

@@ -5,7 +5,8 @@ commercial_reallocation_uniforme = function(k,
                                             n_samp_com,
                                             index_com_i,
                                             loc_x,
-                                            nboat, boats_number){
+                                            sequencesdepeche,
+                                            boats_number){
   
   y_com_i_new = cbind(y_com_i, boats_number)
   y_com_i_new = as.data.frame(y_com_i_new)

### 2021-01-11 Version_M2_10bateaux/Scripts/function/commercial_scientific_14.R

@@ -97,13 +97,13 @@ simu_commercial_scientific <- function(Results,
                                        k,
                                        xlim,
                                        ylim,
-                                       nboat){
+                                       sequencesdepeche){
   #on se fiche de l'argument version ici, on pourrait le supprimer
     
   ################
   ## Simulate data
   ################
-  
+
   set.seed( RandomSeed + i ) # for figures : i = 2
   
   #---------------
@@ -245,7 +245,7 @@ simu_commercial_scientific <- function(Results,
                                                          q1_com,
                                                          q2_com,
                                                          b,
-                                                         nboat)
+                                                         sequencesdepeche)
     
     index_com_i <-  simu_commercial_data_outputs$index_com_i
     y_com_i <- simu_commercial_data_outputs$y_com_i
@@ -258,7 +258,7 @@ simu_commercial_scientific <- function(Results,
     
     if (k>0){ # Si on fait de la reallocation uniforme (cad k = 1)
       y_com_i <- commercial_reallocation_uniforme(k, xlim, ylim, y_com_i, n_samp_com,
-                                                  index_com_i, loc_x, nboat, boats_number)
+                                                  index_com_i, loc_x, sequencesdepeche, boats_number)
     }
     
     
@@ -292,10 +292,10 @@ simu_commercial_scientific <- function(Results,
           facet_wrap(~paste0("b = ", b_set[1]))
       }else if (b==b_set[2]){
         plot_pointsdepecheqte <- plot_pointsdepecheqte +
-          facet_wrap(~paste0("b = ", b_set[1]))
+          facet_wrap(~paste0("b = ", b_set[2]))
       } else {
         plot_pointsdepecheqte <- plot_pointsdepecheqte +
-          facet_wrap(~paste0("b = ", b_set[1]))
+          facet_wrap(~paste0("b = ", b_set[3]))
       }
         
     } else{
@@ -304,13 +304,13 @@ simu_commercial_scientific <- function(Results,
         theme(axis.title = element_blank(),legend.position='none')
       if (b==b_set[1]){
         plot_pointsdepecheqte <- plot_pointsdepecheqte +
-          facet_wrap(~c("b = ", b_set[1]))
+          facet_wrap(~paste0("b = ", b_set[1]))
       }else if (b==b_set[2]){
         plot_pointsdepecheqte <- plot_pointsdepecheqte +
[…]

### 2021-01-11 Version_M2_10bateaux/Scripts/function/plot_simu.R

@@ -38,26 +38,30 @@ Plot_Results <- function(Results,b_set, k){
   Which_plot = which(Results[,"type_b"] == "est_b")
 
   # Handle data to have pretty graph (a bit dirty...)
-  Results %>%
-    filter(Data_source == "scientific_only") -> Results_sc_only
-  Results_sc_only <- Results_sc_only[which(is.na(Results_sc_only$b_true)),]
+  # LA j'ai mis la partie scientifique en commentaire
+  # Results %>%
+  #   filter(Data_source == "scientific_only") -> Results_sc_only
+  # Results_sc_only <- Results_sc_only[which(is.na(Results_sc_only$b_true)),]
   Results %>%
     filter(Data_source != "scientific_only") -> Results_com
   
-  Results_sc_only$b_true <- b_set[1]
-  df_1 <- Results_sc_only
-  Results_sc_only <- do.call(rbind.data.frame, lapply(2:length(b_set),function(i){
-    df_1$b_true <- b_set[i]
-    Results_sc_only <- rbind(Results_sc_only,df_1)
-  }))
+  # Results_sc_only$b_true <- b_set[1]
+  # df_1 <- Results_sc_only
+  # Results_sc_only <- do.call(rbind.data.frame, lapply(2:length(b_set),function(i){
+  #   df_1$b_true <- b_set[i]
+  #   Results_sc_only <- rbind(Results_sc_only,df_1)
+  # }))
 
-  Results_2 <- rbind(Results_sc_only,Results_com)
+  # Results_2 <- rbind(Results_sc_only,Results_com)
+  Results_2 <- rbind(Results_com)
   Results_2$b_est[which(is.na(Results_2[,"b_est"]))] <- 0
   Results_plot <- Results_2[which(Results_2[,"b_est"]<10),]
   
   # Plots
   
-  Results_plot$Data_source <- factor(Results_plot$Data_source, levels = c("scientific_only", "commercial_only", "scientific_commercial"))
+  # La aussi petite modif pour enlever les données scientifiques
+  # Results_plot$Data_source <- factor(Results_plot$Data_source, levels = c("scientific_only", "commercial_only", "scientific_commercial"))
+  Results_plot$Data_source <- factor(Results_plot$Data_source, levels = c("commercial_only"))
   
   #construction du boxplot pour le biais de l'abondance
   RelBias_N <- ggplot()+
[…]

### 2021-01-11 Version_M2_10bateaux/Scripts/function/simu_commercial_data.R

@@ -47,8 +47,8 @@ simu_commercial_data <- function(loc_x,
                                  q1_com,
                                  q2_com,
                                  b,
-                                 nboat){
-  
+                                 sequencesdepeche){
+
   #initialisation
   index_com_i_2 = c() #index_com_i : sampled cell for each observation
   y_com_i_2 = c() # y_com_i : catch data
@@ -71,17 +71,23 @@ simu_commercial_data <- function(loc_x,
     Int_ps[loc_x$y[q],loc_x$x[q]] <- exp(beta0_fb + b*Strue_fb[q])
   }
   
-  # On dit qu'il y a 10 bateaux et 3000 points de peche au total
-  # Donc chaque bateau est associé à 300 points de peche
   
-  # On commence par définir le centre de la zone de peche de chacun des 10 bateaux
-  # Ce point de centre de zone est défini selon un processus de poisson inhomogène qui dépend du champ latent
+  # On commence par définir les centres de peche
+  # Ces points de centres de zone sont définis selon un processus de poisson inhomogène qui dépend du champ latent
   win_df <- owin(xrange=c(0.5,grid_dim['x'] + 0.5), yrange=c(0.5,grid_dim['y'] + 0.5),ystep  = 1,xstep=1)
-  X <- rpoint(nboat,as.im(Int_ps,win_df)) # Simulation des 10 points centraux
-  # X correspond donc aux coordonnées des 10 "points centraux" des zones de peche de chaque bateau
+  X <- rpoint(sequencesdepeche*zonespersequence,as.im(Int_ps,win_df)) # Simulation des points centraux
+  # X correspond donc aux coordonnées des sequencesdepeche*zonespersequence "points centraux" des zones de peche de chaque bateau
   
-  # Visualisation des 10 centres des zones de peche de chaque bateau
-  centres = as.data.frame(cbind(x = X$x, y = X$y, boats = seq(1, nboat)))
+  # Visualisation des sequencesdepeche*zonespersequence centres des zones de peche de chaque bateau
+  boats = numeric()
+  for (q in 1:sequencesdepeche)
+  {
+    boats = c(boats, rep(q, zonespersequence))
+  }
+  
+  ### si choix des centres au hasard, c'est ici qu'il faut modifier
+  
+  centres = as.data.frame(cbind(x = X$x, y = X$y, boats = boats))
[…]

### 2021-01-11 Version_M2_10bateaux/Scripts/simu_main_script.R

@@ -10,7 +10,7 @@ source("Scripts/function/load_packages.R")
 #-----------------------------------------------------------
 
 # Simulation name
-simu_name = "Test2"
+simu_name = "Test_plusieurszonesparbateau"
 
 #---------------------
 # Simulation scenarios
@@ -73,7 +73,8 @@ n_strate <- 9
 
 ## Commercial data
 # Number of commercial samples
-n_samp_com = 3000
+# n_samp_com = 3000
+n_samp_com = 150
 # observation error 
 logSigma_com = log(1)
 # zero-inflation parameter
@@ -90,7 +91,8 @@ b_set = c(0,1,3)
 #--------------------
 
 # Data sources feeding the model
-Data_source = c("scientific_commercial","scientific_only","commercial_only")
+# Data_source = c("scientific_commercial","scientific_only","commercial_only")
+Data_source = c("commercial_only") # On ne fait plus que le modele commercial
 
 # b estimated or b fixed to 0 in estimation
 EM_set = c("fix_b","est_b")
@@ -111,7 +113,7 @@ counter <- 1
 # index of the first iteration
 i0 <- 1
 # Number of simulation
-n_sim = 1
+n_sim = 10
 
 RandomSeed = 123456
 
@@ -119,7 +121,8 @@ RandomSeed = 123456
 reallocation = c(0,1) # Reallocation = 0 si pas de reallocation, = 1 si reallocation
 
 # Nombre de bateaux ?
-nboat = 10
+sequencesdepeche = 5
+zonespersequence = 3
 
 for (k in reallocation) {
 # Reallocation uniforme ?
@@ -236,7 +239,7 @@ for (k in reallocation) {
                                       k,
                                       xlim,
                                       ylim,
-                                      nboat)
+                                      sequencesdepeche)
     
     Results <- res[[1]]
     List_param <- res[[2]]
@@ -302,17 +305,17 @@ figure_MSPE
 # On sauve les 3 graphes de métriques de performance
 if(! dir.exists(simu_file_metriques)) dir.create(simu_file_metriques)
 
-path = paste0(simu_file_metriques, "biaisabondance_boats", nboat, ".png")
+path = paste0(simu_file_metriques, "biaisabondance_boats", sequencesdepeche, ".png")
 png(file = path, width = 1100, height = 500)
 plot(figure_biaisabondance)
 dev.off()
 
[…]

### 2021-01-25 Version_M2_plusieurszones/Scripts/function/plot_results.R
Dernieres modifs

@@ -506,3 +506,102 @@ generation_metriques = function(path_base,
   plot(plot_MSPE)
   dev.off()
 }
+
+
+
+
+
+
+indicateurs = function(path_base,
+                       simu_file,
+                       table,
+                       b,
+                       k,
+                       biaisb,
+                       biaisN,
+                       MSPE){
+  
+  ### On recupère les données
+  
+  path = paste0(path_base, simu_file, "Results.RData")
+  load(path)
+  
+  # On filtre pour avoir seulement les données du b et du k voulus
+  
+  Results %>%
+    filter(b_true == b) -> Results
+  if (k == reallocation[1]){
+    Results = Results %>% filter(reallocation == 0)
+  } else {
+    Results = Results %>% filter(reallocation == 1)
+  }
+  
+  ### Travail sur le tableau Results (issu de plot_simu.R)
+  
+  Results[,"OnBoundary"]=rep(0,nrow(Results))
+  Results[which(abs(Results[,"b_est"])>10),"OnBoundary"]=1
+  Results[,"Converge"]=Results[,"OnBoundary"]+Results[,"Convergence"]
+  Results[which(Results[,"Converge"]==2),"Converge"]=1
+  Converge_table = summaryBy(Converge~b_true+Data_source,data=Results,FUN=sum)
+  WhichFailed <- which(Results[,"Convergence"]!=0)
+  Results_cvg.failed <- Results[WhichFailed,]
+  WhichDone = which(Results[,"Convergence"]==0)
+  Results=Results[WhichDone,]
+  Results[,"RelBias_N"]=(Results[,"N_est"]-Results[,"N_true"])/Results[,"N_true"]
+  Results[,"Bias_b"]=(Results[,"b_est"]-Results[,"b_true"]) / ifelse(Results[,"b_true"] != 0,Results[,"b_true"],1)
+  Which_plot = which(Results[,"type_b"] == "est_b")
+  Results$b_est[which(is.na(Results[,"b_est"]))] <- 0
+  Results <- Results[which(Results[,"b_est"]<10),]
+  Results$Data_source <- factor(Results$Data_source, levels = c("commercial_only"))
+  
+  
[…]

### 2021-01-25 Version_M2_plusieurszones/Scripts/function/simu_latent_field.R

@@ -43,8 +43,10 @@ simu_latent_field <- function(loc_x,
                               SD_eta){
   
   # 1st param : continuous variable, 4 nexts : strata_1, strata_2, strata_3, strata_4
-  beta[1] = runif(1,-.5,.5)
-
+  # beta[1] = runif(1,-.5,.5)
+  beta[1] = 0.5
+  # On passe beta[1] à 0.5 pour avoir un champ latent plus "cohérent"
+  
   # Simulate using Matérn covariance
   Cov_x = sim_GF_Matern(loc_x, nu, range, SD_x^2)[[1]]$y
 

### 2021-01-25 Version_M2_plusieurszones/Scripts/simu_main_script.R

@@ -10,7 +10,7 @@ source("Scripts/function/load_packages.R")
 #-----------------------------------------------------------
 
 # Simulation name
-simu_name = "Simu_complete_10801"
+simu_name = "SimuComplete"
 
 #---------------------
 # Simulation scenarios
@@ -27,7 +27,7 @@ grid_dim = c("x"=grid_side_x, "y"=grid_side_y)
 n_cells = grid_dim[1]*grid_dim[2]
 
 # Intercepts and covariates of the latent field
-beta0 = 2 # intercept
+beta0 = 3 # intercept
 beta = c(0,0,0,0,0) # covariates
 # parameters covariates vector : 1st param : continuous variable, 4 nexts : strata_1, strata_2, strata_3, strata_4
 # La premiere covariable est continue et a une structure spatiale
@@ -40,9 +40,10 @@ beta = c(0,0,0,0,0) # covariates
 
 # Correlation structure of the continuous covariate 
 simu_tool <- "MaternCov" # 2 way to simulate latent fields : "RandomFields", "MaternCov"
-range = sqrt(prod(grid_dim))/(5)*2 # --> MaternCov 
+range = sqrt(prod(grid_dim))/(5)*2 # --> MaternCov
 nu = 1  #portée de la covariable continue
-SD_x = 0.5 # ecart type/ variance marginale du processus de la var continue
+SD_x = 2 # ecart type/ variance marginale du processus de la var continue
+# Initialement on avait SD_x = 0.5, la on passe a SD_x = 2 pour avoir plus de variabilité dans le champ latent
 
 #effets spatiaux aléatoires
 SD_delta = SD_eta= 0.5
@@ -69,7 +70,7 @@ q1_sci <- 1
 # Relative catchability
 q2_sci <- 1
 # parameter controling size of the strata
-n_strate <- 9 
+n_strate <- 9
 
 ## Commercial data
 # Number of commercial samples
@@ -284,7 +285,7 @@ dyn.unload( dynlib(paste0(TmbFile,"inst/executables/com_x_sci_data_14_scientific
 
[…]

### 2021-01-25 Scripts/simu_main_script1.R
modification 1

@@ -59,7 +59,6 @@ q1_sci <- 1
 q2_sci <- 1
 # parameter controling size of the strata
 n_strate <- 9 
-sci_sampling <- "0" # Deprecated
 
 ## Commercial data
 # Number of commercial samples
@@ -139,6 +138,7 @@ TMB::compile(paste0(TmbFile,"inst/executables/com_x_sci_data_14_scientific_comme
 #-------------------------------------------------------------
 #------------------- Simulation/Estimation loop --------------
 #-------------------------------------------------------------
+library(TMB)
 
 ################
 # source Scripts
@@ -178,13 +178,6 @@ if(restart_after_crash == T){
 dyn.load( dynlib(paste0(TmbFile,"inst/executables/com_x_sci_data_14_scientific_commercial_simple") ) )
 #dyn.unload( dynlib(paste0(TmbFile,"inst/executables/",Version,"_scientific_commercial") ) )
 ## loop
-
-latent_fields_simu="simulate"
-Alpha=2
-zone_without_fishing=F
-b_constraint=2
-DataObs=2
-
 for(i in i0:n_sim){
 
     res <- simu_commercial_scientific(Results,
@@ -198,19 +191,19 @@ for(i in i0:n_sim){
                                       beta,
                                       simu_tool,
                                       no_eta,
-                                      SpatialScale = NA,
+                                      SpatialScale,
                                       SpatialScale_eta,
                                       range,
                                       range_eta,
                                       nu,
                                       SD_x,
                                       SD_delta,
-                                      SD_xfb = NA,
+                                      SD_xfb,
                                       SD_eta,
                                       beta0_fb,
-                                      beta_fb = NA,
+                                      beta_fb,
                                       DataObs,
-                                      zero.infl_model  = 2,
+                                      zero.infl_model,
                                       nb_par_zinfl,
                                       n_samp_sci,
                                       logSigma_sci,
@@ -227,15 +220,15 @@ for(i in i0:n_sim){
                                       Data_source,
                                       Samp_process,
                                       EM,
-                                      weights_com = 1,
-                                      commercial_obs = 1,
+                                      weights_com,
+                                      commercial_obs,
                                       catchability_random,
                                       b_constraint,
-                                      Spatial_model = "SPDE",
-                                      Use_REML = FALSE,
+                                      Spatial_model,
+                                      Use_REML,
                                       Alpha,
                                       RandomSeed,
-                                      Version = NA,
+                                      Version,
                                       TmbFile,
                                       ignore.uncertainty,
                                       counter,

### 2021-01-25 scripts/function/commercial_scientific_14.R

@@ -276,8 +276,8 @@ simu_commercial_scientific <- function(Results,
     ## Fit model
     ############
     # Create the SPDE/GMRF model, (kappa^2-Delta)(tau x) = W:
-    mesh = inla.mesh.create( loc_x[,c('x','y')] )
-    spde <- (inla.spde2.matern(mesh, alpha=Alpha)$param.inla)[c("M0","M1","M2")]  # define sparse matrices for parametrisation of precision matrix
+    # mesh = inla.mesh.create( loc_x[,c('x','y')] )
+    # spde <- (inla.spde2.matern(mesh, alpha=Alpha)$param.inla)[c("M0","M1","M2")]  # define sparse matrices for parametrisation of precision matrix
     
     # Loop on alternative configuration models
     for(Estimation_model in Data_source){
@@ -288,17 +288,17 @@ simu_commercial_scientific <- function(Results,
       ## Fit Model
       ############
       skip_to_next <- F
-      tryCatch({fit_IM_res <- fit_IM(Estimation_model_i,
-                                     Samp_process,
-                                     EM,
-                                     TmbFile,
-                                     ignore.uncertainty,
-                                     c_com_x,
-                                     y_com_i,
-                                     index_com_i,
-                                     y_sci_i,
-                                     index_sci_i,
-                                     Cov_x)
+      fit_IM_res <- fit_IM(Estimation_model_i,
+                           Samp_process,
+                           EM,
+                           TmbFile,
+                           ignore.uncertainty,
+                           c_com_x,
+                           y_com_i,
+                           index_com_i,
+                           y_sci_i,
+                           index_sci_i,
+                           Cov_x)
       
       SD <- fit_IM_res$SD
       Report <- fit_IM_res$Report
@@ -380,8 +380,8 @@ simu_commercial_scientific <- function(Results,
                          Opt_par = Opt,
                          SD = SD,
                          Report = Report)
-
-
+      
+      
       
       Results[counter,"N_true"]=sum(Strue_x)
       Results[counter,"LogLik"]=-Opt$objective
@@ -393,9 +393,9 @@ simu_commercial_scientific <- function(Results,
       MSPE_S_2_df <- cbind(loc_x,Strue_x,Report$S_x) %>%
         dplyr::mutate(S_x = Report$S_x) %>%
[…]

### 2021-01-26 Rapport/Rmarkdown_Synthese_jet_2.Rmd
Majoration champ latent, SD minimal champ latent, ajout situation reference, analyse structuration et reallocation, modifs valeurs de b, maps champ latent estime, modifs graphes quantite pechee

---
title: "PROJET INGENIEUR <br/> <br/> Etude de l'influence de la réallocation des captures commerciales  sur la performance d'un modèle d'estimation de la distribution spatiale de la sole commune"

author: "<br/> Océane GUITTON, Chloé TELLIER, Juliette THEOLEYRE"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
# I - Introduction

Les travaux de Alglave et al. ont permis de développer un modèle permettant d’inférer la distribution spatiale des espèces d’intérêt halieutique.

L'objet précis est d'**estimer la distribution spatiale de densités de sole commune dans le golfe de Gascogne**. A terme, l'objectif de la thèse est de décrire la dynamique spatio-temporelle des soles à petite échelle et d'identifier leurs zones fonctionnelles.

Le modèle intègre des données scientifiques et des données commerciales idéales géolocalisées, et prédit l'abondance à partir de ces deux types de données. 
Ce modèle permet notamment de prendre en compte le biais introduit par le comportement de ciblage des pêcheurs vis-à-vis de la ressource : c'est l’échantillonnage préférentiel des données commerciales.

## A) Les données

Les **données scientifiques** sont constituées de relevés de quantités pechées à 50 localisations données, obtenues grâce à un design d'échantillonage aléatoire standardisé. Elles sont issues d'une enquête realisée tous les ans en fin d'année et ne peuvent donc pas être utilisés pour estimer la variabilité au cours de l'année. Ces données sont fiables, géolocalisées et sont utilisées pour fournir des indices d’abondance non biaisés (cf pas d'échantillonage préférentiel) ainsi que pour estimer le stock biologique plutôt que pour d’écrire la distribution des espèces.

Les **données commerciales** sont une source d’information valorisable pour décrire la distribution spatiale des espèces d’intérêt halieutique. La donnée commerciale est caractérisée par une quantité pêchée en un point de pêche. 
Ces données géolocalisées couvrent souvent une surface bien plus importante que les données scientifiques mais dépendent d'un échantillonage préférentiel qu'il est nécessaire de prendre en compte. En effet, les pêcheurs ciblent les zones de forte abondance en ressource, et si on ne prend pas en compte ce phénomène, cela peut mener à des estimations de la distribution et de l’abondance de l’espèce biaisées. 
L'ensemble de ces données commerciales constituent un plan à haute résolution de CPUE (catch per unit effort), qui présente des quantités pêchées aux différents points de pêche.

## B) Le modèle

Le modèle proposé par Alglave et al. est un modèle hiérarchique intégré qui peut utiliser soit uniquement les données scientifiques, soit uniquement les données commerciales, soit les deux types de données simultanément. Dès lors que les données commerciales sont intégrées, l'échantillonage préférentiel peut être pris en compte.

Le modèle itère de nombreux schémas de simulation/estimation, afin de comparer les trois scenari possibles (données scientifiques uniquement, données commerciales uniquement, les deux types de données simultanément) selon différentes métriques de performance. Plus précisément, l'idée est d’évaluer la performance des scénari et d’estimer l’influence de chacun des deux types de données dans l’inférence.

Le modèle présente les *trois composantes* suivantes :

- **Champ Latent S** 
Le champs latent est la densité de biomasse / densité de poissons. Il est modélisé selon champ spatial gaussien (plus précisément « latent log gaussian ») sur un domaine discret.

Cette densité dépend de covariables (par exemple le type de sol, la profondeur) et d'un effet spatial aléatoire. Le champ latent est exprimé en fonction d'un intercept (alpha_s ou $\alpha_{s}$), de covariables environnementales et des paramètres (betas_s ou $\beta_{s}$) qui leur sont associés (traduisant la relation entre les especes et leur habitat (effets fixes)), et enfin de l'effet spatial aleatoire qui traduit la correlation spatiale des captures ($\delta(x)$).

\begin{equation}
S(x)=\exp(\alpha_{S}+\Gamma_{S}(x)^{T} \cdot \beta_{S}+\delta(x))
\end{equation}

\begin{equation}
    \delta(x) \sim G R F(0, M(x , x': \rho))
\end{equation}

- **Processus d'échantillonage** (sampling process)

Désormais, nous nous intéressons au **processus d'échantillonage**, c'est à dire à la **génération des points de pêche** (et de leur distribution spatiale) selon un processus de poisson non homogène. L'intensité des points de pêche dans une zone donnée dépend de l'abondance du champ latent à cet endroit.

Le design d’échantillonage des données scientifiques est indépendant du champ latent S (processus d'échantillonage aléatoire) tandis que le design d’échantillonage des données commerciales est préférentiel et dépend de la connaissance à priori par les pêcheurs de la distribution spatiale de la densité de poisson (champ latent S).

La localisation des points de pêche peut également dépendre d'autres effets incarnés par des covariables (comme la connaissance de la présence d'espèces non ciblées (cf prises accessoires), les technologies disposées, les contraintes logistiques etc) et enfin d'un effet spatial aléatoire.

On fait l'hypothèse que les bateaux d'une même flotte présentent un comportement homogène et partagent donc les mêmes paramètres.

On modélise la distribution spatiale des points de pêche par un processus de Poisson non homogène (l’intensité  $\lambda_{f}(x)$ des points de pêche dans une zone donnée dépend de l’abondance du champ latent à cet endroit-là).

On note $X_{com f}$ les points spatiaux où les bateaux de la flotte f ont peché (ou plutôt ont été identifiés comme étant en train de pecher). $X_{com f}$ est considéré comme un "poisson point process" avec une variation spatiale d'intensité $\lambda_{f}(x)$.
Cette insensité $\lambda_{f}(x)$ est exprimée en fonction d'un intercept ($\alpha_{Xf}$), d'un paramètre $b_{f}$ (qui traduit un échantillonage préférentiel : si $b_{f}$ vaut 0 on ne prend pas en compte l'échantillonage préférentiel, si $b_{f}$ est différent de 0 on prend en compte l'échantillonage préférentiel), du champ latent S(x) au point considéré, de covariables et de leurs paramètres associés ($\beta_{Xf}$), et enfin de l'effet spatial aléatoire ($\eta{f}(x)$).

\begin{equation}
\mathrm{X}_{com \,f} \sim \mathcal{IPP}(\lambda_f(x))
\end{equation}

\begin{equation}
\lambda_f(x)=\exp\left(\alpha_{X\,f}+b_f . \log(S(x))+\Gamma_{X}(x)^{T} \cdot \beta_{X\,f}+\eta_f(x)\right)
\end{equation}

- **Zero inflated observation model**

Conditionnellement au champs latent S et au processus d’échantillonage commercial $X_{com f}$,  les données scientifiques et commerciales sont modélisées selon des modèles zero-inflated spatialement homogènes. 
L'objectif est d'attribuer à chaque point de pêche préalablement déterminé avec le processus d'échantillonage la quantité de poissons pêchés, sachant qu'il est possible qu'en certains points de pêche, aucun poisson n'ai été pêché.

On détermine tout d'abord quels sont les points où la pêche a été 'sans succès' et ceux où elle s'est révélée être fructueuse puis on attribue à chaque point de pêche 'fructueux' la quantité de poisson pêchée. Pour cela, on fait appel à un modèle zero-inflated.

La vraisemblance d’une capture 'sans succès' (abscence de capture) est modélisée selon une distribution de Bernoulli de parametre $exp(-e^{\xi_f} \cdot \mu_f(x_{i}))$, où $\mu_f(x_{i})$ est la capture positive attendue au point $x_{i}$ et $\xi_f$ est le paramètre controllant l'intensité de pêches 'sans succès' pour la flotte f. La capture positive attendue $\mu_f(x_{i})$ est définie par l'expression suivante : $\mu_f(x_i) = q_f\cdot S(x_{i})$, où $q_f$ est le paramètre de capturabilité relative spécifique à chaque flotte et configuration.

La vraisemblance d’une capture 'fructueuse' $y_i$ (présence de capture) est définie comme le produit de la probabilité d'obtenir une capture fructueuse $(1-exp(-e^{\xi_f} \cdot \mu_f(x_{i})))$ et d'une distribution lognormale positive continue L de moyenne attendue $\frac{\mu_f(x_i)}{(1-exp(-e^{\xi_f} \cdot \mu_f(x_i)))}$ et d'écart type $\sigma_f$.

On considère le modèle zero-inflated spatialement homogène suivant :

\begin{equation}
\operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i) \right) =
\left\{
    \begin{array}{ll}
        \exp \left(- e^{\xi_f} \cdot \mu_f(x_{i})\right) & \text { if } y_{i}=0 \\
        \left(1-\exp \left(-e^{\xi_f} \cdot \mu_f(x_{i})\right)\right) \cdot \operatorname{L}\left(y_{i },\frac{\mu_f(x_{i})}{\left(1 - \exp \left(-e^{\xi_f} \cdot \mu_f(x_{i})\right)\right)},\sigma_f^{2} \right) & \text { if } y_{i} > 0
    \end{array}
\right.
\end{equation}

On dispose dès lors de la quantité pêchée $y_i$ en chaque point de pêche.

## C) Problématisation

Les **données commerciales** intégrées dans le modèle d'Alglave et al. sont en fait des données fictives idéales et n'existent pas.

En effet, les pêcheurs ne déclarent en réalité ni la localisation de leurs points de pêche, ni la quantité de poisson qu'ils y pêchent. Chaque bateau de pêche déclare seulement chaque jour la quantité totale pêchée à l'échelle d'un **carré statistique**, un carré statistique étant une zone fictive délimitée et référencée (il en existe plusieurs au sein du golfe de Gascogne). Ces quantités de pêche déclarées chaque jour par les pêcheurs à l'échelle de chaque carré statistique sont ce que l'on appelle les **données de capture (ou de déclaration)** (ou données des carnets de bord). Ces données de capture, contrairement aux données commerciales idéales sont non géolocalisées.

Parallèlement à ces données de capture, pour un bateau de pêche donné et un jour donné, on dispose des **données VMS (Vessel Monitoring System) de localisation**. 
L'idée est de déterminer quelles sont les *données VMS de localisation* qui correspondent à une activité de pêche afin de déterminer des **points de pêche**, la plupart du temps à l'aide de données de vitesse des bateaux.

Les **données commerciales réellement disponnibles** (quantités pêchées à l'échelle du carré statistique) **sont donc moins précises que les données commerciales intégrées dans le modèle d'Alglave et al.** (quantités pêchées $y_i$ à l'échelle de chaque point de pêche).

Les données de déclarations (logbooks) peuvent être croisées avec les données de position des navires (VMS) pour obtenir la distribution des captures à une résolution spatio-temporelle très fine.

Afin de pouvoir utiliser une information à l'échelle de chaque point de pêche dans le cadre de l'estimation de la distribution spatiale de densités de poissons, on cherche à obtenir une quantité pêchée $y_i$ en chaque point de pêche au sein du carré statistique, grâce à de la **réallocation uniforme**.

Ainsi, pour chaque bateau de pêche et pour une journée donnée, on **réalloue uniformément** les données de captures déclarées pour un carré statistique à l'ensemble des points de pêches identifiés du carré (à l'aide des données VMS).

L'objet de notre projet ingénieur est de traiter la problématique de cette **réallocation uniforme**.

La **réallocation uniforme** consiste à diviser la quantité pêchée déclarée à l'échelle du carré statistique par un bateau donné (et pour un jour donné) par le nombre de points de ses points de pêche identifiés dans ce carré statistique : on obtient alors la quantité pêchée $y_i$, qui est la **même en chacun des points de pêche du carré statistique considéré**.

On ne sait pas à quel point cette réallocation des captures pourrait déteriorer les performances du modèle.

L'objectif est de **comprendre et quantifier l'effet de la réallocation sur les estimations**. On atteint cet objectif par une étude par simulation.

On se demande alors :

<span style="color:green">*La réallocation produit - elle un biais dans l'estimation des coefficients du modèle, et notamment dans l'estimation de l'abondance des poissons ? Si oui, dans quelle mesure ?*</span>

<span style="color:green">*Dans quelle mesure la situation idéale fictive, utilisée jusqu'alors par le modèle de la thèse, avec des quantitées pêchées $y_i$ différentes en chacun des points de pêche du carré statistique, est-elle éloignée de la situation réelle où l'on dispose de données recensées moins précises ?*</span>

On peut faire l'hypothèse que la réallocation entrainera une moins bonne estimation des coefficients et un biais plus important des métriques de performance : en effet, les $y_i$ après réallocation sont des données moins précises que les $y_i$ sans réallocation.

# II - Matériel et méthode

On se place dans un cadre simplifié et suffisant pour étudier l'effet de la réallocation uniforme.

On se place à l'échelle d' **un seul carré statistique**, incarné par une grille discrète fictive de taille 25*25 qui comporte 625 cellules. L'étude à l'échelle d'un carré statistique pourra être généralisée et étendue à l'ensemble des carrés statistiques du Golfe de Gascogne.

On décrit dans cette partie le modèle simplifié utilisé pour étudier l'impact de la réallocation uniforme, ainsi que les deux étapes clés du modèle : la simulation des données puis l'estimation des paramètres.

## A) Modèle simplifié

On **simplifie le modèle hiérarchique intégré** de la manière suivante :

- on ne considère pour le champ latent S qu'une covariable continue qui a une structure spatiale.

- on ne prend pas en compte les effets spatiaux aléatoire du champs latent S ($\delta(x)$) et de l'inténsité $\lambda_{f}(x)$ du processus d'échantillonage ($\eta{f}(x)$).

- la localisation des points de pêche ne dépend pas d'autres effets incarnés par des covariables (prises accessoires, technologies disposées, les contraintes logistiques etc)

Ainsi, les **équations finales à considérer dans le cadre de notre application** sont les suivantes :

\begin{equation}
S(x)=\exp(\alpha_{S}+\Gamma_{S}(x)^{T} \cdot \beta_{S})
\end{equation}

\begin{equation}
\mathrm{X}_{com \,f} \sim \mathcal{IPP}(\lambda_f(x))
\end{equation}

\begin{equation}
\lambda_f(x)=\exp\left(\alpha_{X\,f}+b_f . \log(S(x))\right)
\end{equation}

\begin{equation}
\operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i) \right) =
\left\{
    \begin{array}{ll}
        \exp \left(- e^{\xi_f} \cdot \mu_f(x_{i})\right) & \text { if } y_{i}=0 \\
        \left(1-\exp \left(-e^{\xi_f} \cdot \mu_f(x_{i})\right)\right) \cdot \operatorname{L}\left(y_{i },\frac{\mu_f(x_{i})}{\left(1 - \exp \left(-e^{\xi_f} \cdot \mu_f(x_{i})\right)\right)},\sigma_f^{2} \right) & \text { if } y_{i} > 0
    \end{array}
\right.
\end{equation}

Finalement, la **simplification du modèle** peut être résumée de la manière suivante : **on ne prend en compte qu'une seule covariable pour le champ latent et il n'y a pas de corrélation spatiale dans l'abondance et l'échantillonage**.

## B) Etape de simulation

On considère uniquement des **données commerciales**, le nombre de points de pêche scientifiques à l'échelle d'un carré statistique étant très faible. Le nombre de données commerciales (qui est égal au nombre de points de pêche) générées est de **150** (ordre de grandeur raisonnable à l'échelle d'un carré statistique).

Par ailleurs, on cherche à estimer les conséquences de la prise en compte ou non de l’**échantillonage préférentiel** sur les capacités du modèle. Pour cela, on introduit le paramètre **b** qui peut prendre trois valeurs : 0 (pas d'échantillonage préférentiel), 3 (échantillonage préférentiel moyen) et 10 (échantillonage préférentiel important).

En outre, on définit une **séquence de pêche** comme une **combinaison bateau de pêche / jour de pêche donnée**.

La **séquence de pêche** est l'échelle de déclaration. 
En terme de données commerciales réelles, une *séquence de pêche* correspond à plusieurs captures effectives déclarées (une par carré statistique du Golfe de Gascogne).
En terme de données commerciales fictives, une *séquence de pêche* présente une capture $y_i$ en chacun de ses points de pêche identifiés.

Le modèle hiérarchique intégré va simuler des **données commerciales fictives idéales** pour chacune des *séquences de pêche* considérées. Ces **quantités capturées $y_i$ idéales en chaque point de pêche** sont nommées **captures $y_i$ exactes**.

On introduit les paramètres **P**, **Z**, **X**, **W** et **V**.

On définit **P** comme le **nombre de séquences de pêche** considérées qui peut valoir 2, 5 et 10.

On définit **Z** comme le **nombre moyen de zones de pêche visitées par séquence de pêche** qui peut valoir 1, 3 et 5.

Toutes les séquences de pêches ne visitent pas forcément le même nombre de zones de pêche. Par exemple, si le nombre moyen de zones de pêche visitées Z vaut 3 et qu'on considère 2 séquences de pêche, l'une des séquences de pêche peut visiter 2 zones et l'autre 4 zones.

On définit **X** comme la **combinaison de P et Z considérée**.

On définit **W** comme le **nombre de zones de pêches visitées par l'ensemble des séquences de pêche**, et donc le **nombre total de centres de zones de pêche** à générer. Pour un couple (P,Z) fixé, **W** est le produit de P et Z : on a { W = P . Z }.

On définit **V** comme le **nombre de points de pêche par zone de pêche pour chacune des séquences de pêche**. **V** est fixe et on a {V = 150 / W}.

Pour un couple (P,Z) fixé, les W **centres de zones de pêche sont générés selon un processus de poisson non homogène**. De la sorte, l'intensité des centres de zones de pêche dans une zone donnée du carré statistique dépend de l'abondance du champ latent à cet endroit. Plus b est grand (et donc plus l'échantillonage préférentiel est pris en compte), plus les centres de zones de pêches seront propices à être situés dans les zones du carré statistique de forte abondance en poissons.

Ensuite, les centres de zones de pêche sont attribués à l'ensemble des séquences de pêche : tout d'abord, chacune des séquences de pêche se voit attribuer un centre puis le reste des centres sont attribués aléatoirement aux séquences de pêche. Ainsi, toutes les séquences de pêche ont au moins une zone de pêche, et le nombre de zones de pêche par séquence de pêche varie.

Une fois les centres de zones de pêche générés et attribués, on définit une **zone de pêche rectangulaire** autour de chacun des centres de zone de pêche, et les ** V points de pêche sont générés dans chacune des zones de pêche selon un processus de poisson non homogène**. Ainsi, la localisation des points de pêche au sein même des zones de pêche peut également dépendre de l'abondance de poissons.

On considère le **tableau** suivant qui présente l'ensemble des situations possibles pour les couples de (P,Z), ainsi que **deux schémas** illustrant le cas particulier de deux couples (P,Z).

![](C:/Users/ocean/Documents/Agrocampus_Ouest/3A/projet_ingenieur/projet_M2_sole/projet_M2_sole/Version_M2_plusieurszones/Photos_rapport/schema.png)

COMPLETER L'ESTIMATION DES SCENARIOS : A FAIRE

A ce stade, **le modèle génère les *données commerciales fictives idéales* ou  *captures $y_i$ exactes* évoquées précedemment pour chacun des 150 points de pêches totaux**.

La situation fictive idéale, dont on ne dispose pas en réalité, vient d'être simulée. Se pose alors la question de la réallocation uniforme pour se ramener à la situation réelle.

## C) Etape de réallocation uniforme

On dispose désormais des **quantitées de poissons capturées $y_i$ exactes en chacun des 150 points de pêche**.

Si on ne fait pas de réallocation uniforme, ce sont ces **captures $y_i$ exactes**, disponibles pour chacun des points de pêche, qui sont ensuite utilisées pour l'**étape d'estimation** du modèle. Il faut noter que ces **captures $y_i$ exactes** sont potentiellement différentes pour tous les points de pêche du carré statistique.

Si on fait de la réallocation uniforme pour se ramener à la situation réelle, on va disposer en chacun des points de pêche de **captures $y_i$ réallouées**. **Ces *captures $y_i$ réallouées* seront les mêmes pour tous les points de pêche appartenant à la même séquence de pêche**.

Le protocole de réallocation est détaillé ci-dessous.

On introduit le paramètre **k** afin d'étudier l'influence de la **réallocation uniforme**:

- si **k = 0**, on se place dans une configuration **sans réallocation** 
- si **k = 1**, on se place dans une configuration **avec réallocation**

Dans le cas **avec réallocation**, toutes les *captures $y_i$ exactes* obtenues pour les points de pêche d'une séquence de pêche sont **sommées** : on obtient la **quantité pêchée à l'échelle du carré statistique** et on se ramène ainsi à la situation réelle. On divise ensuite cette quantité pêchée à l'échelle du carré statistique par le nombre de points de pêches de la séquence de pêche au sein du carré et on obtient les **captures $y_i$ réallouées**. 
Les *captures $y_i$ réallouées* sont ensuite utilisés par le modèle pour réaliser l'**étape d'estimation**.

## D) Etape d'estimation

La phase d'**estimation** du modèle se fait à l'aide du **package TMB (Template Model Builder)**, qui s’est révélé très efficace pour maximiser la vraisemblance des modèles hiérarchiques à effets aléatoires. 
Le package TMB permet d'estimer les paramètres par la méthode du maximum de vraisemblance. En présence de variables cachées, la vraisemblance n'est pas directement disponible et doit être calculée par intégration, ce qui est en général long. Dans TMB, cette intégration ets remplacée par une approximation de Laplace. Une fois la vraisemblance calculée, la recherche du maximum de vraisemblance se fait grâce à des algorithmes de descente de gradient, impliquant le calcul des dérivées première et seconde du gradient par différenciation automatique.

## E) Objectifs

**L'objectif** est d' **étudier trois métriques de performance**, plus précisément le biais de l'abondance, le biais de b (paramètre controlant l'échantillonage préférentiel) et le MSPE (A COMPLETER : LES EXPLIQUER ETC, CF PLAN). 
Ces métriques sont déja utilisés dans les travaux d'Algave et al. 
On cherche à comparer les paramètres estimés à l'aide du package TMB avec ceux initialement fixés pour la simulation des données. 
Le modèle hiérarchique intégré renvoit une représentation graphique (boxplot) des valeurs du MSPE et des biais de l'abondance et de b sur les 100 schémas de simulation/estimation.

On étudie **100 schémas de simulation/estimation** (incarnés par le paramètre **i** qui varie de 1 à 100) afin de capturer la variabilité des métriques de performance.

Finalement, **pour chacune des 100 simulations, pour chacune des trois valeurs de b (0, 5 et 20), pour chacune des deux valeurs de k (0 pour 'sans réallocation' et 1 pour 'avec réallocation') et pour chacune des 9 combinaisons de (P,Z) possibles (9 valeurs de X possibles), on simule les données puis on estime les paramètres et on renvoie les graphiques pour les trois métriques de performance**.

Pour une seule séquence de pêche, on connait donc les zones de pêche visitées au sein du carré statistique, les points de pêche (leur nombre V et leur localisation) au sein de chaque zone de pêche et les quantitées pêchées $y_i$ en chacun des points de pêche (**captures $y_i$ exactes** si pas de réallocation et **captures $y_i$ réallouées** si réallocation).

Dès lors, on peut se poser la question suivante :

<span style="color:green">*Peut on déterminer des situations critiques en terme de nombre de séquences de pêche et de nombre moyen de zones de pêche par séquence de pêche, pour lesquelles la réallocation uniforme n'est plus raisonnable, dans la mesure ou elle introduit un biais trop important dans l'estimation des paramètres, des métriques de performances ?*</span>

On peut faire l'hypothèse qu'un couple de (P,Z) tel que P est faible (peu de séquences de pêche) et Z est grand (nombre de zones de pêche par séquence de pêche important en moyenne) incarnera une situation critique.

# III - Focus sur un exemple

VOIR POUR METTRE CET EXEMPLE EN ANNEXE

Voici ci-dessous un exemple afin de mieux illuster les différentes étapes de la phase de simulation des données.

Considérons l'exemple où **P = 5** et **Z = 3** et donc où **W = 15** ( = P . Z = 5 . 3) et **V = 10** ( = 150 / W = 150 / 15)

Finalement, dans cet exemple, nous avons **5 séquences de pêche, 3 zones de pêche par séquence de pêche en moyenne, 15 zones de pêche au total et 10 points de pêche par zone de pêche**.

On considère notre grille de taille 25 * 25 (625 cellules).

On y simule tout d'abord le champ latent (densité de poissons), qui dépend uniquement de la covariable continue dans notre cas.

Voici ci-dessous la **Carte du champ latent**.

![](C:/Users/ocean/Documents/Agrocampus_Ouest/3A/projet_ingenieur/projet_M2_sole/projet_M2_sole/Version_M2_plusieurszones/Photos_rapport/latentfield_i1.png)

Les 15 centres de zones de pêche sont ensuite générés selon le processus de poisson non homogène puis attribués à chacune des 5 séquences de pêche.

Voici ci-dessous la **Carte de localisation des centres de pêche** au sein de notre grille d'étude. La légende rapporte l'appartenance des centres à sa séquence de pêche.

![](C:/Users/ocean/Documents/Agrocampus_Ouest/3A/projet_ingenieur/projet_M2_sole/projet_M2_sole/Version_M2_plusieurszones/Photos_rapport/centres_i1_x5.png)

On remarque sur la carte que pour cet exemple :

- la séquence de pêche 1 a 3 zones de pêche
- la séquence de pêche 2 a 5 zones de pêche 
- la séquence de pêche 3 a 2 zones de pêche
- la séquence de pêche 4 a 1 zone de pêche
- la séquence de pêche 5 a 3 zones de pêche

Ainsi, les séquences de pêche n'ont pas toutes le même nombre de zones de pêche.

Ensuite, 10 points de pêche sont simulés pour chacune des 15 zones de pêche.

Voici ci-dessous la **Carte de localisation des points de pêche au sein de chaque zone de pêche**. La légende rapporte l'appartenance des points de pêche à leur centre et à leur séquence de pêche.

![](C:/Users/ocean/Documents/Agrocampus_Ouest/3A/projet_ingenieur/projet_M2_sole/projet_M2_sole/Version_M2_plusieurszones/Photos_rapport/pecheperboat_i1_x5.png)

Puis, les points de pêche sont agrégés à l'échelle de la cellule de la grille la plus proche.

Voici ci-dessous la **Carte de localisation des cellules pêchées**. La légende rapporte l'appartenance des cellules pêchées à leur séquence de pêche. (( On a ici un probléme de recouvrement de couleur que nous n'avons pas encore réglé)).

![](C:/Users/ocean/Documents/Agrocampus_Ouest/3A/projet_ingenieur/projet_M2_sole/projet_M2_sole/Version_M2_plusieurszones/Photos_rapport/pechecell_i1_x5.png)

Enfin, pour toutes les cellules pêchées, une quantité pêchée $y_i$ est calculée, et ce pour les deux situations suivantes : sans réallocation et avec réallocation.

Voici ci-dessous la **Carte des quantités pêchées par cellules, dans les situations "sans réallocation" et "avec réallocation"**. La légende rapporte l'appartenance des cellules pêchées à leur séquence de pêche.

![](C:/Users/ocean/Documents/Agrocampus_Ouest/3A/projet_ingenieur/projet_M2_sole/projet_M2_sole/Version_M2_plusieurszones/Photos_rapport/pecheqte_i1_x5.png)

Dans le cas où il n'y a pas de réallocation, les quantités pêchées sont différentes pour tous les points de pêche de la grille. Dans le cas où il y a réallocation, les quantités pêchées sont les mêmes pour toutes les cellules ayant été pêchées par la même séquence de pêche.

Ces $y_i$ simulés sont les données commerciales qui vont servir de source à l'estimation des paramètres du modèle et donc au calcul des biais des métriques de performance.

# IV - Résultats et interprétations

### 2021-01-26 Version_M2_plusieurszones/Scripts/function/commercial_scientific_14.R

@@ -98,13 +98,14 @@ simu_commercial_scientific <- function(Results,
                                        xlim,
                                        ylim,
                                        sequencesdepeche,
-                                       zonespersequence){
+                                       zonespersequence,
+                                       taillezone){
 
   ################
   ## Simulate data
   ################
 
-  set.seed( RandomSeed + i ) # for figures : i = 2
+  set.seed( RandomSeed + i )
   
   #---------------
   # Construct grid
@@ -151,25 +152,39 @@ simu_commercial_scientific <- function(Results,
   # beta (fixed parameters for the abundance distribution equation), 
   # delta_x (random effect for the abundance distribution equation), 
   # eta_x (random effect for the sampling process equation)
-  simu_latent_field_outputs <- simu_latent_field(loc_x,
-                                                 latent_fields_simu,
-                                                 latent_field,
-                                                 scientific_data,
-                                                 beta0,
-                                                 beta,
-                                                 simu_tool,
-                                                 range,
-                                                 nu,
-                                                 SD_x,
-                                                 SD_delta,
-                                                 SD_eta)
+  validation = 0
+  wronglatentfield = 1
+  
+  while (validation == 0)
+  {
+    simu_latent_field_outputs <- simu_latent_field(loc_x,
+                                                   latent_fields_simu,
+                                                   latent_field,
+                                                   scientific_data,
+                                                   beta0,
+                                                   beta,
+                                                   simu_tool,
+                                                   range,
+                                                   nu,
+                                                   SD_x,
+                                                   SD_delta,
+                                                   SD_eta)
+    Strue_x <- simu_latent_field_outputs$Strue_x
+    # Contient 625 lignes et une colonne correspondant aux valeurs du champ latent en chacun des points
+    if (max(Strue_x) < 400 & sd(Strue_x) > 10){
+      validation = 1
+    } else{
+      wronglatentfield = wronglatentfield + 1
[…]

### 2021-01-26 Version_M2_plusieurszones/Scripts/function/plot_results.R

@@ -29,10 +29,14 @@ generation_map = function(path_base,
   } else if (x==8){
     sequencesdepeche=10
     zonespersequence=3
-  } else {
+  } else if (x==9){
     sequencesdepeche=10
     zonespersequence=5
+  } else{ # Cas ou x = 10 et donc on est dans le cas de reference
+    sequencesdepeche = 1
+    zonespersequence = 1
   }
+  
   
   for (b in b_set){ # On va travailler successivement sur les differentes valeurs de b
     for (k in reallocation){ # On va travailler successivement sur le cas sans reallocation et avec reallocation
@@ -48,16 +52,42 @@ generation_map = function(path_base,
       
       ### On crée les maps
       
-      # 1ère map : champ latent
-      plot_latentfield = ggplot(Strue_x_2) + geom_point(aes(x,y,col=Champ_latent), size=4) + theme_bw() +
-        scale_color_gradient2(midpoint = mean(Strue_x_2$Champ_latent), low = "#E6F2FC", mid = "#62B4FC",
+      # 1ère map : champ latent reel
+      plot_truelatentfield = ggplot(Strue_x_2) + geom_point(aes(x,y,col=Champ_latent_reel), size=4) + theme_bw() +
+        scale_color_gradient2(midpoint = mean(Strue_x_2$Champ_latent_reel), low = "#E6F2FC", mid = "#62B4FC",
                               high = "#02182C", space = "Lab" ) +
         labs(color= "Champ latent") +
         xlim(0.5, 25.5) +
         ylim(0.5, 25.5) +
-        labs(title = paste0("Representation du champ latent pour i = ", i))
+        labs(title = paste0("Representation du champ latent reel pour i = ", i))
+
+      # 2ème map : champ latent estime
       
-      # 2ème map : centres des zones de peche
+      # D'abord on refait le champ latent reel mais sans titre et avec un facet_wrap
+      plot_latentfield = ggplot(Strue_x_2) + geom_point(aes(x,y,col=Champ_latent_reel), size=4) + theme_bw() +
+        scale_color_gradient2(midpoint = mean(Strue_x_2$Champ_latent_reel), low = "#E6F2FC", mid = "#62B4FC",
+                              high = "#02182C", space = "Lab" ) +
+        labs(color= "Champ latent") +
+        xlim(0.5, 25.5) +
+        ylim(0.5, 25.5) +
+        facet_wrap(~"Reel") +
[…]

### 2021-01-26 Version_M2_plusieurszones/Scripts/function/simu_commercial_data.R

@@ -48,7 +48,8 @@ simu_commercial_data <- function(loc_x,
                                  q2_com,
                                  b,
                                  sequencesdepeche,
-                                 zonespersequence){
+                                 zonespersequence,
+                                 taillezone){
 
   # Initialisation
   index_com_i_2 = c() # index_com_i : sampled cell for each observation
@@ -110,7 +111,7 @@ simu_commercial_data <- function(loc_x,
     # pour l'objet owin, n_samp_com/(sequencesdepeche*zonespersequence) pour le rpoint,
     # et on passe dans le rpoint Int_ps qui correspond
     # aux lambdas qui dépendent du champ latent
-    win_df_boat <- owin(xrange=c(max(0.5, X$x[centre]-2), min(25.5, X$x[centre]+2)), yrange=c(max(0.5, X$y[centre]-2), min(25.5, X$y[centre]+2)), ystep=1, xstep=1)
+    win_df_boat <- owin(xrange=c(max(0.5, X$x[centre]-taillezone), min(25.5, X$x[centre]+taillezone)), yrange=c(max(0.5, X$y[centre]-taillezone), min(25.5, X$y[centre]+taillezone)), ystep=1, xstep=1)
     X_boat <- rpoint(n_samp_com/(sequencesdepeche*zonespersequence), as.im(Int_ps, win_df_boat))
     # On ajoute les coordonnées de peche obtenues pour le bateau
     boats_x = c(boats_x, X_boat$x)

### 2021-01-26 Version_M2_plusieurszones/Scripts/simu_main_script.R

@@ -83,7 +83,7 @@ q1_com <- 1
 # Relative catchability
 q2_com <- 1
 # Levels of preferential sampling
-b_set = c(0,5,20)
+b_set = c(0,1,3)
 
 
 
@@ -119,11 +119,12 @@ n_sim = 100
 RandomSeed = 123456
 
 # Reallocation uniforme ?
-reallocation = c(0,1) # Reallocation = 0 si pas de reallocation, = 1 si reallocation
+reallocation = c(0,1)
+# Reallocation = 0 si pas de reallocation et structuration, = 1 si reallocation et structuration
 
 # Nombre de bateaux ?
-sequencesdepeche_vect = c(2, 5, 10)
-zonespersequence_vect = c(1, 3, 5)
+sequencesdepeche_vect = c(2, 5, 10, 1)
+zonespersequence_vect = c(1, 3, 5, 1)
 PZ = list(c(sequencesdepeche_vect[1], zonespersequence_vect[1]),
           c(sequencesdepeche_vect[1], zonespersequence_vect[2]),
           c(sequencesdepeche_vect[1], zonespersequence_vect[3]),
@@ -132,7 +133,8 @@ PZ = list(c(sequencesdepeche_vect[1], zonespersequence_vect[1]),
           c(sequencesdepeche_vect[2], zonespersequence_vect[3]),
           c(sequencesdepeche_vect[3], zonespersequence_vect[1]),
           c(sequencesdepeche_vect[3], zonespersequence_vect[2]),
-          c(sequencesdepeche_vect[3], zonespersequence_vect[3]))
+          c(sequencesdepeche_vect[3], zonespersequence_vect[3]),
+          c(sequencesdepeche_vect[4], zonespersequence_vect[4]))
 # x=1 : P=2 et Z=1
 # x=2 : P=2 et Z=3
 # x=3 : P=2 et Z=5
@@ -142,7 +144,7 @@ PZ = list(c(sequencesdepeche_vect[1], zonespersequence_vect[1]),
 # x=7 : P=10 et Z=1
 # x=8 : P=10 et Z=3
 # x=9 : P=10 et Z=5
-
+# x=10 : P=1 et Z=1
 
 
 ################
@@ -225,12 +227,18 @@ dyn.load( dynlib(paste0(TmbFile,"inst/executables/com_x_sci_data_14_scientific_c
 for(i in i0:n_sim){
[…]

### 2021-01-27 Version_M2_plusieurszones/Scripts/function/simu_commercial_data.R
modification processus d'observation

@@ -214,7 +214,7 @@ simu_commercial_data <- function(loc_x,
       # y_com_i <- rlnorm(1,meanlog = log(exp_catch), sdlog = exp(logSigma_com))
       # shape <- exp(logSigma_com)^-2
       # y_com_i <-  rgamma(1,shape=shape,scale= exp_catch * exp(logSigma_com)^2)
-      y_com_i <- exp(rnorm(1,mean = log(exp_catch) - exp(logSigma_com)^2/2, sd = exp(logSigma_com)))
+      y_com_i <- exp(rnorm(1,mean = log(exp_catch/prob_encount) - exp(logSigma_com)^2/2, sd = exp(logSigma_com)))
     }else{
       y_com_i <- 0
     }

### 2021-01-27 Version_M2_plusieurszones/Scripts/function/commercial_scientific_14.R
modification simu commerciale

@@ -474,7 +474,7 @@ simu_commercial_scientific <- function(Results,
     
     
     # save data
-    if(! dir.exists(simu_file)) dir.create(simu_file)
+    if(! dir.exists(simu_file)) dir.create(simu_file,recursive = T)
     save(List_param, file = paste0(simu_file,"/List_param_",counter,".RData"))
     save(Results, file = paste0(simu_file,"/Results.RData"))
     
@@ -488,7 +488,7 @@ simu_commercial_scientific <- function(Results,
   Strue_x_2 = cbind(Strue_x_2, "Champ_latent_estime" = champ_latent_estime)
   map_results = list(Strue_x_2, centres, peche_com_old, pointsdepeche_com)
   # On sauvegarde cette liste dans un workspace
-  if(! dir.exists(simu_file)) dir.create(simu_file)
+  if(! dir.exists(simu_file)) dir.create(simu_file,recursive = T)
   path = paste0(simu_file,"/map_results_i",i,"_b",b,"_k",k,"_x",x,".RData")
   save(map_results, file = path)
   

### 2021-01-27 Version_M2_plusieurszones/Scripts/function/simu_commercial_data.R

@@ -249,20 +249,31 @@ simu_commercial_data <- function(loc_x,
   
   b_com_i_2 <- factor(b_com_i_2)
   
-  # Visualisation des points de peche de chaque bateau dans sa zone et de la quantité pechee
-  peche_com_boat = peche_com_old
-  peche_com_boat$x = round(peche_com_boat$x)
-  peche_com_boat$y = round(peche_com_boat$y)
-  peche_com_boat$ncell = rep(0, length(peche_com_boat$boats))
-  for (j in 1:length(peche_com_boat$boats))
-  {
-    peche_com_boat[j, "ncell"] = loc_x$cell[which(loc_x[, "x"] == peche_com_boat[i, "x"] & loc_x[, "y"] == peche_com_boat[j, "y"])]
-  }
-  qte_pechee = as.data.frame(cbind(ncell = index_com_i_2, y_com_i = y_com_i))
-  peche_com_boat = peche_com_boat[order(peche_com_boat$ncell),]
-  qte_pechee = qte_pechee[order(qte_pechee$ncell),]
-  peche_com_boat = cbind(peche_com_boat, y_com_i = qte_pechee$y_com_i)
+  # # Visualisation des points de peche de chaque bateau dans sa zone et de la quantité pechee
+  # peche_com_boat = peche_com_old
+  # peche_com_boat$x = round(peche_com_boat$x)
+  # peche_com_boat$y = round(peche_com_boat$y)
+  # peche_com_boat$ncell = rep(0, length(peche_com_boat$boats))
+  # for (j in 1:length(peche_com_boat$boats))
+  # {
+  #   peche_com_boat[j, "ncell"] = loc_x$cell[which(loc_x[, "x"] == peche_com_boat[i, "x"] & loc_x[, "y"] == peche_com_boat[j, "y"])]
+  # }
+  # qte_pechee = as.data.frame(cbind(ncell = index_com_i, y_com_i = y_com_i))
+  # peche_com_boat = peche_com_boat[order(peche_com_boat$ncell),]
+  # qte_pechee = qte_pechee[order(qte_pechee$ncell),]
+  # peche_com_boat = cbind(peche_com_boat, y_com_i = qte_pechee$y_com_i)
+
   
[…]

### 2021-02-19 2021_realloc/02-model.Rmd
Proposition to explore new model

# A new model for commrecial data

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(purrr)
```

As mentionned in the introduction, every  commercial catch is reported by statistical area/vessel/gear and correspond potentially to several VMS ping.
Let's denote $D_j$  a reported catch associated to $|\mathcal{P}_j|$ VMS ping at positions $\left\lbrace x_i, i\in \mathcal{P}_j \right\rbrace$. $D_j$ is the sum of all individuals catch at the $n_j$ visited fishing points:

\begin{equation}
  D_{j} = \sum_{i \in  C_j} Y_i,
  (\#eq:logbook)
\end{equation}

$C_j$ begin the vector of all fishing events associated with the logbook entry $j$.

This diffculty is currently addressed by reallocating the total catch to individual fishing events with a simple proportionality rule:
$$\tilde{Y_i} = \frac{1}{|C_j|} D_j.$$

$\tilde{Y}_i$ is the assumed to follow the observation model defined  \@ref(eq:logbook).

A more satisfying solution would consist in deriving the distribution of $D_j$ from the distribution of $Y_i$. However the sum of a mixture of delta lognormal random variables  do not resume in a nice known distribution.

An alternative would be to fit a delta lognormal model whose proportion of zero and the two first moments coincide i.e

\begin{align}
P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_i = 0 \vert S, X),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi S(x_i)}\right \rbrace}.
\end{align}

The expected catch biomass is defined by $E(D_j) = \sum_{i\in \mathcal{P}_j} E(C_i Z_i).$

As $C_i$ and $Z_i$ are assumed to be independant

\begin{align}
E(D_j) & = \sum_{i\in \mathcal{P}_j} E(C_i Z_i),\nonumber \\
& = \sum_{i\in \mathcal{P}_j} (1-p_i) \frac{\mu(x_i)}{1-p_i}, \nonumber \\
& = \sum_{i\in \mathcal{P}_j} \mu(x_i). (\#eq:newobs)
\end{align}

Regarding the variance

$$Var(D_j)  = \sum_{i\in \mathcal{P}_j} Var(C_i Z_i).$$

We are the lead to compute the Variance of $C_i Z_i$.

\begin{align*}
Var(C_i Z_i) & = E( var(C_i Z_i \vert C_i) )  + Var( E(C_i Z_i \vert C_i) ),\\
& = Var(Z_i) E(C_i^2) + Var ( E(Z_i) C_i),\\
& = Var(Z_i) (1-p_i) + E(Z_i)^2 p_i (1-p_i)
\end{align*}

## Numerical illustration

### Precising the notation 
As mentioned in Equation \@ref(eq:LNv3),  $Y\sim \mathcal{LN}(\mu, \sigma^2)$ stands  a lognormal distribution such that $E(Y) = \mu$ and $Var(log(Y)) = \sigma^2$.

An alternative  specification denotes $Y \sim LN(\rho, \sigma)$ if $E(log(Y))=\rho$ and $Var(log(Y)) = \sigma^2$

The code below is an helper to go back and forth between the different parametrizations.
```{r lognomral_moments}

LN_variance_v1 <- function(rho, sigma){
  ( exp(sigma^2) - 1 ) * exp(2* rho  + sigma^2)
}

LN_mean_v1 <- function(rho, sigma){
   exp( rho + sigma^2/2 )
}

LN_rho2mu <- function(rho, sigma){
   LN_mean_v1(rho, sigma)
}

LN_mu2rho <- function(mu, sigma){
   log(mu) - sigma^2 /2
}

LN_variance_v3 <- function(mu, sigma){
  mu^2* (exp(sigma^2)-1)
}

```

### The sum of lognormal random variables is reasonnably log normal

#### n = 2

```{r sum_lognormal}
set.seed(1234)
n <- 2
rho <- rnorm(n, mean = 0, sd = 1 )
sigma <- 01
```

We are ready to look at the sum of `r n` lognormal random variables with mean `r round(LN_mean_v1(rho, sigma),3)`  and variance `r  round(LN_variance_v1(rho, sigma),3)` .

```{r sample_2_lognorm, echo = FALSE}
n_sim <- 1000
dta <- rho %>% map(function(x){ exp(rnorm(n = n_sim, mean = x, sd = sigma))}) %>% 
   map_df(
         ~ tibble(x = .x), 
         .id = "dist"
     )

sum_dta <- dta %>% group_by(dist) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = dist,    names_glue = "dist_{dist}",
 values_from =x) %>% 
  mutate( D = rowSums(.[grep("dist_", names(.))]))

p <- sum_dta %>% ggplot() + geom_histogram(aes(x=D, y = ..density..))

mu_sum <- sum(LN_rho2mu(rho, sigma))
S2_sum <- sum(LN_variance_v1(rho, sigma))
sigma_sum = sqrt( log(S2_sum/mu_sum^2 +1))

sum_density_LN <- tibble(x  = seq(0.001, max(sum_dta$D)), 0.1) %>% 
  mutate( log_x= log(x), dens = dnorm(log_x, 
                         mean=LN_mu2rho(mu_sum, sigma_sum), 
                         sd = sigma_sum)/x)

p_LN <- p + geom_line(data= sum_density_LN, aes(x=x, y =dens), col = 'blue')
sum_density_N <- tibble(x  = seq(0.001, max(sum_dta$D), 0.1)) %>% 
  mutate(dens = dnorm(x, 
                         mean=mu_sum, 
                         sd = sqrt(LN_variance_v3(mu_sum, sigma))))

p_N <- p_LN + geom_line(data= sum_density_N, aes(x=x, y =dens), col = 'red')
p_N
```

### n = 5

```{r sum_lognormal_n5}
n <- 5
rho <- rnorm(n, mean = 0, sd = 1 )
sigma <- 01
```

We are ready to look at the sum of `r n` lognormal random variables with mean `r round(LN_mean_v1(rho, sigma),3)`  and variance `r  round(LN_variance_v1(rho, sigma),3)` .

```{r sample_2_lognorm_n5, echo = FALSE}
n_sim <- 1000
dta <- rho %>% map(function(x){ exp(rnorm(n = n_sim, mean = x, sd = sigma))}) %>% 
   map_df(
         ~ tibble(x = .x), 
         .id = "dist"
     )

sum_dta <- dta %>% group_by(dist) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = dist,    names_glue = "dist_{dist}",
 values_from =x) %>% 
  mutate( D = rowSums(.[grep("dist_", names(.))]))

p <- sum_dta %>% ggplot() + geom_histogram(aes(x=D, y = ..density..))

mu_sum <- sum(LN_rho2mu(rho, sigma))
S2_sum <- sum(LN_variance_v1(rho, sigma))
sigma_sum = sqrt( log(S2_sum/mu_sum^2 +1))

sum_density_LN <- tibble(x  = seq(0.001, max(sum_dta$D)), 0.1) %>% 
  mutate( log_x= log(x), dens = dnorm(log_x, 
                         mean=LN_mu2rho(mu_sum, sigma_sum), 
                         sd = sigma_sum)/x)

p_LN <- p + geom_line(data= sum_density_LN, aes(x=exp(log_x), y =dens), col = 'blue')

sum_density_N <- tibble(x  = seq(0.001, max(sum_dta$D), 0.1)) %>% 
  mutate(dens = dnorm(x, 
                         mean=mu_sum, 
                         sd = sqrt(LN_variance_v3(mu_sum, sigma))))

p_N <- p_LN + geom_line(data= sum_density_N, aes(x=x, y =dens), col = 'red')
p_N
```

The expected value equals  `r mu_sum`, while the mean equals  `r mean(sum_dta$D)`. The variance equals `r LN_variance_v3(mu_sum, sigma_sum)` and the empiric variance equals `r var(sum_dta$D)`.

### What happens with the addittion of some zero inflation

### 2021-02-19 2021_realloc/_book/2021_realloc.tex
Proposition to explore new model

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{book}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Accounting for commercial data without reallocationg the catches.},
  pdfauthor={MP Etienne},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}

\newcommand{\point}{\mathcal{P}}
\usepackage[]{natbib}
\bibliographystyle{apalike}

\title{Accounting for commercial data without reallocationg the catches.}
\author{MP Etienne}
\date{2021-02-19}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{1}
\tableofcontents
}
\hypertarget{model-presentation}{%
\chapter{Model presentation}\label{model-presentation}}

\hypertarget{log-normal-parametrization}{%
\section{Log Normal parametrization}\label{log-normal-parametrization}}

Many different parametrizations of the log normal exist, We have to choose the most convenient for our purpose

\hypertarget{version-1}{%
\subsection{Version 1}\label{version-1}}

\begin{equation} 
  D \sim \mathcal{LN} (\rho, \sigma^2) \Longleftrightarrow D =e^{\rho + \sigma N}, \ N\sim \mathcal{N}(0,1) .
  \label{eq:LNv1}
\end{equation}

In this parametrization
\[E(D) = e^{\rho +\frac{\sigma^2}{2}},\quad Var(D) = \left(e^{\sigma^2}-1\right) e^{2\rho + \sigma^2}\]

\hypertarget{version-2}{%
\subsection{Version 2}\label{version-2}}

\begin{equation} 
  D \sim \mathcal{LN} (\nu, \sigma^2) \Longleftrightarrow D =\mu e^{\sigma N}, \ N\sim \mathcal{N}(0,1) .
  \label{eq:LNv2}
\end{equation}

In this parametrization
\[E(D) = \nu e^{\frac{\sigma^2}{2}},\quad Var(D) = \nu^2 \left(e^{\sigma^2}-1\right) e^{ \sigma^2}\]

\hypertarget{version-3-the-mean-of-the-log-normal-is-the-actual-parameter}{%
\subsection{Version 3, the mean of the log normal is the actual parameter}\label{version-3-the-mean-of-the-log-normal-is-the-actual-parameter}}

\begin{equation} 
  D \sim \mathcal{LN} (\mu, \sigma^2) \Longleftrightarrow D =\mu e^{\sigma N- \frac{\sigma^2}{2}}, \ N\sim \mathcal{N}(0,1) .
  \label{eq:LNv3}
\end{equation}

In this parametrization
\[E(D) = \mu ,\quad Var(D) = \mu^2 \left(e^{\sigma^2}-1\right). \]
With this parametrization,

\[D \sim\mathcal{LN}\left ( E(D), \ln\left\lbrace Var(D)/E(D)^2 +1\right\rbrace\right ).\]
We will use this last parametrization throughout the paper.

\hypertarget{model}{%
\section{Model}\label{model}}

The model proposed by Baptiste has a hierarchical structure with three main components

\begin{itemize}
\tightlist
\item
  the biomass latent field \(S\), representing the spatial repartition of biomass:
\end{itemize}

\begin{equation} 
  S(x)  = \exp{ \left \lbrace \alpha_S + \Gamma_S(x)^\intercal \beta_S +\delta(x)\right \rbrace}
  \label{eq:latent}
\end{equation}

\begin{itemize}
\tightlist
\item
  the preferential sampling density, defined conditionnally to \(S\):
\end{itemize}

\begin{equation} 
  \lambda(x)  = \exp{ \left \lbrace \alpha_X + b \log{\left \lbrace S(x)\right\rbrace} + \Gamma_X(x)^\intercal \beta_X +\eta(x)\right \rbrace}
  \label{eq:pref}
\end{equation}

\begin{itemize}
\tightlist
\item
  and finally, the observations are defined at the VMS ping scale. \(Y_i\) stands for the CPUE at location \(x_i\) and is defined conditionnaly on the Poisson process and the biomass latent field \(S\):
\end{itemize}

\begin{align*} 
  Y_i  &= C_i Z_i,    \label{eq:obs}\\ 
  C_i &\sim \mathcal{B}(1-p_i), \mbox{ with } p_i = \exp{ \left \lbrace  - e^{\xi \mu(x_i) }\right \rbrace}, \nonumber  \\
  Z_i &\sim \mathcal{LN}( \frac{\mu(x_i)}{1-p_i}, \sigma^2),\nonumber \\
  \mu(x_i) &= q S(x_i).\nonumber \\
\end{align*}

\hypertarget{potential-limitation}{%
\section{Potential limitation}\label{potential-limitation}}

The information regarding the catch is not known at the fishing point level (VMS ping) but at the statistical unit, as recorded in the logbook.
The available information on the commercial catch is the declaration \(D_j\) at the statistical unit level and the VMS ping location \((x_{j1}, \ldots x_{jn_j})\). \(D_j\) is the sum of all individuals catch at the \(n_j\) visited fishing points:

\begin{equation*}
  D_{j} = \sum_{i \in  \mathcal{P}_j} Y_i,
\end{equation*}

\(\mathcal{P}_j\) being the vector of all fishing events associated with the logbook entry \(j\).

This diffculty is currently addressed by reallocating the total catch to individual fishing events with a simple proportionality rule:
\[\tilde{Y_i} = \frac{1}{|\mathcal{P}_j|} D_j.\]

This is not completely satisfactory as it might introduce some bias is the total biomass estimation as well as its spatial repetition.

\hypertarget{is-is-a-real-limitation}{%
\section{Is is a real limitation ?}\label{is-is-a-real-limitation}}

A strategy to answer this question would be

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Simulation study to understand when this reallocation is problematic
\item
  A new observation model at the statitical unit level
\item
  Does this new model improves the estimation ?
\end{enumerate}

\hypertarget{effects}{%
\chapter{A simulation study to explore the effects of the reallocation}\label{effects}}

Put the stats project here

\hypertarget{a-new-model-for-commrecial-data}{%
\chapter{A new model for commrecial data}\label{a-new-model-for-commrecial-data}}

As mentionned in the introduction, every commercial catch is reported by statistical area/vessel/gear and correspond potentially to several VMS ping.
Let's denote \(D_j\) a reported catch associated to \(|\mathcal{P}_j|\) VMS ping at positions \(\left\lbrace x_i, i\in \mathcal{P}_j \right\rbrace\). \(D_j\) is the sum of all individuals catch at the \(n_j\) visited fishing points:

\begin{equation}
  D_{j} = \sum_{i \in  C_j} Y_i,
  \label{eq:logbook}
\end{equation}

\(C_j\) begin the vector of all fishing events associated with the logbook entry \(j\).

This diffculty is currently addressed by reallocating the total catch to individual fishing events with a simple proportionality rule:
\[\tilde{Y_i} = \frac{1}{|C_j|} D_j.\]

\(\tilde{Y}_i\) is the assumed to follow the observation model defined \eqref{eq:logbook}.

A more satisfying solution would consist in deriving the distribution of \(D_j\) from the distribution of \(Y_i\). However the sum of a mixture of delta lognormal random variables do not resume in a nice known distribution.

An alternative would be to fit a delta lognormal model whose proportion of zero and the two first moments coincide i.e

\begin{align}
P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_i = 0 \vert S, X),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi S(x_i)}\right \rbrace}.
\end{align}

The expected catch biomass is defined by \(E(D_j) = \sum_{i\in \mathcal{P}_j} E(C_i Z_i).\)

As \(C_i\) and \(Z_i\) are assumed to be independant

\begin{align}
E(D_j) & = \sum_{i\in \mathcal{P}_j} E(C_i Z_i),\nonumber \\
& = \sum_{i\in \mathcal{P}_j} (1-p_i) \frac{\mu(x_i)}{1-p_i}, \nonumber \\
& = \sum_{i\in \mathcal{P}_j} \mu(x_i). \label{eq:newobs}
\end{align}

Regarding the variance

\[Var(D_j)  = \sum_{i\in \mathcal{P}_j} Var(C_i Z_i).\]

We are the lead to compute the Variance of \(C_i Z_i\).

\begin{align*}
Var(C_i Z_i) & = E( var(C_i Z_i \vert C_i) )  + Var( E(C_i Z_i \vert C_i) ),\\
& = Var(Z_i) E(C_i^2) + Var ( E(Z_i) C_i),\\
& = Var(Z_i) (1-p_i) + E(Z_i)^2 p_i (1-p_i)
\end{align*}

\hypertarget{numerical-illustration}{%
\section{Numerical illustration}\label{numerical-illustration}}

\hypertarget{precising-the-notation}{%
\subsection{Precising the notation}\label{precising-the-notation}}

As mentioned in Equation \eqref{eq:LNv3}, \(Y\sim \mathcal{LN}(\mu, \sigma^2)\) stands a lognormal distribution such that \(E(Y) = \mu\) and \(Var(log(Y)) = \sigma^2\).

An alterntive commun specificication denotes \(Y \sim LN(\rho, \sigma)\) if \(E(log(Y))=\rho\) and \(Var(log(Y)) = \sigma^2\)

Look at the the sum of n lognormal distribution

The code below is an helper to go back and forth between the different parametrizations.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{log_normal_variance2 <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(rho, sigma)\{}
\NormalTok{  ( }\KeywordTok{exp}\NormalTok{(sigma}\OperatorTok{^}\DecValTok{2}\NormalTok{) }\OperatorTok{-}\StringTok{ }\DecValTok{1}\NormalTok{ ) }\OperatorTok{*}\StringTok{ }\KeywordTok{exp}\NormalTok{(}\DecValTok{2}\OperatorTok{*}\StringTok{ }\NormalTok{rho  }\OperatorTok{+}\StringTok{ }\NormalTok{sigma}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
\NormalTok{\}}

\NormalTok{LN_rho2mu <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(rho, sigma)\{}
   \KeywordTok{exp}\NormalTok{(rho}\OperatorTok{+}\StringTok{ }\NormalTok{sigma}\OperatorTok{^}\DecValTok{2} \OperatorTok{/}\DecValTok{2}\NormalTok{ )}
\NormalTok{\}}

\NormalTok{LN_mu2rho <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(mu, sigma)\{}
   \KeywordTok{log}\NormalTok{(mu) }\OperatorTok{-}\StringTok{ }\NormalTok{sigma}\OperatorTok{^}\DecValTok{2} \OperatorTok{/}\DecValTok{2}
\NormalTok{\}}

\NormalTok{log_normal_variance1 <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(mu, sigma)\{}
  \KeywordTok{log_normal_variance2}\NormalTok{(}\KeywordTok{LN_mu2rho}\NormalTok{(mu, sigma), sigma)}
\NormalTok{\}}

\NormalTok{log_normal_mean2 <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(rho, sigma)\{}
   \KeywordTok{exp}\NormalTok{( rho }\OperatorTok{+}\StringTok{ }\NormalTok{sigma}\OperatorTok{^}\DecValTok{2}\OperatorTok{/}\DecValTok{2}\NormalTok{ )}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{n-2}{%
\subsection{n = 2}\label{n-2}}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1234}\NormalTok{)}
\NormalTok{n <-}\StringTok{ }\DecValTok{2}
\NormalTok{rho <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(n, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \DecValTok{1}\NormalTok{ )}
\NormalTok{sigma <-}\StringTok{ }\DecValTok{01}
\end{Highlighting}
\end{Shaded}

We are ready to look at the sum of 2 lognormal random variables with mean 0.493, 2.176 and variance 0.418, 8.135 .

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n_sim <-}\StringTok{ }\DecValTok{1000}
\NormalTok{dta <-}\StringTok{ }\NormalTok{rho }\OperatorTok{%>%}\StringTok{ }\KeywordTok{map}\NormalTok{(}\ControlFlowTok{function}\NormalTok{(x)\{ }\KeywordTok{exp}\NormalTok{(}\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =}\NormalTok{ n_sim, }\DataTypeTok{mean =}\NormalTok{ x, }\DataTypeTok{sd =}\NormalTok{ sigma))\}) }\OperatorTok{%>%}\StringTok{ }
\StringTok{   }\KeywordTok{map_df}\NormalTok{(}
         \OperatorTok{~}\StringTok{ }\KeywordTok{data_frame}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ .x), }
         \DataTypeTok{.id =} \StringTok{"dist"}
\NormalTok{     ) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: `data_frame()` is deprecated as of tibble 1.1.0.
## Please use `tibble()` instead.
## This warning is displayed once every 8 hours.
## Call `lifecycle::last_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p <-}\StringTok{ }\NormalTok{dta }\OperatorTok{%>%}\StringTok{ }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{x, }\DataTypeTok{y=}\NormalTok{..density.., }\DataTypeTok{fill =}\NormalTok{ dist), }\DataTypeTok{position=}\StringTok{"identity"}\NormalTok{, )}

\NormalTok{p}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
\end{verbatim}

\includegraphics{2021_realloc_files/figure-latex/sample_2_lognorm-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sum_dta <-}\StringTok{ }\NormalTok{dta }\OperatorTok{%>%}\StringTok{ }\KeywordTok{group_by}\NormalTok{(dist) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{id =} \KeywordTok{row_number}\NormalTok{()) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ dist,    }\DataTypeTok{names_glue =} \StringTok{"dist_\{dist\}"}\NormalTok{,}
 \DataTypeTok{values_from =}\NormalTok{x) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{( }\DataTypeTok{D =} \KeywordTok{rowSums}\NormalTok{(.[}\KeywordTok{grep}\NormalTok{(}\StringTok{"dist_"}\NormalTok{, }\KeywordTok{names}\NormalTok{(.))]))}

\NormalTok{sum_dta }\OperatorTok{%>%}\StringTok{ }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{D), }\DataTypeTok{position=}\StringTok{"identity"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
\end{verbatim}

\includegraphics{2021_realloc_files/figure-latex/sample_2_lognorm-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu_sum <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{LN_rho2mu}\NormalTok{(rho, sigma))}
\NormalTok{S2_sum <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\KeywordTok{log_normal_variance2}\NormalTok{(rho, sigma))}
\NormalTok{sigma_sum =}\StringTok{ }\KeywordTok{sqrt}\NormalTok{( S2_sum}\OperatorTok{/}\NormalTok{mu_sum}\OperatorTok{^}\DecValTok{2} \OperatorTok{+}\DecValTok{1}\NormalTok{)}

\NormalTok{sum_density <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x  =} \KeywordTok{seq}\NormalTok{(}\FloatTok{0.001}\NormalTok{, }\KeywordTok{max}\NormalTok{(sum_dta}\OperatorTok{$}\NormalTok{D), }\FloatTok{0.1}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{log_x =} \KeywordTok{log}\NormalTok{(x), }
         \DataTypeTok{dens =} \KeywordTok{dnorm}\NormalTok{(log_x, }
                         \DataTypeTok{mean=}\KeywordTok{LN_mu2rho}\NormalTok{(mu_sum, sigma_sum), }
                         \DataTypeTok{sd =}\NormalTok{ sigma_sum))}

\NormalTok{p }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{ sum_density, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{x, }\DataTypeTok{y =}\NormalTok{dens))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
\end{verbatim}

\includegraphics{2021_realloc_files/figure-latex/sample_2_lognorm-3.pdf}

La moyenne théorique attendue est 2.6689604 et la moyenne empirique vaut 2.6173234, la variance théorique vaut 57.2084894 et la variance empirique vaut 7.0105572.

  \bibliography{book.bib,packages.bib}

\end{document}

### 2021-02-19 2021_realloc/index.Rmd
Proposition to explore new model

--- 
title: "Accounting for commercial data without reallocationg the catches."
author: "MP Etienne"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "Working at the statistical unit scale."
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(purrr)
```

# Model presentation

## Log Normal parametrization

Many different parametrizations of the log normal exist, We have to choose the most convenient for our purpose

### Version 1

\begin{equation} 
  D \sim \mathcal{LN} (\rho, \sigma^2) \Longleftrightarrow D =e^{\rho + \sigma N}, \ N\sim \mathcal{N}(0,1) .
  (\#eq:LNv1)
\end{equation}

In this parametrization 
$$E(D) = e^{\rho +\frac{\sigma^2}{2}},\quad Var(D) = \left(e^{\sigma^2}-1\right) e^{2\rho + \sigma^2}$$

### Version 2

\begin{equation} 
  D \sim \mathcal{LN} (\nu, \sigma^2) \Longleftrightarrow D =\mu e^{\sigma N}, \ N\sim \mathcal{N}(0,1) .
  (\#eq:LNv2)
\end{equation}

In this parametrization 
$$E(D) = \nu e^{\frac{\sigma^2}{2}},\quad Var(D) = \nu^2 \left(e^{\sigma^2}-1\right) e^{ \sigma^2}$$

### Version 3, the mean of the log normal is the actual parameter

\begin{equation} 
  D \sim \mathcal{LN} (\mu, \sigma^2) \Longleftrightarrow D =\mu e^{\sigma N- \frac{\sigma^2}{2}}, \ N\sim \mathcal{N}(0,1) .
  (\#eq:LNv3)
\end{equation}

In this parametrization 
$$E(D) = \mu ,\quad Var(D) = \mu^2 \left(e^{\sigma^2}-1\right). $$
With this parametrization,

$$D \sim\mathcal{LN}\left ( E(D), \ln\left\lbrace Var(D)/E(D)^2 +1\right\rbrace\right ).$$
We will use this last parametrization throughout the paper.

## Model

The model proposed by Baptiste has a hierarchical structure with three main components

- the biomass latent field $S$, representing the spatial repartition of biomass:

\begin{equation} 
  S(x)  = \exp{ \left \lbrace \alpha_S + \Gamma_S(x)^\intercal \beta_S +\delta(x)\right \rbrace}
  (\#eq:latent)
\end{equation}

- the preferential sampling density, defined conditionnally to $S$:

\begin{equation} 
  \lambda(x)  = \exp{ \left \lbrace \alpha_X + b \log{\left \lbrace S(x)\right\rbrace} + \Gamma_X(x)^\intercal \beta_X +\eta(x)\right \rbrace}
  (\#eq:pref)
\end{equation}

- and finally, the observations are defined at the VMS ping scale. $Y_i$ stands for the CPUE at location $x_i$ and is defined conditionnaly on  the Poisson process and the biomass latent field $S$:

\begin{align*} 
  Y_i  &= C_i Z_i,    (\#eq:obs)\\ 
  C_i &\sim \mathcal{B}(1-p_i), \mbox{ with } p_i = \exp{ \left \lbrace  - e^{\xi \mu(x_i) }\right \rbrace}, \nonumber  \\
  Z_i &\sim \mathcal{LN}( \frac{\mu(x_i)}{1-p_i}, \sigma^2),\nonumber \\
  \mu(x_i) &= q S(x_i).\nonumber \\
\end{align*}

## Potential limitation

The information regarding the catch is not known at the fishing point level (VMS ping) but at the statistical unit, as recorded in the logbook.
The available information on the commercial catch is the declaration $D_j$  at the statistical unit level and the VMS ping location $(x_{j1}, \ldots x_{jn_j})$. $D_j$ is the sum of all individuals catch at the $n_j$ visited fishing points:

\begin{equation*}
  D_{j} = \sum_{i \in  \mathcal{P}_j} Y_i,
\end{equation*}

$\mathcal{P}_j$ being the vector of all fishing events associated with the logbook entry $j$.

This diffculty is currently addressed by reallocating the total catch to individual fishing events with a simple proportionality rule:
$$\tilde{Y_i} = \frac{1}{|\mathcal{P}_j|} D_j.$$

This is not completely satisfactory as it might introduce some bias is the total biomass estimation as well as its spatial repetition.

## Is is a real limitation ?

A strategy to answer this question would be

1. Simulation study to understand when this reallocation is problematic
2. A new observation model at the statitical unit level
3. Does this new model improves the estimation ?

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

### 2021-02-21 2021_realloc/02-model.Rmd
Aggregated model derivation

@@
 P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_i = 0 \vert S, X),\nonumber \\
-                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi S(x_i)}\right \rbrace}.
+                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi S(x_i)}\right \rbrace} = \pi_j.
 \end{align}
 
-The expected catch biomass is defined by $E(D_j) = \sum_{i\in \mathcal{P}_j} E(C_i Z_i).$
-
+The expected catch biomass is defined by $E(D_j) = \sum_{i\in \mathcal{P}_j} E(C_i Z_i),$ while the expected positive catch biomass is defined by 
+$$E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_i Z_i\vert \exists i_0\in\mathcal{P}_j , C_{i_0}=1)$$
 As $C_i$ and $Z_i$ are assumed to be independant
 
+\begin{align*}
+E(D_j\vert D_j > 0) & = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / P\left ( D_j > 0\right ),   \\
+& = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / \left (1-\pi_j\right).   \\
+\end{align*}
+
+As $E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } ) = E(D_j ),$
+
 \begin{align}
-E(D_j) & = \sum_{i\in \mathcal{P}_j} E(C_i Z_i),\nonumber \\
-& = \sum_{i\in \mathcal{P}_j} (1-p_i) \frac{\mu(x_i)}{1-p_i}, \nonumber \\
-& = \sum_{i\in \mathcal{P}_j} \mu(x_i). (\#eq:newobs)
+E(D_j\vert D_j > 0) & = \left (1- \pi_j\right)^{-1} E(D_j)   ,   \nonumber \\
+& \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_i Z_i),\nonumber \\
+& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_i) \frac{\mu(x_i)}{1-p_i}, \nonumber \\
+& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} \mu(x_i). (\#eq:newobs)
 \end{align}
@@
 
-$$Var(D_j)  = \sum_{i\in \mathcal{P}_j} Var(C_i Z_i).$$
+$$Var(D_j \vert D_j >0)  = E(D_j^2 \vert D_j >0)- E(D_j \vert D_j >0)^2.$$
+
+But $E(D_j^2 \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 1_{\left \lbrace D_j >0\right\rbrace}) = (1-\pi_j)^{-1} E(D_j^2 )$ and
+
+$E(D_j \vert D_j >0)^2 = ((1-\pi_j)^{-1} E(D_j 1_{\left \lbrace D_j >0\right\rbrace}))^2 =  (1-\pi_j)^{-2} E(D_j)^2$.
+
+So that
+
+$$Var(D_j \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 ) - (1-\pi_j)^{-2} E(D_j)^2 = (1-\pi_j)^{-1} Var(D_j) - \frac{\pi_j}{(1-\pi_j)^2} E(D_j)^2.$$
+
+As the $(Y_i )_{i\in \mathcal{p}_j}$ are independent, $Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_i)$.
 
@@
 \begin{align*}
-Var(C_i Z_i) & = E( var(C_i Z_i \vert C_i) )  + Var( E(C_i Z_i \vert C_i) ),\\
-& = Var(Z_i) E(C_i^2) + Var ( E(Z_i) C_i),\\
-& = Var(Z_i) (1-p_i) + E(Z_i)^2 p_i (1-p_i)
+Var(C_i Z_i) & = E( C_i^2 Z_i^2 ) - E(C_i Z_i)^2,\\
+& =  E( C_i^2) E(Z_i^2 ) -E(C_i)^2 E(Z_i)^2,\\
+& = p_i E(Z_i^2 ) - p_i^2 E(Z_i)^2 = p_i ( Var(Z_i) + (1-p_i) E(Z_i)^2),\\
+& = p_i\left( \frac{\mu(x_i)^2 }{(1-p_i)^2} (e^{\sigma^2}-1) + (1-p_i) \frac{\mu(x_i)^2 }{(1-p_i)^2}\right),\\
+& = \frac{p_i}{(1-p_i)^2} \mu(x_i)^2  \left(e^{\sigma^2} -p_i \right)  
 \end{align*}

### 2021-02-23 Projet_Vannes/Code R rectifié.R
Create Code R rectifié.R

@@ -1,0 +1,326 @@
+load("C:/Users/Anis/Documents/Projet d'étude/Documents d'étude/tacsatEflalo_2.RData")
+tacsatEflalo_2= dplyr::filter(tacsatEflalo_2, SI_STATE == 1)
+
+
+install.packages("rnaturalearth")
+install.packages("sf")
+install.packages("mapdata")
+install.packages("ggplot2")
+install.packages("ggspatial")
+install.packages("rgeos")
+library(rnaturalearth)
+library(sf)
+library(mapdata)
+library(ggplot2)
+library(ggspatial)
+library(rgeos)
+
+# setwd("C:/R_projects/phd_zfh_baptiste_alglave/examples/map/scientific_data")
+setwd("C:/Users/Anis/Documents/Projet d'étude/scientific_data")
+# C:\Users\Anis\Documents\Projet d'étude\scientific_data.zip
+
+load("res/ptsurvey_wgs84_full.RData")
+load("map/strata_ZN.RData")
+load("map/strata_ZCC.RData")
+load("map/strata_ZCL.RData")
+load("map/strata_ZS.RData")
+
+mapBase <- ne_countries(scale = "medium", returnclass = "sf")
+# mapBase <- map("worldHires", fill = T, plot = F)
+# mapBase <- st_as_sf(mapBase) # %>% filter(ID %in% c("France","Spain"))
+
+# for (j in 23:50)
+# {
+plot <- ggplot() +
+  geom_sf(data = strata_ZN,
+          alpha = 0.35,
+          size = 0.75,
+          color="#7FC97F",
+          fill ="#7FC97F")+
+  geom_sf(data = strata_ZS,
+          alpha = 0.35,
+          size = 0.75,
+          color="#FFED6F",
+          fill="#FFED6F")+
+  geom_sf(data = strata_ZCC,
+          alpha = 0.35,
+          size = 0.75,
+          color="white",
+          fill="white")+
+  geom_sf(data = strata_ZCL,
+          alpha = 0.35,
+          size = 0.75,
+          color="#EF8A62",
+          fill="#EF8A62")+
+  geom_sf(data = mapBase)+
+  
+  geom_point(data = tacsatEflalo_2,
+             aes(x=SI_LONG,y=SI_LATI,col = tacsatEflalo_2[,35]))+
[…]

### 2021-08-23 Version_M2_plusieurszones/Scripts/In_progress/plot_maps.R
modif estimation model + parallel
Add the reallocation component to the estimation model.
Parallelize simulations.

@@ -1,0 +1,84 @@
+
+test <- Obj$report()
+
+list_test <- list(y_com_i = Data$y_com_i,
+                  E_com=test$E_com,
+                  log_pi_j=test$log_pi_j,
+                  pi_j=test$pi_j,
+                  sum_mu=test$sum_mu,
+                  Var_D_com=test$Var_D_com,
+                  p_i=test$p_i,
+                  
+                  Var_Dsup0_com=test$Var_Dsup0_com,
+                  E_D_com=test$E_D_com,
+                  encounterprob_com=test$encounterprob_com,
+                  Sigma_D=test$Sigma_D,
+                  log_Sigma_D=test$log_Sigma_D)
+
+data.frame(index_com_i = index_com_i,
+           E_com = list_test$E_com,
+           p_i = list_test$p_i,
+           sum_mu = list_test$sum_mu,
+           Var_D_com = list_test$Var_D_com,
+           
+           E_D_com = list_test$E_D_com,
+           pi_j = list_test$pi_j,
+           log_pi_j = list_test$log_pi_j,
+           
+           y_com_i = list_test$y_com_i
+           
+           )
+
+# Var_D_com
+test$p_i / (1 - test$p_i)^2 * test$E_com^2 * ((exp(exp(logSigma_com)^2)) - test$p_i)
+
+
+## Diagnose model
+plot(Strue_x,Report$S_x)
+
+loc_x_2 <- loc_x %>%
+        mutate(Strue_x = Strue_x,
+               obs_ref = fit_IM_res_ref$Report$S_x,
+               obs_non_aggreg = fit_IM_res_non_aggreg$Report$S_x,
+               obs_aggreg = fit_IM_res_aggreg$Report$S_x
+               )
+
+plot_1 <- ggplot(loc_x_2)+
+        geom_point(aes(x=x,y=y,col=Strue_x),size=4,shape=15)+
+        scale_color_distiller(palette = "Spectral")+
+        theme(aspect.ratio = 1,
+              legend.title = element_blank(),
+              plot.title = element_text(hjust = 0.5,size = 12,face = "bold"))+
+        ggtitle("Simulated latent field")+
+        xlab("")+ylab("")
+
+plot_2 <- ggplot(loc_x_2)+
+        geom_point(aes(x=x,y=y,col=obs_ref),size=4,shape=15)+
+        scale_color_distiller(palette = "Spectral")+
+        theme(aspect.ratio = 1,
+              legend.title = element_blank(),
[…]

### 2021-08-23 Version_M2_plusieurszones/Scripts/function/commercial_scientific_14.R

@@ -87,6 +87,7 @@ simu_commercial_scientific <- function(Results,
                                        Data_source,
                                        Samp_process,
                                        EM,
+                                       aggreg_obs,
                                        RandomSeed,
                                        Version,
                                        TmbFile,
@@ -99,7 +100,12 @@ simu_commercial_scientific <- function(Results,
                                        ylim,
                                        sequencesdepeche,
                                        zonespersequence,
-                                       taillezone){
+                                       taillezone,
+                                       i0,
+                                       i1,
+                                       n_fact,
+                                       cluster_nb,
+                                       n_nodes){
 
   ################
   ## Simulate data
@@ -273,8 +279,8 @@ simu_commercial_scientific <- function(Results,
     y_com_i <- commercial_reallocation_uniforme(k, xlim, ylim, y_com_i, n_samp_com,
                                                 index_com_i, loc_x, sequencesdepeche, boats_number)
   }
-  # plot(Strue_x[index_com_i],(y_com_i),xlab="S_x",ylab="commercial obs. (reallocated)")
-  # plot(y_com_i,y_com_i_ref,xlab="commercial obs. (true)",ylab="commercial obs. (reallocated)")
+  plot(Strue_x[index_com_i],(y_com_i), xlab="S_x",ylab="commercial obs. (reallocated)")
+  plot(x = y_com_i_ref,y= y_com_i,col = boats_number,xlab="true commercial obs.",ylab="reallocated commercial obs.")
   
   
   # Représentation graphique des donnees commerciales (4 graphes)
@@ -339,6 +345,7 @@ simu_commercial_scientific <- function(Results,
                          index_com_i,
                          y_sci_i,
                          index_sci_i,
+                         aggreg_obs,
                          boats_number,
                          as.matrix(Cov_x[,1]))
     
@@ -398,7 +405,12 @@ simu_commercial_scientific <- function(Results,
                                 index_com_i=index_com_i,
                                 y_com_i=y_com_i,
                                 # b_com_i=b_com_i,
-                                c_com_x=c_com_x)
+                                c_com_x=c_com_x,
[…]

### 2021-08-23 Version_M2_plusieurszones/Scripts/function/fit_IM.R

@@ -48,6 +48,7 @@ fit_IM <- function(Estimation_model_i = 1,
                    index_com_i,
                    y_sci_i,
                    index_sci_i,
+                   aggreg_obs,
                    boats_number,
                    Cov_x){
   
@@ -75,7 +76,8 @@ fit_IM <- function(Estimation_model_i = 1,
                    'zero.infl_model' = 2,  # (DEPRECATED)
                    'commercial_obs' = 1, # (DEPRECATED)
                    'b_constraint' = 2, # (DEPRECATED)
-                   'catchability_random' = 0)  # (DEPRECATED)
+                   'catchability_random' = 0, # (DEPRECATED)
+                   'aggreg_obs' = ifelse(aggreg_obs == T,1,0)) 
   
   ## Data & Params
   Map = list() #liste vide : correspond aux paramètres fixés dans l'estimation
@@ -151,36 +153,20 @@ fit_IM <- function(Estimation_model_i = 1,
     Map[["k_com"]][1] <- NA
     Map[["k_com"]] <- factor(Map[["k_com"]])
     
-  }else if(Estimation_model_i == 5){ # Commercial model with aggregated data at the level of fishing sequence
+  }
+  
+  if(aggreg_obs == T){ # Commercial model with aggregated data at the level of fishing sequence
     
     # commercial model (commercial_only)
-    
     aggreg_data_df <- data.frame(y_com_i= y_com_i,seq = boats_number) %>%
       group_by(seq) %>%
       dplyr::summarise(y_com_i = sum(y_com_i))
     
+    Data[["y_com_i"]] = aggreg_data_df$y_com_i
+    Data[["seq_com_i"]] = boats_number - 1
     
-    Data = list( "Options_vec"=Options_vec,
-                 "c_com_x"=c_com_x,
-                 "y_com_i"=aggreg_data_df$y_com_i,
-                 "index_com_i"=index_com_i-1,
-                 "Cov_xj"=cbind(1,Cov_x),
-                 "Cov_xk"=array(1,c(list(nrow(Cov_x),1))),
-                 "q2_sci" =  1,
-                 "q2_com" = 1,
-                 "seq" = boats_number)
-    
-    Params = list("beta_j"=rep(0,ncol(Data$Cov_xj)), # linear predictor for abundance 
-                  "beta_k"=0, # intercept of fishin intensity
[…]

### 2021-08-23 Version_M2_plusieurszones/Scripts/function/simu_commercial_data.R

@@ -115,8 +115,11 @@ simu_commercial_data <- function(loc_x,
     # et on passe dans le rpoint Int_ps qui correspond
     # aux lambdas qui dépendent du champ latent
     win_df_boat <- owin(xrange=c(max(0.5, X$x[centre]-taillezone), min(25.5, X$x[centre]+taillezone)), yrange=c(max(0.5, X$y[centre]-taillezone), min(25.5, X$y[centre]+taillezone)), ystep=1, xstep=1)
-    n_samp <- round(n_samp_com/(sequencesdepeche*zonespersequence))+1 # + 1 --> sinon il manque des échantillons
+    n_samp <- round(n_samp_com/(sequencesdepeche*zonespersequence)) # + 1 --> sinon il manque des échantillons
     X_boat <- rpoint(n_samp, as.im(Int_ps, win_df_boat))
+    
+    # if()
+    
     # On ajoute les coordonnées de peche obtenues pour le bateau
     boats_x = c(boats_x, X_boat$x)
     boats_y = c(boats_y, X_boat$y)
@@ -124,11 +127,23 @@ simu_commercial_data <- function(loc_x,
     
   }
   
-  # sample n_samp_com in boats_y, boats_x, boats
-  index_com <- sample(1:(length(boats_y)),n_samp_com,replace = FALSE)
-  boats_x <- boats_x[index_com]
-  boats_y <- boats_y[index_com]
-  boats <- boats[index_com]
+  ## Si les chiffres tombent pas ronds, il manquera qql points de données 
+  ## --> on complète et les pings restant sont réalloués de facon aléatoire au bateaux
+  if(length(boats) < n_samp_com){
+    
+    n_samp_comp <- n_samp_com - length(boats)
+    X_boat_comp <- rpoint(n_samp_com - length(boats), as.im(Int_ps, win_df_boat))
+    boats_x = c(boats_x, X_boat_comp$x)
+    boats_y = c(boats_y, X_boat_comp$y)
+    boats = c(boats, sample(boats,n_samp_comp,replace = T))
+
+  }
+  
+  # # sample n_samp_com in boats_y, boats_x, boats
+  # index_com <- sample(1:(length(boats_y)),n_samp_com,replace = FALSE)
+  # boats_x <- boats_x[index_com]
+  # boats_y <- boats_y[index_com]
+  # boats <- boats[index_com]
 
   # On a donc :
   # boats_x les coordonnées sur l'axe des x des 150 points de peche

### 2021-08-23 Version_M2_plusieurszones/Scripts/inst/executables/com_x_sci_data_14_scientific_commercial_simple.R

@@ -1,7 +1,7 @@
 
       library( TMB )
       dyn.load( dynlib('Scripts/inst/executables/com_x_sci_data_14_scientific_commercial_simple') )
-      setwd('C:/R_projects/projet_M2_stats')
+      setwd('C:/R_projects/projet_M2_sole/Version_M2_plusieurszones')
       load( 'All_inputs.RData' )
       Obj = MakeADFun(data=All_inputs[['data']], parameters=All_inputs[['parameters']], random=All_inputs[['random']], All_inputs[['Other_inputs']])
       save( Obj, file='Obj.RData')

### 2021-08-23 Version_M2_plusieurszones/Scripts/inst/executables/com_x_sci_data_14_scientific_commercial_simple.cpp

@@ -149,10 +149,12 @@ Type objective_function<Type>::operator() (){
 
     DATA_VECTOR( c_com_x );       	// Response (count) for each observation i (commercial data)
     DATA_VECTOR( y_com_i );       	// Response (0:not surveyed, 1:surveyed) for each site (commercial data)
+    DATA_FACTOR( seq_com_i );       	// Response (0:not surveyed, 1:surveyed) for each site (commercial data)
+    
     DATA_FACTOR( index_com_i );
    	DATA_VECTOR( q2_com );
     DATA_MATRIX( Cov_xk );  //design matrix for sampling of commercial data
-
+    
 
     //////////////
     // Parameters
@@ -167,7 +169,8 @@ Type objective_function<Type>::operator() (){
     /////////////////
 
     int n_com_i = y_com_i.size();
-
+    int n_ping_i = seq_com_i.size();
+    
 
     /////////////////////
     // Observation model
@@ -176,26 +179,102 @@ Type objective_function<Type>::operator() (){
     if( Options_vec(9) == 1){
       
       vector<Type> Sigma_com = exp(logSigma_com);
-      vector<Type> E_com(n_com_i);
+      vector<Type> E_com(n_ping_i);
+      
       vector<Type> encounterprob_com(n_com_i);
       vector<Type> log_notencounterprob_com(n_com_i);
       
-      // Zero-inflated gamma distribution
-      for(int i=0; i<n_com_i; i++){
-        if( !isNA(y_com_i(i)) ){
-          
-          E_com(i) = q2_com(0) * 1 * S_x(index_com_i(i));
-          
-          // proba of encountering
-          encounterprob_com(i) = ( 1.0 - exp(-1 * E_com(i) * exp(q1_com(0))) );
-          log_notencounterprob_com(i) = -1 * E_com(i) * exp(q1_com(0));
-          
-          jnll_comp(0) -= dzinflognorm(y_com_i(i), log(E_com(i))-log(encounterprob_com(i)), encounterprob_com(i), log_notencounterprob_com(i), Sigma_com(0), true);
-          // jnll_comp(0) -= dzinfgamma(y_com_i(i), E_com(i)/encounterprob_com(i), encounterprob_com(i), log_notencounterprob_com(i), Sigma_com(0), true);
-          
[…]

### 2021-08-23 Version_M2_plusieurszones/Scripts/simu_main_script.R

@@ -3,475 +3,516 @@
 ####################
 # B. Alglave based on Conn et al. (2017)
 
-source("Scripts/function/load_packages.R")
-
-#-----------------------------------------------------------
-#-------------- Simulation/estimation settings -------------
-#-----------------------------------------------------------
+# source("Scripts/function/load_packages.R")
+library(parallel)
 
 # Simulation name
 simu_name = "SimuComplete"
 
-#---------------------
-# Simulation scenarios
-#---------------------
-
-###
-## Latent field
-###
-
-# Grid dimension
-grid_side_x <- 25
-grid_side_y <- 25
-grid_dim = c("x"=grid_side_x, "y"=grid_side_y)
-n_cells = grid_dim[1]*grid_dim[2]
-
-# Intercepts and covariates of the latent field
-beta0 = 3 # intercept
-beta = c(0,0,0,0,0) # covariates
-# parameters covariates vector : 1st param : continuous variable, 4 nexts : strata_1, strata_2, strata_3, strata_4
-# La premiere covariable est continue et a une structure spatiale
-# Afin de simuler cette corrélation spatiale on a besoin de deux paramètres :
-# - la portée de cette corrélation (nu) (+ la portée est grande et + la distance sur 
-#   laquelle les points sont corrélées entre eux est importante, et donc plus la 
-#   distance est grande, plus il faut "s'éloigner" pour avoir des points indépendants)
-# - la variance marginale de ce processus (écart type SD_x) (cad à quel point ce signal varie)
-
-
-# Correlation structure of the continuous covariate 
-simu_tool <- "MaternCov" # 2 way to simulate latent fields : "RandomFields", "MaternCov"
-range = sqrt(prod(grid_dim))/(5)*2 # --> MaternCov
-nu = 1  #portée de la covariable continue
-SD_x = 2 # ecart type/ variance marginale du processus de la var continue
-# Initialement on avait SD_x = 0.5, la on passe a SD_x = 2 pour avoir plus de variabilité dans le champ latent
-
-#effets spatiaux aléatoires
-SD_delta = SD_eta= 0.5
-
-
-###
-## Commercial sampling process
-###
-
-# intercept pour l'intensite lambda
-beta0_fb = 2 
-
-###
-## Observation process
-###
-
-## Scientific data
-# Number of scientific data
[…]

### 2021-09-19 2021_realloc/test.Rmd
correction estimation model
slight mistake in the expression of the moments

--- 2021-02-21 2021_realloc/02-model.Rmd
@@
+---
+title: "Untitled"
+author: "Baptiste Alglave"
+date: "15/09/2021"
+output: html_document
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
 # A new model for commrecial data
@@
 ```{r include=FALSE}
-library(tidyverse)
+library(dplyr)
+library(tidyr)
 library(ggplot2)
@@
 An alternative would be to fit a delta lognormal model whose proportion of zero and the two first moments coincide i.e
+
+## Fitting a Delta Log normal distribution for the catch
+
+Recals that at the VMS ping level $Y_i = C_i Z_i, $
+
+with
+
+$$C_i \sim  mathcal{B}(p_i), \quad p_i = $$
+
+### the probability of an empty catch
 
@@
 & =  E( C_i^2) E(Z_i^2 ) -E(C_i)^2 E(Z_i)^2,\\
-& = p_i E(Z_i^2 ) - p_i^2 E(Z_i)^2 = p_i ( Var(Z_i) + (1-p_i) E(Z_i)^2),\\
-& = p_i\left( \frac{\mu(x_i)^2 }{(1-p_i)^2} (e^{\sigma^2}-1) + (1-p_i) \frac{\mu(x_i)^2 }{(1-p_i)^2}\right),\\
-& = \frac{p_i}{(1-p_i)^2} \mu(x_i)^2  \left(e^{\sigma^2} -p_i \right)  
+& = (1- p_i) E(Z_i^2 ) - (1-p_i)^2 E(Z_i)^2 = (1-p_i) ( Var(Z_i)  + p_i E(Z_i)^2),\\
+& = (1-p_i)\left( \mu(x_i)^2 (e^{\sigma^2}-1) + p_i \mu(x_i)^2 \right),\\
+& = (1-p_i) \mu(x_i)^2 \left ( e^{\sigma^2}-1  + p_i \right),\\
 \end{align*}

### 2021-09-19 Version_M2_plusieurszones/Scripts/In_progress/plot_from_list_param.R

@@ -1,0 +1,118 @@
+#----------------------
+## Plot from List_param
+#----------------------
+
+folder_name <- "C:/Users/balglave/Desktop/com_x_sci_data_14_scientific_commercial_simple-2021-09-03_16_24_00_SimuTest/"
+
+for(i in 1:10){
+  load(paste0(folder_name,"/",i,"/Results",i,".RData"))
+  if(i==1) Results_2 <- Results
+  if(i!=1) Results_2 <- rbind(Results,Results_2)
+}
+library(stringr)
+for(i in 1){
+  
+  # Load Results
+  load(paste0(folder_name,"/",i,"/Results",i,".RData"))
+  
+  # Load List_param
+  list_file_i <- list.files(paste0(folder_name,"/",i))
+  list_file_i <- list_file_i[which(str_detect(list_file_i,"List_param"))]
+  
+  Results$aggreg_obs <- NA
+  Results$sequencesdepeche <- NA
+  Results$zonespersequence <- NA
+  Results$n_samp_com <- NA
+  Results$rho_S <- NA
+  Results$beta1_true <- NA
+  Results$beta1_est <- NA
+  
+  # Run over List_param
+  for(counter in Results$counter){
+    print(paste0("i: ",i," | counter: ",counter))
+    load(paste0(folder_name,"/",i,"/List_param_",counter,".RData"))
+    Results$aggreg_obs[which(Results$counter == counter)] <- List_param$data.info$aggreg_obs
+    Results$sequencesdepeche[which(Results$counter == counter)] <- List_param$data.info$sequencesdepeche
+    Results$zonespersequence[which(Results$counter == counter)] <- List_param$data.info$zonespersequence
+    Results$n_samp_com[which(Results$counter == counter)] <- List_param$data.info$n_samp_com
+    Results$rho_S[which(Results$counter == counter)] <- cor(List_param$data.info$Strue_x,List_param$Report$S_x)
+    Results$beta1_true[which(Results$counter == counter)] <- List_param$data.info$beta[1]
+    Results$beta1_est[which(Results$counter == counter)] <- List_param$Report$beta_j[2]
+  }
+  
+  # Build full Results
+  if(i==1) Results_2 <- Results
+  if(i!=1) Results_2 <- rbind(Results,Results_2)
+  
+}
+
+library(doBy)
+summaryBy(Convergence~
+            sequencesdepeche+
+            zonespersequence+
+            aggreg_obs+
[…]

### 2021-09-19 Version_M2_plusieurszones/Scripts/function/commercial_scientific_14.R

@@ -490,6 +490,7 @@ simu_commercial_scientific <- function(Results,
     
     Results[counter, "x"] = x
     Results[counter, "reallocation"] = k
+    Results[counter, "aggreg_obs"] = aggreg_obs
     
     Results[counter,"core"]=cluster_nb
     

### 2021-09-19 Version_M2_plusieurszones/Scripts/function/fit_IM.R

@@ -185,7 +185,7 @@ fit_IM <- function(Estimation_model_i = 1,
     # Map[["logSigma_com"]] <- factor(NA)
     
     Params$par_b <- b
-    Map[["par_b"]] <- factor(NA)
+    # Map[["par_b"]] <- factor(NA)
   }
   
   #-----------

### 2021-09-19 Version_M2_plusieurszones/Scripts/inst/executables/com_x_sci_data_14_scientific_commercial_simple.cpp

@@ -39,10 +39,14 @@ Type square(Type x){
   return pow(x,2.0);
 }
 
-// dzinfgamma, shape = 1/CV^2, scale = mean*CV^2
-
+// square-root
 template<class Type>
+Type sqrt(Type x){
+  return pow(x,0.5);
+}
 
+// dzinfgamma, shape = 1/CV^2, scale = mean*CV^2
+template<class Type>
 Type dzinfgamma(Type x, Type posmean, Type encounter_prob, Type log_notencounter_prob, Type cv, int give_log=false){
   Type Return;
   if(x==0){
@@ -130,8 +134,8 @@ Type objective_function<Type>::operator() (){
   ///////////////
 
   // Predicted densities
-  vector<Type> S_x(n_x);
-  vector<Type> debug_j(n_j);
+  vector<Type> S_x(n_x);S_x.setZero();
+  vector<Type> debug_j(n_j);debug_j.setZero();
   vector<Type> linpredS_x = Cov_xj * beta_j;
   for(int s=0; s<n_x; s++){
     S_x(s) = exp(linpredS_x(s) );
@@ -179,10 +183,10 @@ Type objective_function<Type>::operator() (){
     if( Options_vec(9) == 1){
       
       vector<Type> Sigma_com = exp(logSigma_com);
-      vector<Type> E_com(n_ping_i);
+      vector<Type> E_com(n_ping_i);E_com.setZero();
       
-      vector<Type> encounterprob_com(n_com_i);
-      vector<Type> log_notencounterprob_com(n_com_i);
+      vector<Type> encounterprob_com(n_com_i);encounterprob_com.setZero();
+      vector<Type> log_notencounterprob_com(n_com_i);log_notencounterprob_com.setZero();
       
       ///////////////////////////////////
       // Non-aggregated observation model
@@ -207,13 +211,13 @@ Type objective_function<Type>::operator() (){
       ///////////////////////////////
       // Aggregated observation model
       ///////////////////////////////
-      vector<Type> E_D_com(n_com_i);
-      vector<Type> log_pi_j(n_com_i);
-      vector<Type> pi_j(n_com_i);
-      vector<Type> sum_mu(n_com_i);
-      vector<Type> Var_D_com(n_com_i);
-      vector<Type> Var_Dsup0_com(n_com_i);
-      vector<Type> p_i(n_ping_i);
[…]

### 2021-09-19 Version_M2_plusieurszones/Scripts/simu_main_script.R

@@ -241,7 +241,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   
   ## Results dataframe
   n_cov = 5
-  colnames_Results <- c("counter","sim","b_true","Data_source","type_b","Alpha","b_est","ObsMod","Sigma_com_true","Sigma_sci_true","Sigma_com","Sigma_sci","n_samp_com","n_samp_sci","N_true","N_est","SD_N","Convergence","LogLik","MSPE_S","k", "x","reallocation")
+  colnames_Results <- c("counter","sim","b_true","Data_source","type_b","Alpha","b_est","ObsMod","Sigma_com_true","Sigma_sci_true","Sigma_com","Sigma_sci","n_samp_com","n_samp_sci","N_true","N_est","SD_N","Convergence","LogLik","MSPE_S","k", "x","reallocation","aggreg_obs")
   
   Results = data.frame(matrix(NA,1,length(colnames_Results)))
   colnames(Results)=colnames_Results
@@ -289,7 +289,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   
   ## loop
   # 1er niveau de boucle : on repete chaque simulation n_sim fois pour avoir de la variabilité
-  for(i in i0:n_sim){
+  for(i in i0:i1){
     # 2ème niveau de boucle : on teste chaque valeur de b
     for (b in b_set){
       

### 2022-01-04 Version_M2_plusieurszones/draft/paper_3/3_formulas.Rmd
case study and simulations

---
title: "**Formulas for the model built on catch declarations**"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notations

We model observations at the level of catch declarations $D_j$ as a sum of $Y_{ij}$ lognormal observations each one realised at one fishing position $x_{ij}$.

We note :

- $\mathcal{P}_j$ : the vector of all fishing positions related to the $j^th$ declaration

- $j \in \{1,...,n\}$ with $n$ the number of declarations

- $i \in \{1,...,m_j\}$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.

$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{ij}}$$

# Alternative observation models accounting for reallocation

## First assumption about convolution of lognormal random variables

We assume a sum of lognormal random variables is still lognormal and that $D_j \sim \mathcal{LN}(M(D_j),\Phi(D_j))$ with $M(.)$ being a function representing the central value of $D_j$ and $\Phi(.)$ the variance function of $D_j$.

## Reparameterization of the lognormal distribution

The standard lognormal distribution is written as :

$D \sim \mathcal{LN}(\rho;\sigma^2)$ with $D=e^{\rho+\sigma N}$ and $N \sim \mathcal{N}(0,1)$

In this case, $E(D)=e^{\rho + \frac{\sigma^2}{2}}$ and $Var(D)=(e^{\sigma^2}-1)e^{2 \rho + \sigma^2}$.

We choose to slightly reparameterize the lognormal distribution so that $E(D)$ and $Var(D)$ have more simple expressions.

Let's define $\rho = ln(\mu) - \frac{sigma^2}{2}$, then :

- $D=\mu e^{\sigma N - \frac{sigma^2}{2}}$

- $E(D)=\mu$

- $Var(D)=\mu^2(e^{\sigma^2} - 1) \Leftrightarrow	\sigma^2=ln(\frac{Var(D)}{E(D)^2} + 1)$

Then, $D \sim\mathcal{LN}(M(D),\Phi(D))$ with $M(D)=E(D)$ and $\Phi(D)=ln(\frac{Var(D)}{E(D)^2} + 1)$.

## $D_j$ probability distribution and moments

We have to express the probability distribution of $D_j$ and its moments as a function of $Y_{ij}$ with $i \in \mathcal{P}_j$.

Let's assume $Y_{ij} = C_{ij}.Z_{ij}$ is a zero-inflated lognormal distribution with $C_{ij}$ a binary random variable and $Z_{ij}$ a lognormal random variable.

$$C_{ij} \sim \mathcal{B}(1-p_{ij})$$ 
with $p_{ij}=exp(-e^\xi .S(x_{ij}))$ the probability to obtain a zero value

$$Z_{ij} \sim \mathcal{LN}(\frac{S(x_{ij})}{1-p_{ij}},\sigma^2)$$

Here, $Y_{ij}$, $C_{ij}$ and $Z_{ij}$ are observations of a latent field $S(x_{ij})$ at a sampled point $x_{ij}$.

### Probability of obtaining a zero declaration

\begin{align*}
P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_{ij} = 0 \vert S, X),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{ij})\right \rbrace} = \pi_j.
\end{align*}

### Expectancy of a positive declaration

Following calculations are supposed to be conditionnal on $S$ and $X$.

$$E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_{ij} Z_{ij}\vert \exists i\in\mathcal{P}_j , C_{ij}=1)$$

As $C_{ij}$ and $Z_{ij}$ are assumed to be independant.

\begin{align*}
E(D_j\vert D_j > 0) & = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / P\left ( D_j > 0\right ),   \\
& = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / \left (1-\pi_j\right).   \\
\end{align*}

As $E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } ) = E(D_j ),$

\begin{align*}
E(D_j\vert D_j > 0) & = \left (1- \pi_j\right)^{-1} E(D_j)   ,   \nonumber \\
& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_{ij} Z_{ij}),\nonumber \\
& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_{ij}) \frac{S(x_{ij})}{1-p_{ij}}, \nonumber \\
& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} S(x_{ij}).
\end{align*}

### Variance of a positive declaration

$$Var(D_j \vert D_j >0)  = E(D_j^2 \vert D_j >0)- E(D_j \vert D_j >0)^2.$$

$$E(D_j^2 \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 1_{\left \lbrace D_j >0\right\rbrace}) = (1-\pi_j)^{-1} E(D_j^2 )$$

$$E(D_j \vert D_j >0)^2 = ((1-\pi_j)^{-1} E(D_j 1_{\left \lbrace D_j >0\right\rbrace}))^2 =  (1-\pi_j)^{-2} E(D_j)^2$$

And then,

$$Var(D_j \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 ) - (1-\pi_j)^{-2} E(D_j)^2 = (1-\pi_j)^{-1} Var(D_j) - \frac{\pi_j}{(1-\pi_j)^2} E(D_j)^2.$$

As the $(Y_{ij})_{i\in \mathcal{P}_j}$ are independent, $Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_{ij}) = \sum_{i \in\mathcal{P}_j} Var(C_{ij}.Z_{ij})$.

\begin{align*}
Var(C_{ij} Z_{ij}) & = E( C_{ij}^2 Z_{ij}^2 ) - E(C_{ij} Z_{ij})^2,\\
& =  E( C_{ij}^2) E(Z_{ij}^2 ) -E(C_{ij})^2 E(Z_{ij})^2,\\
& = (1-p_{ij}) E(Z_{ij}^2 ) - (1-p_{ij})^2 E(Z_{ij})^2,\\
& = (1-p_{ij})(Var(Z_{ij})+E(Z_{ij})^2)-(1-p_{ij})^2E(Z_{ij})^2,\\
& = \frac{S(x_{ij})^2}{1-p_{ij}}(e^{\sigma^2}-1)+\frac{S(x_{ij})^2}{1-p_{ij}}-S(x_{ij})^2,\\
& = \frac{S(x_{ij})^2}{1-p_{ij}}(e^{\sigma^2}-(1-p_{ij}))
\end{align*}

### Conclusion

$$P(D_j = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{ij})\right \rbrace} = \pi_j$$

$$D_j \vert D_j > 0 \sim \mathcal{LN}(E(D_j \vert D_j > 0); ln(\frac{Var(D_j \vert D_j > 0)}{E(D_j \vert D_j > 0)^2} + 1))$$

$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{ij})}{1-\pi_j}$$

$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{ij})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$

$$Var(Y_{ij})=\frac{S(x_{ij})^2}{1-p_{ij}}(e^{\sigma^2}-(1-p_{ij}))$$

### Simulations

```{r,echo=F,fig.asp=1,fig.width=8,fig.height=2,fig.align='center'}

obs_model <- "zerolognorm"
q1 <- 2 # zero-inflation parameter (when increases, nb of zero increases)
SD_obs <- 0.1  # observation error (standard deviation for lognormal distribution, CV for gamma)
n_obs <- 10 # number of observations per fishing sequence (number of observations over which the catches are summed)
n_sim <- 1000 # number of replicates for each 'mu' value
source("r/3_simulations.R")

```

### 2022-01-04 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
case study and simulations

---
title             : "Can we use reallocated logbooks data to map species distribution habitats?"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["paper_3.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
header-includes:
   - \usepackage{graphicx}

output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

**To do:**

- **email to Kasper**

- **redaction of M&M**

- **run single square simulations**

- **run multiple square simulations**

# Material and methods

## Model

See `3_formulas.rmd`.

## Simulation-estimation

### Single-square simulations

How do we simulate the covariate?

How to justify the number of fishing sequence and the number of VMS pings?

### Multiple-squares simulations

How do we simulate the covariate?

## Case-study: sole of the Bay of Biscay

# Results

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2022-01-04 Version_M2_plusieurszones/draft/paper_3/paper_3.tex
case study and simulations

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The title},
  pdfauthor={First Author1 \& Ernst-August Doelle1,2},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Title}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{graphicx}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi

\title{The title}
\author{First Author\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
\date{}

\authornote{

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

The authors made the following contributions. First Author: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing.

Correspondence concerning this article should be addressed to First Author, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

\begin{document}
\maketitle

\textbf{To do:}

\begin{itemize}
\item
  \textbf{email to Kasper}
\item
  \textbf{redaction of M\&M}
\item
  \textbf{run single square simulations}
\item
  \textbf{run multiple square simulations}
\end{itemize}

\hypertarget{material-and-methods}{%
\section{Material and methods}\label{material-and-methods}}

\hypertarget{model}{%
\subsection{Model}\label{model}}

See \texttt{3\_formulas.rmd}.

\hypertarget{simulation-estimation}{%
\subsection{Simulation-estimation}\label{simulation-estimation}}

\hypertarget{single-square-simulations}{%
\subsubsection{Single-square simulations}\label{single-square-simulations}}

How do we simulate the covariate?

How to justify the number of fishing sequence and the number of VMS pings?

\hypertarget{multiple-squares-simulations}{%
\subsubsection{Multiple-squares simulations}\label{multiple-squares-simulations}}

How do we simulate the covariate?

\hypertarget{case-study-sole-of-the-bay-of-biscay}{%
\subsection{Case-study: sole of the Bay of Biscay}\label{case-study-sole-of-the-bay-of-biscay}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}

\endgroup

\end{document}

### 2022-01-04 Version_M2_plusieurszones/Scripts/In_progress/plot_case_study.R

@@ -1,0 +1,177 @@
+
+#######
+## Maps
+#######
+## Fitted data
+# ggplot()+
+#   geom_sf(data=points_sf_2,alpha=0.5,size=0.5)+
+#   geom_point(data=HH_obsmer_2,aes(x=x,y=y),col="orange")+
+#   geom_point(data=ptsurvey_wgs84_2,aes(x=x,y=y),col="red")+
+#   geom_sf(data=st_geometry(ICES_rect_2),alpha=0)+
+#   geom_sf(data=st_geometry(st_as_sf(survey_layer_3)),
+#           size=1.5,col="darkgrey",
+#           alpha=0)+
+#   geom_sf(data = mapBase)+
+#   coord_sf(xlim = c(-6,0), ylim = c(43,48),
+#            expand = FALSE)+
+#   scale_color_distiller(palette="Spectral")+
+#   theme(panel.background = element_rect(fill = "skyblue"))+
+#   xlab("")+ylab("")
+
+load("results/case_study/no.realloc_int_df.RData")
+
+no.realloc_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
+no.realloc_plot <- ggplot(no.realloc_df)+
+  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
+  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
+  ggtitle("Integrated model",subtitle = "Commercial likelihood on Yi")+
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.subtitle = element_text(hjust = 0.5))+
+  geom_sf(data = mapBase)+
+  coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
+  xlab("")+ylab("")
+
+load("results/case_study/realloc_int_df.RData")
+
+realloc_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
+realloc_plot <- ggplot(realloc_df)+
+  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
+  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
+  ggtitle("Integrated model",subtitle = "Commercial likelihood on Dj")+
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.subtitle = element_text(hjust = 0.5))+
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/In_progress/plot_from_list_param.R

@@ -48,15 +48,32 @@ for(i in 1:10){
 
 library(doBy)
 library(gt)
-summaryBy(Convergence~
+Results_conv <- Results_2
+Results_conv$one <- 1
+Results_conv <- Results_conv %>% 
+  filter(b_true == 0)
+summaryBy(Convergence+one~
             sequencesdepeche+
             # zonespersequence+
             aggreg_obs+
             # b_true+
             n_samp_com+
             reallocation,
-          data=Results_2,
+          data=Results_conv,
           FUN=sum) %>%
+  dplyr::select(n_samp_com,
+                sequencesdepeche,
+                reallocation,
+                aggreg_obs,
+                Convergence.sum,
+                one.sum) %>%
+  dplyr::rename(fishing_sequence = sequencesdepeche,
+                data_realloc = reallocation,
+                lkl_level = aggreg_obs) %>%
+  mutate(lkl_level = ifelse(lkl_level == T,"Dj","Yi"),
+         data_realloc = ifelse(data_realloc == 1,"Yes","No"),
+         perc_convergence = round((1 - Convergence.sum / one.sum)*100,digits = 3)) %>%
+  dplyr::select(-Convergence.sum,-one.sum) %>%
   gt() %>%
   tab_options(
     summary_row.background.color = "#ACEACE80",
@@ -79,6 +96,9 @@ summaryBy(Convergence~
     column_labels.font.weight = "bold"
   )
 
+
+Results <- Results_2 %>%
+  filter(b_true == 0 & Convergence == 0)
 Results[,"RelBias_N"]=(Results[,"N_est"]-Results[,"N_true"])/Results[,"N_true"]
 Results[,"RelBias_beta"]=(Results[,"beta1_est"]-Results[,"beta1_true"])/Results[,"beta1_true"]
 Results[,"Bias_b"]=(Results[,"b_est"]-Results[,"b_true"]) / ifelse(Results[,"b_true"] != 0,Results[,"b_true"],1)
@@ -89,11 +109,8 @@ Results <- Results %>%
 Results$zonespersequence <- as.factor(Results$zonespersequence)
 Results$b_true <- as.factor(Results$b_true)
 
-Results_2 <- Results %>%
-  filter(b_true != 1 & b_true != 3) # & n_samp_com != 50 & relloc_aggreg != "realloc.sim:1_realloc.est:0")
-
 RelBias_N_plot <- ggplot()+
-  geom_boxplot(data = Results_2,
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/In_progress/plot_real_simu.R

@@ -1,0 +1,239 @@
+
+library(rnaturalearth)
+library(sf)
+library(mapdata)
+mapBase <- map("worldHires", fill = T, plot = F)
+mapBase <- st_as_sf(mapBase) %>% filter(ID %in% c("France","Spain"))
+ref_capt <- "k_com"
+
+load(file="results/real_simu/simu.data.RData")
+S_x <- simu.data$S_x
+index_i <- simu.data$index_i
+y_sci_i <- simu.data$y_sci_i
+y_i <- simu.data$y_i
+boats_i <- simu.data$boats_i
+cov_x <- simu.data$cov_x
+index_sci_i <- simu.data$index_sci_i
+
+load(file="results/real_simu/no.realloc_int_df.RData")
+res_no.realloc_int <- fit_IM_res
+load(file="results/real_simu/sci_df.RData")
+res_sci <- fit_IM_res
+load(file="results/real_simu/no.realloc_com_df.RData")
+res_no.realloc_com <- fit_IM_res
+load(file="results/real_simu/realloc_int_df.RData")
+res_realloc_int <- fit_IM_res
+load(file="results/real_simu/realloc_com_df.RData")
+res_realloc_com <- fit_IM_res
+
+samp_sci <- data.frame(cell = index_sci_i) %>%
+        dplyr::group_by(cell) %>%
+        dplyr::summarise(n_sci=n()) %>%
+        inner_join(loc_x[,c("cell","x","y")])
+
+samp_com <- data.frame(cell = index_i) %>%
+        dplyr::group_by(cell) %>%
+        dplyr::summarise(n_com=n()) %>%
+        inner_join(loc_x[,c("cell","x","y")])
+
+
+#######
+## Maps
+#######
+
+## Sampling locations/effort
+library(viridis)
+pal <- inferno(100)
+sampling_plot <- ggplot()+
+        geom_point(data=samp_com,aes(x=x,y=y,col=n_com),shape=15,size=2)+
+        geom_point(data=samp_sci,aes(x=x,y=y),col="red",size=1)+
+        # scale_color_gradientn(colours = pal) + 
+        scale_color_distiller(palette = "RdBu",limits=c(0,NA))+
+        ggtitle("Sample points",subtitle = "Red dots: scientific samples, scale color: commercial samples")+
+        theme_bw()+
+        theme(plot.title = element_text(hjust = 0.5),
+              plot.subtitle = element_text(hjust = 0.5,size=10))+
+        geom_sf(data = mapBase)+
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/function/extract_covariates.R

@@ -1,0 +1,290 @@
+############################################
+## Extract covariates from shapefiles/raster
+############################################
+
+#---------
+# Sediment
+#---------
+#' @title extract_substrate_rasterCelticBoB()
+#' @param cov_file file where are located covariates
+#' @param gridpoint grid points
+#' @param gridpoint_cov grid points with covariates
+#' @param aggreg_substrate if T, substrate are aggregated
+#' 
+#' @return gridpoint_cov gridpoint with substrate levels
+
+# cov_file <- "E:/data2/Envir.Data/"
+# aggreg_substrate = T
+extract_substrate_rasterCelticBoB <- function(cov_file,gridpoint,gridpoint_cov,aggreg_substrate){
+  # Extract from raster and project in the grid
+  str_name<-paste0(cov_file,'Substrate.Habitat.Atl.Bob.tif')
+  imported_raster=raster(str_name)
+  substrate <- raster::extract(stack(imported_raster), coordinates(gridpoint))
+  substratepoint.sp <- gridpoint
+  substratepoint.sp@data <- cbind(substratepoint.sp@data,substrate)
+  colnames(substratepoint.sp@data)[which(colnames(substratepoint.sp@data)=="Substrate.Habitat.Atl.Bob")] <- "Substrate.Code"
+  # spplot(substratepoint.sp,"Substrate.Code",cex = 0.01)
+  
+  # numbers --> sediment name
+  Substrate.names <- c("Fine mud",
+                       "Rock or other hard substrata",
+                       "Sandy mud or Muddy sand",
+                       "Sand",
+                       "Muddy sand",
+                       "Seabed",
+                       "Mixed sediment",
+                       "Sandy mud",
+                       "Coarse substrate",
+                       "Sediment",
+                       "Fine mud or Sandy mud or Muddy sand")
+  
+  Substrate.Code <- c(1:length(Substrate.names))
+  Substrate.Code_df <-  data.frame(Substrate.names = Substrate.names,
+                                   Substrate.Code = Substrate.Code)
+  
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/function/fit_IM.R

@@ -51,9 +51,20 @@ fit_IM <- function(Estimation_model_i = 1,
                    aggreg_obs,
                    boats_number,
                    Cov_x,
+                   ref_level = NULL,
                    lf_param = "cov",
                    spde = NULL,
-                   mesh = NULL
+                   mesh = NULL,
+                   n_cells=n_cells,
+                   cov_est = F,
+                   Params_step.est=NULL,
+                   Map_step.est=NULL,
+                   ObsM=F,
+                   y_ObsM_i = NULL,
+                   index_ObsM_i = NULL,
+                   sampling="simulate",
+                   landings=T,
+                   quadratic_cov=NULL
                    ){
   
   #----------------------
@@ -82,11 +93,15 @@ fit_IM <- function(Estimation_model_i = 1,
                    'b_constraint' = 2, # (DEPRECATED)
                    'catchability_random' = 0, # (DEPRECATED)
                    'aggreg_obs' = ifelse(aggreg_obs == T,1,0),
-                   'lf_param'=ifelse(lf_param == "cov",0,1)
+                   'lf_param'=ifelse(lf_param == "cov",0,1),
+                   'ObsM'=ifelse(ObsM == F,0,1),
+                   'landings'=ifelse(landings == F,0,1)
                    )
   
   ## Data & Params
   Map = list() #liste vide : correspond aux paramètres fixés dans l'estimation
+  
+  if(cov_est == F) Cov_x <- as.matrix(rep(0,n_cells))
   
   #Pour chacune des trois configurations de modele (1,2 ou 3 ie les deux, scienti
   #only ou commer only) on définit les données d'entrée, les paramètres à estimer 
@@ -122,10 +137,12 @@ fit_IM <- function(Estimation_model_i = 1,
     #beta_k : effets des covariables qui jouent sur l’échantillonnage, mais ici on n’en prend pas en compte (cf formule 4 du papier de Baptiste)
     #k_com = 1 #capturabilite relative
     
-    Map[["k_com"]] <- seq(1:(length(Params$k_com))) # first k is for scientific data
-    Map[["k_com"]][1] <- NA # reference level is the first fleet
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/inst/executables/com_x_sci_data_14_scientific_commercial_simple.R

@@ -1,8 +1,0 @@
-
-      library( TMB )
-      dyn.load( dynlib('Scripts/inst/executables/com_x_sci_data_14_scientific_commercial_simple') )
-      setwd('C:/R_projects/projet_M2_sole/Version_M2_plusieurszones')
-      load( 'All_inputs.RData' )
-      Obj = MakeADFun(data=All_inputs[['data']], parameters=All_inputs[['parameters']], random=All_inputs[['random']], All_inputs[['Other_inputs']])
-      save( Obj, file='Obj.RData')
-    

### 2022-01-04 Version_M2_plusieurszones/Scripts/inst/executables/com_x_sci_data_14_scientific_commercial_simple.cpp

@@ -166,160 +166,194 @@ Type objective_function<Type>::operator() (){
 
   
   //////////////////////////////////// Commercial /////////////////////////////////////////
-  
-  
   if(Options_vec(5) == 3 | Options_vec(5) == 1){
-
+    
     ////////
     // Data
     ////////
-
     DATA_VECTOR( c_com_x );       	// Response (count) for each observation i (commercial data)
-    DATA_VECTOR( y_com_i );       	// Response (0:not surveyed, 1:surveyed) for each site (commercial data)
-    DATA_FACTOR( seq_com_i );       	// Response (0:not surveyed, 1:surveyed) for each site (commercial data)
-    
-    DATA_FACTOR( index_com_i );
-   	DATA_VECTOR( q2_com );
     DATA_MATRIX( Cov_xk );  //design matrix for sampling of commercial data
+    DATA_VECTOR( q2_com );
     
-
     //////////////
     // Parameters
     //////////////
-
     PARAMETER_VECTOR(q1_com);
     PARAMETER_VECTOR(logSigma_com);
-	  PARAMETER_VECTOR( k_com );
-
-	  /////////////////
-    // derived values
-    /////////////////
-
-    int n_com_i = y_com_i.size();
-    int n_ping_i = seq_com_i.size();
+    PARAMETER_VECTOR( k_com );
+    vector<Type> Sigma_com = exp(logSigma_com);
     
 
-    /////////////////////
-    // Observation model
-    ////////////////////
-    
-    if( Options_vec(9) == 1){
+    //////////////////////////////// Landings observations ////////////////////////////////////
+    if( Options_vec(15) == 1 ){
+
+      ////////
+      // Data
+      ////////
+      DATA_VECTOR( y_com_i );       	// Response (0:not surveyed, 1:surveyed) for each site (commercial data)
+      DATA_FACTOR( seq_com_i );       	// Response (0:not surveyed, 1:surveyed) for each site (commercial data)
+      DATA_FACTOR( index_com_i );
+
+      /////////////////
+      // derived values
+      /////////////////
+      int n_com_i = y_com_i.size();
+      int n_ping_i = seq_com_i.size();
       
-      vector<Type> Sigma_com = exp(logSigma_com);
-      vector<Type> E_com(n_ping_i);E_com.setZero();
-      
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/reallocation_real_simu.R

@@ -3,27 +3,31 @@
 ##################################################
 ## B. Alglave
 
-memory.size(1e+9)
 
 library(cowplot)
 library(dplyr)
 library(ggplot2)
 library(raster)
 library(rnaturalearth)
+library(RNetCDF)
 library(sf)
 library(spatstat)
 library(stringr)
 library(TMB)
+library(tidyr)
 
-folder_data <- "C:/R_projects/phd_zfh_baptiste_alglave"
+set.seed(2) # 3
+
+folder_phd_codes <- "C:/R_projects/phd_zfh_baptiste_alglave"
 folder_project <- "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones"
 
-setwd(folder_data)
+setwd(folder_phd_codes)
 
 ##--------------
 ## Load raw data
 ##--------------
 ## Map info
+domain_shapefile <- "ORHAGO" # "ORHAGO", "EVHOE"
 grid_projection <- "+proj=longlat +datum=WGS84"
 mapBase <- ne_countries(scale = "medium", returnclass = "sf")
 grid_xmin <- -6 # Coordinates of study area 
@@ -33,17 +37,30 @@ grid_ymax <- 48
 grid_limit <- extent(c(grid_xmin,grid_xmax,grid_ymin,grid_ymax))
 Resol_map <- 0.05
 
-## Time interval
-year_y <- 2018
-month_m <- 11
+# Species name
+species_to_plot <- "Solea_solea"
+# species_to_plot <- "Merluccius_merluccius"
+
+# Time step
+if(species_to_plot == "Solea_solea"){
+  year_y <- 2017
+  month_m <- 11
+  quarter_q <- 4
+}else if(species_to_plot == "Merluccius_merluccius"){
+  year_y <- 2018
+  month_m <- 11
+  quarter_q <- 4
+}
 
 ## data file name
-sampling <- "simulate" # "simulate": simulate sampling, "from_tacsateflalo": take data points from tacsateflalo
+sampling <- "from_tacsateflalo" # "simulate": simulate sampling, "from_tacsateflalo": take data points from tacsateflalo
 commercial_data_folder <- 'data/raw/vms_logbooks/bob_cs/MergeTACSAT_EFLALO/'
+ObsMer_file <- 'data/raw/SIH/obsmer'
+ObsM <- T
 
 ## If simulate data points
 # n_samp_com <- 5000
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/simu_main_script.R

@@ -7,7 +7,7 @@
 library(parallel)
 
 # Simulation name
-simu_name = "SimuTest"
+simu_name = "SimuRef"
 
 # Parallel configuration
 n_nodes <- 8
@@ -120,7 +120,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   # Relative catchability
   q2_com <- 1
   # Levels of preferential sampling
-  b_set = c(0,1,3)
+  b_set = c(0)
   
   
   
@@ -139,7 +139,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   aggreg_obs = F
   
   # if 1, sampling process is accounted for in estimation
-  Samp_process = 1
+  Samp_process = 0
   
   # TMB model version
   TmbFile = "Scripts/"
@@ -285,7 +285,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   
   # load TMB model
   dyn.load( dynlib(paste0(TmbFile,"inst/executables/com_x_sci_data_14_scientific_commercial_simple") ) )
-  # dyn.unload( dynlib(paste0(TmbFile,"inst/executables/",Version,"_scientific_commercial") ) )
+  # dyn.unload( dynlib(paste0(TmbFile,"inst/executables/com_x_sci_data_14_scientific_commercial_simple") ) )
   
   ## loop
   # 1er niveau de boucle : on repete chaque simulation n_sim fois pour avoir de la variabilité

### 2022-01-04 Version_M2_plusieurszones/Scripts/source/load_domain.R

@@ -1,0 +1,134 @@
+##############
+## Load domain
+##############
+
+setwd(folder_phd_codes)
+
+if(domain_shapefile == "ORHAGO"){
+  
+  strata_file <- "data/raw/shapefile/QGIS_ORHAGO/shape/" 
+  
+  zone_centre_large <- sf::st_read(paste0(strata_file,"zone centre au large.shp"))
+  zone_centre_cote <-sf::st_read(paste0(strata_file,"zone centre cote.shp"))
+  zone_nord <- sf::st_read(paste0(strata_file,"zone nord.shp"))
+  zone_sud <- sf::st_read(paste0(strata_file,"zone sud.shp"))
+  
+  ## Orhago polygon
+  survey_layer <- st_union(zone_centre_large,zone_centre_cote)
+  survey_layer <- st_union(survey_layer,zone_nord)
+  survey_layer <- st_union(survey_layer,zone_sud)
+  survey_layer_2 <- nngeo::st_remove_holes(survey_layer)
+  survey_layer_3 <- sf::as_Spatial(survey_layer_2)
+  crs(survey_layer_3) <- grid_projection
+  study_domain_2 <- survey_layer_3
+  
+  ## Matrices to plot maps
+  grid <- raster(grid_limit)
+  res(grid) <- Resol_map
+  crs(grid) <- grid_projection
+  gridpolygon <- rasterToPolygons(grid)
+  gridpolygon$layer <- c(1:length(gridpolygon$layer))
+  
+  raster_to_point <- rasterToPoints(grid)
+  datapoint <- SpatialPointsDataFrame(coords=raster_to_point,
+                                      data=data.frame(layer = 1:nrow(raster_to_point)),
+                                      proj4string=crs(grid_projection))
+  
+  crs(datapoint) <- crs(study_domain_2) <- grid_projection
+  datapoint_2 <- over(datapoint,study_domain_2)
+  datapoint_2 <- cbind(datapoint_2,layer = datapoint@data$layer)
+  datapoint_2 <- datapoint_2[which(!is.na(datapoint_2[,1])),]
+  ## Add coordinates to gridpoint
+  tmp <- cbind(datapoint@data, coordinates(datapoint))
+  colnames(tmp) <- c(names(datapoint@data), "x", "y")
+  datapoint_3 <- inner_join(datapoint_2,tmp) %>%
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/source/shape_ObsMer.R

@@ -1,0 +1,48 @@
+###############
+## Shape ObsMer
+###############
+
+HH_obsmer_2 <- HH_obsmer %>%
+  mutate(y=(latIni+latFin)/2,x=(lonIni+lonFin)/2) %>%
+  dplyr::select(-latIni,-latFin,-lonIni,-lonFin) %>%
+  mutate(month=as.numeric(str_sub(date,6,7)))
+
+HH_obsmer_2$quarter <- NA
+HH_obsmer_2$quarter[which(HH_obsmer_2$month %in% c(1,2,3))] <- 1
+HH_obsmer_2$quarter[which(HH_obsmer_2$month %in% c(4,5,6))] <- 2
+HH_obsmer_2$quarter[which(HH_obsmer_2$month %in% c(7,8,9))] <- 3
+HH_obsmer_2$quarter[which(HH_obsmer_2$month %in% c(10,11,12))] <- 4
+
+HH_obsmer_2 <- HH_obsmer_2 %>%
+  filter(str_detect(foCatEu6,"OTB_CEP_>=70_0|OTB_DEF_>=70_0")) %>%
+  filter(year == year_y) %>%
+  filter(y<48 & quarter == quarter_q) # month == month_m)
+
+SL_obsmer_2 <- SL_obsmer %>%
+  mutate(Species = str_replace_all(spp," ","_")) %>%
+  filter(Species == species_to_plot)
+
+ObsMer_df <- left_join(HH_obsmer_2,SL_obsmer_2) %>%
+  dplyr::group_by_all() %>%
+  dplyr::summarise(wt=sum(wt)) %>%
+  mutate(Catch = ifelse(is.na(wt),0,wt)) %>%
+  mutate(Catch = Catch/1000,
+         foDur = foDur / 60) %>%
+  mutate(CPUE = Catch/foDur)
+
+# convert the points to an sf object
+ObsMer_sf <- st_as_sf(ObsMer_df, coords = c("x", "y"))
+
+# set CRS for the points to be the same as shapefile
+st_crs(ObsMer_sf) <- st_crs(grid_projection)
+
+ObsMer_sf_2 <- ObsMer_sf[st_intersects(ObsMer_sf,gridpolygon_sf) %>% lengths > 0,]
+ObsMer_sf_2 <- st_join(ObsMer_sf_2,gridpolygon_sf)
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/source/shape_VMS_logbooks.R

@@ -1,0 +1,100 @@
+#########################
+## Shape 'VMS x logbooks'
+#########################
+
+tacsatEflalo_2 <- tacsatEflalo %>%
+  filter(SI_LATI < 48 &  SI_LONG > -6 & SI_STATE == 1) %>%
+  filter(str_detect(LE_MET_level6,"OTB|OTT|PTM|GTR|OTM|PS"))
+
+# convert the points to an sf object
+points_sf <- st_as_sf(tacsatEflalo_2, coords = c("SI_LONG", "SI_LATI"))
+
+# set CRS for the points to be the same as shapefile
+st_crs(points_sf) <- st_crs(grid_projection)
+
+# cross with statistical rectangles
+points_sf <- points_sf[st_intersects(points_sf,ICES_rect_2) %>% lengths > 0,]
+points_sf <- st_join(points_sf,ICES_rect_2)
+
+points_sf <- points_sf %>%
+  mutate(seq_id = paste0(VE_REF,"_",SI_DATE,"_",LE_GEAR,"_",ICESNAME))
+
+# ## Number of pings per fishing sequence
+# points_sf  <- points_sf %>% filter(SI_STATE == 1)
+# points_df <- points_df %>% filter(SI_STATE == 1)
+# 
+# nb_ping_seq <- as.data.frame(table(points_sf$seq_id)) %>%
+#   group_by(Freq) %>%
+#   dplyr::summarise(n = n()) %>%
+#   dplyr::rename(ping_per_seq = Freq) %>%
+#   filter(ping_per_seq <= 24) %>%
+#   mutate(freq = n / sum(n))
+# 
+# ggplot(nb_ping_seq,aes(x=ping_per_seq,y=freq))+
+#   geom_point()+
+#   geom_line()+
+#   ylim(0,NA)
+# save(data=nb_ping_seq,file="results/exploratory_analysis/reallocation/nb_ping_seq.RData")
+
+nb_ping_gear <- as.data.frame(table(points_sf$VE_REF,points_sf$LE_GEAR)) %>%
+  filter(Freq != 0)
+
+vessel_vec <- c("103","157914","14868","151271") # OTB_DEF, GTR, OTT, PTM
+
+points_sf <- points_sf %>%
+  # filter(VE_REF %in% vessel_vec) %>%
+  mutate(year = as.numeric(str_sub(SI_DATE,7,10)),
+         month = as.numeric(str_sub(SI_DATE,4,5)))
+
+points_sf$quarter <- NA
[…]

### 2022-01-04 Version_M2_plusieurszones/Scripts/source/shape_scientific.R

@@ -1,0 +1,24 @@
+########################
+## Shape scientific data
+########################
+ptsurvey_wgs84_2 <- ptsurvey_wgs84 %>%
+  group_by_at(setdiff(names(ptsurvey_wgs84),c("Sex","CatchWgt_mature","CatCatchWgt","scientificname","CatchWgt_spp"))) %>%
+  dplyr::summarise(CatchWgt_spp = sum(CatchWgt_spp)) %>%
+  dplyr::rename(x=long,y=lati) %>%
+  ungroup %>%
+  dplyr::select(-layer) %>%
+  filter(Year == year_y)
+
+ptsurvey_wgs84_sf <- st_as_sf(ptsurvey_wgs84_2,coords = c("x","y"),crs=grid_projection)
+ptsurvey_wgs84_sf_2 <- ptsurvey_wgs84_sf[st_intersects(ptsurvey_wgs84_sf,gridpolygon_sf) %>% lengths > 0,]
+ptsurvey_wgs84_sf_2 <- st_join(ptsurvey_wgs84_sf_2,gridpolygon_sf)
+ptsurvey_wgs84_sf_3 <- ptsurvey_wgs84_sf_2 %>%
+  data.frame
+ptsurvey_wgs84_sf_4 <- inner_join(ptsurvey_wgs84_sf_3,loc_x[,c("layer","cell","x","y")])
+
+y_sci_i <- ptsurvey_wgs84_sf_4$CatchWgt_spp
+index_sci_i <- ptsurvey_wgs84_sf_4$cell
+
+## Covariate matrix
+cov_x <- loc_x %>%
+  dplyr::select_at(vars(starts_with("bathy") | starts_with("substr")))

### 2022-01-04 Version_M2_plusieurszones/Scripts/source/simulate_real.R

@@ -1,0 +1,157 @@
+########################
+## Shape scientific data
+########################
+
+area_df <- loc_x %>%
+  group_by(ICESNAME) %>%
+  dplyr::summarise(area=n()) %>%
+  mutate(perc_area = area/(sum(area))) %>%
+  mutate(ICESNAME = factor(ICESNAME)) %>%
+  arrange(ICESNAME)
+
+## Sequential sampling (not that there is no preferential sampling)
+# First select a statistical rectangle
+sample_full_rect <- F
+zonesize <- 0.2
+n_zone <- 1 # c(1,3,5)
+samp_df <- do.call(rbind,lapply(1:n_seq_com, function(j){
+  
+  ## Select a statistical rectangle
+  rect_stat <- sample(x = area_df$ICESNAME,
+                      size = 1,
+                      replace=T,
+                      prob =  area_df$perc_area)
+  
+  loc_x_rect <- loc_x %>%
+    filter(ICESNAME == rect_stat)
+  
+  if(sample_full_rect == T){
+    
+    ## Sample within the statistical rectangle
+    samp_cell <- sample(x = loc_x_rect$cell,
+                        size = n_ping_per_seq,
+                        replace=T)
+    
+  }else{
+    
+    samp_cell <- do.call(c,lapply(1:n_zone, function(j){
+      ## Sample within the statistical rectangle
+      center <- sample(x = loc_x_rect$cell,
+                       size = 1,
+                       replace=T)
+      
+      loc_x_zone <- loc_x_rect[which(loc_x_rect$x < loc_x_rect$x[which(loc_x_rect$cell == center)] + zonesize &
+                                       loc_x_rect$x > loc_x_rect$x[which(loc_x_rect$cell == center)] - zonesize &
+                                       loc_x_rect$y < loc_x_rect$y[which(loc_x_rect$cell == center)] + zonesize &
+                                       loc_x_rect$y > loc_x_rect$y[which(loc_x_rect$cell == center)] - zonesize),]
+      
+      ## Sample within the statistical rectangle
+      samp_cell <- sample(x = loc_x_zone$cell,
[…]

### 2022-01-04 Version_M2_plusieurszones/draft/paper_3/r/3_simulations.R

@@ -1,0 +1,210 @@
+
+## Simulate observations
+#-----------------------
+
+# obs_model <- "gamma" # "gamma", "zerolognorm" (zero-inflated lognormal distribution)
+
+par(mfrow = c(3,2))
+for(set_sim in 1:3){
+  
+  sum_y_i <- c()
+  log_pi_j <- c()
+  sum_mu <- c()
+  Var_D <- c()
+  E_D <- c()
+  Var_Dsup0 <- c()
+  Sigma_D <- c()
+  decl_j <- c()
+  
+  sum_shape <- c()
+  
+  mu <- runif(1,min=0,max=2)
+  
+  for(i in 1:n_sim){
+    
+    ## Data simulation
+    if(obs_model == "zerolognorm"){
+      y_i <- do.call(c,lapply(1:n_obs, function(j){
+        exp_catch <- mu
+        prob_encount <- 1-exp(- exp(q1) * exp_catch)
+        abs_i <- rbinom(1,1,prob_encount)
+        if(abs_i>0){
+          y_i <- exp(rnorm(1,mean = log(exp_catch/prob_encount) - SD_obs^2/2, sd = SD_obs))
+        }else{
+          y_i <- 0
+        }
+      }))
+    }else if(obs_model == "gamma"){
+      y_i <- do.call(c,lapply(1:n_obs, function(j){
+        exp_catch <- mu
+        prob_encount <- 1-exp(- exp(q1) * exp_catch)
+        abs_i <- rbinom(1,1,prob_encount)
+        y_i <- rgamma(1, shape = 1/SD_obs^2, scale = exp_catch*SD_obs^2)
+      }))
+    }
+    
+    # y_i <- Data$y_i
+    sum_y_i <- c(sum_y_i,sum(y_i))
+    
+    ## Moments
+    if(obs_model == "zerolognorm"){
+      log_pi_j <- c(log_pi_j,sum(rep(-1 * mu * exp(q1),n_obs)))
+      
+      pi_j <- exp(log_pi_j)
+      p_i <- rep(exp(-1 * mu * exp(q1)),n_obs)
+      
+      if(sum(y_i) == 0){
+        
+        Var_D <- c(Var_D,0)
[…]

### 2022-01-06 Version_M2_plusieurszones/Scripts/In_progress/plot_from_list_param.R
paper_3

@@ -2,8 +2,12 @@
 ## Plot from List_param
 #----------------------
 
-folder_name <- "C:/Users/balglave/Desktop/com_x_sci_data_14_scientific_commercial_simple-2021-09-19_15_07_00_SimuTest"
+folder_name <- "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/com_x_sci_data_14_scientific_commercial_simple-2022-01-04_18_55_00_SimuRef"
+folder_name <- "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/com_x_sci_data_14_scientific_commercial_simple-2022-01-05_15_15_00_SimuRef"
 
+#-------------
+## If parallel
+#-------------
 for(i in 1:10){
   load(paste0(folder_name,"/",i,"/Results",i,".RData"))
   if(i==1) Results_2 <- Results
@@ -46,9 +50,60 @@ for(i in 1:10){
   
 }
 
+#----------------
+## If no parallel
+#----------------
+load(paste0(folder_name,"/Results.RData"))
+Results_2 <- Results
+
+# Load List_param
+list_file_i <- list.files(paste0(folder_name,"/",i))
+list_file_i <- list_file_i[which(str_detect(list_file_i,"List_param"))]
+  
+Results$aggreg_obs <- NA
+Results$sequencesdepeche <- NA
+Results$zonespersequence <- NA
+Results$n_samp_com <- NA
+Results$rho_S <- NA
+Results$beta1_true <- NA
+Results$beta1_est <- NA
+Results$alpha <- NA
+Results$beta <- NA
+Results$gamma <- NA
+
+# Run over List_param
+for(counter in Results$counter){
+  print(paste0("counter: ",counter))
+  load(paste0(folder_name,"/List_param_",counter,".RData"))
+  Results$aggreg_obs[which(Results$counter == counter)] <- List_param$data.info$aggreg_obs
+  Results$sequencesdepeche[which(Results$counter == counter)] <- List_param$data.info$sequencesdepeche
+  Results$zonespersequence[which(Results$counter == counter)] <- List_param$data.info$zonespersequence
+  Results$n_samp_com[which(Results$counter == counter)] <- List_param$data.info$n_samp_com
[…]

### 2022-01-06 Version_M2_plusieurszones/Scripts/function/commercial_scientific_14.R

@@ -161,32 +161,47 @@ simu_commercial_scientific <- function(Results,
   # beta (fixed parameters for the abundance distribution equation), 
   # delta_x (random effect for the abundance distribution equation), 
   # eta_x (random effect for the sampling process equation)
-  validation = 0
-  wronglatentfield = 1
+
+  simu_latent_field_outputs <- simu_latent_field(loc_x,
+                                                 latent_fields_simu,
+                                                 latent_field,
+                                                 scientific_data,
+                                                 beta0,
+                                                 beta,
+                                                 simu_tool,
+                                                 range,
+                                                 nu,
+                                                 SD_x,
+                                                 SD_delta,
+                                                 SD_eta)
+  Strue_x <- simu_latent_field_outputs$Strue_x
   
-  while (validation == 0)
-  {
-    simu_latent_field_outputs <- simu_latent_field(loc_x,
-                                                   latent_fields_simu,
-                                                   latent_field,
-                                                   scientific_data,
-                                                   beta0,
-                                                   beta,
-                                                   simu_tool,
-                                                   range,
-                                                   nu,
-                                                   SD_x,
-                                                   SD_delta,
-                                                   SD_eta)
-    Strue_x <- simu_latent_field_outputs$Strue_x
-    # Contient 625 lignes et une colonne correspondant aux valeurs du champ latent en chacun des points
-    if (max(Strue_x) < 400 & sd(Strue_x) > 10){
-      validation = 1
-    }else{
-      wronglatentfield = wronglatentfield + 1
-      set.seed( RandomSeed - wronglatentfield )
-    }
-  }
+  # validation = 0
+  # wronglatentfield = 1
+  # 
+  # while (validation == 0)
+  # {
+  #   simu_latent_field_outputs <- simu_latent_field(loc_x,
+  #                                                  latent_fields_simu,
+  #                                                  latent_field,
+  #                                                  scientific_data,
+  #                                                  beta0,
+  #                                                  beta,
+  #                                                  simu_tool,
[…]

### 2022-01-06 Version_M2_plusieurszones/Scripts/function/simu_latent_field.R

@@ -44,9 +44,8 @@ simu_latent_field <- function(loc_x,
   
   # 1st param : continuous variable, 4 nexts : strata_1, strata_2, strata_3, strata_4
   # beta[1] = runif(1,-.5,.5)
-  beta[1] = 0.5
-  # On passe beta[1] à 0.5 pour avoir un champ latent plus "cohérent"
-  
+  beta[1] = 2
+
   # Simulate using Matérn covariance
   Cov_x = sim_GF_Matern(loc_x, nu, range, SD_x^2)[[1]]$y
 

### 2022-01-06 Version_M2_plusieurszones/Scripts/simu_main_script.R

@@ -10,8 +10,9 @@ library(parallel)
 simu_name = "SimuRef"
 
 # Parallel configuration
-n_nodes <- 8
-n_sim <- 8
+n_nodes <- 1
+n_sim <- 100
+cluster_nb <- 1
 run_datarmor <- F
 cl <- makeCluster(n_nodes, outfile = "")
 for(cluster_nb in 1:n_nodes){
@@ -22,6 +23,7 @@ for(cluster_nb in 1:n_nodes){
 x_parallel1 <- clusterEvalQ(cl,{
   
   library(dplyr)
+  library(INLA)
   library(TMB)
   library(stringr)
   library(spatstat)
@@ -64,7 +66,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   n_cells = grid_dim[1]*grid_dim[2]
   
   # Intercepts and covariates of the latent field
-  beta0 = 3 # intercept
+  beta0 = 2 # intercept
   beta = c(0,0,0,0,0) # covariates
   # parameters covariates vector : 1st param : continuous variable, 4 nexts : strata_1, strata_2, strata_3, strata_4
   # La premiere covariable est continue et a une structure spatiale
@@ -79,7 +81,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   simu_tool <- "MaternCov" # 2 way to simulate latent fields : "RandomFields", "MaternCov"
   range = sqrt(prod(grid_dim))/(5)*2 # --> MaternCov
   nu = 1  #portée de la covariable continue
-  SD_x = 2 # ecart type/ variance marginale du processus de la var continue
+  SD_x = 0.5 # ecart type/ variance marginale du processus de la var continue
   # Initialement on avait SD_x = 0.5, la on passe a SD_x = 2 pour avoir plus de variabilité dans le champ latent
   
   #effets spatiaux aléatoires
@@ -114,9 +116,9 @@ x_parallel1 <- clusterEvalQ(cl,{
   # n_samp_com = 3000
   n_samp_com = 3000
   # observation error 
-  logSigma_com = log(0.1)
+  logSigma_com = log(1.000001)
   # zero-inflation parameter
-  q1_com <- -3
+  q1_com <- -1
   # Relative catchability
   q2_com <- 1
   # Levels of preferential sampling

### 2022-01-07 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3

@@
 ```{r setup, include = FALSE}
-library("papaja")
+library(cowplot)
+library(doBy)
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+library(kableExtra)
+library(papaja)
 ```
@@
 
+How many simulations? 100 or more?
+
 How do we simulate the covariate?
 
-How to justify the number of fishing sequence and the number of VMS pings?
+Still some lack of convergence.
+
+Number of pings | nu:
+1|10
+10|100
+100|1000
+
+Size within a fishing sequence: radius = 3
+
+Number of fishing zones: 1|3|5
 
@@
 # Results
+
+## Single square analysis
+
+```{r}
+
+load("res/Results_full_single_square.RData")
+
+## Table of convergence
+Results_conv <- Results
+Results_conv$one <- 1
+Results_conv_2 <- summaryBy(Convergence+one~
+                              sequencesdepeche+
+                              # zonespersequence+
+                              aggreg_obs+
+                              # b_true+
+                              n_samp_com+
+                              reallocation,
+                            data=Results_conv,
+                            FUN=sum) %>%
+  dplyr::select(n_samp_com,
+                sequencesdepeche,
+                # zonespersequence,
+                reallocation,
+                aggreg_obs,
+                Convergence.sum,
+                one.sum) %>%
+    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dj","Yi"),
+           reallocation = ifelse(reallocation == 1,"Yes","No"),
+           perc_convergence = round((1 - Convergence.sum / one.sum)*100,digits = 3))
+
+Results_conv_2$n_samp_com <- as.integer(Results_conv_2$n_samp_com)
+Results_conv_2$sequencesdepeche <- as.integer(Results_conv_2$sequencesdepeche)
+
+Results_conv_2 <- Results_conv_2 %>%
+  dplyr::rename(`Nb samples` = n_samp_com,
+                `Fishing sequence` = sequencesdepeche,
+                `Reallocation` = reallocation,
+                `Likelihood level` = aggreg_obs,
+                `Convergence (%)` = perc_convergence) %>%
+  dplyr::select(-Convergence.sum,-one.sum)
+
+knitr::kable(Results_conv_2,booktabs = T,align = "c",
+                  caption = "Percentage of convergence per simulation/model configuration")
+
+```
+
+```{r}
+
+## Plot Performance metric
+Results <- Results %>%
+  filter(b_true == 0 & Convergence == 0)
+Results[,"RelBias_N"]=(Results[,"N_est"]-Results[,"N_true"])/Results[,"N_true"]
+Results[,"RelBias_beta"]=(Results[,"beta1_est"]-Results[,"beta1_true"])/Results[,"beta1_true"]
+Results[,"Bias_b"]=(Results[,"b_est"]-Results[,"b_true"]) / ifelse(Results[,"b_true"] != 0,Results[,"b_true"],1)
+
+Results <- Results %>%
+  mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dj","Yi")))
+
+Results$zonespersequence <- as.factor(Results$zonespersequence)
+Results$b_true <- as.factor(Results$b_true)
+Results$relloc_aggreg <- factor(Results$relloc_aggreg,
+                                levels = c("Reallocation: No -  Likelihood: Yi",
+                                           "Reallocation: Yes -  Likelihood: Yi",
+                                           "Reallocation: Yes -  Likelihood: Dj"))
+
+Results_2 <- Results %>%
+  filter(n_samp_com > 10)
+
+RelBias_N_plot <- ggplot()+
+  geom_boxplot(data = Results_2,
+               aes(x = zonespersequence,
+                   y = RelBias_N,
+                   fill = relloc_aggreg))+
+  geom_hline(yintercept = 0,linetype="dashed")+
+  theme_bw()+
+  ylab("Relative bias of biomass")+
+  facet_wrap(.~factor(n_samp_com))+
+  theme(legend.title = element_blank(),
+        aspect.ratio = 1,
+        legend.position = 'none')+
+  ylim(-1,2)+
+  ylab("")+xlab("")
+
+RelBias_beta_plot <- ggplot()+
+  geom_boxplot(data = Results_2,
+               aes(x = zonespersequence,
+                   y = RelBias_beta,
+                   fill = relloc_aggreg))+
+  geom_hline(yintercept = 0,linetype="dashed")+
+  theme_bw()+
+  ylab("Relative bias of beta1")+
+  facet_wrap(.~factor(n_samp_com))+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1)+
+  ylab("")+xlab("")
+
+MSPE_S_plot <- ggplot()+
+  geom_boxplot(data = Results_2,
+               aes(x = zonespersequence,
+                   y = log(MSPE_S),
+                   fill = relloc_aggreg))+
+  theme_bw()+
+  # ylab("MSPE")+
+  facet_wrap(.~factor(n_samp_com))+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1)+
+  ylab("")+xlab("")
+
+legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="right",legend.title = element_blank())))
+
+Perf.metric <- plot_grid(RelBias_N_plot,
+                      RelBias_beta_plot,
+                      MSPE_S_plot,
+                      legend,
+                      ncol = 1,rel_heights = c(1,1,1,0.25))
+
+ggsave("images/Perf.metric_single_square.png",width = 7.5/1.5, height = 15/1.5)
+
+```
 

### 2022-01-07 Version_M2_plusieurszones/draft/paper_3/paper_3.tex
paper_3

@@
 \hypersetup{
-  pdftitle={The title},
+  pdftitle={Can we use reallocated logbooks data to map species distribution habitats?},
   pdfauthor={First Author1 \& Ernst-August Doelle1,2},
@@
 \makeatother
-\shorttitle{Title}
+\shorttitle{SHORTTITLE}
 \keywords{keywords\newline\indent Word count: X}
@@
 
-\title{The title}
+\title{Can we use reallocated logbooks data to map species distribution habitats?}
 \author{First Author\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
@@
 
+How many simulations? 100 or more?
+
 How do we simulate the covariate?
 
-How to justify the number of fishing sequence and the number of VMS pings?
+Still some lack of convergence.
+
+Number of pings \textbar{} nu:
+1\textbar10
+10\textbar100
+100\textbar1000
+
+Size within a fishing sequence: radius = 3
+
+Number of fishing zones: 1\textbar3\textbar5
 
@@
 \section{Results}\label{results}}
+
+\hypertarget{single-square-analysis}{%
+\subsection{Single square analysis}\label{single-square-analysis}}
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-1}Percentage of convergence for each simulation/model configuration}
+\centering
+\begin{tabular}[t]{ccccc}
+\toprule
+Nb samples & Fishing sequence & Reallocation & Likelihood level & Convergence (\%)\\
+\midrule
+10 & 1 & No & Yi & 99.668\\
+10 & 1 & Yes & Yi & 0.333\\
+10 & 1 & Yes & Dj & 0.000\\
+100 & 10 & No & Yi & 100.000\\
+100 & 10 & Yes & Yi & 100.000\\
+\addlinespace
+100 & 10 & Yes & Dj & 92.000\\
+1000 & 100 & No & Yi & 100.000\\
+1000 & 100 & Yes & Yi & 100.000\\
+1000 & 100 & Yes & Dj & 97.333\\
+\bottomrule
+\end{tabular}
+\end{table}
 

### 2022-01-07 Version_M2_plusieurszones/Scripts/simu_main_script.R

@@ -167,7 +167,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   # Reallocation = 0 si pas de reallocation et structuration, = 1 si reallocation et structuration
   
   # Nombre de bateaux ?
-  n_samp_com_vect = c(50, 100, 1000)
+  n_samp_com_vect = c(10, 100, 1000)
   sequencesdepeche_vect = c(1, 10, 100)
   zonespersequence_vect = c(1, 3, 5)
   NPZ <- expand.grid(sequencesdepeche = sequencesdepeche_vect,

### 2022-01-21 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
Start model

@@
 ## Model
+
+Commercial data have been proven to bring valuable information to estimate harvested fish spatial distribution [@pennino_fishery-dependent_2016].
+Unlike scientific survey data and with the exceptions of fishing vessels with on on board observers, commercial catch are rarely precisely geolocated.
+Mostly, catch data are reported at some statistical or management unit, i.e all each vessel reports for each day and each statistical unit the total catch in the logbook.
+
+Many works [@alglave2021integrated] combine VMS data and logbook declaration to associate catch to spatial locations.
+In a nutshell, a classification rule (for example based on speed limits) is used to identify fishing VMS ping and, for each vessel, each day, an average catch per ping is defined as the  declared  catch divided by the total number of fishing pings.
+Each fishing pings is then associated to this average catch.
+This process is referred to as catch reallocation in the following.
 

### 2022-03-16 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
Merge branch 'main' of https://github.com/MarieEtienne/projet_M2_sole

--- 2022-01-07 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
@@
 ---
-title             : "Can we use reallocated logbooks data to map species distribution habitats?"
+title             : "Going from coarse landings data to fine scale species distribution"
 
@@
    - \usepackage{graphicx}
+   - \usepackage{caption}
 
-output            : papaja::apa6_pdf
+output            :
+ papaja::apa6_pdf
+    # reference_docx: mystyles_number_test.docx
 ---
@@
 library(kableExtra)
+library(latex2exp)
+library(mapdata)
+library(scales)
 library(papaja)
+library(rnaturalearth)
+library(sf)
 ```
@@
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+
+mapBase <- map("worldHires", fill = T, plot = F)
+mapBase <- st_as_sf(mapBase) %>% filter(ID %in% c("France","Spain"))
 ```
@@
 
+References for change of support issues in integrated modelling:
+
+https://esajournals.onlinelibrary.wiley.com/doi/epdf/10.1002/ecy.2710
+
+https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12793
+
+https://www.sciencedirect.com/science/article/pii/S0169534719302551#b0010
+
+https://www.sciencedirect.com/science/article/pii/S0006320721001993
+
 # Material and methods
 
-## Model
+We base our approach on the model developped by @alglave_combining_2022 and build on these developpments to provide a solution to overcome the issues related to change of support in our specific context.
 
-See `3_formulas.rmd`.
+## Defining the problem from a modelling point of view
+
+Let's define the latent field $S$ the punctual observations $Y$.
+$S$ is a spatial Gaussian Field (GF) defined in space such as $S(x)=\mu+\beta .
+\Gamma + \delta(x)$.
+$\beta .
+\Gamma$ is the covariate term with $\Gamma$ the matrix of deisgn of the coavariates and $\beta$ the effect of the covariates.
+$\delta(x)$ is GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$).
+
+Punctual observations $Y_i$ are conditionnal on the latent field values at fishing location $x_i$ for observation $i$.
+Observations are supposed to follow some distribution $\mathcal{L}_Y$.
+In our case, data are  positive continuous zero inflated data and they are modelled through a zero-inflated lognormal model described in supplementary material (SM) previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.
+
+$$Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma)$$
+
+All observations belong to a catch declaration D_j such as:
+
+$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$$
+
+$j \in [\![1,n]\!$ is the declaration index with $n$ the number of declarations.
+$\mathcal{P}_j$ is the vector of all fishing positions related to the $j^{th}$ declaration.
+The punctual observations $Y_i$ are indexed through $i \in [\![1,m_j]\!$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.
+
+In standard processing, we know $D_j$ through logbooks data, $x_i$ through VMS data and $D_j$ are reallocated uniformly on related $x_i$ so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$.
+The likelihood is then directly computed on reallocated $Y^*_i$ assuming these are the exact punctual observations.
+This strongly simplifies the actual process of observation and have most likely repercussion on model performance.
+
+To overcome such limitation, an alternative is to consider that only the catch declarations are observed, but they can inform of the punctual observations $Y_i$ which are not-observed (i.e.
+they are latent variables).
+
+Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations $D_j | S_{\mathcal{P}_j},\mathcal{P}_j \sim \mathcal{L}_D( S_{\mathcal{P}_j},\xi,\sigma)$ with $S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))$.
+
+Our approach is to match the 2 first moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$ i.e. $Y_i$ probability distribution) and $\mathcal{L}_D$.
+
+First, as at the punctual level, the declarations can be decomposed in 2 components: (1) the probability to obtain a zero value and, if the declaration is positive, (2) the probability to obtain a certain declaration value.
+
+(1) We derive the probability to obtain a zero-declaration $P(D_j = 0 \vert S_{\mathcal{P}_j},\mathcal{P}_j)$ by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 \vert S(x_i), x_i)$ to all fishing points $x_i \in \mathcal{P}_j$ as $P(D_j = 0 \vert S_{\mathcal{P}_j},\mathcal{P}_j) = \prod_{i \in \mathcal{P}_j} P(Y_i=0 \vert S(x_i), x_i)$.
+
+(2) If the declarations is positive, based on the moments of $\mathcal{L}_Y$, we derive $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$ (see Table ...).
+Assuming $D_j|D_j>0$ follows some probability distribution $\mathcal{L}_{D_j|D_j>0}$, we make the two first moments of the $\mathcal{L}_{D_j|D_j>0}$ coincides with $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$.
+Considering $\mathcal{L}_{D_j|D_j>0}$ has a lognormal distribution, such link is done through the formulas $\mu_D = E(D|D_j>0)$ for the mean component of $\mathcal{L}_{D_j|D_j>0}$ and $\sigma_D = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$ for its variance component.
+
+Equations of $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$ are given by:
+
+$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
+$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
+$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$
+
+All calculations to obtain these formulas are available in SM.
+
+![(\#fig:ParBiasSingle) Reallocation process.](images/realloc.png)
+
+```{r}
+
+# save(data=gridpolygon_sf_2,file="res/simu_lf.RData")
+# 
+# plot_1 <- ggplot(gridpolygon_sf_2)+
+#   geom_sf(aes(fill=Strue_x),alpha=0.9)+
+#   scale_fill_distiller(palette = "Spectral")+
+#   theme_void()+theme(legend.position = "none")
+# 
+# ggsave("images/simu_lf.png",width = 7.5,height = 7.5,dpi = 500)
+
+```
 
@@
 
-```{r}
+```{r,warning=F}
 
@@
 knitr::kable(Results_conv_2,booktabs = T,align = "c",
-                  caption = "Percentage of convergence per simulation/model configuration")
+                  caption = "Percentage of convergence per simulation/model configuration at the level of a single rectangle")
 
@@
 Results[,"Bias_b"]=(Results[,"b_est"]-Results[,"b_true"]) / ifelse(Results[,"b_true"] != 0,Results[,"b_true"],1)
+Results[,"Bias_sigma_com"]=(Results[,"sigma_com_est"]-Results[,"sigma_com_true"]) / Results[,"sigma_com_true"]
+Results[,"Bias_q1_com"]=(Results[,"q1_com_est"]-Results[,"q1_com_true"]) / Results[,"q1_com_true"]
+Results[,"Bias_intercept"]=(Results[,"intercept_est"]-Results[,"intercept_true"]) / Results[,"intercept_true"]
 
@@
 
+MSPE_S_plot <- ggplot()+
+  geom_boxplot(data = Results_2,
+               aes(x = zonespersequence,
+                   y = mspe,
+                   fill = relloc_aggreg),
+               outlier.colour = NA)+
+  theme_bw()+
+  facet_wrap(.~factor(n_samp_com))+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1)+
+  ylab("MSPE")+
+  xlab("")+
+  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
+                labels = trans_format("log10", math_format(10^.x)))+
+  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
+                                "Reallocation: Yes -  Likelihood: Yi",
+                                "Reallocation: Yes -  Likelihood: Dj"),
+                     values=c("gold","#F8766D","#00BA38"))
+
+Beta_plot <- ggplot()+
+  geom_boxplot(data = Results_2,
+               aes(x = zonespersequence,
+                   y = beta1_est,
+                   fill = relloc_aggreg),
+               outlier.colour = NA)+
+  geom_hline(yintercept = 0,linetype="dashed")+
+  geom_hline(yintercept = 2,col="red",linetype="dashed")+
+  theme_bw()+
+  facet_wrap(.~factor(n_samp_com))+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1)+
+  # ylim(-1,1)+ 
+  ylab(parse(text = TeX('$\\beta_S$')))+
+  xlab("")+
+  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
+                                "Reallocation: Yes -  Likelihood: Yi",
+                                "Reallocation: Yes -  Likelihood: Dj"),
+                     values=c("gold","#F8766D","#00BA38"))
+
 RelBias_N_plot <- ggplot()+
@@
                    y = RelBias_N,
-                   fill = relloc_aggreg))+
+                   fill = relloc_aggreg),
+               outlier.colour = NA)+
   geom_hline(yintercept = 0,linetype="dashed")+
   theme_bw()+
-  ylab("Relative bias of biomass")+
   facet_wrap(.~factor(n_samp_com))+
@@
         legend.position = 'none')+
-  ylim(-1,2)+
-  ylab("")+xlab("")
+  ylim(-1,1)+
+  ylab("Relative bias of biomass")+
+  xlab("")+
+  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
+                                "Reallocation: Yes -  Likelihood: Yi",
+                                "Reallocation: Yes -  Likelihood: Dj"),
+                     values=c("gold","#F8766D","#00BA38"))
 
-RelBias_beta_plot <- ggplot()+
-  geom_boxplot(data = Results_2,
-               aes(x = zonespersequence,
-                   y = RelBias_beta,
-                   fill = relloc_aggreg))+
+legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="right",legend.title = element_blank())))
+
+Perf.metric_single_square <- plot_grid(MSPE_S_plot,
+                                       Beta_plot,
+                                       # RelBias_N_plot,
+                                       legend,
+                                       ncol = 1,
+                                       rel_heights = c(1,1,0.25))
+
+ggsave("images/Perf.metric_single_square.png",width = 5*1.25, height = 6.5*1.25)
+
+Results_3 <- Results_2 %>%
+  filter(n_samp_com == 1000)
+
+q1_plot <- ggplot()+
+  geom_boxplot(data = Results_3,
+               aes(x = relloc_aggreg,
+                   y = q1_com_est,
+                   fill = relloc_aggreg),
+               outlier.colour = NA)+
+  geom_hline(yintercept = 0,linetype="dashed")+
+  geom_hline(yintercept = -1,col="red",linetype="dashed")+
+  theme_bw()+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1,
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        plot.title = element_text(hjust = 0.5))+
+  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
+                               "Reallocation: Yes -  Likelihood: Yi",
+                               "Reallocation: Yes -  Likelihood: Dj"),
+                    values=c("gold","#F8766D","#00BA38"))+
+  xlab("")+ylab(parse(text = TeX('$\\xi$')))+
+  ggtitle("Zero-inflation parameter")
+
+sigma_plot <- ggplot()+
+  geom_boxplot(data = Results_3,
+               aes(x = relloc_aggreg,
+                   y = sigma_com_est,
+                   fill = relloc_aggreg),
+               outlier.colour = NA)+
   geom_hline(yintercept = 0,linetype="dashed")+
   theme_bw()+
-  ylab("Relative bias of beta1")+
-  facet_wrap(.~factor(n_samp_com))+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1,
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        plot.title = element_text(hjust = 0.5))+
+  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
+                               "Reallocation: Yes -  Likelihood: Yi",
+                               "Reallocation: Yes -  Likelihood: Dj"),
+                    values=c("gold","#F8766D","#00BA38"))+
+  xlab("")+ylab(parse(text = TeX('$\\sigma^2$')))+
+  ggtitle("Observation variance parameter")+
+  geom_hline(yintercept = 1,col="red",linetype="dashed")
+
+intercept_plot <- ggplot()+
+  geom_boxplot(data = Results_3,
+               aes(x = relloc_aggreg,
+                   y = intercept_est,
+                   fill = relloc_aggreg),
+               outlier.colour = NA)+
+  geom_hline(yintercept = 0,linetype="dashed")+
+  theme_bw()+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1,
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        plot.title = element_text(hjust = 0.5))+
+  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
+                               "Reallocation: Yes -  Likelihood: Yi",
+                               "Reallocation: Yes -  Likelihood: Dj"),
+                    values=c("gold","#F8766D","#00BA38"))+
+  xlab("")+ylab(parse(text = TeX('$\\mu$')))+
+  ggtitle("Intercept of the latent field")+
+  geom_hline(yintercept = 2,col="red",linetype="dashed")
+
+par_plot <- plot_grid(q1_plot,sigma_plot,intercept_plot,nrow = 1,align = "hv")
+
+legend <- as_ggplot(cowplot::get_legend(q1_plot+theme(legend.position="bottom",legend.title = element_blank())))
+
+par_plot <- plot_grid(par_plot,legend,rel_heights = c(1,0.08),ncol=1)
+
+ggsave("images/par_plot_single_square.png",width = 10.5, height = 4)
+
+```
+
+![(\#fig:PerfMetricSingle) Performance metric for single-square simulations.
+columns: number of commercial pings.
+x-axis: number of zones visited within each declaration.
+'Reallocation: ', data are or are not reallocated in simulations.
+'Likelihood: ', the likelihood is computed on reallocated observations $Y_i$ or on catch declarations $D_j$.
+Gold: golden standard.
+Red: actual situation ($Y_i$ level).
+Green: alternative model ($D_j$ level).](images/Perf.metric_single_square.png)
+
+Reallocation has a major effect on predictions and estimates accuracy (Figure \@ref(fig:PerfMetricSingle)).
+Reallocating data conduct to a 10 to 200 times decrease of predictions accuracy when working at punctual observations level (MSPE gold compared to red boxplots) and decreases as the number of visited zones within a declaration increases.
+It also leads to the loss of the species habitat relationship; $\beta_S$ is biased an tends towards 0 as the number of fishing zones within a declaration increases.
+Increasing the number of samples only increases the accuracy of the estimates and spatial predictions but overall spatial predictions remain inaccurate and $\beta_S$ remain biased.
+Regarding other parameters, the zero-inflation parameter is over-estimated (i.e.
+when reallocating, the quantity of data is under-estimated), the observation variance is underestimated (i.e.
+the data is estimated to be less noisy than they actually are) and the interecept of the latent field is slightly over-estimated (Figure \@ref(fig:ParBiasSingle)).
+
+Working at the level of catch declarations $D_j$ allows to recover the species-habitat relationship and to improve the accuracy of spatial predictions even though accuracy is not as accurate as the golden-standard (Figure \@ref(fig:PerfMetricSingle)).
+Furthermore, the zero-inflation parameter is unbiased when the likelihood is built on catch declarations.
+Other parameters (observation variance, intercept) are also better  estimated even though they remain slightly biased (Figure \@ref(fig:ParBiasSingle)).
+This alternative model have some convergence difficulties (Table 1) as 8% of the model did not converged when sample size is medium (100 pings) and only 3% did not when sample size is large (1000 pings).
+
+![(\#fig:ParBiasSingle) Parameters relative bias for single-square simulations.
+'Reallocation: ', data are or are not reallocated in simulations.
+'Likelihood: ', the likelihood is computed on reallocated observations $Y_i$ or on catch declarations $D_j$.
+Gold: golden standard.
+Red: actual situation ($Y_i$ level).
+Green: alternative model ($D_j$ level).
+Only the simulations with 1000 observations are represented.
+Black line: zero value.
+Red line: parameter true value.](images/par_plot_single_square.png)
+
+## Multiple square analysis
+
+```{r}
+
+load("res/Results_full_multi_square.RData")
+
+Results_conv <- Results_2
+Results_conv$one <- 1
+
+Results_conv_2 <- summaryBy(converge+one~
+                              Model+
+                              aggreg_obs,
+                            data=Results_conv,
+                            FUN=sum) %>%
+  dplyr::select(Model,
+                aggreg_obs,
+                converge.sum,
+                one.sum) %>%
+    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dj","Yi"),
+           perc_convergence = round((1 - converge.sum / one.sum)*100,digits = 3))
+
+Results_conv_2 <- Results_conv_2 %>%
+  dplyr::rename(`Likelihood level` = aggreg_obs,
+                `Convergence (%)` = perc_convergence) %>%
+  dplyr::select(-converge.sum,-one.sum)
+
+Results_conv_2$Model <- factor(Results_conv_2$Model,levels = c("Scientific model", "Commercial model", "Integrated model"))
+
+Results_conv_2 <- Results_conv_2[-which(Results_conv_2$Model == "Scientific model" &
+                                          Results_conv_2$`Likelihood level` == "Dj"),]
+Results_conv_2$`Likelihood level`[which(Results_conv_2$Model == "Scientific model" &
+                       Results_conv_2$`Likelihood level` == "Yi")] <- " "
+knitr::kable(Results_conv_2,booktabs = T,align = "c",
+                  caption = "Percentage of convergence per simulation/model configuration at the level of several ICES rectangles")
+
+```
+
+```{r}
+
+Results_plot <- Results_2 %>%
+  filter(converge == 0) %>%
+  filter(simu_type == "Unsampled Rectangles")
+
+Results_plot[,"RelBias_N"]=(Results_plot[,"N_est"]-Results_plot[,"N_true"])/Results_plot[,"N_true"]
+Results_plot[,"RelBias_N.2"]=(Results_plot[,"N_est.2"]-Results_plot[,"N_true.2"])/Results_plot[,"N_true.2"]
+Results_plot[,"RelBias_beta"]=(Results_plot[,"beta1_est"]-Results_plot[,"beta1_true"])/Results_plot[,"beta1_true"]
+
+Results_plot$obs <- NA
+Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
+Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi"
+Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific data"
+Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi","Dj","Scientific data"))
+
+Results_plot$Model[which(Results_plot$Model == "Scientific model")] <- "Scientific"
+Results_plot$Model[which(Results_plot$Model == "Commercial model")] <- "Commercial"
+Results_plot$Model[which(Results_plot$Model == "Integrated model")] <- "Integrated"
+
+# Results_plot_fullArea <- Results_plot
+# Results_plot_UnsampArea <- Results_plot
+
+beta_plot <- ggplot()+
+  geom_boxplot(data = Results_plot,
+               aes(x = Model,
+                   y = beta1_est,
+                   fill = obs))+
+  geom_hline(yintercept = 0,linetype="dashed")+
+  theme_bw()+
   theme(legend.title = element_blank(),
@@
         aspect.ratio = 1)+
-  ylab("")+xlab("")
+  # facet_wrap(.~simu_type)+
+  ylab(TeX("$\\beta_S$")) +
+  xlab("")+
+  geom_hline(yintercept = 2,col="red",linetype="dashed")+
+  scale_color_manual(breaks = c("Scientific model",
+                                "Yi",
+                                "Dj"),
+                     values=c("#619CFF","#F8766D","#00BA38"))
 
 MSPE_S_plot <- ggplot()+
-  geom_boxplot(data = Results_2,
-               aes(x = zonespersequence,
-                   y = log(MSPE_S),
-                   fill = relloc_aggreg))+
+  geom_boxplot(data = Results_plot,
+               aes(x = Model,
+                   y = mspe,
+                   fill = obs))+
   theme_bw()+
-  # ylab("MSPE")+
-  facet_wrap(.~factor(n_samp_com))+
   theme(legend.title = element_blank(),
@@
         aspect.ratio = 1)+
-  ylab("")+xlab("")
+  # facet_wrap(.~simu_type)+
+  ylab("MSPE")+xlab("")+
+  scale_color_manual(breaks = c("Scientific model",
+                                "Yi",
+                                "Dj"),
+                     values=c("#619CFF","#F8766D","#00BA38"))+
+  ylim(0,NA)
 
-legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="right",legend.title = element_blank())))
+Range_plot <- ggplot()+
+  geom_boxplot(data = Results_plot,
+               aes(x = Model,
+                   y = Range,
+                   fill = obs))+
+  geom_hline(yintercept = 0,linetype="dashed")+
+  theme_bw()+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1)+
+  # facet_wrap(.~simu_type)+
+  ylab("Range")+
+  xlab("")+
+  geom_hline(yintercept = 0.6,col="red",linetype="dashed")+
+  scale_color_manual(breaks = c("Scientific model",
+                                "Yi",
+                                "Dj"),
+                     values=c("#619CFF","#F8766D","#00BA38"))+
+  ylim(-0.1,2)
 
-Perf.metric <- plot_grid(RelBias_N_plot,
-                      RelBias_beta_plot,
-                      MSPE_S_plot,
+legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="bottom",legend.title = element_blank())))
+
+pl_plot <- plot_grid(MSPE_S_plot,
+                     beta_plot,
+                     Range_plot,
+                     nrow = 1,
+                     align="hv")
+
+Perf.metric_multi_square <- plot_grid(pl_plot,
                                       legend,
-                      ncol = 1,rel_heights = c(1,1,1,0.25))
+                                      ncol = 1,
+                                      rel_heights = c(1,0.15))
 
-ggsave("images/Perf.metric_single_square.png",width = 7.5/1.5, height = 15/1.5)
+ggsave("images/Perf.metric_multiple_square.png",width = 12, height = 4)
+
+load("res/lkl_prof_beta.RData")
 
 ```
+
+The scientific-based model provides species-habitat relationship and range estimates that are unbiased.
+Whether the model is built on $Y_i$ or on $D_j$, the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models.
+Comparing this to the maps obtained from Figure \@ref(fig:MapSeveral), we highlight that the commercial-based models logically miss the hotspot at 2.5°W - 46.75°N as commercial data do not cover this hotspot, while scientific data provide predictions that do not well capture the local hotposts.
+Integrating scientific and commercial data allows to (1) capture the hotspot missed by commercial data through scientific data and and (2) better capture the local correlation structures through the dense commercial data.
+
+Furtermore, consistently with single-square simulations, working at the level of punctual observations ($Y_i$) conduct to a loss in both the species-habitat relationship and predictions accuracy (Figure \@ref(fig:PerfMetricSeveral)) compared to the model built on commercial declarations ($D_j$).
+
+Interestingly, reallocating does not affect only the species-habitat relationship but also the spatial autocorrelation terms such as the range parameter.
+When working at the $Y_i$ level, the range parameter is biased, while it is not when working at the $D_j$ level.
+
+This is a consequence of the loss of the species-habitat relationship: when working at $Y_i$ level part of the variability related to the covariate effect is captured by the random effect and then the range parameter captures both autocorrelation related to the actual random effect and to the covariate.
+
+Working at the level of catch declarations in estimation allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
+This is evidenced in Figure \@ref(fig:MapSeveral) where the model based on $Y_i$ provide smoothed maps and do not capture the relatively small scale patterns that are shaped by the covariate effect.
+On the other hand the model built on $D_j$ (and the scientific-based model too) better capture and disentangle the covariate effect and the spatial random effect and then provide predictions that better emphasize to the small-scale patterns of species distribution.
+
+However, this goes with some difficulty in convergence as only 75% of the model built on catch declarations converge.
+
+![(\#fig:PerfMetricSeveral) Performance metric for several rectangles simulations.
+columns: commercial data coverage.
+x-axis: likelihood level.
+$1^{st}$ row, red line: true value of $\beta_S$.](images/Perf.metric_multiple_square.png)
+
+![(\#fig:MapSeveral) Relative distribution of simulated/estimated biomass field.
+A: Simulated biomass field with scientific samples (red) and rectangles that have not been sampled by commercial data (transparent rectangles).
+B: simulated biomass field.
+C: biomass field from the scientific-based model.
+$Y_i$: model fitted at the punctual observation level (D, E).
+$D_j$: model fitted at the declaration level (F, G).](images/Map_multi_square.png)
+
+<!-- see the file 'Scripts/In_progress/plot_multiple_square_2.R' -->
+
+## Real case study
+
+```{r}
+
+## Load data
+load("res/sci_df.RData")
+
+load("res/realloc_rest_df.RData")
+
+load("res/no.realloc_df.RData")
+
+load("res/est_par_df_full_3.RData")
+
+## Plot maps
+no.realloc_plot <- ggplot(no.realloc_df)+
+  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
+  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
+  ggtitle("Integrated model",subtitle = "Commercial likelihood on Yi")+
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.subtitle = element_text(hjust = 0.5))+
+  geom_sf(data = mapBase)+
+  coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
+  xlab("")+ylab("")
+
+realloc_rest_plot <- ggplot(realloc_rest_df)+
+  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
+  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
+  ggtitle("Integrated model",subtitle = "Commercial likelihood on Dj")+
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.subtitle = element_text(hjust = 0.5))+
+  geom_sf(data = mapBase)+
+  coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
+  xlab("")+ylab("")
+
+sci_plot <- ggplot(sci_df)+
+  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
+  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
+  ggtitle("Scientific model")+
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.subtitle = element_text(hjust = 0.5))+
+  geom_sf(data = mapBase)+
+  coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
+  xlab("")+ylab("")
+
+# plot(log(no.realloc_df$S_x),log(sci_df$S_x))
+# cor(log(no.realloc_df$S_x),log(sci_df$S_x),method = "spearman")
+# plot(log(realloc_rest_df$S_x),log(sci_df$S_x))
+# cor(log(realloc_rest_df$S_x),log(sci_df$S_x),method = "spearman")
+
+case_study_plot <- plot_grid(sci_plot,no.realloc_plot,realloc_rest_plot,ncol=3,align="hv")
+
+ggsave("images/case_study_plot.png",width = 12, height = 4)
+
+## Parameter estimates
+est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi","Scientific model"))
+est_par_df_full_3 <- est_par_df_full_3 %>%
+  filter(par_names != "substr_Sand_Coarse_substrate")
+par_plot <- ggplot(est_par_df_full_3, aes(y=par_val, x=par_names))+
+  geom_point(
+    aes(color = lkl),
+    position = position_dodge(0.5),
+    size=2
+  )+
+  geom_errorbar(aes(ymin = CI.inf, ymax = CI.sup, color = lkl),
+                position = position_dodge(0.5),width=0.4
+  )+
+  scale_color_manual(breaks = c("Scientific model",
+                                "Yi",
+                                "Dj"),
+                     values=c("#619CFF","#F8766D","#00BA38"))+
+  geom_hline(yintercept=0, linetype="dashed", color = "red",alpha=0.4)+
+  xlab("")+ylab("")+
+  theme_bw()+
+  theme(legend.title = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 1)+
+  coord_flip()+
+  scale_x_discrete(labels=c('k_com'=parse(text = TeX('$k_{com}$')),
+                            'Sigma_com'=parse(text = TeX('$\\sigma_{com}$')),
+                            'q1_com'=parse(text = TeX('$\\xi_{com}$')),
+                            'Sigma_sci'=parse(text = TeX('$\\sigma_{sci}$')),
+                            'q1_sci'=parse(text = TeX('$\\xi_{sci}$')),
+                            'Range'="Range",
+                            'MargSD'="Marginale variance",
+                            'substr_Mud_sediment'=parse(text = TeX('$\\beta_{S}$')),
+                            'intercept'=parse(text = TeX('$\\mu$'))))
+
+legend <- as_ggplot(cowplot::get_legend(
+  par_plot+
+    theme(legend.position="right",
+          legend.title = element_blank()
+          )))
+
+par_plot <- plot_grid(par_plot,legend,ncol=1,rel_heights = c(1,0.25),align="hv")
+
+ggsave("images/par_plot.png",width = 7.5, height = 6)
+
+```
+
+![(\#fig:CaseStudyPar) Parameters obtained with scientific-based model, the integrated model built on punctual observations $Y_i$ and the integrated model built on catch declarations $D_j$.](images/par_plot.png)
+
+![(\#fig:CaseStudyMap) Maps obtained from (left) the scientific-based model, (center) the integrated model built on punctual observations $Y_i$, (right) the integrated model built on catch declarations $D_j$.](images/case_study_plot.png)
+
+Passing from punctual observations to catch declarations modifies the overall distribution patterns and some of the parameters estimates and related confidence intervals.
+Consistently with simulations, model built on $Y_i$ tend to provide biased estimates and under-estimate the uncertainty related to other estimates compared with the scientific model and the model buit on $D_j$.
+Mainly, the effect of substrate is recovered in the integrated model built on $D_j$ (Figure \@ref(fig:CaseStudyMap)).
+The zero-inflation parameter $\xi$ is revised downwards (i.e.
+there are actually more zero-values than in the reallocated data) while the observation variance of commercial data is revised upwards (i.e.
+the commercial data are more noisy than expected when building likelihood on $Y_i$).
+In the model built on $Y_i$, the confidence interval of some parameters (specifically $\beta_S$, the marginal variance, the range, $\xi_{com}$, $\sigma_{com}$) are very narrow while in the model built on $D_j$ confidence intervals are larger.
+
+Furthermore, when comparing the scientific-based model estimates and the integrated model built on $D_j$, some parameters are better estimated.
+For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on $D_j$ substrate is significant and the confidence interval is smaller.
+Other parameters are more precisely estimated such as the marginal variance, the range, $\xi_{sci}$ and $\sigma_{sci}$.
+
+On the contrary, other parameters do not seems well estimated in either $Y_i$ or $D_j$ models.
+For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on $Y_i$ and revised downards when working on $D_j$.
+This is consistent with simulations results, see Figure \@ref(fig:ParBiasSingle).
+
+Regarding the maps of the spatial predictions, working at $D_j$ level strongly modify the model biomass field compared with the $Y_i$ model.
+Mainly, the effect of the covariate have a sharper effect on species distribution.
+Overall, the strength of the hotspots are revised and the map of the model fitted on $D_j$ is less homogeneous than the map of the model built on $Y_i$.
+
+Finally, the $D_j$ model built on commercial data only does not converge (while the one built on $Y_i$ does) emphasizing the model build on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
+Interestingly, apart from the species-habitat relationship that are similar between the scientific-based model and the integrated model built on $D_j$, maps are not similar emphasizing commercial data brings information to the spatial predictions and better allow to capture local spatial patches of biomass.
 

### 2022-03-16 Version_M2_plusieurszones/Scripts/In_progress/plot_case_study.R

@@ -19,6 +19,7 @@
 #   xlab("")+ylab("")
 
 load("results/case_study/no.realloc_int_df.RData")
+# save(data=fit_IM_res,file="results/case_study/no.realloc_int_df.RData")
 
 no.realloc_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
 no.realloc_plot <- ggplot(no.realloc_df)+
@@ -32,21 +33,22 @@ no.realloc_plot <- ggplot(no.realloc_df)+
   coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
   xlab("")+ylab("")
 
-load("results/case_study/realloc_int_df.RData")
-
-realloc_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
-realloc_plot <- ggplot(realloc_df)+
-  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
-  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-  ggtitle("Integrated model",subtitle = "Commercial likelihood on Dj")+
-  theme_bw()+
-  theme(plot.title = element_text(hjust = 0.5),
-        plot.subtitle = element_text(hjust = 0.5))+
-  geom_sf(data = mapBase)+
-  coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
-  xlab("")+ylab("")
+# load("results/case_study/realloc_int_df.RData")
+# 
+# realloc_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
+# realloc_plot <- ggplot(realloc_df)+
+#   geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
+#   scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
+#   ggtitle("Integrated model",subtitle = "Commercial likelihood on Dj")+
+#   theme_bw()+
+#   theme(plot.title = element_text(hjust = 0.5),
+#         plot.subtitle = element_text(hjust = 0.5))+
+#   geom_sf(data = mapBase)+
+#   coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
+#   xlab("")+ylab("")
 
 load("results/case_study/realloc_int_df_rest.RData")
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/In_progress/plot_case_study_lkl_prof.R

@@ -1,0 +1,74 @@
+
+
+## Scientific model
+load("C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/case_study/sci_df.RData")
+prof_sci <- tmbprofile(obj = fit_IM_res$Obj,lincomb = c(0,1,0,0,0,0))
+# x11();plot(prof_sci)
+
+prof_sci_2 <- prof_sci %>%
+  mutate(Model = "Scientific model") %>%
+  mutate(value_2 = (value - mean(value)) / sd(value) ) %>%
+  mutate(lkl = "Yi")
+
+prof_sci_2$value_2 <- prof_sci_2$value - min(prof_sci_2$value)
+
+## No realloc
+load("C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/case_study/no.realloc_int_df.RData")
+prof_no.realloc <- tmbprofile(obj = fit_IM_res$Obj,lincomb = c(0,1,0,0,0,0,0,0,0))
+# x11();plot(prof_no.realloc)
+
+prof_no.realloc_2 <- prof_no.realloc %>%
+  mutate(Model = "Integrated model") %>%
+  mutate(value_2 = (value - mean(value)) / sd(value) ) %>%
+  mutate(lkl = "Yi")
+
+prof_no.realloc_2$value_2 <- prof_no.realloc_2$value - min(prof_no.realloc_2$value)
+
+## Reallocation and re-estimation
+load("C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/case_study/realloc_int_df_rest.RData")
+prof_realloc <- tmbprofile(obj = fit_IM_res$Obj,lincomb = c(0,1,0,0,0,0,0,0))
+# x11();plot(prof_no.realloc)
+
+prof_realloc_2 <- prof_realloc %>%
+  mutate(Model = "Integrated model") %>%
+  mutate(value_2 = (value - min(value))) %>%
+  mutate(lkl = "Dj")
+
+prof_realloc_2$value_2 <- prof_realloc_2$value - min(prof_realloc_2$value)
+
+## Test
+prof_test <- tmbprofile(obj = fit_IM_res$Obj,lincomb = c(0,1,0,0,0,0,0,0,0))
+# x11();plot(prof_no.realloc)
+
+prof_test_2 <- prof_test %>%
+  mutate(Model = "Test") %>%
+  mutate(value_2 = (value - min(value))) %>%
+  mutate(lkl = "Yi")
+
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/In_progress/plot_from_list_param.R

@@ -5,6 +5,8 @@
 folder_name <- "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/com_x_sci_data_14_scientific_commercial_simple-2022-01-04_18_55_00_SimuRef"
 folder_name <- "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/com_x_sci_data_14_scientific_commercial_simple-2022-01-05_15_15_00_SimuRef"
 
+library(HistogramTools)
+
 #-------------
 ## If parallel
 #-------------
@@ -57,7 +59,7 @@ load(paste0(folder_name,"/Results.RData"))
 Results_2 <- Results
 
 # Load List_param
-list_file_i <- list.files(paste0(folder_name,"/",i))
+list_file_i <- list.files(paste0(folder_name,"/"))
 list_file_i <- list_file_i[which(str_detect(list_file_i,"List_param"))]
   
 Results$aggreg_obs <- NA
@@ -70,18 +72,48 @@ Results$beta1_est <- NA
 Results$alpha <- NA
 Results$beta <- NA
 Results$gamma <- NA
+Results$mspe <- NA
+Results$mspe <- NA
+Results$sigma_com_true <- NA
+Results$sigma_com_est <- NA
+Results$q1_com_est <- NA
+Results$q1_com_true <- NA
+Results$intercept_est <- NA
+Results$intercept_true <- NA
 
 # Run over List_param
 for(counter in Results$counter){
+  
   print(paste0("counter: ",counter))
   load(paste0(folder_name,"/List_param_",counter,".RData"))
+  
+  ## Config simu
+  Results$aggreg_obs[which(Results$counter == counter)] <- List_param$data.info$aggreg_obs
   Results$aggreg_obs[which(Results$counter == counter)] <- List_param$data.info$aggreg_obs
   Results$sequencesdepeche[which(Results$counter == counter)] <- List_param$data.info$sequencesdepeche
   Results$zonespersequence[which(Results$counter == counter)] <- List_param$data.info$zonespersequence
   Results$n_samp_com[which(Results$counter == counter)] <- List_param$data.info$n_samp_com
+  
+  ## Coefficient de correlation
   Results$rho_S[which(Results$counter == counter)] <- cor(List_param$data.info$Strue_x,List_param$Report$S_x)
+
+  ## MSPE
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/In_progress/plot_multiple_square.R

@@ -1,0 +1,384 @@
+#----------------------------------
+## Load results for multiple square
+#----------------------------------
+library(HistogramTools)
+library(gt)
+
+
+folder_name <- "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/severa_rect-2022-01-11_17_16_00_SimuTest"
+folder_name <- "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/severa_rect-2022-01-12_13_03_00_SimuUnsampledRect"
+folder_name <- "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/severa_rect-2022-01-13_11_14_00_SimuFullArea"
+
+folder_c <- c("C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/severa_rect-2022-01-12_13_03_00_SimuUnsampledRect",
+              "C:/R_projects/projet_M2_sole/Version_M2_plusieurszones/results/severa_rect-2022-01-13_11_14_00_SimuFullArea")
+
+for(folder_i in folder_c){
+  
+  load(paste0(folder_i,"/Results.RData"))
+  
+  # Load List_param
+  list_file_i <- list.files(paste0(folder_i,"/"))
+  list_file_i <- list_file_i[which(str_detect(list_file_i,"List_param"))]
+  
+  Results$rho_S <- NA
+  Results$MAPE <- NA
+  Results$mspe <- NA
+  Results$alpha <- NA
+  Results$beta <- NA
+  Results$gamma <- NA
+  Results$converge <- NA
+  Results$N_est.2 <- NA
+  Results$N_true.2 <- NA
+  Results$Range <- NA
+  Results$SD_Range <- NA
+  Results$MargeSD <- NA
+  Results$SD_MargeSD <- NA
+  Results$sigma_com_true <- NA
+  Results$sigma_com_est <- NA
+  Results$q1_com_est <- NA
+  Results$q1_com_true <- NA
+  Results$intercept_est <- NA
+  Results$intercept_true <- NA
+  
+  
+  if(str_detect(folder_i,"UnsampledRect")) simu_type <- "Unsampled Rectangles"
+  if(str_detect(folder_i,"FullArea")) simu_type <- "All area sampled"
+  
+  # Run over List_param
+  for(counter in Results$counter){
+    
+    print(paste0("simu_type: ", simu_type," | counter: ",counter))
+    load(paste0(folder_i,"/List_param_",counter,".RData"))
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/In_progress/plot_multiple_square_2.R

@@ -1,0 +1,430 @@
+##############################################################
+## Multiple square simulations - Plot single simulation ouputs
+##############################################################
+
+compute_profile <- F
+save_data <- T
+
+load("draft/paper_3/res/list_input.RData")
+
+loc_x<-list_input$loc_x
+S_x<-list_input$S_x
+c_com_x<-list_input$c_com_x
+y_i2<-list_input$y_i2
+index_i<-list_input$index_i
+y_sci_i<-list_input$y_sci_i
+index_sci_i<-list_input$index_sci_i
+boats_i<-list_input$boats_i
+cov_x_2<-list_input$cov_x_2
+ref_level<-list_input$ref_level
+spde<-list_input$spde
+mesh<-list_input$mesh
+n_cells<-list_input$n_cells
+sampling<-list_input$sampling
+
+# ## To build list input
+# list_input <- list(loc_x=loc_x,
+#                    S_x=S_x,
+#                    c_com_x=c_com_x,
+#                    y_i=y_i,
+#                    y_i2=y_i2,
+#                    index_i=index_i,
+#                    y_sci_i=y_sci_i,
+#                    index_sci_i=index_sci_i,
+#                    boats_i=boats_i,
+#                    cov_x_2=cov_x_2,
+#                    ref_level=ref_level,
+#                    spde=spde,
+#                    mesh=mesh,
+#                    n_cells=n_cells,
+#                    sampling=sampling)
+# save(data=list_input,file="draft/paper_3/res/list_input.RData")
+
+
+## Simulated latent field
+simu_df <- data.frame(loc_x,S_x=S_x)
+
+if(save_data == T) save(data=simu_df,file="draft/paper_3/res/simu_df.RData")
+
+## Simulated data
+sci.data_df <- data.frame(cell = index_sci_i,
+                          y_sci = y_sci_i) %>%
+  inner_join(loc_x[,c("x","y","cell","ICESNAME")])
+
+com.data_df <- data.frame(cell = index_i,
+                          y_com = y_i2) %>%
+  inner_join(loc_x[,c("x","y","cell","ICESNAME")]) %>%
+  group_by(x,y,cell,ICESNAME) %>%
+  dplyr::summarise(n=n())
+
+## Scientific data
+fit_IM_res <- fit_IM(Estimation_model_i = 2,
+                     Samp_process = 0,
+                     EM = "fix_b",
+                     TmbFile = "Scripts/",
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/function/fit_IM.R

@@ -404,15 +404,10 @@ fit_IM <- function(Estimation_model_i = 1,
   #-------------------------
   
   # SD  --> very long with catchability and I got NANs
-  
-  #dans ce if (on rentre dans la boucle si l'algorithme a convergé),
-  #on calcule la matrice de variance covariance de notre modèle, on récupère les 
-  #écarts types des paramètres/grandeurs du modèle
   if(Converge==0){
     Report = Obj$report()
-    SD = sdreport( Obj,ignore.parm.uncertainty = F)
-    SD$unbiased$value = c("total_abundance"=Report$total_abundance)
-    
+    SD = sdreport( Obj,ignore.parm.uncertainty = F,bias.correct = T)
+
   }else{SD = NULL}
   Opt[["run_time"]] = Sys.time()-Start_time
   

### 2022-03-16 Version_M2_plusieurszones/Scripts/inst/executables/com_x_sci_data_14_scientific_commercial_simple.cpp

@@ -149,7 +149,7 @@ Type objective_function<Type>::operator() (){
   Q_delta = Q_spde(spde, exp(logkappa(0)));
   vector<Type> delta_x( n_nodes );delta_x.setZero();
   delta_x = deltainput_x / exp(logtau(0)); // Transform random effects
-  if(Options_vec(13)==1) jnll_comp(0) = GMRF(Q_delta)(deltainput_x);
+  if(Options_vec(13)==1) jnll_comp(3) = GMRF(Q_delta)(deltainput_x);
   
 
   ////////////////
@@ -330,7 +330,7 @@ Type objective_function<Type>::operator() (){
           encounterprob_E_ObsM(i) = ( 1.0 - exp(-1 * E_ObsM(i) * exp(q1_com(0))) );
           log_notencounterprob_ObsM(i) = -1 * E_ObsM(i) * exp(q1_com(0));
           
-          jnll_comp(0) -= dzinflognorm(y_ObsM_i(i), log(E_ObsM(i))-log(encounterprob_E_ObsM(i)), encounterprob_E_ObsM(i), log_notencounterprob_ObsM(i), Sigma_com(0), true);
+          jnll_comp(4) -= dzinflognorm(y_ObsM_i(i), log(E_ObsM(i))-log(encounterprob_E_ObsM(i)), encounterprob_E_ObsM(i), log_notencounterprob_ObsM(i), Sigma_com(0), true);
           // jnll_comp(0) -= dzinfgamma(y_com_i(i), E_com(i)/encounterprob_com(i), encounterprob_com(i), log_notencounterprob_com(i), Sigma_com(0), true);
           
         }
@@ -423,7 +423,7 @@ Type objective_function<Type>::operator() (){
         encounterprob_sci(i) = ( 1.0 - exp(-1 * E_sci(i) * exp(q1_sci(0)) ));
         log_notencounterprob_sci(i) = -1 * E_sci(i) * exp(q1_sci(0));
 
-        jnll_comp(0) -= dzinflognorm(y_sci_i(i), log(E_sci(i))-log(encounterprob_sci(i)), encounterprob_sci(i), log_notencounterprob_sci(i), Sigma_sci, true);
+        jnll_comp(2) -= dzinflognorm(y_sci_i(i), log(E_sci(i))-log(encounterprob_sci(i)), encounterprob_sci(i), log_notencounterprob_sci(i), Sigma_sci, true);
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/reallocation_real_simu.R

@@ -3,6 +3,7 @@
 ##################################################
 ## B. Alglave
 
+memory.size(1e5)
 
 library(cowplot)
 library(dplyr)
@@ -30,7 +31,7 @@ setwd(folder_phd_codes)
 domain_shapefile <- "ORHAGO" # "ORHAGO", "EVHOE"
 grid_projection <- "+proj=longlat +datum=WGS84"
 mapBase <- ne_countries(scale = "medium", returnclass = "sf")
-grid_xmin <- -6 # Coordinates of study area 
+grid_xmin <- -6 # Coordinates of study area
 grid_xmax <- 0
 grid_ymin <- 42
 grid_ymax <- 48
@@ -53,15 +54,19 @@ if(species_to_plot == "Solea_solea"){
 }
 
 ## data file name
-sampling <- "from_tacsateflalo" # "simulate": simulate sampling, "from_tacsateflalo": take data points from tacsateflalo
+sampling <- "simulate" # "simulate": simulate sampling, "from_tacsateflalo": take data points from tacsateflalo
+simu_name <- "SimuFullArea"
+
+# file name for saving outputs
+Start_time.tot = round(Sys.time(), units = "mins")
+Start_time.tot_2 <- str_replace_all(Start_time.tot, " ", "_")
+Start_time.tot_2 <- str_replace_all(Start_time.tot_2, ":", "_")
+simu_file <- paste0("results/severa_rect-",Start_time.tot_2,"_",simu_name,"/")
+
 commercial_data_folder <- 'data/raw/vms_logbooks/bob_cs/MergeTACSAT_EFLALO/'
 ObsMer_file <- 'data/raw/SIH/obsmer'
 ObsM <- T
 
-## If simulate data points
-# n_samp_com <- 5000
-n_seq_com <- 300
-n_ping_per_seq <- 10
 
 ## Load ICES rectangle shapefile
 ICES_rect <- st_read("data/raw/shapefile/ices_rect_stats/ICES_Statistical_Rectangles_Eco.shp")
@@ -111,7 +116,7 @@ useless_levels <- c("NA","Seabed")
 normalize_covariates <- T
 
 ## Reshape datapoint_cov
-latent_field_cov <- c("substr","bathy") # ,"SST","salinity", 
+latent_field_cov <- c("bathy") # ,"substr","SST","salinity", 
 discret_cov <- c("aggreg_substrate","strata") # "bathy"
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/source/load_commercial_data.R

@@ -1,0 +1,32 @@
+#######################
+## Load commercial data
+#######################
+
+## Load tacsatEflalo
+setwd(folder_phd_codes)
+filter_with_Eflalo <- F
+source("r/real_data/load_data/extract_commercial_data.R")
+
+## Load ObsMer
+HH_obsmer <- read.csv(paste0(ObsMer_file,"/HH_OBSMER_VIII_VII_2005_2019_230320.csv"),sep = ";",header = T)
+HL_obsmer <- read.csv(file = paste0(ObsMer_file,"/HL_OBSMER_VIII_VII_2005_2019_230320.csv"),sep = ";",header = T)
+SL_obsmer <- read.csv(file = paste0(ObsMer_file,"/SL_OBSMER_VIII_VII_2005_2019_230320.csv"),sep = ";",header = T)
+TR_obsmer <- read.csv(file = paste0(ObsMer_file,"/TR_OBSMER_VIII_VII_2005_2019_230320.csv"),sep = ";",header = T)
+
+## Load scientific data
+if(species_to_plot == "Solea_solea") load("data/tidy/fitting_intermediate/bob_cs/Solea_solea/OTB/with_size/mature_survey_data.RData")
+if(species_to_plot == "Merluccius_merluccius") load("data/tidy/fitting_intermediate/bob_cs/Merluccius_merluccius/OTB/with_size/mature_survey_data.RData")
+
+setwd(folder_project)
+
+## 'VMS x logbooks'
+#------------------
+source("Scripts/source/shape_VMS_logbooks.R")
+
+## ObsMer data
+#-------------
+source("Scripts/source/shape_ObsMer.R")
+
+## Scientific data
+#-----------------
+source("Scripts/source/shape_scientific.R")

### 2022-03-16 Version_M2_plusieurszones/Scripts/source/plot_outputs.R

@@ -1,0 +1,331 @@
+################################################
+## Plot outputs from mutliple square simulations
+################################################
+## B. Alglave
+
+
+#----------------------------------------
+## Plot latent field and covariate effect
+#----------------------------------------
+if(plot_outputs){
+  
+  if(sampling == "simulate"){
+    
+    ## Plot simulations outputs
+    plot(x=log(S_x),
+         y=log(Report$S_x),
+         main="Estimated S_x vs. simulated S_x",
+         xlab="Simulated",
+         ylab="Estimated")
+    mtext(text = paste0("Est_mod: ",Data_source[Estimation_model_i],
+                        " Realloc: ",aggreg_obs,
+                        " Converge: ",ifelse(opt$convergence==0,T,F)),
+          cex=0.75)
+    
+    for(i in 1:2){
+      if(i==1){
+        S_plot <- S_x
+        main_title <- "Simulated S_x"
+        breaks_S <- quantile(S_plot,probs = seq(0,1,length.out=101))
+        breaks_S <- unique(breaks_S)
+        S_plot.ref <- S_x
+        S_interval <- cut(S_plot,breaks = 100,include.lowest = T)
+        pal_col <- rev(colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))(100))
+        # pal_col <- rev(heat.colors(100))
+      }else if(i==2){
+        S_plot <- Report$S_x
+        main_title <- "Estimated S_x"
+        breaks_S <- quantile(S_plot,probs = seq(0,1,length.out=101))
+        breaks_S <- unique(breaks_S)
+        S_plot.ref <- S_x
+        S_interval <- cut(S_plot,breaks = 100,include.lowest = T)
+        pal_col <- rev(colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))(100))
+      }
+      S_interval <- cut(S_plot,breaks = breaks_S, include.lowest = T)
+      S_xint <- data.frame(S_interval = S_interval, cell=1:n_cells)
+      col_S_xint <- data.frame(S_interval=levels(S_interval),col=pal_col)
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/source/save_outputs.R

@@ -1,0 +1,111 @@
+###############################
+## Save outputs of simumlations
+###############################
+## B. Alglave
+
+simu.est.data <- list(S_x=S_x,
+                  index_i=index_i,
+                  y_sci_i=y_sci_i,
+                  y_i=y_i,
+                  boats_i=boats_i,
+                  cov_x=cov_x,
+                  index_sci_i=index_sci_i,
+                  intercept=intercept,
+                  beta1=beta1,
+                  SD_delta=SD_delta,
+                  range_delta=range_delta,
+                  q1_com=q1,
+                  Sigma_com=SD_obs,
+                  q1_sci=q1_sci,
+                  Sigma_sci=Sigma_sci,
+                  nu = nu,
+                  range_cov=range_cov,
+                  SD_cov=SD_cov,
+                  q1=q1,
+                  n_seq_com=n_seq_com,
+                  n_ping_per_seq=n_ping_per_seq,
+                  zonesize=zonesize,
+                  n_zone=n_zone,
+                  sample_full_rect=sample_full_rect,
+                  k_sim=k_sim,
+                  unsampled_rect=unsampled_rect,
+                  q1_sci=q1_sci,
+                  n_samp_sci=n_samp_sci,
+                  k_com=1,
+                  aggreg_obs=aggreg_obs,
+                  Estimation_model=Estimation_model_i,
+                  counter=counter,
+                  sim=i)
+
+
+
+# index_sci_i=index_sci_i,
+# c_sci_x=c_sci_x,
+# y_sci_i=y_sci_i,
+# b=b,
+# index_com_i=index_com_i,
+# y_com_i=y_com_i,
+# # b_com_i=b_com_i,
+# c_com_x=c_com_x,
+# aggreg_obs=aggreg_obs,
+# k=k,
+# sequencesdepeche=sequencesdepeche,
+# zonespersequence=zonespersequence,
+# taillezone=taillezone,
+# 
+List_param <- list(data.info = simu.est.data,
+                   Opt_par = Opt,
+                   SD = SD,
+                   Report = Report)
+
+if(counter == 1){
+  
+  Results <- data.frame(counter=counter,
+                        sim=i,
+                        Estimation_model=Estimation_model_i,
+                        aggreg_obs=aggreg_obs,
+                        n_seq_com=n_seq_com,
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/source/simulate_real.R

@@ -9,11 +9,18 @@ area_df <- loc_x %>%
   mutate(ICESNAME = factor(ICESNAME)) %>%
   arrange(ICESNAME)
 
+## Sampling is null in 1/3 of the statistical rectangles
+if(unsampled_rect) area_df$perc_area[sample(x = 1:nrow(area_df),size = 1/3 * nrow(area_df),replace = F)] <- 0
+
+# area_df <- points_df %>%
+#   group_by(ICESNAME) %>%
+#   dplyr::summarise(area=n()) %>%
+#   mutate(perc_area = area/(sum(area))) %>%
+#   mutate(ICESNAME = factor(ICESNAME)) %>%
+#   arrange(ICESNAME)
+
 ## Sequential sampling (not that there is no preferential sampling)
 # First select a statistical rectangle
-sample_full_rect <- F
-zonesize <- 0.2
-n_zone <- 1 # c(1,3,5)
 samp_df <- do.call(rbind,lapply(1:n_seq_com, function(j){
   
   ## Select a statistical rectangle
@@ -73,9 +80,13 @@ c_com_x <- samp_df %>%
   c() %>% unlist() %>%
   as.matrix()
 
+# test <- cbind(loc_x,c_com_x)
+# 
+# ggplot(test)+
+#   geom_point(aes(x=long,y=lati,col=c_com_x))+scale_color_distiller(palette = "Spectral")
+
+
 ## Simulate catch observations (zero-inflated lognormal distribution)
-q1 <- -1
-SD_obs <- exp(0.3)
 y_i <- do.call(c,lapply(1:length(index_i), function(j){
   exp_catch <- S_x[index_i[j]]
   prob_encount <- 1-exp(- exp(q1) * exp_catch)
@@ -88,7 +99,6 @@ y_i <- do.call(c,lapply(1:length(index_i), function(j){
 }))
 
 ## Reallocation
-k_sim <- 1
 if(k_sim==1){ ## Reallocate data
   mean_y_i <- aggregate(x = y_i, 
                         by = list(unique.values = boats_i),
@@ -119,7 +129,6 @@ if(k_sim==1){ ## Reallocate data
 
 index_sci_i <- c()
 n_cells <- nrow(loc_x)
-n_samp_sci <- 75
 nb_hauls_strata <- loc_x %>% 
   mutate(value=1) %>%
   group_by(strata) %>%
[…]

### 2022-03-16 Version_M2_plusieurszones/Scripts/source/step_estimation.R

@@ -1,0 +1,122 @@
+#################################
+## Step estimation initialization
+#################################
+## B. Alglave
+
+if(Estimation_model_i == 1){
+  
+  Obj_step.est <- Obj
+  init_val <- Obj_step.est$env$last.par.best
+  Params <- fit_IM_res$Params
+  Map <- fit_IM_res$Map
+  
+  #-------------------
+  ## PARAMETERS UPDATE
+  #-------------------
+  ## Observation parameters
+  Params$logSigma_com <- init_val[which(names(init_val) == "logSigma_com")]+0.00001
+  Params$logSigma_sci <- init_val[which(names(init_val) == "logSigma_sci")]
+  Params$q1_sci <- init_val[which(names(init_val) == "q1_sci")]
+  Params$q1_com <- init_val[which(names(init_val) == "q1_com")]
+  
+  ## Covariates
+  Params$beta_j[which(!is.na(Map$beta_j))] <- init_val[which(names(init_val) == "beta_j")]
+  
+  ## Spatial random effect structure
+  if("logtau" %in% names(init_val)) Params$logtau <- init_val[which(names(init_val) == "logtau")]
+  if("logkappa" %in% names(init_val)) Params$logkappa <- init_val[which(names(init_val) == "logkappa")]
+  Params$deltainput_x <- init_val[which(names(init_val) == "deltainput_x")]
+  
+  if(step.est>0) aggreg_obs <- T
+  
+  #---------------------------------------
+  # COVARIATES - 1st PARAMETER TO LET FREE
+  #---------------------------------------
+  Map[["beta_j"]] <- as.character(Map[["beta_j"]])
+  if(step.est == 1) Map[["beta_j"]][1] <- factor(NA)
+  if(step.est > 1) Map$beta_j[1] <- 1
+  Map[["beta_j"]] <- factor(Map[["beta_j"]])
+  
+  #-----------------------------------------
+  # CATCHABILITY - 2nd PARAMETER TO LET FREE
+  #-----------------------------------------
+  if("k_com" %in% names(init_val)) Params$k_com <- init_val[which(names(init_val) == "k_com")]
+  if("k_sci" %in% names(init_val)) Params$k_sci <- init_val[which(names(init_val) == "k_sci")]
+  
+  if(step.est %in% c(1,2)){
+    Map[["k_com"]] <- factor(NA)
+    Map[["k_sci"]] <- factor(NA)
+  }
[…]

### 2022-03-16 Version_M2_plusieurszones/draft/paper_3/paper_3.tex
Update paper_3.tex

@@
 \hypersetup{
-  pdftitle={Can we use reallocated logbooks data to map species distribution habitats?},
+  pdftitle={Going from coarse landings data to fine scale species distribution},
   pdfauthor={First Author1 \& Ernst-August Doelle1,2},
@@
 \usepackage{graphicx}
+\usepackage{caption}
 \ifxetex
@@
 
-\title{Can we use reallocated logbooks data to map species distribution habitats?}
+\title{Going from coarse landings data to fine scale species distribution}
 \author{First Author\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
@@
 
+References for change of support issues in integrated modelling:
+
+\url{https://esajournals.onlinelibrary.wiley.com/doi/epdf/10.1002/ecy.2710}
+
+\url{https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12793}
+
+\url{https://www.sciencedirect.com/science/article/pii/S0169534719302551\#b0010}
+
+\url{https://www.sciencedirect.com/science/article/pii/S0006320721001993}
+
 \hypertarget{material-and-methods}{%
@@
 
-\hypertarget{model}{%
-\subsection{Model}\label{model}}
+We base our approach on the model developped by Alglave et al.
+(2022) and build on these developpments to provide a solution to overcome the issues related to change of support in our specific context.
 
-See \texttt{3\_formulas.rmd}.
+\hypertarget{defining-the-problem-from-a-modelling-point-of-view}{%
+\subsection{Defining the problem from a modelling point of view}\label{defining-the-problem-from-a-modelling-point-of-view}}
+
+Let's define the latent field \(S\) the punctual observations \(Y\).
+\(S\) is a spatial Gaussian Field (GF) defined in space such as \(S(x)=\mu+\beta .
+\Gamma + \delta(x)\).
+\(\beta .
+\Gamma\) is the covariate term with \(\Gamma\) the matrix of deisgn of the coavariates and \(\beta\) the effect of the covariates.
+\(\delta(x)\) is GF capturing spatial correlation (\(\delta \sim \mathcal{N}(0,\Sigma)\)).
+
+Punctual observations \(Y_i\) are conditionnal on the latent field values at fishing location \(x_i\) for observation \(i\).
+Observations are supposed to follow some distribution \(\mathcal{L}_Y\).
+In our case, data are positive continuous zero inflated data and they are modelled through a zero-inflated lognormal model described in supplementary material (SM) previously introduced by Thorson (2018) and already used in Alglave et al.
+(2022).
+
+\[Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma)\]
+
+All observations belong to a catch declaration D\_j such as:
+
+\[D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}\]
+
+\(j \in [\![1,n]\!\) is the declaration index with \(n\) the number of declarations.
+\(\mathcal{P}_j\) is the vector of all fishing positions related to the \(j^{th}\) declaration.
+The punctual observations \(Y_i\) are indexed through \(i \in [\![1,m_j]\!\) with \(m_j\) the number of fishing positions belonging to the \(j^{th}\) declaration.
+
+In standard processing, we know \(D_j\) through logbooks data, \(x_i\) through VMS data and \(D_j\) are reallocated uniformly on related \(x_i\) so that derived punctual observations \(Y^*_i\) are computed as \(Y^*_i=D_j/m_j\).
+The likelihood is then directly computed on reallocated \(Y^*_i\) assuming these are the exact punctual observations.
+This strongly simplifies the actual process of observation and have most likely repercussion on model performance.
+
+To overcome such limitation, an alternative is to consider that only the catch declarations are observed, but they can inform of the punctual observations \(Y_i\) which are not-observed (i.e.~they are latent variables).
+
+Such way, we define some distribution \(\mathcal{L}_D\) as the distribution of the catch declarations \(D_j | S_{\mathcal{P}_j},\mathcal{P}_j \sim \mathcal{L}_D( S_{\mathcal{P}_j},\xi,\sigma)\) with \(S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))\).
+
+Our approach is to match the 2 first moments of \(D_j\) (obtained from the moments of \(\mathcal{L}_Y\) i.e.~\(Y_i\) probability distribution) and \(\mathcal{L}_D\).
+
+First, as at the punctual level, the declarations can be decomposed in 2 components: (1) the probability to obtain a zero value and, if the declaration is positive, (2) the probability to obtain a certain declaration value.
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\item
+  We derive the probability to obtain a zero-declaration \(P(D_j = 0 \vert S_{\mathcal{P}_j},\mathcal{P}_j)\) by simply multiplying the probability to obtain a zero-punctual observation \(P(Y_i=0 \vert S(x_i), x_i)\) to all fishing points \(x_i \in \mathcal{P}_j\) as \(P(D_j = 0 \vert S_{\mathcal{P}_j},\mathcal{P}_j) = \prod_{i \in \mathcal{P}_j} P(Y_i=0 \vert S(x_i), x_i)\).
+\item
+  If the declarations is positive, based on the moments of \(\mathcal{L}_Y\), we derive \(E(D_j \vert D_j > 0)\) and \(Var(D_j \vert D_j > 0)\) (see Table \ldots).
+Assuming \(D_j|D_j>0\) follows some probability distribution \(\mathcal{L}_{D_j|D_j>0}\), we make the two first moments of the \(\mathcal{L}_{D_j|D_j>0}\) coincides with \(E(D_j \vert D_j > 0)\) and \(Var(D_j \vert D_j > 0)\).
+Considering \(\mathcal{L}_{D_j|D_j>0}\) has a lognormal distribution, such link is done through the formulas \(\mu_D = E(D|D_j>0)\) for the mean component of \(\mathcal{L}_{D_j|D_j>0}\) and \(\sigma_D = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)\) for its variance component.
+\end{enumerate}
+
+Equations of \(E(D_j \vert D_j > 0)\) and \(Var(D_j \vert D_j > 0)\) are given by:
+
+\[E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}\]
+\[Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2\]
+\[Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))\]
+
+All calculations to obtain these formulas are available in SM.
+
+\begin{figure}
+\centering
+\includegraphics{images/realloc.png}
+\caption{\label{fig:ParBiasSingle} Reallocation process.}
+\end{figure}
 
@@
 
-\caption{\label{tab:unnamed-chunk-1}Percentage of convergence for each simulation/model configuration}
+\caption{\label{tab:unnamed-chunk-2}Percentage of convergence per simulation/model configuration at the level of a single rectangle}
 \centering
@@
 
+\begin{figure}
+\centering
+\includegraphics{images/Perf.metric_single_square.png}
+\caption{\label{fig:PerfMetricSingle} Performance metric for single-square simulations.
+columns: number of commercial pings.
+x-axis: number of zones visited within each declaration.
+\enquote{Reallocation:}, data are or are not reallocated in simulations.
+\enquote{Likelihood:}, the likelihood is computed on reallocated observations \(Y_i\) or on catch declarations \(D_j\).
+Gold: golden standard.
+Red: actual situation (\(Y_i\) level).
+Green: alternative model (\(D_j\) level).}
+\end{figure}
+
+Reallocation has a major effect on predictions and estimates accuracy (Figure \ref{fig:PerfMetricSingle}).
+Reallocating data conduct to a 10 to 200 times decrease of predictions accuracy when working at punctual observations level (MSPE gold compared to red boxplots) and decreases as the number of visited zones within a declaration increases.
+It also leads to the loss of the species habitat relationship; \(\beta_S\) is biased an tends towards 0 as the number of fishing zones within a declaration increases.
+Increasing the number of samples only increases the accuracy of the estimates and spatial predictions but overall spatial predictions remain inaccurate and \(\beta_S\) remain biased.
+Regarding other parameters, the zero-inflation parameter is over-estimated (i.e.~when reallocating, the quantity of data is under-estimated), the observation variance is underestimated (i.e.~the data is estimated to be less noisy than they actually are) and the interecept of the latent field is slightly over-estimated (Figure \ref{fig:ParBiasSingle}).
+
+Working at the level of catch declarations \(D_j\) allows to recover the species-habitat relationship and to improve the accuracy of spatial predictions even though accuracy is not as accurate as the golden-standard (Figure \ref{fig:PerfMetricSingle}).
+Furthermore, the zero-inflation parameter is unbiased when the likelihood is built on catch declarations.
+Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \ref{fig:ParBiasSingle}).
+This alternative model have some convergence difficulties (Table 1) as 8\% of the model did not converged when sample size is medium (100 pings) and only 3\% did not when sample size is large (1000 pings).
+
+\begin{figure}
+\centering
+\includegraphics{images/par_plot_single_square.png}
+\caption{\label{fig:ParBiasSingle} Parameters relative bias for single-square simulations.
+\enquote{Reallocation:}, data are or are not reallocated in simulations.
+\enquote{Likelihood:}, the likelihood is computed on reallocated observations \(Y_i\) or on catch declarations \(D_j\).
+Gold: golden standard.
+Red: actual situation (\(Y_i\) level).
+Green: alternative model (\(D_j\) level).
+Only the simulations with 1000 observations are represented.
+Black line: zero value.
+Red line: parameter true value.}
+\end{figure}
+
+\hypertarget{multiple-square-analysis}{%
+\subsection{Multiple square analysis}\label{multiple-square-analysis}}
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-4}Percentage of convergence per simulation/model configuration at the level of several ICES rectangles}
+\centering
+\begin{tabular}[t]{ccc}
+\toprule
+Model & Likelihood level & Convergence (\%)\\
+\midrule
+Commercial model & Yi & 100.000\\
+Commercial model & Dj & 75.377\\
+Integrated model & Yi & 100.000\\
+Integrated model & Dj & 76.382\\
+Scientific model &  & 100.000\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+## Warning: Removed 7 rows containing non-finite values (stat_boxplot).
+\end{verbatim}
+
+The scientific-based model provides species-habitat relationship and range estimates that are unbiased.
+Whether the model is built on \(Y_i\) or on \(D_j\), the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models.
+Comparing this to the maps obtained from Figure \ref{fig:MapSeveral}, we highlight that the commercial-based models logically miss the hotspot at 2.5°W - 46.75°N as commercial data do not cover this hotspot, while scientific data provide predictions that do not well capture the local hotposts.
+Integrating scientific and commercial data allows to (1) capture the hotspot missed by commercial data through scientific data and and (2) better capture the local correlation structures through the dense commercial data.
+
+Furtermore, consistently with single-square simulations, working at the level of punctual observations (\(Y_i\)) conduct to a loss in both the species-habitat relationship and predictions accuracy (Figure \ref{fig:PerfMetricSeveral}) compared to the model built on commercial declarations (\(D_j\)).
+
+Interestingly, reallocating does not affect only the species-habitat relationship but also the spatial autocorrelation terms such as the range parameter.
+When working at the \(Y_i\) level, the range parameter is biased, while it is not when working at the \(D_j\) level.
+
+This is a consequence of the loss of the species-habitat relationship: when working at \(Y_i\) level part of the variability related to the covariate effect is captured by the random effect and then the range parameter captures both autocorrelation related to the actual random effect and to the covariate.
+
+Working at the level of catch declarations in estimation allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
+This is evidenced in Figure \ref{fig:MapSeveral} where the model based on \(Y_i\) provide smoothed maps and do not capture the relatively small scale patterns that are shaped by the covariate effect.
+On the other hand the model built on \(D_j\) (and the scientific-based model too) better capture and disentangle the covariate effect and the spatial random effect and then provide predictions that better emphasize to the small-scale patterns of species distribution.
+
+However, this goes with some difficulty in convergence as only 75\% of the model built on catch declarations converge.
+
+\begin{figure}
+\centering
+\includegraphics{images/Perf.metric_multiple_square.png}
+\caption{\label{fig:PerfMetricSeveral} Performance metric for several rectangles simulations.
+columns: commercial data coverage.
+x-axis: likelihood level.
+\(1^{st}\) row, red line: true value of \(\beta_S\).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{images/Map_multi_square.png}
+\caption{\label{fig:MapSeveral} Relative distribution of simulated/estimated biomass field.
+A: Simulated biomass field with scientific samples (red) and rectangles that have not been sampled by commercial data (transparent rectangles).
+B: simulated biomass field.
+C: biomass field from the scientific-based model.
+\(Y_i\): model fitted at the punctual observation level (D, E).
+\(D_j\): model fitted at the declaration level (F, G).}
+\end{figure}
+
+\hypertarget{real-case-study}{%
+\subsection{Real case study}\label{real-case-study}}
+
+\begin{figure}
+\centering
+\includegraphics{images/par_plot.png}
+\caption{\label{fig:CaseStudyPar} Parameters obtained with scientific-based model, the integrated model built on punctual observations \(Y_i\) and the integrated model built on catch declarations \(D_j\).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{images/case_study_plot.png}
+\caption{\label{fig:CaseStudyMap} Maps obtained from (left) the scientific-based model, (center) the integrated model built on punctual observations \(Y_i\), (right) the integrated model built on catch declarations \(D_j\).}
+\end{figure}
+
+Passing from punctual observations to catch declarations modifies the overall distribution patterns and some of the parameters estimates and related confidence intervals.
+Consistently with simulations, model built on \(Y_i\) tend to provide biased estimates and under-estimate the uncertainty related to other estimates compared with the scientific model and the model buit on \(D_j\).
+Mainly, the effect of substrate is recovered in the integrated model built on \(D_j\) (Figure \ref{fig:CaseStudyMap}).
+The zero-inflation parameter \(\xi\) is revised downwards (i.e.~there are actually more zero-values than in the reallocated data) while the observation variance of commercial data is revised upwards (i.e.~the commercial data are more noisy than expected when building likelihood on \(Y_i\)).
+In the model built on \(Y_i\), the confidence interval of some parameters (specifically \(\beta_S\), the marginal variance, the range, \(\xi_{com}\), \(\sigma_{com}\)) are very narrow while in the model built on \(D_j\) confidence intervals are larger.
+
+Furthermore, when comparing the scientific-based model estimates and the integrated model built on \(D_j\), some parameters are better estimated.
+For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on \(D_j\) substrate is significant and the confidence interval is smaller.
+Other parameters are more precisely estimated such as the marginal variance, the range, \(\xi_{sci}\) and \(\sigma_{sci}\).
+
+On the contrary, other parameters do not seems well estimated in either \(Y_i\) or \(D_j\) models.
+For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on \(Y_i\) and revised downards when working on \(D_j\).
+This is consistent with simulations results, see Figure \ref{fig:ParBiasSingle}.
+
+Regarding the maps of the spatial predictions, working at \(D_j\) level strongly modify the model biomass field compared with the \(Y_i\) model.
+Mainly, the effect of the covariate have a sharper effect on species distribution.
+Overall, the strength of the hotspots are revised and the map of the model fitted on \(D_j\) is less homogeneous than the map of the model built on \(Y_i\).
+
+Finally, the \(D_j\) model built on commercial data only does not converge (while the one built on \(Y_i\) does) emphasizing the model build on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
+Interestingly, apart from the species-habitat relationship that are similar between the scientific-based model and the integrated model built on \(D_j\), maps are not similar emphasizing commercial data brings information to the spatial predictions and better allow to capture local spatial patches of biomass.
+
 \hypertarget{discussion}{%
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-alglave_combining_2022}{}%
+Alglave, B., Rivot, E., Etienne, M.-P., Woillez, M., Thorson, J.
+T., \& Vermard, Y.
+(2022).
+Combining scientific survey and commercial catch data to map fish distribution.
+\emph{ICES Journal of Marine Science}, fsac032.
+\url{https://doi.org/10.1093/icesjms/fsac032}
+
+\leavevmode\hypertarget{ref-thorson_three_2018}{}%
+Thorson, J.
+T.
+(2018).
+Three problems with the conventional delta-model for biomass sampling data, and a computationally efficient alternative.
+\emph{Canadian Journal of Fisheries and Aquatic Sciences}, \emph{75}(9), 1369--1382.
 

### 2022-03-16 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd

@@
 ```
+Commercial data have been proven to bring valuable information to estimate harvested fish spatial distribution [@pennino_fishery-dependent_2016].
+Unlike scientific survey data and with the exceptions of fishing vessels with on on board observers, commercial catch are rarely precisely geolocated.
+Mostly, catch data are reported at some statistical or management unit, i.e all each vessel reports for each day and each statistical unit the total catch in the logbook.
+
+Many works [@alglave2021integrated] combine VMS data and logbook declaration to associate catch to spatial locations.
+In a nutshell, a classification rule (for example based on speed limits) is used to identify fishing VMS ping and, for each vessel, each day, an average catch per ping is defined as the  declared  catch divided by the total number of fishing pings.
+Each fishing pings is then associated to this average catch.
+This process is referred to as catch reallocation in the following.
+>>>>>>> a68a28decddeaa4f316defe756ead059dc53ecd6
 

### 2022-03-29 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3

--- 2022-03-16 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
@@
 author: 
-  - name          : "First Author"
+  - name          : "Author1"
     affiliation   : "1"
@@
       - Writing - Review & Editing
-  - name          : "Ernst-August Doelle"
+  - name          : "Author2"
     affiliation   : "1,2"
@@
   - id            : "1"
-    institution   : "Wilhelm-Wundt-University"
+    institution   : "Instit1"
   - id            : "2"
-    institution   : "Konstanz Business School"
+    institution   : "Instit2"
 
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
+  Test
 
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
+  Test
 
@@
 
-**To do:**
-
-- **email to Kasper**
-
-- **redaction of M&M**
-
-- **run single square simulations**
-
-- **run multiple square simulations**
-
 References for change of support issues in integrated modelling:
@@
 
-We base our approach on the model developped by @alglave_combining_2022 and build on these developpments to provide a solution to overcome the issues related to change of support in our specific context.
+We base our approach on the model developped by @alglave_combining_2022.
+In this framework, all observations were supposed to be realised at the same scale.
+Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and potentially combine these data with other (more accurate) datasets such as scientific data.
 
@@
 
-Let's define the latent field $S$ the punctual observations $Y$.
-$S$ is a spatial Gaussian Field (GF) defined in space such as $S(x)=\mu+\beta .
-\Gamma + \delta(x)$.
+**Definition of the variables**
+
+Let's define the latent field $S$ and the punctual observations $Y$.
+$S$ is a spatial log-Gaussian Field (GF) defined as $\log(S(x))=\mu+\beta .
+\Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)).
+We consider a discrete study domain where $x \in [\![1,n]\!]$ stands for the discrete locations.
 $\beta .
-\Gamma$ is the covariate term with $\Gamma$ the matrix of deisgn of the coavariates and $\beta$ the effect of the covariates.
-$\delta(x)$ is GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$).
+\Gamma(x)$ is the covariate term with $\Gamma$ the matrix of design of the covariates and $\beta$ the effect of the covariates.
+$\delta(x)$ is a GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$).
 
-Punctual observations $Y_i$ are conditionnal on the latent field values at fishing location $x_i$ for observation $i$.
+All fishing locations $x_i$ are supposed to be known through VMS data.
+At each fishing location $x_i$, fishermen realize a catch $Y_i$ conditionnally on the latent field values (referred hereafter as a punctual catch).
 Observations are supposed to follow some distribution $\mathcal{L}_Y$.
-In our case, data are  positive continuous zero inflated data and they are modelled through a zero-inflated lognormal model described in supplementary material (SM) previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.
+In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.
 
-$$Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma)$$
+$$Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma^2)$$
 
-All observations belong to a catch declaration D_j such as:
+$\mathcal{L}_Y$ is decomposed it 2 parts:
+
+- the probability to obtain a zero-catch which is modelled as a Bernoulli variable with probability $p_{i}=\exp(-e^\xi .S(x_{i}))$.
+$\xi$ is a parameter controlling zero-inflation.
+When $\xi$ increases, the amount of zero in the data decreases.
+
+- if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM...)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$.
+
+$$
+\operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i)\right) =
+\left\{
+    \begin{array}{ll}
+        p_i & \text { if } y_{i}=0 \\
+        \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0
+    \end{array}
+\right.
+$$
+
+Fishermen must declare what they catch at the level of the statistical rectangle.
+Declarations are daily recorded in logbooks data.
+The declaration (denoted $D_j$) is a summation of all the $Y_i$ realized on the fishing positions belonging to the catch declaration $D_j$ and they can be expressed as:
 
@@
 
-$j \in [\![1,n]\!$ is the declaration index with $n$ the number of declarations.
-$\mathcal{P}_j$ is the vector of all fishing positions related to the $j^{th}$ declaration.
-The punctual observations $Y_i$ are indexed through $i \in [\![1,m_j]\!$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.
+$\mathcal{P}_j$ is the vector of the fishing observations belonging to declaration $D_j$.
+The punctual observations $Y_i$ are indexed through $i \in [\![1,m_j]\!]$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.
 
-In standard processing, we know $D_j$ through logbooks data, $x_i$ through VMS data and $D_j$ are reallocated uniformly on related $x_i$ so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$.
-The likelihood is then directly computed on reallocated $Y^*_i$ assuming these are the exact punctual observations.
+**Catch reallocation: uniform reallocation or model-based reallocation**
+
+In standard processing, $D_j$ (from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$.
+This is what we call uniform (or proportional) reallocation.
+In this case, inference of species distribution is directly computed through reallocated $Y^*_i$ assuming these are the exact punctual observations.
 This strongly simplifies the actual process of observation and have most likely repercussion on model performance.
 
-To overcome such limitation, an alternative is to consider that only the catch declarations are observed, but they can inform of the punctual observations $Y_i$ which are not-observed (i.e.
-they are latent variables).
+To overcome such limitation, an alternative is to consider that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables).
+Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations:
 
-Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations $D_j | S_{\mathcal{P}_j},\mathcal{P}_j \sim \mathcal{L}_D( S_{\mathcal{P}_j},\xi,\sigma)$ with $S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))$.
+$$D_j | \mathcal{P}_j,S_{\mathcal{P}_j} \sim \mathcal{L}_D( S_{\mathcal{P}_j},\xi,\sigma^2)$$
 
-Our approach is to match the 2 first moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$ i.e. $Y_i$ probability distribution) and $\mathcal{L}_D$.
+with $S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))$, $\xi$ the zero-inflation parmeter and $\sigma^2$ the variance parameter.
+As $D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$, it is possible to relate $D_j$, $Y_i$ and $S_{\mathcal{P}_j}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$.
+Our approach is to match the moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_j$.
+This is what we call model-based reallocation.
+This model will be referred as the $D_j$ model.
+In contrast, the model fitted on punctual observations $Y_i$ will be referred as the $Y^*_i$ model.
 
-First, as at the punctual level, the declarations can be decomposed in 2 components: (1) the probability to obtain a zero value and, if the declaration is positive, (2) the probability to obtain a certain declaration value.
+In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | S_{\mathcal{P}_j},\mathcal{P}_j$.
 
-(1) We derive the probability to obtain a zero-declaration $P(D_j = 0 \vert S_{\mathcal{P}_j},\mathcal{P}_j)$ by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 \vert S(x_i), x_i)$ to all fishing points $x_i \in \mathcal{P}_j$ as $P(D_j = 0 \vert S_{\mathcal{P}_j},\mathcal{P}_j) = \prod_{i \in \mathcal{P}_j} P(Y_i=0 \vert S(x_i), x_i)$.
+**Matching $\mathcal{L}_Y$ and $\mathcal{L}_D$**
 
-(2) If the declarations is positive, based on the moments of $\mathcal{L}_Y$, we derive $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$ (see Table ...).
-Assuming $D_j|D_j>0$ follows some probability distribution $\mathcal{L}_{D_j|D_j>0}$, we make the two first moments of the $\mathcal{L}_{D_j|D_j>0}$ coincides with $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$.
-Considering $\mathcal{L}_{D_j|D_j>0}$ has a lognormal distribution, such link is done through the formulas $\mu_D = E(D|D_j>0)$ for the mean component of $\mathcal{L}_{D_j|D_j>0}$ and $\sigma_D = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$ for its variance component.
+First, as at the punctual level ($Y$), the declarations can be decomposed in 2 components: (1) the probability to obtain a zero-value declaration and, if the declaration is positive, (2) the probability to obtain some declaration value $d_j$.
 
-Equations of $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$ are given by:
+(1) We derive the probability to obtain a zero-declaration $P(D_j = 0)$ by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.
+
+\begin{align*}
+P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
+\end{align*}
+
+(2) If the declaration is positive, we assume the probability to obtain catch $d_j$ follows a continous distribution $\mathcal{L}_{D_j|D_j>0}$ with mean component $\mu_j$ and variance component $\sigma_j$.
+From $\mathcal{L}_Y$ we can derive $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$ (given below) and re-express $\mu_j$ and $\sigma_j$ as a function of these two moments.
 
@@
 
+If we consider $\mathcal{L}_{D_j|D_j>0}$ is lognormal (with parameterization described in SM...) then, $\mu_j = E(D|D_j>0)$ and $\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$.
 All calculations to obtain these formulas are available in SM.
+By fitting the model on declaration $D_j$ rather than reallocated catch $Y_i$ and linking $\mathcal{L}_D$ to $\mathcal{L}_Y$, we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation.
 
-![(\#fig:ParBiasSingle) Reallocation process.](images/realloc.png)
+The inference is realized through Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].
+
+![(\#fig:SchemeRealloc) Reallocation process.](images/realloc.png)
 
@@
 
-### Single-square simulations
+To evaluate what bring the performance of the new approach, we assess models performance through simulation-estimation.
 
-How many simulations? 100 or more?
+First, we conduct basic simulations in order to explore the base properties of the framework and the effect of some key variables on model performance.
+We assume the study domain is a single statistical rectangle, only commercial data feed the model and the expression of the latent field in both simulation and estimation is simplified (the variability only depends on one covariate, not on a spatial random term).
+These simulations will be referred as single-square simulations.
 
-How do we simulate the covariate?
+Then, we extend the simulation-estimation study to several rectangles to get closer to a real case study configuration.
+We add a spatial random effect in the latent field (for both simulation and estimation) and we also simulate scientific data (in addition to commercial data) to explore the contribution of both datasets in inference.
+These simulations will be referred as multiple-square simulations.
 
-Still some lack of convergence.
+In these two sets of simulation-estimation, the covariate is modelled as a continous GRF that we suppose known at each point of the grid.
+The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$.
+Regarding commercial data, the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for a single declaration in real data.
+All parameterizations are detailed in the table ....
 
-Number of pings | nu:
-1|10
-10|100
-100|1000
+Each time we compare several model configurations and evaluate these in regards to 2 metrics:
 
-Size within a fishing sequence: radius = 3
+- the MSPE which quantifies the accuracy of the spatial predictions of the latent field over the spatial domain ($n$ is the number of locations over the grid).
 
-Number of fishing zones: 1|3|5
+$$MSPE=\frac{\sum_x (S(x) - \hat{S}(x) ) ^2}{n}$$
 
-### Multiple-squares simulations
+- the parameter of the species habitat relationship $\beta_S$ compared with its estimate $\hat \beta_S$.
 
-How do we simulate the covariate?
+**Single-square simulations**
+
+Two important variables may affect the accuracy of model outputs: the amount of commercial data and the number of fishing zones explored within one single declaration.
+In these simulations-estimations, we intend to explore the effect of these 2 variables.
+
+First, the amount of commercial data; it is expected that when increasing the amount of data the accuracy of the predictions will be improved.
+In simulations, we progressively increase the number of fishing positions (10, 100 and 1000 fishing pings).
+As there are 10 pings per declaration, this means that the number of declarations increases respectively (1, 10 and 100 declarations).
+
+Furthermore, the samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle).
+Note that we make a distinction between fishing zones and statistical rectangles: fishing zones are included in a statistical rectangle and there can be several fishing zones within a statistical rectangles.
+Consequently, a declaration can mix catches that have been realized in zones of high biomass and in zones of low biomass.
+For instance, a declaration can aggregate data from several fishing operations that occured the same day within the same statistical rectangle, but that were realized on 2 distinct types of fishing grounds (and then on two distinct habitats).
+Reallocating uniformly the declaration will strongly homogeneise the actual catch, will mix up information from to different grounds and then may lead to a strong loss of accuracy in model outputs.
+It is expected that when the number of fishing zones within a declaration increases, the accuracy of the model outputs decreases.
+To assess the effect of such process, we simulated the pings of a fishing declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones.
+
+Concretely, for each fishing declaration the fishing pings are sampled in 2 steps: first the centroid of the zones are randomly sampled over the statistical rectangle (here the simulation domain) and then the fishing positions are randomly sampled within the radius of the centroid of the zones.
+The zone size was chosen to be a square of side 7.
+This is a realistic order of magnitude if we consider each zone is one fishing operation and that for each fishing operation the distance of the operation will not exceed 30km.
+
+<!--  One fishing operation is about 3 or 4 hours at a speed of 4 nodes (7.4 km/h) so it makes a distance between 22 and 29 km.
+If we consider the side of the study domain (here a 25 x 25 grid) is $1° \approx 80km$ (as France is at 45° of latitude), a 7 cell zone is equivalent to a 22 km zone.
+-->
+
+At each fishing position, a catch is realized conditionnally on the value of the latent field and follow the probability distribution described above ($\mathcal{L}_y$).
+
+We compare three simulation/estimation configurations:
+
+- a golden standard configuration: observations are known exactly and the $Y^*_i$ model is fitted to the exact data.
+
+- a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations.
+We then fit the $Y^*_i$ model to the reallocated data ($Y_i^*$).
+
+- a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the $D_j$ model to the data (i.e. the one working at the level of $D_j$).
+
+In addition to the metrics we introduced to compare model configuration ($MSPE$ and species-habitat parameter $\beta_S$), we also assess the simulated/estimated values for the intercept $\mu$, the observation variance parameter $\sigma^2$ and the zero-inflation parameter $\xi$.
+
+**Multiple-square simulations**
+
+In this case, we extend the study to more than one single rectangle, we simulate scientific data and include these in inference.
+Finally, in addition to the covariate, we also simulate/estimate a spatial random effect in the latent field as a Matérn GRF.
+
+The study area is based on the case study; it includes the whole coast of the Bay of Biscay (Figure \@ref(fig:MapSeveral)A).
+To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration).
+Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area.
+Similarly to the single-square simulations, the sampling of the commercial points of a declaration are realized in 3 steps.
+(1) The declaration is randomly affected to one of the ICES rectangle.
+(2) The centroid of the fishing zone is randomly sampled within this statistical rectangle.
+(3) The 10 fishing locations are randomly sampled within the fishing zone.
+The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km.
+Note that we do not explore the effect of exploring several zones within the same declaration as it is already done in the single-square simulations.
+
+100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A).
+Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table ...).
+
+We compare several model configurations:
+
+- to assess what bring our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.
+
+- to assess the information brought by each data source, we compare models built on scientific data only (scientific-based models), models built on commercial data only (commercial-based models) and models combining both data sources (integrated models).
+
+In addition to the 2 metrics introduced at the beginning of the section ($MSPE$ and species-habitat parameter $\beta_S$), we also compare the simulated/estimated values of the range parameter.
 
 ## Case-study: sole of the Bay of Biscay
+
+To illustrate our method on a real case study, we applied the approach to the common sole of the Bay of Biscay.
+VMS-logbooks data were extracted for the bottom trawlers fleet (OTB).
+The methods to cross VMS-logbooks data and to filter the fleet is already extensively described in the previous papers [@alglave_combining_2022] and is not developped further here.
+Scientific data were extracted from the DATRAS database for the Orhago beam trawl survey [@biais_gerard_orhago_2003;@ices_report_2018].
+To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - @ices_report_2018-1).
+To illustrate the method, we compare the outputs of (1) the scientific-based model, (2) the integrated model fitted to scientific data and reallocated commercial data ($Y^*_i$ model) and (3) the integrated model fitted to scientific and declarations data ($D_j$ model).
+
+The $D_j$ model faced congvergence issues (some of the parameters were hardly estimated e.g.
+the range parameter).
+To ease convergence, we integrated in the analysis onboard observer data for the same fleet.
+They are considered as exact commercial data of catches.
+Then, integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
+observation variance and zero-inflation parameter of commercial data).
+
+Furthermore, we adopt a step estimation procedure to initialize the optimization algorithm for the $D_j$ model.
+We first fit the $Y^*_i$ model and use the estimates of the model to initialize the optimization of the $D_j$ model.
+We eventually fix the parameters that are hard to estimate in first steps of optimization (intercept $\mu$, covariate effect $\beta_S$, range and marginal variance) and let them free iteratively.
 
@@
 Results_conv_2 <- Results_conv_2 %>%
-  dplyr::rename(`Nb samples` = n_samp_com,
-                `Fishing sequence` = sequencesdepeche,
+  dplyr::rename(`Fishing positions` = n_samp_com,
+                `Declarations` = sequencesdepeche,
                 `Reallocation` = reallocation,
@@
 Gold: golden standard.
-Red: actual situation ($Y_i$ level).
-Green: alternative model ($D_j$ level).](images/Perf.metric_single_square.png)
+Gold: golden standard.
+Red: uniform reallocation situation.
+Green: model-based reallocation ($D_j$ model).
+Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the $D_j$ model.](images/Perf.metric_single_square.png)
 
 Reallocation has a major effect on predictions and estimates accuracy (Figure \@ref(fig:PerfMetricSingle)).
-Reallocating data conduct to a 10 to 200 times decrease of predictions accuracy when working at punctual observations level (MSPE gold compared to red boxplots) and decreases as the number of visited zones within a declaration increases.
-It also leads to the loss of the species habitat relationship; $\beta_S$ is biased an tends towards 0 as the number of fishing zones within a declaration increases.
-Increasing the number of samples only increases the accuracy of the estimates and spatial predictions but overall spatial predictions remain inaccurate and $\beta_S$ remain biased.
-Regarding other parameters, the zero-inflation parameter is over-estimated (i.e.
-when reallocating, the quantity of data is under-estimated), the observation variance is underestimated (i.e.
-the data is estimated to be less noisy than they actually are) and the interecept of the latent field is slightly over-estimated (Figure \@ref(fig:ParBiasSingle)).
+Reallocating data conduct to a 10 to 200 times decrease in accuracy for spatial predictions when fitting the model to reallocated catch (MSPE gold compared to red boxplots).
+Accuracy decreases as the number of visited zones within a declaration increases.
+Besides, $\beta_S$ estimates are biased and reallocation leads to the loss of the species-habitat relationship as the number of fishing zones (within a declaration) increases ($\beta_S$ estimates get closer to 0).
+Increasing the number of samples does not improve inference.
+Regarding other parameters estimates, the zero-inflation parameter ($\xi$) is over-estimated (i.e.
+when uniformly reallocating commercial data, the quantity of positive observations is under-estimated), the observation variance ($\sigma$) is underestimated (i.e.
+the data is estimated to be less noisy than they actually are) and the intercept of the latent field ($\mu$) is slightly over-estimated (Figure \@ref(fig:ParBiasSingle)).
 
-Working at the level of catch declarations $D_j$ allows to recover the species-habitat relationship and to improve the accuracy of spatial predictions even though accuracy is not as accurate as the golden-standard (Figure \@ref(fig:PerfMetricSingle)).
-Furthermore, the zero-inflation parameter is unbiased when the likelihood is built on catch declarations.
+The $D_j$ model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \@ref(fig:PerfMetricSingle)).
+Still, the $D_j$ model outputs are not as accurate as the ones of the golden-standard.
+Furthermore, the zero-inflation parameter is unbiased when the model is fitted to catch declarations $D_j$.
 Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \@ref(fig:ParBiasSingle)).
@@
 Gold: golden standard.
-Red: actual situation ($Y_i$ level).
-Green: alternative model ($D_j$ level).
-Only the simulations with 1000 observations are represented.
+Red: uniform reallocation situation.
+Green: model-based reallocation ($D_j$ model).
+Only the simulations with 1000 fishing positionss are represented.
 Black line: zero value.
@@
 
-The scientific-based model provides species-habitat relationship and range estimates that are unbiased.
-Whether the model is built on $Y_i$ or on $D_j$, the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models.
-Comparing this to the maps obtained from Figure \@ref(fig:MapSeveral), we highlight that the commercial-based models logically miss the hotspot at 2.5°W - 46.75°N as commercial data do not cover this hotspot, while scientific data provide predictions that do not well capture the local hotposts.
-Integrating scientific and commercial data allows to (1) capture the hotspot missed by commercial data through scientific data and and (2) better capture the local correlation structures through the dense commercial data.
+Whether the model is fitted at the punctual level or at the declaration level, the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models.
+This can be well evidenced when looking at Figure \@ref(fig:MapSeveral).
+Integrating scientific and commercial data allows to (1) capture the hotspot missed by commercial data through scientific data and (2) better capture the local correlation structures through the dense commercial data.
 
-Furtermore, consistently with single-square simulations, working at the level of punctual observations ($Y_i$) conduct to a loss in both the species-habitat relationship and predictions accuracy (Figure \@ref(fig:PerfMetricSeveral)) compared to the model built on commercial declarations ($D_j$).
+Furthermore, consistently with single-square simulations, uniform reallocation conducts to a loss in both the predictions accuracy and the species-habitat relationship (Figure \@ref(fig:PerfMetricSeveral)) compared to the model built on commercial declarations ($D_j$).
 
-Interestingly, reallocating does not affect only the species-habitat relationship but also the spatial autocorrelation terms such as the range parameter.
-When working at the $Y_i$ level, the range parameter is biased, while it is not when working at the $D_j$ level.
-
-This is a consequence of the loss of the species-habitat relationship: when working at $Y_i$ level part of the variability related to the covariate effect is captured by the random effect and then the range parameter captures both autocorrelation related to the actual random effect and to the covariate.
-
-Working at the level of catch declarations in estimation allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
-This is evidenced in Figure \@ref(fig:MapSeveral) where the model based on $Y_i$ provide smoothed maps and do not capture the relatively small scale patterns that are shaped by the covariate effect.
-On the other hand the model built on $D_j$ (and the scientific-based model too) better capture and disentangle the covariate effect and the spatial random effect and then provide predictions that better emphasize to the small-scale patterns of species distribution.
-
-However, this goes with some difficulty in convergence as only 75% of the model built on catch declarations converge.
+Interestingly, uniform reallocation does not affect only the species-habitat relationship but also the spatial autocorrelation terms such as the range parameter.
+The model fitted on reallocated data ($Y_i^*$) provides biased range estimates while the $D_j$ model provides unbiased estimates.
+This is a consequence of the loss of the species-habitat relationship: when uniformly reallocating declarations, part of the variability related to the covariate effect is captured by the random effect.
+Consequently, the range parameter captures both the autocorrelation related to the actual random effect and to the covariate.
+The $D_j$ model allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
+This is evidenced in Figure \@ref(fig:MapSeveral) where the model fitted to reallocated catch $Y_i^*$ provides smoothed maps and does not captures the relatively small scale patterns that are shaped by the covariate.
+On the other hand the $D_j$ model (as the scientific-based model) better capture and disentangle the covariate effect and the spatial random effect and then provides predictions that better fit to the small-scale patterns of the species distribution.
+However, this goes with some difficulty in convergence as only 75% of the model built on catch declarations converge (Table 2).
 
@@
 $Y_i$: model fitted at the punctual observation level (D, E).
-$D_j$: model fitted at the declaration level (F, G).](images/Map_multi_square.png)
+$D_j$: $D_j$ model (F, G).](images/Map_multi_square.png)
 
@@
 
-![(\#fig:CaseStudyPar) Parameters obtained with scientific-based model, the integrated model built on punctual observations $Y_i$ and the integrated model built on catch declarations $D_j$.](images/par_plot.png)
+![(\#fig:CaseStudyPar) Parameters obtained with scientific-based model, the integrated model fitted on reallocated catch $Y_i^*$ and the integrated model fitted on catch declarations $D_j$.](images/par_plot.png)
 
-![(\#fig:CaseStudyMap) Maps obtained from (left) the scientific-based model, (center) the integrated model built on punctual observations $Y_i$, (right) the integrated model built on catch declarations $D_j$.](images/case_study_plot.png)
+![(\#fig:CaseStudyMap) Maps obtained from (left) the scientific-based model, (center) the integrated model fitted on punctual observations $Y_i^*$, (right) the integrated model fitted on catch declarations $D_j$.](images/case_study_plot.png)
 
-Passing from punctual observations to catch declarations modifies the overall distribution patterns and some of the parameters estimates and related confidence intervals.
-Consistently with simulations, model built on $Y_i$ tend to provide biased estimates and under-estimate the uncertainty related to other estimates compared with the scientific model and the model buit on $D_j$.
-Mainly, the effect of substrate is recovered in the integrated model built on $D_j$ (Figure \@ref(fig:CaseStudyMap)).
+Consistently with simulations, the $D_j$ model modifies some of the parameters estimates and revises the spatial pattern of the species distribution compared with the $Y^*_i$ model.
+In particular, the substrate effect is recovered in the $D_j$ model and fall in the same range of values of the scientific-based estimate (Figure \@ref(fig:CaseStudyMap)).
 The zero-inflation parameter $\xi$ is revised downwards (i.e.
@@
 the commercial data are more noisy than expected when building likelihood on $Y_i$).
-In the model built on $Y_i$, the confidence interval of some parameters (specifically $\beta_S$, the marginal variance, the range, $\xi_{com}$, $\sigma_{com}$) are very narrow while in the model built on $D_j$ confidence intervals are larger.
 
-Furthermore, when comparing the scientific-based model estimates and the integrated model built on $D_j$, some parameters are better estimated.
+In addition, uncertainty is also revised when fitting the model at the declaration level.
+For instance, when comparing the $Y^*_i$ model to the $D_j$ model, the confidence intervals of $\beta_S$, the marginal variance, the range, $\xi_{com}$, $\sigma_{com}$ are much wider.
+This emphasizes that uncertainty is probably underestimated in the $Y^*_i$ model compared with the $D_j$ model.
+Now, when comparing the scientific-based model and the integrated $D_j$ model, some parameters are better estimated than when there is only scientific data feeding the model.
 For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on $D_j$ substrate is significant and the confidence interval is smaller.
-Other parameters are more precisely estimated such as the marginal variance, the range, $\xi_{sci}$ and $\sigma_{sci}$.
 
-On the contrary, other parameters do not seems well estimated in either $Y_i$ or $D_j$ models.
+On the contrary, other parameters do not seems be well estimated in either the $Y_i^*$ or the $D_j$ models.
 For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on $Y_i$ and revised downards when working on $D_j$.
@@
 
-Regarding the maps of the spatial predictions, working at $D_j$ level strongly modify the model biomass field compared with the $Y_i$ model.
-Mainly, the effect of the covariate have a sharper effect on species distribution.
-Overall, the strength of the hotspots are revised and the map of the model fitted on $D_j$ is less homogeneous than the map of the model built on $Y_i$.
+Regarding the maps of the species distribution, fitting the model at the $D_j$ level strongly modifies the model biomass field compared with the $Y_i^*$ model.
+In particular, the substrate covariate have a sharper effect on species distribution.
+Furthermore, the intensity of the hotspots are revised when fitting the model on $D_j$.
 
-Finally, the $D_j$ model built on commercial data only does not converge (while the one built on $Y_i$ does) emphasizing the model build on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
-Interestingly, apart from the species-habitat relationship that are similar between the scientific-based model and the integrated model built on $D_j$, maps are not similar emphasizing commercial data brings information to the spatial predictions and better allow to capture local spatial patches of biomass.
+Finally, the $D_j$ model fitted only on commercial data does not converge (while the one built on $Y_i^*$ does) emphasizing the model fitted on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
 

### 2022-03-29 Version_M2_plusieurszones/draft/paper_3/paper_3.tex
paper_3

--- 2022-01-07 Version_M2_plusieurszones/draft/paper_3/paper_3.tex
@@
 \hypersetup{
-  pdftitle={Can we use reallocated logbooks data to map species distribution habitats?},
-  pdfauthor={First Author1 \& Ernst-August Doelle1,2},
+  pdftitle={Going from coarse landings data to fine scale species distribution},
+  pdfauthor={Author11 \& Author21,2},
   pdfkeywords={keywords},
@@
 
-\title{Can we use reallocated logbooks data to map species distribution habitats?}
-\author{First Author\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
+\title{Going from coarse landings data to fine scale species distribution}
+\author{Author1\textsuperscript{1} \& Author2\textsuperscript{1,2}}
 \date{}
@@
 
-Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+Test
 
-Enter author note here.
+The authors made the following contributions. Author1: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Author2: Writing - Review \& Editing.
 
-The authors made the following contributions. First Author: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing.
-
-Correspondence concerning this article should be addressed to First Author, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
+Correspondence concerning this article should be addressed to Author1, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Instit1\\\textsuperscript{2} Instit2}
 
 \abstract{
-One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.
-
-Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.
-
-One sentence clearly stating the \textbf{general problem} being addressed by this particular study.
-
-One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).
-
-Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
-
-One or two sentences to put the results into a more \textbf{general context}.
-
-Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
+Test
 }
@@
 
-\textbf{To do:}
+References for change of support issues in integrated modelling:
 
-\begin{itemize}
-\item
-  \textbf{email to Kasper}
-\item
-  \textbf{redaction of M\&M}
-\item
-  \textbf{run single square simulations}
-\item
-  \textbf{run multiple square simulations}
-\end{itemize}
+\url{https://esajournals.onlinelibrary.wiley.com/doi/epdf/10.1002/ecy.2710}
+
+\url{https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12793}
+
+\url{https://www.sciencedirect.com/science/article/pii/S0169534719302551\#b0010}
+
+\url{https://www.sciencedirect.com/science/article/pii/S0006320721001993}
 
@@
 
-\hypertarget{model}{%
-\subsection{Model}\label{model}}
+We base our approach on the model developped by Alglave et al. (2022). In this framework, all observations were supposed to be realised at the same scale.
 
-See \texttt{3\_formulas.rmd}.
+Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and combine these data with other (more accurate) datasets such as scientific data.
+
+In the following we describe the processes from a modeling perspective in order to clearly define the problem we adress.
+
+\hypertarget{defining-the-problem-from-a-modelling-point-of-view}{%
+\subsection{Defining the problem from a modelling point of view}\label{defining-the-problem-from-a-modelling-point-of-view}}
+
+\textbf{Definition of the variables}
+
+Let's define the latent field \(S\) and the punctual observations \(Y\).
+\(S\) is a spatial log-Gaussian Field (GF) defined as \(\log(S(x))=\mu+\beta .
+\Gamma(x) + \delta(x)\) (Figure \label{fig:ParBiasSingle}).
+We consider a discrete study domain where \(x \in [\![1,n]\!]\) stands for the discrete locations.
+\(\beta .
+\Gamma(x)\) is the covariate term with \(\Gamma\) the matrix of design of the covariates and \(\beta\) the effect of the covariates.
+\(\delta(x)\) is a GF capturing spatial correlation (\(\delta \sim \mathcal{N}(0,\Sigma)\)).
+
+All fishing locations \(x_i\) are supposed to be known through VMS data.
+At each fishing location \(x_i\), fishermen realize a catch \(Y_i\) conditionnally on the latent field values.
+Observations are supposed to follow some distribution \(\mathcal{L}_Y\).
+In our case, data are zero-inflated positive continuous data and so they are modelled through a zero-inflated lognormal model previously introduced by Thorson (2018) and already used in Alglave et al.
+(2022).
+
+\[Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma)\]
+
+\(\mathcal{L}_Y\) is decomposed it 2 parts:
+
+\begin{itemize}
+\item
+  the probability to obtain a zero-catch which is modelled as Bernoulli variable with probability \(p_{i}=\exp(-e^\xi .S(x_{i}))\).
+\item
+  if the catch is positive, the probability to obtain some value \(y_i\) is modelled through a continous distribution \(\operatorname{L}\) (here a lognormal distribution - its parameterization is described in SM\ldots)) with mean component \(\frac{S(x_{i})}{1-p_{i}}\) and variance term \(\sigma^2\).
+\end{itemize}
+
+\[
+\operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i)\right) =
+\left\{
+    \begin{array}{ll}
+        p_i & \text { if } y_{i}=0 \\
+        \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0
+    \end{array}
+\right.
+\]
+
+Fishermen must declare what they catch at the level of the statistical rectangle.
+The declaration \(D_j\) is then a summation of all the \(Y_i\) realized on the fishing positions belonging to the catch declaration \(D_j\) (\(j\) is the declaration index).
+We denote the vector of the fishing observations belonging to declaration \(D_j\) as \(\mathcal{P}_j\).
+Then, \(D_j\) can be expressed as:
+
+\[D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}\]
+
+The punctual observations \(Y_i\) are indexed through \(i \in [\![1,m_j]\!]\) with \(m_j\) the number of fishing positions belonging to the \(j^{th}\) declaration.
+\(D_j\) are commonly recorded in logbooks data.
+
+\textbf{Catch reallocation: uniform reallocation or model-based reallocation}
+
+\(D_j\) are known through logbooks data, \(x_i\) are known through VMS data.
+Consequently, in standard processing \(D_j\) are reallocated uniformly on related \(x_i\) so that derived punctual observations \(Y^*_i\) are computed as \(Y^*_i=D_j/m_j\).
+This is what we call uniform (or proportional) reallocation.
+In this case, inference of species distribution is directly computed through reallocated \(Y^*_i\) assuming these are the exact punctual observations.
+This strongly simplifies the actual process of observation and have most likely repercussion on model performance.
+
+To overcome such limitation, an alternative is to consider that only the catch declarations \(D_j\) are observed while \(Y_i\) are not (they are latent variables).
+
+Such way, we define some distribution \(\mathcal{L}_D\) as the distribution of the catch declarations:
+
+\[D_j | S_{\mathcal{P}_j},\mathcal{P}_j \sim \mathcal{L}_D( S_{\mathcal{P}_j},\xi,\sigma)\]
+
+with \(S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))\).
+As \(D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}\), it is possible to relate \(D_j\), \(Y_i\) and \(S_{\mathcal{P}_j}\) through the distribution \(\mathcal{L}_Y\) and \(\mathcal{L}_D\).
+Our approach is to match the moments of \(D_j\) (obtained from the moments of \(\mathcal{L}_Y\) i.e.~the probability distribution of \(Y\)) and \(\mathcal{L}_D\) and then to infer \(S(x)\) and \(Y_i\) from \(D_j\).
+This is what we call model-based reallocation.
+The models based on declaration \(D_j\) will be referred as the declaration-based models.
+
+In the following, \(Y\) and \(D\) will be assumed conditionnal over the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as \(Y_i\) or \(D_j\) instead of \(Y_i \vert x_i, S(x_i)\) and \(D_j | S_{\mathcal{P}_j},\mathcal{P}_j\).
+
+\textbf{Matching \(\mathcal{L}_Y\) and \(\mathcal{L}_D\)}
+
+First, as at the \(Y\) level, the declarations can be decomposed in 2 components: (1) the probability to obtain a zero value declaration and, if the declaration is positive, (2) the probability to obtain some declaration value \(d_j\).
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\tightlist
+\item
+  We derive the probability to obtain a zero-declaration \(P(D_j = 0)\) by simply multiplying the probability to obtain a zero-punctual observation \(P(Y_i=0 )\) to all fishing points \(x_i \in \mathcal{P}_j\).
+\end{enumerate}
+
+\begin{align*}
+P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
+\end{align*}
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  If the declaration is positive, we assume the probability to obtain catch \(d_j\) follows a continous distribution \(\mathcal{L}_{D_j|D_j>0}\) with mean component \(\mu_j\) and variance component \(\sigma_j\).
+These are expressed as a function of the two first moments of \(\mathcal{L}_Y\).
+In other words, from \(\mathcal{L}_Y\) we can derive \(E(D_j \vert D_j > 0)\) and \(Var(D_j \vert D_j > 0)\).
+If we consider \(\mathcal{L}_{D_j|D_j>0}\) is lognormal (with parameterization described in SM \ldots) then, \(\mu_j = E(D|D_j>0)\) and \(\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)\).
+\end{enumerate}
+
+The expressions of \(E(D_j \vert D_j > 0)\) and \(Var(D_j \vert D_j > 0)\) are given by:
+
+\[E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}\]
+\[Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2\]
+\[Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))\]
+
+All calculations to obtain these formulas are available in SM.
+By linking \(\mathcal{L}_Y\) and \(\mathcal{L}_D\), we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation.
+
+The inference procedure is realized through Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation (Kristensen, Nielsen, Berg, Skaug, \& Bell, 2016).
+
+\begin{figure}
+\centering
+\includegraphics{images/realloc.png}
+\caption{\label{fig:ParBiasSingle} Reallocation process.}
+\end{figure}
 
@@
 
-\hypertarget{single-square-simulations}{%
-\subsubsection{Single-square simulations}\label{single-square-simulations}}
+To evaluate what bring the performance of the new approach, we assess model performance through simulation-estimation.
 
-How many simulations? 100 or more?
+First, we limit the study to simaultions over a single square.
+Only commercial data feed the model and the expression of the latent field in both simulation and estimation is simplified (the variability only depends on one covariate, not on a spatial random term) in order to explore the base properties of the model.
 
-How do we simulate the covariate?
+Then, we extend the simulation-estimation study to several rectangles to get closer to a real case study configuration.
+We add a spatial random effect in the latent field (for both simulation and estimation) and, in addition to commercial data, we also simulate scientific data to explore the contribution of both datasets in inference.
 
-Still some lack of convergence.
+In these two steps, the covariate is modelled as a continous GRF that we suppose known at each point of the grid.
+The covariate effect is fixed to \(\beta_S=2\) and the intercept is also fixed to \(\mu=2\).
+The random effect (for the several rectangle simulation) is modelled as a Matérn GRF, but is supposed unkwnon over the area.
+Regarding commercial data, to fit the real data the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for one declaration.
+All parameterizations are detailed in the table \ldots{} .
 
-Number of pings \textbar{} nu:
-1\textbar10
-10\textbar100
-100\textbar1000
+Each time we compare several model configurations and evaluate these in regards to 2 metrics:
 
-Size within a fishing sequence: radius = 3
+\begin{itemize}
+\tightlist
+\item
+  the MSPE which quantifies the accuracy of the spatial predictions of the latent field over the spatial domain.
+\end{itemize}
 
-Number of fishing zones: 1\textbar3\textbar5
+\[MSPE=\frac{\sum_x (S(x) - \hat{S}(x) ) ^2}{n}\]
+
+\begin{itemize}
+\tightlist
+\item
+  the parameter of the species habitat relationship \(\beta_S\) compared with its estimate \(\hat \beta_S\).
+\end{itemize}
+
+\textbf{Single-square simulations}
+
+Two important variables may affect the accuracy of model outputs.
+
+First, the amount of commercial data; it is expected that when increasing the amount of data the accuracy of the predictions will be improved.
+In simulations, we progressively increase the number of fishing positions (10, 100 and 1000 fishing pings).
+As there are 10 pings per declaration, this means that the number of declarations increases respectively (1, 10 and 100 declarations).
+
+Furthermore, the samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle).
+Note that we make a distinction between fishing zones and statistical rectangles, fishing zones are included in a statistical rectangle and there can be several fishing zones within a statistical rectangles.
+In other words, a declaration can mix catches that have been realized in zones of high biomass and in zones of low biomass.
+For instance, a declaration can aggregate data from several fishing operations that occured the same day within the same statistical rectangle, but that were realized on 2 distinct types of fishing grounds (and then on two distinct habitats).
+Reallocating uniformly the declaration will strongly homogeneise the actual catch, will mix up information from to different locations and then may lead to a strong loss of accuracy in model outputs.
+It is expected that when the number of fishing zones within a declaration increases, the accuracy of the model outputs decreases.
+To assess this effect of such process, we simulated the pings of a fishing declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones.
+
+Concretely, for each fishing declaration the fishing pings are sampled in 2 steps: first the centroid of the zones are randomly sampled over the statistical rectangle (here the simulation domain) and then the fishing positions are sampled within the radius of the centroid of the zones.
+The zone size was chosen to be a square of side 7.
+This is a realistic order of magnitude if we consider each zone is one fishing operation and that for each fishing operation the distance of the operation will not exceed 30km.
+
+At each fishing position a catch is realized conditionnally on the value of the latent field at the related positions and follow the probability distribution described above (\(\mathcal{L}_y\)).
+
+We compare three simulation/estimation configurations:
+
+\begin{itemize}
+\item
+  a golden standard configuration: observations are known exactly and the standard model is fitted to the data.
+\item
+  a configuration corresponding to the actual situation (uniform reallocation): simulated catch are reallocated over fishing locations, the standard model is fitted to the data (i.e.~the one working at the level of \(Y_i^*\)).
+\item
+  a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations, the alternative model is fitted to the data (i.e.~the one working at the level of \(D_j\)).
+\end{itemize}
+
+In addition to the metrics we introduced to compare model configuration (\(MSPE\) and species-habitat parameter \(\beta_S\)), we also assess the simulated/estimated values for the intercept \(\mu\), the observation variance parameter \(\sigma^2\) and the zero-inflation parameter \(\xi\).
 
@@
 
-How do we simulate the covariate?
+In this case, we extend the study to more than one single rectangle, we simulate scientific data and include these in inference.
+Finally, in addition to the covariate, we also simulate/estimate a spatial random effect in the latent field.
+
+The study area is based on the case study; it includes the whole coast of the Bay of Biscay (Figure \ref{fig:MapSeveral}A).
+To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration).
+Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area.
+Similarly to the single-square simulations, the sampling of the commercial points of a declaration is realized in 3 steps.
+(1) The declarations is affected to one of the ICES rectangle.
+(2) A point is sampled within this statistical rectangle to sample the fishing zone in which the fishing points of the declaration will be sampled.
+(3) The fishing locations are randomly sampled within a fishing zone around this fishing point.
+The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km.
+Note that we do not explore the effect of exploring several zones within the same declaration as it is already studied in the single-square simulations.
+
+100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (similar to the Orhago survey - see the case study).
+Scientific observations are simulated following the observation equations of \(\mathcal{L}_Y\) with specific parameters for scientific data (see Table \ldots).
+
+We compare several model configurations:
+
+\begin{itemize}
+\item
+  to assess what bring our alternative approach, we compare models built on reallocated commercial catch data \(Y_i^*\) versus models built on catch declaration data \(D_j\).
+\item
+  to assess the information brought by each data source, we compare models built on scientific data only (scientific-based models), models built on commercial data only (commercial-based models) and models combining both data sources (integrated models).
+\end{itemize}
+
+In addition to the 2 metrics introduced at the begginning of the section (\(MSPE\) and species-habitat parameter \(\beta_S\)), we also compare the simulated/estimated values of the range parameter.
 
@@
 \subsection{Case-study: sole of the Bay of Biscay}\label{case-study-sole-of-the-bay-of-biscay}}
+
+To illustrate our method on a real case study, we applied the approach to the common sole of the Bay of Biscay.
+VMS-logbooks data were extracted for the bottom trawlers fleet (OTB).
+The methods to cross VMS-logbooks data and to filter the fleet is already extensively described in the previous papers (Alglave et al., 2022) and is not developped further here.
+Scientific data were extracted from the DATRAS database for the Orhago beam trawl survey (Gérard, 2003; ICES, 2018b).
+To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - ICES (2018a)).
+To illustrate the method, we compare the outputs (1) of the scientific-based model, (2) of the integrated model based on scientific data and reallocated commercial data (\(Y_i^*\) model) and (3) the integrated model based on catch declarations data (\(D_j\) model).
+
+The model based on catch declarations faced congvergence issues (some of the parameters were hardly estimated e.g.~the range parameter).
+To ease convergence, we integrated in the analysis onboard observer data for the same fleet.
+They are considered as exact commercial data of catches.
+Then, integrating these data allow to have direct information on \(Y_i\) and then to better estimate the observation equations parameters (i.e.~observation variance and zero-inflation parameter of commercial data).
+
+Furthermore, we adopt a step estimation procedure for the parameters that are hardly estimated (intercept \(\mu\), covariate effect \(\beta_S\), range and marginal variance).
+We first fit the model on reallocated data \(Y_i^*\) to have a first estimate of these parameters.
+Then, we fit the \(D_j\) model while fixing the initial values of the model with the previously fitted model and we fix the parameters hard to estimate with the estimates of the previously fitted model.
+Last, we let free all the parameters while setting the initial values with the previosuly fitted models.
 
@@
 \toprule
-Nb samples & Fishing sequence & Reallocation & Likelihood level & Convergence (\%)\\
+Fishing positions & Declarations & Reallocation & Likelihood level & Convergence (\%)\\
 \midrule
@@
 
+\begin{figure}
+\centering
+\includegraphics{images/Perf.metric_single_square.png}
+\caption{\label{fig:PerfMetricSingle} Performance metric for single-square simulations.
+columns: number of commercial pings.
+x-axis: number of zones visited within each declaration.
+\enquote{Reallocation:}, data are or are not reallocated in simulations.
+\enquote{Likelihood:}, the likelihood is computed on reallocated observations \(Y_i\) or on catch declarations \(D_j\).
+Gold: golden standard.
+Gold: golden standard.
+Red: uniform reallocation situation.
+Green: model-based reallocation (declaration-based model).
+Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the declaration-based model.}
+\end{figure}
+
+Reallocation has a major effect on predictions and estimates accuracy (Figure \ref{fig:PerfMetricSingle}).
+Reallocating data conduct to a 10 to 200 times decrease of predictions accuracy when working at punctual observations (\(Y_i\)) level (MSPE gold compared to red boxplots).
+Accuracy decreases as the number of visited zones within a declaration increases.
+It also leads to the loss of the species-habitat relationship as the \(\beta_S\) estimates are biased; furthermore, estimates tend towards 0 as the number of fishing zones within a declaration increases.
+Increasing the number of samples does not improve inference.
+Regarding other parameters estimates (\(\mu\),\(\xi\),\(\sigma\)), the zero-inflation parameter (\(\xi\)) is over-estimated (i.e.~when uniformly reallocating commercial data, the quantity of positive observations is under-estimated), the observation variance (\(\sigma\)) is underestimated (i.e.~the data is estimated to be less noisy than they actually are) and the intercept of the latent field (\(\mu\)) is slightly over-estimated (Figure \ref{fig:ParBiasSingle}).
+
+The declaration-based model (\(D_j\)) allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \ref{fig:PerfMetricSingle}).
+Still, the declaration-based model outputs are not as accurate as the ones of the golden-standard.
+Furthermore, the zero-inflation parameter is unbiased when the likelihood is built on catch declarations \(D_j\).
+Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \ref{fig:ParBiasSingle}).
+This alternative model have some convergence difficulties (Table 1) as 8\% of the model did not converged when sample size is medium (100 pings) and only 3\% did not when sample size is large (1000 pings).
+
+\begin{figure}
+\centering
+\includegraphics{images/par_plot_single_square.png}
+\caption{\label{fig:ParBiasSingle} Parameters relative bias for single-square simulations.
+\enquote{Reallocation:}, data are or are not reallocated in simulations.
+\enquote{Likelihood:}, the likelihood is computed on reallocated observations \(Y_i\) or on catch declarations \(D_j\).
+Gold: golden standard.
+Red: uniform reallocation situation.
+Green: model-based reallocation (declaration-based model).
+Only the simulations with 1000 fishing positionss are represented.
+Black line: zero value.
+Red line: parameter true value.}
+\end{figure}
+
+\hypertarget{multiple-square-analysis}{%
+\subsection{Multiple square analysis}\label{multiple-square-analysis}}
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-4}Percentage of convergence per simulation/model configuration at the level of several ICES rectangles}
+\centering
+\begin{tabular}[t]{ccc}
+\toprule
+Model & Likelihood level & Convergence (\%)\\
+\midrule
+Commercial model & Yi & 100.000\\
+Commercial model & Dj & 75.377\\
+Integrated model & Yi & 100.000\\
+Integrated model & Dj & 76.382\\
+Scientific model &  & 100.000\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+## Warning: Removed 7 rows containing non-finite values (stat_boxplot).
+\end{verbatim}
+
+The scientific-based model provides species-habitat relationship and range estimates that are unbiased.
+Whether the model is built on \(Y_i\) or on \(D_j\), the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models.
+This can be well evidenced when looking at Figure \ref{fig:MapSeveral}.
+Integrating scientific and commercial data allows to (1) capture the hotspot missed by commercial data through scientific data and (2) better capture the local correlation structures through the dense commercial data.
+
+Furthermore, consistently with single-square simulations, uniform reallocation conducts to a loss in both the predictions accuracy and the species-habitat relationship (Figure \ref{fig:PerfMetricSeveral}) compared to the model built on commercial declarations (\(D_j\)).
+
+Interestingly, uniform reallocation does not affect only the species-habitat relationship but also the spatial autocorrelation terms such as the range parameter.
+The model fitted on reallocated data (\(Y_i^*\)) provides biased range estimates while the declaration-based model provides unbiased estimates.
+This is a consequence of the loss of the species-habitat relationship: when uniformly reallocating declarations, part of the variability related to the covariate effect is captured by the random effect and then the range parameter captures both autocorrelation related to the actual random effect and to the covariate.
+
+Working the declaration-based model allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
+This is evidenced in Figure \ref{fig:MapSeveral} where the model based on reallocated catch \(Y_i^*\) provides smoothed maps and do not captures the relatively small scale patterns that are shaped by the covariate effect.
+On the other hand the declaration-based model (and the scientific-based model too) better capture and disentangle the covariate effect and the spatial random effect and then provide predictions that better fit to the small-scale patterns of the species distribution.
+However, this goes with some difficulty in convergence as only 75\% of the model built on catch declarations converge (Table 2).
+
+\begin{figure}
+\centering
+\includegraphics{images/Perf.metric_multiple_square.png}
+\caption{\label{fig:PerfMetricSeveral} Performance metric for several rectangles simulations.
+columns: commercial data coverage.
+x-axis: likelihood level.
+\(1^{st}\) row, red line: true value of \(\beta_S\).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{images/Map_multi_square.png}
+\caption{\label{fig:MapSeveral} Relative distribution of simulated/estimated biomass field.
+A: Simulated biomass field with scientific samples (red) and rectangles that have not been sampled by commercial data (transparent rectangles).
+B: simulated biomass field.
+C: biomass field from the scientific-based model.
+\(Y_i\): model fitted at the punctual observation level (D, E).
+\(D_j\): declaration-based model (F, G).}
+\end{figure}
+
+\hypertarget{real-case-study}{%
+\subsection{Real case study}\label{real-case-study}}
+
+\begin{figure}
+\centering
+\includegraphics{images/par_plot.png}
+\caption{\label{fig:CaseStudyPar} Parameters obtained with scientific-based model, the integrated model fitted on reallocated catch \(Y_i^*\) and the integrated model fitted on catch declarations \(D_j\).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{images/case_study_plot.png}
+\caption{\label{fig:CaseStudyMap} Maps obtained from (left) the scientific-based model, (center) the integrated model fitted on punctual observations \(Y_i^*\), (right) the integrated model fitted on catch declarations \(D_j\).}
+\end{figure}
+
+Passing from punctual observations \(Y_i^*\) to catch declarations \(D_j\) modifies the overall distribution patterns.
+Besides, some of the parameters estimates are modified consistently with simulations.
+The model fitted on \(Y_i^*\) tend to provide biased estimates and under-estimate the uncertainty related to other estimates compared with the scientific-based model and the declaration-based model.
+In particular, the substrate effect is recovered in both models (Figure \ref{fig:CaseStudyMap}).
+The zero-inflation parameter \(\xi\) is revised downwards (i.e.~there are actually more zero-values than in the reallocated data) while the observation variance of commercial data is revised upwards (i.e.~the commercial data are more noisy than expected when building likelihood on \(Y_i\)).
+In addition, related confidence intervals (CI) are also revised when fitting the model at the declaration level.
+Specifically, the CI from \(\beta_S\), the marginal variance, the range, \(\xi_{com}\), \(\sigma_{com}\) are very narrow in the model fitted to reallocated catch while in the model fitted to \(D_j\) confidence intervals are larger.
+This emphasizes that uncertainty is probably underestimated in the model fitted to reallocated catch compared with the declaration-based model.
+
+Now, when comparing the scientific-based model and the integrated model fitted on \(D_j\), some parameters are better estimated.
+For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on \(D_j\) substrate is significant and the confidence interval is smaller.
+Other parameters are more precisely estimated such as the marginal variance, the range, \(\xi_{sci}\) and \(\sigma_{sci}\).
+
+On the contrary, other parameters do not seems be well estimated in either the \(Y_i^*\) or the \(D_j\) models.
+For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on \(Y_i\) and revised downards when working on \(D_j\).
+This is consistent with simulations results, see Figure \ref{fig:ParBiasSingle}.
+
+Regarding the maps of the spatial predictions, fitting the model at the \(D_j\) level strongly modifies the model biomass field compared with the \(Y_i^*\) model.
+In particular, the effect of the covariate have a sharper effect on species distribution.
+Furthermore, the intensity of the hotspots are revised when fitting the model on \(D_j\).
+
+Finally, the \(D_j\) model fitted only on commercial data does not converge (while the one built on \(Y_i\) does) emphasizing the model build on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
+
 \hypertarget{discussion}{%
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-alglave_combining_2022}{}%
+Alglave, B., Rivot, E., Etienne, M.-P., Woillez, M., Thorson, J.
+T., \& Vermard, Y.
+(2022).
+Combining scientific survey and commercial catch data to map fish distribution.
+\emph{ICES Journal of Marine Science}, fsac032.
+\url{https://doi.org/10.1093/icesjms/fsac032}
+
+\leavevmode\hypertarget{ref-biais_gerard_orhago_2003}{}%
+Gérard, B. (2003). ORHAGO. \url{https://doi.org/10.18142/23}
+
+\leavevmode\hypertarget{ref-ices_report_2018-1}{}%
+ICES. (2018a). \emph{Report of the Working Group for the Bay of Biscay and the Iberian Waters Ecoregion (WGBIE)} (p. 642). Copenhagen, Denmark.
+
+\leavevmode\hypertarget{ref-ices_report_2018}{}%
+ICES. (2018b). \emph{Report of the Working Group on Beam Trawl Surveys (WGBEAM)} (p. 121). Galway, Ireland.
+
+\leavevmode\hypertarget{ref-kristensen_tmb_2016}{}%
+Kristensen, K., Nielsen, A., Berg, C.
+W., Skaug, H., \& Bell, B.
+M.
+(2016).
+TMB: Automatic Differentiation and Laplace Approximation.
+\emph{Journal of Statistical Software}, \emph{70}(1), 1--21.
+\url{https://doi.org/10.18637/jss.v070.i05}
+
+\leavevmode\hypertarget{ref-thorson_three_2018}{}%
+Thorson, J.
+T.
+(2018).
+Three problems with the conventional delta-model for biomass sampling data, and a computationally efficient alternative.
+\emph{Canadian Journal of Fisheries and Aquatic Sciences}, \emph{75}(9), 1369--1382.
 

### 2022-03-29 Version_M2_plusieurszones/pres/GDREcostat_April2022/GDREcostat_2022.Rmd
paper_3

---
title: "Infering species distribution from aggregated spatial data"
subtitle : "Going from coarse landings data to fine scale species distribution"
author : "Baptiste Alglave,  Marie-Pierre Etienne, Youen Vermard,  Mathieu Woillez, Etienne Rivot"

fontsize: 10pt
bibliography : ["ref.bib"]
csl : ices-journal-of-marine-science.csl

output:
  beamer_presentation:
    theme: "Darmstadt"
    colortheme: "default"
    slide_level: 3
    # template: mytemplate.tex
    includes:
      in_header: header-simple.tex
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(rnaturalearth)
library(stringr)
library(sf)
library(raster)
library(maps)
library(mapdata)
library(ggplot2)
library(ggspatial)
library(cowplot)
library(latex2exp)
library(ggpubr)

col_scale <- function(palette = "Spectral", name = "", limits = NULL) {
  scale_colour_distiller(palette = palette,   # spectral colour scale
                         guide = "colourbar", # continuous colour bar
                         name = name,
                         limits = limits)
}

run_codes <- F

grid_projection <- "+proj=longlat +datum=WGS84"

mapBase <- map("worldHires", fill = T, plot = F)
mapBase <- st_as_sf(mapBase) %>% filter(ID %in% c("France","Spain"))

```

# Context

### Spatial data in ecology

<!-- Standardized and high quality data, but low amount of samples -->
<!-- more data, but no standardize sampling plan (e.g. problem related to preferential sampling) -->
<!-- legal and massive data, but low quality, declared at the level of some administrative unit -->

\begin{columns}
\begin{column}{0.1\textwidth}
\end{column}
\begin{column}{0.3\textwidth}
\center {\bf Survey data}
\end{column}
\begin{column}{0.3\textwidth}
\center {\bf Citizen science data}
\end{column}
\begin{column}{0.3\textwidth}
\center {\bf Declaration data}
\end{column}
\end{columns}

\vspace{\baselineskip}

\begin{columns}
\begin{column}{0.1\textwidth}
\center \Huge \textbf{\textcolor{BaptisteBlue}{\ding{58}}}
\normalsize
\end{column}
\begin{column}{0.3\textwidth}
\normalsize
Standardized sampling plan \\
High quality data
\end{column}
\begin{column}{0.3\textwidth}
Access to more data \\
Exact locations usually available
\end{column}
\begin{column}{0.3\textwidth}
Mandatory declaration \\
Massive data
\end{column}
\end{columns}

\vspace{\baselineskip}

\begin{columns}
\begin{column}{0.1\textwidth}
\center \Huge \fontsize{45}{15}{\textbf{\textcolor{BaptisteBlue}{-}}}
\normalsize
\end{column}
\begin{column}{0.3\textwidth}
\normalsize
Low amount of samples
\end{column}
\begin{column}{0.3\textwidth}
Opportunistic (or even preferential) sampling
\end{column}
\begin{column}{0.3\textwidth}
Aggregated at the scale of administrative units
\end{column}
\end{columns}

\begin{columns}
\begin{column}{0.1\textwidth}
\center \footnotesize {\bf Examples}
\end{column}
\begin{column}{0.3\textwidth}
\center
\includegraphics[width=2cm]{images/EVHOE_stations.jpg}
\end{column}
\begin{column}{0.3\textwidth}
\center
\includegraphics[width=3cm]{images/bird_apps.jpg}
\end{column}
\begin{column}{0.3\textwidth}
\center
\includegraphics[width=2.5cm]{images/harvest_picture.PNG}
\end{column}
\end{columns}

\begin{columns}
\begin{column}{0.1\textwidth}
\end{column}
\begin{column}{0.3\textwidth}
\center \scriptsize
EVHOE data, Bay of Biscay \\ (marine ecology)
\end{column}
\begin{column}{0.3\textwidth}
\center \scriptsize
Ebird application \\ (ornithology)
\end{column}
\begin{column}{0.3\textwidth}
\center \scriptsize
Harvest data, Wisconsin \\ (hunting)
\end{column}
\end{columns}

---
\btVFill
\begin{center} \Large
\textbf{How to integrate all these datasources?}
\vspace{\baselineskip}
\large
(especially when they do not have the same spatial resolution)
\normalsize

\btVFill

\end{center}
\onslide<2>{
\footnotesize
\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Some specific proposition in the context of fishery science: \\
\tiny
Alglave Baptiste, Rivot Etienne, Etienne Marie-Pierre, Woillez Mathieu, Thorson James T, Vermard Youen (2022). \textbf{Combining scientific survey and commercial catch data to map fish distribution.} ICES Journal of Marine Science IN PRESS. \url{https://doi.org/10.1093/icesjms/fsac032}
\vspace{\baselineskip}
}

# Material and method

### Commercial catch declarations data in fishery science

```{r,echo=F,fig.align='center',out.width="120%"}

if(run_codes){

  load("map/ICES_rect.RData")
  load("res/points_sf.RData")

  #-------------------
  ## Logbook data plot
  #-------------------
  logbooks_df <- points_sf %>%
    # filter(str_detect(SI_DATE,"11/2018")) %>%
    group_by(ICESNAME_1) %>%
    dplyr::summarise(Catch = sum(LE_KG_Solea_solea)/1000,
                     Effort = sum(INTV)/60)

  logbooks_sf <- st_join(ICES_rect,logbooks_df) %>%
    filter(!is.na(Effort))

  Logbooks_data <- ggplot()+
    geom_sf(data = ICES_rect,fill="skyblue",alpha=0.25)+
    geom_sf(data=logbooks_sf,aes(fill=Catch))+
    geom_sf(data = mapBase)+
    scale_fill_distiller(palette = "Spectral")+
    theme(panel.background = element_rect(fill = "skyblue"),
          plot.title = element_text(hjust=0.5),
          legend.position = "none")+
    geom_sf(col="black")+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    annotation_scale(location = "bl", line_width = .5)+
    annotation_north_arrow(location = "tl", height = unit(0.7, "cm"),
                           width = unit(0.7, "cm"))

  ggsave(Logbooks_data,filename = "images/logbooks_data.png",width = 5,height = 5)

  #---------------
  ## VMS data plot
  #---------------
  ## Aggregate/discretize data
  vms_df <- points_sf %>%
    mutate(long = sf::st_coordinates(.)[,1],
           lati = sf::st_coordinates(.)[,2]) %>%
    as.data.frame %>%
    dplyr::select(-geometry) %>%
    mutate(long = round(long,digits = 2),
           lati = round(lati,digits = 2)) %>%
    group_by(long,lati) %>%
    dplyr::summarise(Effort = sum(INTV)/60)

  vms_sf <- st_as_sf(vms_df,coords = c("long","lati"),crs=grid_projection)

  ## Build polygon
  # Coordinates Bay of Biscay
  grid_xmin <- -6
  grid_xmax <- 0
  grid_ymin <- 42
  grid_ymax <- 48
  grid_limit <- extent(c(grid_xmin,grid_xmax,grid_ymin,grid_ymax))

  # Build grid
  grid <- raster(grid_limit)
  res(grid) <- 0.025
  crs(grid) <- "+proj=longlat +datum=WGS84"
  gridpolygon <- rasterToPolygons(grid)
  gridpolygon$layer <- c(1:length(gridpolygon$layer))
  gridpolygon_sf <- st_as_sf(gridpolygon)

  vms_sf_2 <- st_join(gridpolygon_sf,points_sf)
  vms_sf_3 <- vms_sf_2 %>%
    group_by(layer) %>%
    dplyr::summarise(Effort = sum(INTV)/60)

  vms_sf_3 <- vms_sf_3 %>%
    filter(!is.na(Effort))

  # save(data=vms_sf_3,file="res/vms_sf_3.RData")

  vms_data <- ggplot()+
    geom_sf(data = ICES_rect,fill="skyblue",alpha=0.25)+
    geom_sf(data=vms_sf_3,aes(col=Effort,fill=Effort))+
    geom_sf(data = ICES_rect,alpha=0)+
    geom_sf(data = mapBase)+
    scale_color_distiller(direction = 1,palette = "OrRd")+ # palette = "OrRd"
    scale_fill_distiller(direction = 1,palette = "OrRd")+ # ,palette = "Spectral"
    theme(panel.background = element_rect(fill = "skyblue"),
          plot.title = element_text(hjust=0.5),
          legend.position = "none")+
    geom_sf(col="black")+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)

  ggsave(filename = "images/vms_data.png",width = 5,height = 5)

}

```

\begin{columns}
\begin{column}{0.1\textwidth}
\end{column}
\begin{column}{0.45\textwidth}
\center {\bf Logbook landings} \\
\scriptsize (Sole, trawlers OTB-DEF, 2018) 
\end{column}
\begin{column}{0.45\textwidth}
\center {\bf Fishing locations (VMS)} \\
\scriptsize (Trawlers OTB-DEF, 2018)
\end{column}
\end{columns}

\vspace{\baselineskip}

\begin{columns}
\begin{column}{0.1\textwidth}
\end{column}
\begin{column}{0.45\textwidth}
\includegraphics[width=5cm]{images/logbooks_data.png}
\end{column}
\begin{column}{0.45\textwidth}
\includegraphics[width=5cm]{images/vms_data.png}
\end{column}
\end{columns}

\begin{columns}
\begin{column}{0.1\textwidth}
\center \tiny {\bf Resolution}
\end{column}
\begin{column}{0.45\textwidth}
\center \tiny Catch are daily declared at the resolution of ICES rectangles
\end{column}
\begin{column}{0.45\textwidth}
\center \tiny VMS pings are vessels GPS locations emitted each hour
\end{column}
\end{columns}
\tiny
\onslide<2>{
\begin{columns}
\begin{column}{0.1\textwidth}
\end{column}
\begin{column}{0.45\textwidth}
\end{column}
\begin{column}{0.45\textwidth}
\center \scriptsize \textbf{\textcolor{BaptisteBlue}{\ding{224}}  Refine landings spatial resolution}
\end{column}
\end{columns}
}

### Modelling change of support

\center

\includegraphics[width=11cm]{images/realloc.png}

\onslide<2-3>{

\begin{table}
  \centering \scriptsize
  \begin{tabular}{ M{2cm} M{4cm} M{4cm} }
  \textcolor{BaptisteRed}{\textbf{Actual situation}} & 
  $Y_i | \textcolor{BaptisteBlue}{S}(x_i), x_i \sim \mathcal{L}_Y(\textcolor{BaptisteBlue}{S}(x_i),\xi,\sigma)$          &
  $Y_i=\textcolor{BaptisteRed}{Y_i^*}$ \\
  \end{tabular}
\end{table}

}

\onslide<3>{

\begin{table}
  \centering \scriptsize
  \begin{tabular}{ M{2cm} M{2cm} M{4cm} M{2cm} }
  \textcolor{BaptisteLightGreen}{\textbf{Alternative solution}} & 
  $\textcolor{BaptisteLightGreen}{D_j}=\sum_{\textcolor{BaptisteLightGreen}{i \in \mathcal{P}_j}}{Y_{i}}$          &
  $\textcolor{BaptisteLightGreen}{D_j} | \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\textcolor{BaptisteLightGreen}{\mathcal{P}_j} \sim \mathcal{L}_D( \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\xi,\sigma)$
  & \textcolor{BaptisteBlue}{\ding{224}}  Match $\mathcal{L}_D$ and $\mathcal{L}_Y$ moments \\
  \end{tabular}
\end{table}

}

### Simulation-estimation and case study

```{r,echo=F}

if(run_codes){
  ## Load simulated data
  load("res/simu_df.RData")
  load("res/list_input.RData")

  simu_df_2 <- simu_df %>%
    mutate(Ab = sum(S_x))

  loc_x <- list_input$loc_x
  sci.data_df <- data.frame(cell = list_input$index_sci_i,
                            y_sci = list_input$y_sci_i) %>%
    inner_join(loc_x[,c("x","y","cell","ICESNAME")])

  com.data_df <- data.frame(cell = list_input$index_i,
                            y_com = list_input$y_i2) %>%
    inner_join(loc_x[,c("x","y","cell","ICESNAME")]) %>%
    group_by(x,y,cell,ICESNAME) %>%
    dplyr::summarise(n=n())

  ## Grid data
  grid <- raster(grid_limit)
  res(grid) <- 0.05
  crs(grid) <- "+proj=longlat +datum=WGS84"
  gridpolygon <- rasterToPolygons(grid)
  gridpolygon$layer <- c(1:length(gridpolygon$layer))
  gridpolygon_sf <- st_as_sf(gridpolygon)

  simu_sf <- st_as_sf(simu_df_2,coords = c("long","lati"),crs=grid_projection)

  simu_sf_2 <- st_join(gridpolygon_sf,simu_sf) %>%
    filter(!is.na(S_x))

  ## Plot data
  data_plot <- ggplot()+
    geom_sf(data=simu_sf_2,aes(fill=log(S_x)),col=NA,alpha=0.5)+
    geom_sf(data=simu_sf_2[which(simu_sf_2$ICESNAME %in% com.data_df$ICESNAME),],
            aes(fill=log(S_x),col=log(S_x)))+
    scale_fill_distiller(palette = "Spectral")+
    scale_color_distiller(palette = "Spectral")+
    geom_point(data=sci.data_df,aes(x=x,y=y),fill="#619CFF",shape=21,size=1)+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5),
          legend.title = element_blank(),
          legend.position = "none")+
    geom_sf(data=ICES_rect,alpha=0)+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    xlab("")+ylab("")

  ggsave(filename = "images/data_plot.png",width = 5,height = 5)

}

```

\begin{columns}
\begin{column}{0.5\textwidth}

\center
\textbf{Simulation/estimation}

\end{column}
\begin{column}{0.5\textwidth}

\end{column}
\end{columns}

\tiny \vspace{\baselineskip}

\begin{columns}
\begin{column}{0.5\textwidth}

\begin{columns}
\begin{column}{0.45\textwidth}

\includegraphics[width=3cm]{images/data_plot.png}

\end{column}
\begin{column}{0.55\textwidth}
\scriptsize
\textbf{Simulation}
\tiny
\begin{itemize}
\item \textcolor{BaptisteBlue}{\textbf{Latent field}} (\textcolor{BaptisteOrange}{covariate} + \textcolor{BaptisteBrown}{spatial random effect})
\item \textbf{Commercial data} (3000 samples over 2/3 of the area)
\item \textcolor{BaptisteLightGreen}{\textbf{Reallocation process}} (10 locations per declaration)
\item \textcolor{SurveyBlue}{\textbf{Scientific data}} (100 samples over the whole the area)
\end{itemize}

\end{column}
\end{columns}

\scriptsize
\vspace{\baselineskip}

\onslide<2-4>{

\textbf{Estimation} \\
Comparison of 3 model configurations: \\
\tiny \vspace{\baselineskip} \scriptsize
1/ Model fitted to \textcolor{SurveyBlue}{scientific data} (reference data) \\
\tiny \vspace{\baselineskip} \scriptsize
2/ Integrated model with \textcolor{BaptisteRed}{commercial likelihood built on $Y_i^*$} \\
\tiny \vspace{\baselineskip} \scriptsize
3/ Integrated model with \textcolor{BaptisteLightGreen}{commercial likelihood built on $D_j$} \\
\tiny \vspace{\baselineskip}

Estimation realized through TMB (Template Model Builder) \\
100 runs of simulation-estimation

}

\end{column}
\begin{column}{0.5\textwidth}

\onslide<3-4>{

\scriptsize
\textbf{Model evaluation} \\
1/ Mean square prediction error \\
$$MSPE=\frac{\sum_x (\textcolor{BaptisteBlue}{S(x)} - \textcolor{BaptisteBlue}{\hat{S}(x)} ) ^2}{n}$$

\vspace{\baselineskip}

2/ Covariate effect (or species-habitat relationship): \\
$$\textcolor{BaptisteOrange}{\beta_S} = 2 \, \, \, \text{versus} \, \, \, \textcolor{BaptisteOrange}{\hat{\beta}_S}$$

}

\onslide<4>{

\vspace{\baselineskip}

\normalsize
\begin{center}
\textbf{Case study:} \scriptsize Sole in the Bay of Biscay \\
\end{center}

\begin{columns}
\begin{column}{0.5\textwidth}
\center
\includegraphics[width=2cm]{images/solea_solea.png}
\end{column}
\begin{column}{0.5\textwidth}
\tiny \textcolor{SurveyBlue}{\textbf{Survey data:}} Orhago\\
\textbf{Commercial data:} OTB-DEF traulers (to ease convergence onboard observer data were integrated in fitting)\\
\textbf{Fitted models:} same as simulation\\
\textcolor{BaptisteOrange}{\textbf{Covariate:}} substrate

\end{column}
\end{columns}

}

\end{column}
\end{columns}

# Results

### Simulation-estimation

```{r,echo=F}

if(run_codes){

  ######################
  ## Parameter estimates
  ######################
  load("res/Results_full_multi_square.RData")

  Results_plot <- Results_2 %>%
    filter(converge == 0) %>%
    filter(simu_type == "Unsampled Rectangles")

  Results_plot[,"RelBias_N"]=(Results_plot[,"N_est"]-Results_plot[,"N_true"])/Results_plot[,"N_true"]
  Results_plot[,"RelBias_N.2"]=(Results_plot[,"N_est.2"]-Results_plot[,"N_true.2"])/Results_plot[,"N_true.2"]
  Results_plot[,"RelBias_beta"]=(Results_plot[,"beta1_est"]-Results_plot[,"beta1_true"])/Results_plot[,"beta1_true"]

  Results_plot$obs <- NA
  Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
  Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi"
  Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific model"
  Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi","Dj","Scientific model"))

  Results_plot$Model[which(Results_plot$Model == "Scientific model")] <- "Scientific"
  Results_plot$Model[which(Results_plot$Model == "Commercial model")] <- "Commercial"
  Results_plot$Model[which(Results_plot$Model == "Integrated model")] <- "Integrated"

  Results_plot <- Results_plot %>%
    filter(Model != "Commercial")

  # Results_plot_fullArea <- Results_plot
  # Results_plot_UnsampArea <- Results_plot

  beta_plot <- ggplot()+
    geom_boxplot(data = Results_plot,
                 aes(x = Model,
                     y = beta1_est,
                     fill = obs))+
    geom_hline(yintercept = 0,linetype="dashed")+
    theme_bw()+
    theme(legend.title = element_blank(),
          legend.position = "none",
          aspect.ratio = 1)+
    # facet_wrap(.~simu_type)+
    ylab(TeX("$\\beta_S$")) +
    xlab("")+
    geom_hline(yintercept = 2,col="red",linetype="dashed")+
    scale_color_manual(breaks = c("Scientific model",
                                  "Yi",
                                  "Dj"),
                       values=c("#619CFF","#F8766D","#00BA38"))

  MSPE_S_plot <- ggplot()+
    geom_boxplot(data = Results_plot,
                 aes(x = Model,
                     y = mspe,
                     fill = obs))+
    theme_bw()+
    theme(legend.title = element_blank(),
          legend.position = "none",
          aspect.ratio = 1)+
    # facet_wrap(.~simu_type)+
    ylab("MSPE")+xlab("")+
    scale_color_manual(breaks = c("Scientific model",
                                  "Yi",
                                  "Dj"),
                       values=c("#619CFF","#F8766D","#00BA38"))+
    ylim(0,NA)

  legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="bottom",legend.title = element_blank())))

  pl_plot <- plot_grid(MSPE_S_plot,
                       beta_plot,
                       ncol = 1,
                       legend,
                       align="hv",
                       rel_heights = c(1,1,0.15))

  ggsave(filename = "images/res_simu_plot.png",width = 4,height = 7.5)

  #######
  ## Maps
  #######
  simu_plot <- ggplot(simu_df)+
    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
    scale_color_distiller(palette = "Spectral")+
    ggtitle("Simulated latent field")+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5,face = "bold"),
          plot.subtitle = element_text(hjust = 0.5),
          legend.position = "none")+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    xlab("")+ylab("")

  load("res/fit_IM_res_sci.RData")
  sci_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
  sci_plot <- ggplot(sci_df)+
    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
    scale_color_distiller(palette = "Spectral")+
    ggtitle("Scientific model")+
    theme_bw()+
    theme(plot.title = element_text(colour = "#619CFF",hjust = 0.5,face = "bold"),
          plot.subtitle = element_text(hjust = 0.5),
          legend.position = "none")+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    xlab("")+ylab("")

  load("res/fit_IM_res_int_Yi.RData")
  int_Yi_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
  int_Yi_plot <- ggplot(int_Yi_df)+
    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
    scale_color_distiller(palette = "Spectral")+
    ggtitle("Commercial likelihood on Yi")+
    theme_bw()+
    theme(plot.title = element_text(colour = "#F8766D",hjust = 0.5,face = "bold"),
          plot.subtitle = element_text(hjust = 0.5),
          legend.position = "none")+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    xlab("")+ylab("")

  load("res/fit_IM_res_int_Dj.RData")
  int_Dj_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
  int_Dj_plot <- ggplot(int_Dj_df)+
    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
    scale_color_distiller(palette = "Spectral")+
    ggtitle("Commercial likelihood on Dj")+
    theme_bw()+
    theme(plot.title = element_text(colour = "#00BA38",hjust = 0.5,face = "bold"),
          plot.subtitle = element_text(hjust = 0.5),
          legend.position = "none")+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    xlab("")+ylab("")

  map_simu <- plot_grid(simu_plot,sci_plot,int_Yi_plot,int_Dj_plot)

  ggsave(filename = "images/map_simu_plot.png",width = 6,height = 7.5)

}

```

\center

\begin{columns}
\begin{column}{0.4\textwidth}

\center
\includegraphics[width=4cm]{images/res_simu_plot.png}

\end{column}

\begin{column}{0.6\textwidth}
\center

\includegraphics[width=5.5cm]{images/map_simu_plot.png}
\vspace{\baselineskip}

\end{column}

\end{columns}

### Case study: sole in the Bay of Biscay

```{r,echo=F}

if(run_codes){

  ## Load data
  load("res/sci_df.RData")

  load("res/realloc_rest_df.RData")

  load("res/no.realloc_df.RData")

  load("res/est_par_df_full_3.RData")

  ## Plot maps
  no.realloc_plot <- ggplot(no.realloc_df)+
    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
    ggtitle("Commercial likelihood on Yi")+
    theme_bw()+
    theme(legend.position = "none",
          plot.title = element_text(colour = "#F8766D",hjust = 0.5,face = "bold"))+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    xlab("")+ylab("")

  realloc_rest_plot <- ggplot(realloc_rest_df)+
    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
    ggtitle("Commercial likelihood on Dj")+
    theme_bw()+
    theme(legend.position = "none",
          plot.title = element_text(colour = "#00BA38",hjust = 0.5,face = "bold"))+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    xlab("")+ylab("")

  sci_plot <- ggplot(sci_df)+
    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
    ggtitle("Scientific model")+
    theme_bw()+
    theme(legend.position = "none",
          plot.title = element_text(colour = "#619CFF",hjust = 0.5,face = "bold"))+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    xlab("")+ylab("")

  # plot(log(no.realloc_df$S_x),log(sci_df$S_x))
  # cor(log(no.realloc_df$S_x),log(sci_df$S_x),method = "spearman")
  # plot(log(realloc_rest_df$S_x),log(sci_df$S_x))
  # cor(log(realloc_rest_df$S_x),log(sci_df$S_x),method = "spearman")

  case_study_plot <- plot_grid(NULL,sci_plot,no.realloc_plot,realloc_rest_plot,align="hv")

  ggsave("images/case_study_plot.png",width = 6, height = 7.5)

  ## Parameter estimates
  est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi","Scientific model"))
  est_par_df_full_3 <- est_par_df_full_3 %>%
    filter(par_names != "substr_Sand_Coarse_substrate")
  par_plot <- ggplot(est_par_df_full_3, aes(y=par_val, x=par_names))+
    geom_point(
      aes(color = lkl),
      position = position_dodge(0.5),
      size=2
    )+
    geom_errorbar(aes(ymin = CI.inf, ymax = CI.sup, color = lkl),
                  position = position_dodge(0.5),width=0.4
    )+
    scale_color_manual(breaks = c("Scientific model",
                                  "Yi",
                                  "Dj"),
                       values=c("#619CFF","#F8766D","#00BA38"))+
    geom_hline(yintercept=0, linetype="dashed", color = "red",alpha=0.4)+
    xlab("")+ylab("")+
    theme_bw()+
    ggtitle("Parameters estimates")+
    theme(plot.title = element_text(hjust = 0.5,face = "bold"),
          legend.title = element_blank(),
          legend.position = "bottom",
          aspect.ratio = 1)+
    coord_flip()+
    scale_x_discrete(labels=c('k_com'=parse(text = TeX('$k_{com}$')),
                              'Sigma_com'=parse(text = TeX('$\\sigma_{com}$')),
                              'q1_com'=parse(text = TeX('$\\xi_{com}$')),
                              'Sigma_sci'=parse(text = TeX('$\\sigma_{sci}$')),
                              'q1_sci'=parse(text = TeX('$\\xi_{sci}$')),
                              'Range'="Range",
                              'MargSD'="M. var.",
                              'substr_Mud_sediment'=parse(text = TeX('$\\beta_{S}$')),
                              'intercept'=parse(text = TeX('$\\mu$'))))

  ggsave("images/par_plot.png",width = 6,height = 7.5)

}

```

\center

\begin{columns}
\begin{column}{0.4\textwidth}

\begin{center}
\includegraphics[width=5cm]{images/par_plot.png}
\end{center}
\tiny *n.b.: punctual commercial data (i.e. observer data) needed to be integrated in inference in order to make the model converge and some parmaters needed to be step estimated (e.g. range)

\end{column}

\begin{column}{0.6\textwidth}
\center
\includegraphics[width=5.5cm]{images/case_study_plot.png}
\vspace{\baselineskip}

\end{column}

\end{columns}

# Discussion

### Discussion

\begin{itemize}

\item Framework that integrates \textbf{catch declarations data} (rough resolution of catch) and \textbf{scientific/onboard observer data} (exact locations of catch) \\

\onslide<2-4>{

\textcolor{BaptisteBlue}{\ding{224}} \textbf{Still question:} how to ease convergence ? \\

}

\vspace{\baselineskip}

\onslide<3-4>{

\item \textbf{ Change of support issues } \\

\textbf{\textcolor{BaptisteBlue}{\ding{224}}} not so frequent in spatial ecology, \\

\textbf{\textcolor{BaptisteBlue}{\ding{224}}} but increasing amount of data available that may face such issues \\

}

\vspace{\baselineskip}

\onslide<4>{

\item \textbf{Is it a generic framework ?} \\

\textbf{\textcolor{BaptisteBlue}{\ding{224}}} the overall approach is, \tiny \\ (i.e. modelling observed aggregagetd catches as a convolution of latent exact catches) \\ \vspace{\baselineskip} \normalsize

\textbf{\textcolor{BaptisteBlue}{\ding{224}}} but need to adapt the observation model to the data \tiny \\ (here zeroinflated continuous data)   \\ \vspace{\baselineskip} \normalsize

\textbf{\textcolor{BaptisteBlue}{\ding{224}}} and make the hypothesis that fishing locations ($\textcolor{BaptisteLightGreen}{\mathcal{P}_j}$) are known \scriptsize \\ (in reality $\textcolor{BaptisteLightGreen}{\mathcal{P}_j}$ is a trajectory and the set of fishing points are a realization of this trajectory) \\ \vspace{\baselineskip} \normalsize

}

\end{itemize}

#

\centering \Large \textcolor{BaptisteGrey}{\textbf{Thank you for your attention!}}

\tiny \vspace{\baselineskip}

\begin{figure}
    \begin{tabular}{c c}
    \includegraphics[width=4.5cm]{images/peche_afrique_artisanale.jpg} &
    \includegraphics[width=4.5cm]{images/peche_artisanale.jpg} \\
    \includegraphics[width=4.5cm]{images/peche_hauturiere.jpg} &
    \includegraphics[width=4.5cm]{images/thon_rouge.jpg}
    \end{tabular}
\end{figure}

### 2022-03-29 Version_M2_plusieurszones/pres/GDREcostat_April2022/test.Rmd
paper_3

--- 2022-03-29 Version_M2_plusieurszones/pres/GDREcostat_April2022/GDREcostat_2022.Rmd
@@
 ---
-title: "Infering species distribution from aggregated spatial data"
-subtitle : "Going from coarse landings data to fine scale species distribution"
-author : "Baptiste Alglave,  Marie-Pierre Etienne, Youen Vermard,  Mathieu Woillez, Etienne Rivot"
+title: Peut-on identifier des zones fonctionnelles à l’aide des données de débarquements commerciaux ?
+subtitle : Application à la sole du golfe de Gascogne
+author : Baptiste Alglave, Youen Vermard, Marie-Pierre Etienne,  Mathieu Woillez, Etienne Rivot
 
@@
 library(sf)
-library(raster)
 library(maps)
-library(mapdata)
 library(ggplot2)
@@
 library(cowplot)
-library(latex2exp)
-library(ggpubr)
 
@@
 
-run_codes <- F
-
-grid_projection <- "+proj=longlat +datum=WGS84"
-
-mapBase <- map("worldHires", fill = T, plot = F)
-mapBase <- st_as_sf(mapBase) %>% filter(ID %in% c("France","Spain"))
-
 ```
 
-# Context
-
-### Spatial data in ecology
-
-<!-- Standardized and high quality data, but low amount of samples -->
-<!-- more data, but no standardize sampling plan (e.g. problem related to preferential sampling) -->
-<!-- legal and massive data, but low quality, declared at the level of some administrative unit -->
-
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\end{column}
-\begin{column}{0.3\textwidth}
-\center {\bf Survey data}
-\end{column}
-\begin{column}{0.3\textwidth}
-\center {\bf Citizen science data}
-\end{column}
-\begin{column}{0.3\textwidth}
-\center {\bf Declaration data}
-\end{column}
-\end{columns}
-
-\vspace{\baselineskip}
-
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\center \Huge \textbf{\textcolor{BaptisteBlue}{\ding{58}}}
-\normalsize
-\end{column}
-\begin{column}{0.3\textwidth}
-\normalsize
-Standardized sampling plan \\
-High quality data
-\end{column}
-\begin{column}{0.3\textwidth}
-Access to more data \\
-Exact locations usually available
-\end{column}
-\begin{column}{0.3\textwidth}
-Mandatory declaration \\
-Massive data
-\end{column}
-\end{columns}
-
-\vspace{\baselineskip}
-
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\center \Huge \fontsize{45}{15}{\textbf{\textcolor{BaptisteBlue}{-}}}
-\normalsize
-\end{column}
-\begin{column}{0.3\textwidth}
-\normalsize
-Low amount of samples
-\end{column}
-\begin{column}{0.3\textwidth}
-Opportunistic (or even preferential) sampling
-\end{column}
-\begin{column}{0.3\textwidth}
-Aggregated at the scale of administrative units
-\end{column}
-\end{columns}
-
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\center \footnotesize {\bf Examples}
-\end{column}
-\begin{column}{0.3\textwidth}
-\center
-\includegraphics[width=2cm]{images/EVHOE_stations.jpg}
-\end{column}
-\begin{column}{0.3\textwidth}
-\center
-\includegraphics[width=3cm]{images/bird_apps.jpg}
-\end{column}
-\begin{column}{0.3\textwidth}
-\center
-\includegraphics[width=2.5cm]{images/harvest_picture.PNG}
-\end{column}
-\end{columns}
-
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\end{column}
-\begin{column}{0.3\textwidth}
-\center \scriptsize
-EVHOE data, Bay of Biscay \\ (marine ecology)
-\end{column}
-\begin{column}{0.3\textwidth}
-\center \scriptsize
-Ebird application \\ (ornithology)
-\end{column}
-\begin{column}{0.3\textwidth}
-\center \scriptsize
-Harvest data, Wisconsin \\ (hunting)
-\end{column}
-\end{columns}
-
----
-\btVFill
-\begin{center} \Large
-\textbf{How to integrate all these datasources?}
-\vspace{\baselineskip}
-\large
-(especially when they do not have the same spatial resolution)
-\normalsize
-
-\btVFill
-
-\end{center}
-\onslide<2>{
-\footnotesize
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Some specific proposition in the context of fishery science: \\
-\tiny
-Alglave Baptiste, Rivot Etienne, Etienne Marie-Pierre, Woillez Mathieu, Thorson James T, Vermard Youen (2022).
-\textbf{Combining scientific survey and commercial catch data to map fish distribution.} ICES Journal of Marine Science IN PRESS.
-\url{https://doi.org/10.1093/icesjms/fsac032}
-\vspace{\baselineskip}
-}
-
-# Material and method
-
-### Commercial catch declarations data in fishery science
-
-```{r,echo=F,fig.align='center',out.width="120%"}
-
-if(run_codes){
-
-  load("map/ICES_rect.RData")
-  load("res/points_sf.RData")
-
-  #-------------------
-  ## Logbook data plot
-  #-------------------
-  logbooks_df <- points_sf %>%
-    # filter(str_detect(SI_DATE,"11/2018")) %>%
-    group_by(ICESNAME_1) %>%
-    dplyr::summarise(Catch = sum(LE_KG_Solea_solea)/1000,
-                     Effort = sum(INTV)/60)
-
-  logbooks_sf <- st_join(ICES_rect,logbooks_df) %>%
-    filter(!is.na(Effort))
-
-  Logbooks_data <- ggplot()+
-    geom_sf(data = ICES_rect,fill="skyblue",alpha=0.25)+
-    geom_sf(data=logbooks_sf,aes(fill=Catch))+
-    geom_sf(data = mapBase)+
-    scale_fill_distiller(palette = "Spectral")+
-    theme(panel.background = element_rect(fill = "skyblue"),
-          plot.title = element_text(hjust=0.5),
-          legend.position = "none")+
-    geom_sf(col="black")+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    annotation_scale(location = "bl", line_width = .5)+
-    annotation_north_arrow(location = "tl", height = unit(0.7, "cm"),
-                           width = unit(0.7, "cm"))
-
-  ggsave(Logbooks_data,filename = "images/logbooks_data.png",width = 5,height = 5)
-
-  #---------------
-  ## VMS data plot
-  #---------------
-  ## Aggregate/discretize data
-  vms_df <- points_sf %>%
-    mutate(long = sf::st_coordinates(.)[,1],
-           lati = sf::st_coordinates(.)[,2]) %>%
-    as.data.frame %>%
-    dplyr::select(-geometry) %>%
-    mutate(long = round(long,digits = 2),
-           lati = round(lati,digits = 2)) %>%
-    group_by(long,lati) %>%
-    dplyr::summarise(Effort = sum(INTV)/60)
-
-  vms_sf <- st_as_sf(vms_df,coords = c("long","lati"),crs=grid_projection)
-
-  ## Build polygon
-  # Coordinates Bay of Biscay
-  grid_xmin <- -6
-  grid_xmax <- 0
-  grid_ymin <- 42
-  grid_ymax <- 48
-  grid_limit <- extent(c(grid_xmin,grid_xmax,grid_ymin,grid_ymax))
-
-  # Build grid
-  grid <- raster(grid_limit)
-  res(grid) <- 0.025
-  crs(grid) <- "+proj=longlat +datum=WGS84"
-  gridpolygon <- rasterToPolygons(grid)
-  gridpolygon$layer <- c(1:length(gridpolygon$layer))
-  gridpolygon_sf <- st_as_sf(gridpolygon)
-
-  vms_sf_2 <- st_join(gridpolygon_sf,points_sf)
-  vms_sf_3 <- vms_sf_2 %>%
-    group_by(layer) %>%
-    dplyr::summarise(Effort = sum(INTV)/60)
-
-  vms_sf_3 <- vms_sf_3 %>%
-    filter(!is.na(Effort))
-
-  # save(data=vms_sf_3,file="res/vms_sf_3.RData")
-
-  vms_data <- ggplot()+
-    geom_sf(data = ICES_rect,fill="skyblue",alpha=0.25)+
-    geom_sf(data=vms_sf_3,aes(col=Effort,fill=Effort))+
-    geom_sf(data = ICES_rect,alpha=0)+
-    geom_sf(data = mapBase)+
-    scale_color_distiller(direction = 1,palette = "OrRd")+ # palette = "OrRd"
-    scale_fill_distiller(direction = 1,palette = "OrRd")+ # ,palette = "Spectral"
-    theme(panel.background = element_rect(fill = "skyblue"),
-          plot.title = element_text(hjust=0.5),
-          legend.position = "none")+
-    geom_sf(col="black")+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)
-
-  ggsave(filename = "images/vms_data.png",width = 5,height = 5)
-
-}
-
-```
-
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\end{column}
-\begin{column}{0.45\textwidth}
-\center {\bf Logbook landings} \\
-\scriptsize (Sole, trawlers OTB-DEF, 2018) 
-\end{column}
-\begin{column}{0.45\textwidth}
-\center {\bf Fishing locations (VMS)} \\
-\scriptsize (Trawlers OTB-DEF, 2018)
-\end{column}
-\end{columns}
-
-\vspace{\baselineskip}
-
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\end{column}
-\begin{column}{0.45\textwidth}
-\includegraphics[width=5cm]{images/logbooks_data.png}
-\end{column}
-\begin{column}{0.45\textwidth}
-\includegraphics[width=5cm]{images/vms_data.png}
-\end{column}
-\end{columns}
-
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\center \tiny {\bf Resolution}
-\end{column}
-\begin{column}{0.45\textwidth}
-\center \tiny Catch are daily declared at the resolution of ICES rectangles
-\end{column}
-\begin{column}{0.45\textwidth}
-\center \tiny VMS pings are vessels GPS locations emitted each hour
-\end{column}
-\end{columns}
-\tiny
-\onslide<2>{
-\begin{columns}
-\begin{column}{0.1\textwidth}
-\end{column}
-\begin{column}{0.45\textwidth}
-\end{column}
-\begin{column}{0.45\textwidth}
-\center \scriptsize \textbf{\textcolor{BaptisteBlue}{\ding{224}}  Refine landings spatial resolution}
-\end{column}
-\end{columns}
-}
-
-### Modelling change of support
-
-\center
-
-\includegraphics[width=11cm]{images/realloc.png}
-
-\onslide<2-3>{
-
-\begin{table}
-  \centering \scriptsize
-  \begin{tabular}{ M{2cm} M{4cm} M{4cm} }
-  \textcolor{BaptisteRed}{\textbf{Actual situation}} & 
-  $Y_i | \textcolor{BaptisteBlue}{S}(x_i), x_i \sim \mathcal{L}_Y(\textcolor{BaptisteBlue}{S}(x_i),\xi,\sigma)$          &
-  $Y_i=\textcolor{BaptisteRed}{Y_i^*}$ \\
-  \end{tabular}
-\end{table}
-
-}
-
-\onslide<3>{
-
-\begin{table}
-  \centering \scriptsize
-  \begin{tabular}{ M{2cm} M{2cm} M{4cm} M{2cm} }
-  \textcolor{BaptisteLightGreen}{\textbf{Alternative solution}} & 
-  $\textcolor{BaptisteLightGreen}{D_j}=\sum_{\textcolor{BaptisteLightGreen}{i \in \mathcal{P}_j}}{Y_{i}}$          &
-  $\textcolor{BaptisteLightGreen}{D_j} | \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\textcolor{BaptisteLightGreen}{\mathcal{P}_j} \sim \mathcal{L}_D( \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\xi,\sigma)$
-  & \textcolor{BaptisteBlue}{\ding{224}}  Match $\mathcal{L}_D$ and $\mathcal{L}_Y$ moments \\
-  \end{tabular}
-\end{table}
-
-}
-
-### Simulation-estimation and case study
-
-```{r,echo=F}
-
-if(run_codes){
-  ## Load simulated data
-  load("res/simu_df.RData")
-  load("res/list_input.RData")
-
-  simu_df_2 <- simu_df %>%
-    mutate(Ab = sum(S_x))
-
-  loc_x <- list_input$loc_x
-  sci.data_df <- data.frame(cell = list_input$index_sci_i,
-                            y_sci = list_input$y_sci_i) %>%
-    inner_join(loc_x[,c("x","y","cell","ICESNAME")])
-
-  com.data_df <- data.frame(cell = list_input$index_i,
-                            y_com = list_input$y_i2) %>%
-    inner_join(loc_x[,c("x","y","cell","ICESNAME")]) %>%
-    group_by(x,y,cell,ICESNAME) %>%
-    dplyr::summarise(n=n())
-
-  ## Grid data
-  grid <- raster(grid_limit)
-  res(grid) <- 0.05
-  crs(grid) <- "+proj=longlat +datum=WGS84"
-  gridpolygon <- rasterToPolygons(grid)
-  gridpolygon$layer <- c(1:length(gridpolygon$layer))
-  gridpolygon_sf <- st_as_sf(gridpolygon)
-
-  simu_sf <- st_as_sf(simu_df_2,coords = c("long","lati"),crs=grid_projection)
-
-  simu_sf_2 <- st_join(gridpolygon_sf,simu_sf) %>%
-    filter(!is.na(S_x))
-
-  ## Plot data
-  data_plot <- ggplot()+
-    geom_sf(data=simu_sf_2,aes(fill=log(S_x)),col=NA,alpha=0.5)+
-    geom_sf(data=simu_sf_2[which(simu_sf_2$ICESNAME %in% com.data_df$ICESNAME),],
-            aes(fill=log(S_x),col=log(S_x)))+
-    scale_fill_distiller(palette = "Spectral")+
-    scale_color_distiller(palette = "Spectral")+
-    geom_point(data=sci.data_df,aes(x=x,y=y),fill="#619CFF",shape=21,size=1)+
-    theme_bw()+
-    theme(plot.title = element_text(hjust = 0.5),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.title = element_blank(),
-          legend.position = "none")+
-    geom_sf(data=ICES_rect,alpha=0)+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  ggsave(filename = "images/data_plot.png",width = 5,height = 5)
-
-}
-
-```
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-
-\center
-\textbf{Simulation/estimation}
-
-\end{column}
-\begin{column}{0.5\textwidth}
-
-\end{column}
-\end{columns}
-
-\tiny \vspace{\baselineskip}
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-
-\begin{columns}
-\begin{column}{0.45\textwidth}
-
-\includegraphics[width=3cm]{images/data_plot.png}
-
-\end{column}
-\begin{column}{0.55\textwidth}
-\scriptsize
-\textbf{Simulation}
-\tiny
-\begin{itemize}
-\item \textcolor{BaptisteBlue}{\textbf{Latent field}} (\textcolor{BaptisteOrange}{covariate} + \textcolor{BaptisteBrown}{spatial random effect})
-\item \textbf{Commercial data} (3000 samples over 2/3 of the area)
-\item \textcolor{BaptisteLightGreen}{\textbf{Reallocation process}} (10 locations per declaration)
-\item \textcolor{SurveyBlue}{\textbf{Scientific data}} (100 samples over the whole the area)
-\end{itemize}
-
-\end{column}
-\end{columns}
-
-\scriptsize
-\vspace{\baselineskip}
-
-\onslide<2-4>{
-
-\textbf{Estimation} \\
-Comparison of 3 model configurations: \\
-\tiny \vspace{\baselineskip} \scriptsize
-1/ Model fitted to \textcolor{SurveyBlue}{scientific data} (reference data) \\
-\tiny \vspace{\baselineskip} \scriptsize
-2/ Integrated model with \textcolor{BaptisteRed}{commercial likelihood built on $Y_i^*$} \\
-\tiny \vspace{\baselineskip} \scriptsize
-3/ Integrated model with \textcolor{BaptisteLightGreen}{commercial likelihood built on $D_j$} \\
-\tiny \vspace{\baselineskip}
-
-Estimation realized through TMB (Template Model Builder) \\
-100 runs of simulation-estimation
-
-}
-
-\end{column}
-\begin{column}{0.5\textwidth}
-
-\onslide<3-4>{
-
-\scriptsize
-\textbf{Model evaluation} \\
-1/ Mean square prediction error \\
-$$MSPE=\frac{\sum_x (\textcolor{BaptisteBlue}{S(x)} - \textcolor{BaptisteBlue}{\hat{S}(x)} ) ^2}{n}$$
-
-\vspace{\baselineskip}
-
-2/ Covariate effect (or species-habitat relationship): \\
-$$\textcolor{BaptisteOrange}{\beta_S} = 2 \, \, \, \text{versus} \, \, \, \textcolor{BaptisteOrange}{\hat{\beta}_S}$$
-
-}
-
-\onslide<4>{
-
-\vspace{\baselineskip}
-
-\normalsize
-\begin{center}
-\textbf{Case study:} \scriptsize Sole in the Bay of Biscay \\
-\end{center}
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-\center
-\includegraphics[width=2cm]{images/solea_solea.png}
-\end{column}
-\begin{column}{0.5\textwidth}
-\tiny \textcolor{SurveyBlue}{\textbf{Survey data:}} Orhago\\
-\textbf{Commercial data:} OTB-DEF traulers (to ease convergence onboard observer data were integrated in fitting)\\
-\textbf{Fitted models:} same as simulation\\
-\textcolor{BaptisteOrange}{\textbf{Covariate:}} substrate
-
-\end{column}
-\end{columns}
-
-}
-
-\end{column}
-\end{columns}
-
-# Results
-
-### Simulation-estimation
-
-```{r,echo=F}
-
-if(run_codes){
-
-  ######################
-  ## Parameter estimates
-  ######################
-  load("res/Results_full_multi_square.RData")
-
-  Results_plot <- Results_2 %>%
-    filter(converge == 0) %>%
-    filter(simu_type == "Unsampled Rectangles")
-
-  Results_plot[,"RelBias_N"]=(Results_plot[,"N_est"]-Results_plot[,"N_true"])/Results_plot[,"N_true"]
-  Results_plot[,"RelBias_N.2"]=(Results_plot[,"N_est.2"]-Results_plot[,"N_true.2"])/Results_plot[,"N_true.2"]
-  Results_plot[,"RelBias_beta"]=(Results_plot[,"beta1_est"]-Results_plot[,"beta1_true"])/Results_plot[,"beta1_true"]
-
-  Results_plot$obs <- NA
-  Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
-  Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi"
-  Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific model"
-  Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi","Dj","Scientific model"))
-
-  Results_plot$Model[which(Results_plot$Model == "Scientific model")] <- "Scientific"
-  Results_plot$Model[which(Results_plot$Model == "Commercial model")] <- "Commercial"
-  Results_plot$Model[which(Results_plot$Model == "Integrated model")] <- "Integrated"
-
-  Results_plot <- Results_plot %>%
-    filter(Model != "Commercial")
-
-  # Results_plot_fullArea <- Results_plot
-  # Results_plot_UnsampArea <- Results_plot
-
-  beta_plot <- ggplot()+
-    geom_boxplot(data = Results_plot,
-                 aes(x = Model,
-                     y = beta1_est,
-                     fill = obs))+
-    geom_hline(yintercept = 0,linetype="dashed")+
-    theme_bw()+
-    theme(legend.title = element_blank(),
-          legend.position = "none",
-          aspect.ratio = 1)+
-    # facet_wrap(.~simu_type)+
-    ylab(TeX("$\\beta_S$")) +
-    xlab("")+
-    geom_hline(yintercept = 2,col="red",linetype="dashed")+
-    scale_color_manual(breaks = c("Scientific model",
-                                  "Yi",
-                                  "Dj"),
-                       values=c("#619CFF","#F8766D","#00BA38"))
-
-  MSPE_S_plot <- ggplot()+
-    geom_boxplot(data = Results_plot,
-                 aes(x = Model,
-                     y = mspe,
-                     fill = obs))+
-    theme_bw()+
-    theme(legend.title = element_blank(),
-          legend.position = "none",
-          aspect.ratio = 1)+
-    # facet_wrap(.~simu_type)+
-    ylab("MSPE")+xlab("")+
-    scale_color_manual(breaks = c("Scientific model",
-                                  "Yi",
-                                  "Dj"),
-                       values=c("#619CFF","#F8766D","#00BA38"))+
-    ylim(0,NA)
-
-  legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="bottom",legend.title = element_blank())))
-
-  pl_plot <- plot_grid(MSPE_S_plot,
-                       beta_plot,
-                       ncol = 1,
-                       legend,
-                       align="hv",
-                       rel_heights = c(1,1,0.15))
-
-  ggsave(filename = "images/res_simu_plot.png",width = 4,height = 7.5)
-
-  #######
-  ## Maps
-  #######
-  simu_plot <- ggplot(simu_df)+
-    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral")+
-    ggtitle("Simulated latent field")+
-    theme_bw()+
-    theme(plot.title = element_text(hjust = 0.5,face = "bold"),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.position = "none")+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  load("res/fit_IM_res_sci.RData")
-  sci_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
-  sci_plot <- ggplot(sci_df)+
-    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral")+
-    ggtitle("Scientific model")+
-    theme_bw()+
-    theme(plot.title = element_text(colour = "#619CFF",hjust = 0.5,face = "bold"),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.position = "none")+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  load("res/fit_IM_res_int_Yi.RData")
-  int_Yi_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
-  int_Yi_plot <- ggplot(int_Yi_df)+
-    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral")+
-    ggtitle("Commercial likelihood on Yi")+
-    theme_bw()+
-    theme(plot.title = element_text(colour = "#F8766D",hjust = 0.5,face = "bold"),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.position = "none")+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  load("res/fit_IM_res_int_Dj.RData")
-  int_Dj_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
-  int_Dj_plot <- ggplot(int_Dj_df)+
-    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral")+
-    ggtitle("Commercial likelihood on Dj")+
-    theme_bw()+
-    theme(plot.title = element_text(colour = "#00BA38",hjust = 0.5,face = "bold"),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.position = "none")+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  map_simu <- plot_grid(simu_plot,sci_plot,int_Yi_plot,int_Dj_plot)
-
-  ggsave(filename = "images/map_simu_plot.png",width = 6,height = 7.5)
-
-}
-
-```
-
-\center
-
-\begin{columns}
-\begin{column}{0.4\textwidth}
-
-\center
-\includegraphics[width=4cm]{images/res_simu_plot.png}
-
-\end{column}
-
-\begin{column}{0.6\textwidth}
-\center
-
-\includegraphics[width=5.5cm]{images/map_simu_plot.png}
-\vspace{\baselineskip}
-
-\end{column}
-
-\end{columns}
-
-### Case study: sole in the Bay of Biscay
-
-```{r,echo=F}
-
-if(run_codes){
-
-  ## Load data
-  load("res/sci_df.RData")
-
-  load("res/realloc_rest_df.RData")
-
-  load("res/no.realloc_df.RData")
-
-  load("res/est_par_df_full_3.RData")
-
-  ## Plot maps
-  no.realloc_plot <- ggplot(no.realloc_df)+
-    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-    ggtitle("Commercial likelihood on Yi")+
-    theme_bw()+
-    theme(legend.position = "none",
-          plot.title = element_text(colour = "#F8766D",hjust = 0.5,face = "bold"))+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  realloc_rest_plot <- ggplot(realloc_rest_df)+
-    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-    ggtitle("Commercial likelihood on Dj")+
-    theme_bw()+
-    theme(legend.position = "none",
-          plot.title = element_text(colour = "#00BA38",hjust = 0.5,face = "bold"))+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  sci_plot <- ggplot(sci_df)+
-    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-    ggtitle("Scientific model")+
-    theme_bw()+
-    theme(legend.position = "none",
-          plot.title = element_text(colour = "#619CFF",hjust = 0.5,face = "bold"))+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  # plot(log(no.realloc_df$S_x),log(sci_df$S_x))
-  # cor(log(no.realloc_df$S_x),log(sci_df$S_x),method = "spearman")
-  # plot(log(realloc_rest_df$S_x),log(sci_df$S_x))
-  # cor(log(realloc_rest_df$S_x),log(sci_df$S_x),method = "spearman")
-
-  case_study_plot <- plot_grid(NULL,sci_plot,no.realloc_plot,realloc_rest_plot,align="hv")
-
-  ggsave("images/case_study_plot.png",width = 6, height = 7.5)
-
-  ## Parameter estimates
-  est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi","Scientific model"))
-  est_par_df_full_3 <- est_par_df_full_3 %>%
-    filter(par_names != "substr_Sand_Coarse_substrate")
-  par_plot <- ggplot(est_par_df_full_3, aes(y=par_val, x=par_names))+
-    geom_point(
-      aes(color = lkl),
-      position = position_dodge(0.5),
-      size=2
-    )+
-    geom_errorbar(aes(ymin = CI.inf, ymax = CI.sup, color = lkl),
-                  position = position_dodge(0.5),width=0.4
-    )+
-    scale_color_manual(breaks = c("Scientific model",
-                                  "Yi",
-                                  "Dj"),
-                       values=c("#619CFF","#F8766D","#00BA38"))+
-    geom_hline(yintercept=0, linetype="dashed", color = "red",alpha=0.4)+
-    xlab("")+ylab("")+
-    theme_bw()+
-    ggtitle("Parameters estimates")+
-    theme(plot.title = element_text(hjust = 0.5,face = "bold"),
-          legend.title = element_blank(),
-          legend.position = "bottom",
-          aspect.ratio = 1)+
-    coord_flip()+
-    scale_x_discrete(labels=c('k_com'=parse(text = TeX('$k_{com}$')),
-                              'Sigma_com'=parse(text = TeX('$\\sigma_{com}$')),
-                              'q1_com'=parse(text = TeX('$\\xi_{com}$')),
-                              'Sigma_sci'=parse(text = TeX('$\\sigma_{sci}$')),
-                              'q1_sci'=parse(text = TeX('$\\xi_{sci}$')),
-                              'Range'="Range",
-                              'MargSD'="M. var.",
-                              'substr_Mud_sediment'=parse(text = TeX('$\\beta_{S}$')),
-                              'intercept'=parse(text = TeX('$\\mu$'))))
-
-  ggsave("images/par_plot.png",width = 6,height = 7.5)
-
-}
-
-```
-
-\center
-
-\begin{columns}
-\begin{column}{0.4\textwidth}
-
-\begin{center}
-\includegraphics[width=5cm]{images/par_plot.png}
-\end{center}
-\tiny *n.b.: punctual commercial data (i.e. observer data) needed to be integrated in inference in order to make the model converge and some parmaters needed to be step estimated (e.g. range)
-
-\end{column}
-
-\begin{column}{0.6\textwidth}
-\center
-\includegraphics[width=5.5cm]{images/case_study_plot.png}
-\vspace{\baselineskip}
-
-\end{column}
-
-\end{columns}
-
-# Discussion
-
-### Discussion
-
-\begin{itemize}
-
-\item Framework that integrates \textbf{catch declarations data} (rough resolution of catch) and \textbf{scientific/onboard observer data} (exact locations of catch) \\
-
-\onslide<2-4>{
-
-\textcolor{BaptisteBlue}{\ding{224}} \textbf{Still question:} how to ease convergence ? \\
-
-}
-
-\vspace{\baselineskip}
-
-\onslide<3-4>{
-
-\item \textbf{ Change of support issues } \\
-
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} not so frequent in spatial ecology, \\
-
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} but increasing amount of data available that may face such issues \\
-
-}
-
-\vspace{\baselineskip}
-
-\onslide<4>{
-
-\item \textbf{Is it a generic framework ?} \\
-
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} the overall approach is, \tiny \\ (i.e.
-modelling observed aggregagetd catches as a convolution of latent exact catches) \\ \vspace{\baselineskip} \normalsize
-
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} but need to adapt the observation model to the data \tiny \\ (here zeroinflated continuous data)   \\ \vspace{\baselineskip} \normalsize
-
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} and make the hypothesis that fishing locations ($\textcolor{BaptisteLightGreen}{\mathcal{P}_j}$) are known \scriptsize \\ (in reality $\textcolor{BaptisteLightGreen}{\mathcal{P}_j}$ is a trajectory and the set of fishing points are a realization of this trajectory) \\ \vspace{\baselineskip} \normalsize
-
-}
-
-\end{itemize}
-
-#
-
-\centering \Large \textcolor{BaptisteGrey}{\textbf{Thank you for your attention!}}
+\centering \Large \textcolor{BaptisteGrey}{\textbf{Merci de votre attention !}}
 

### 2022-03-30 Version_M2_plusieurszones/draft/paper_3/3_formulas.Rmd
paper_3

@@
 
-# Notations
+**Notations**
 
-We model observations at the level of catch declarations $D_j$ as a sum of $Y_{ij}$ lognormal observations each one realised at one fishing position $x_{ij}$.
+We model catch declarations $D_j$ (available at coarse resolution through logbooks data) as a sum of $Y_{i}$ punctual observations (which are unknown i.e.
+latent variables) each one realised at one fishing position $x_{i}$ (known through VMS data).
 
-We note :
+We note:
 
-- $\mathcal{P}_j$ : the vector of all fishing positions related to the $j^th$ declaration
+- $\mathcal{P}_j$ : the vector of all fishing positions related to the $j^{th}$ declaration.
 
-- $j \in \{1,...,n\}$ with $n$ the number of declarations
+- $j \in \{1,...,l\}$ with $l$ the number of declarations.
 
@@
 
-$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{ij}}$$
+$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$$
 
-# Alternative observation models accounting for reallocation
+**Reparameterization of the lognormal distribution**
 
-## First assumption about convolution of lognormal random variables
+The lognormal distribution is usually written as $Z \sim \mathcal{LN}(\rho;\sigma^2)$ with $Z=e^{\rho+\sigma N}$ and $N \sim \mathcal{N}(0,1)$.
+In this case, $E(Z)=e^{\rho + \frac{\sigma^2}{2}}$ and $Var(Z)=(e^{\sigma^2}-1)e^{2 \rho + \sigma^2}$.
 
-We assume a sum of lognormal random variables is still lognormal and that $D_j \sim \mathcal{LN}(M(D_j),\Phi(D_j))$ with $M(.)$ being a function representing the central value of $D_j$ and $\Phi(.)$ the variance function of $D_j$.
+We choose to slightly reparameterize the lognormal distribution. Let's define $\rho = ln(\mu) - \frac{sigma^2}{2}$, then:
 
-## Reparameterization of the lognormal distribution
+- $Z=\mu e^{\sigma N - \frac{sigma^2}{2}}$
 
-The standard lognormal distribution is written as :
+- $E(Z)=\mu$
 
-$D \sim \mathcal{LN}(\rho;\sigma^2)$ with $D=e^{\rho+\sigma N}$ and $N \sim \mathcal{N}(0,1)$
+- $Var(Z)=\mu^2(e^{\sigma^2} - 1) \Leftrightarrow	\sigma^2=ln(\frac{Var(Z)}{E(Z)^2} + 1)$
 
-In this case, $E(D)=e^{\rho + \frac{\sigma^2}{2}}$ and $Var(D)=(e^{\sigma^2}-1)e^{2 \rho + \sigma^2}$.
+**$D_j$ probability distribution and moments**
 
-We choose to slightly reparameterize the lognormal distribution so that $E(D)$ and $Var(D)$ have more simple expressions.
+We have to express the probability distribution of $D_j$ and its moments as a function of $Y_{i}$ and its related moments.
+Let's assume $Y_{i} = C_{i}.Z_{i}$ is a zero-inflated lognormal distribution with $C_{i}$ a binary random variable and $Z_{i}$ a lognormal random variable.
 
-Let's define $\rho = ln(\mu) - \frac{sigma^2}{2}$, then :
+$$C_{i} \sim \mathcal{B}(1-p_{i})$$ 
+with $p_{i}=exp(-e^\xi .S(x_{i}))$ the probability to obtain a zero value.
 
-- $D=\mu e^{\sigma N - \frac{sigma^2}{2}}$
+$$Z_{i} \sim \mathcal{LN}(\frac{S(x_{i})}{1-p_{i}},\sigma^2)$$
 
-- $E(D)=\mu$
+Here, $Y_{i}$, $C_{i}$ and $Z_{i}$ are observations of a latent field $S(x_{i})$ at a sampled point $x_{i}$.
 
-- $Var(D)=\mu^2(e^{\sigma^2} - 1) \Leftrightarrow	\sigma^2=ln(\frac{Var(D)}{E(D)^2} + 1)$
-
-Then, $D \sim\mathcal{LN}(M(D),\Phi(D))$ with $M(D)=E(D)$ and $\Phi(D)=ln(\frac{Var(D)}{E(D)^2} + 1)$.
-
-## $D_j$ probability distribution and moments
-
-We have to express the probability distribution of $D_j$ and its moments as a function of $Y_{ij}$ with $i \in \mathcal{P}_j$.
-
-Let's assume $Y_{ij} = C_{ij}.Z_{ij}$ is a zero-inflated lognormal distribution with $C_{ij}$ a binary random variable and $Z_{ij}$ a lognormal random variable.
-
-$$C_{ij} \sim \mathcal{B}(1-p_{ij})$$ 
-with $p_{ij}=exp(-e^\xi .S(x_{ij}))$ the probability to obtain a zero value
-
-$$Z_{ij} \sim \mathcal{LN}(\frac{S(x_{ij})}{1-p_{ij}},\sigma^2)$$
-
-Here, $Y_{ij}$, $C_{ij}$ and $Z_{ij}$ are observations of a latent field $S(x_{ij})$ at a sampled point $x_{ij}$.
-
-### Probability of obtaining a zero declaration
+**Probability of obtaining a zero declaration**
 
 \begin{align*}
-P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_{ij} = 0 \vert S, X),\nonumber \\
-                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{ij})\right \rbrace} = \pi_j.
+P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0 \vert S, X),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
 \end{align*}
 
-### Expectancy of a positive declaration
+**Expectancy of a positive declaration**
 
@@
 
-$$E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_{ij} Z_{ij}\vert \exists i\in\mathcal{P}_j , C_{ij}=1)$$
-
-As $C_{ij}$ and $Z_{ij}$ are assumed to be independant.
+$$E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}\vert \exists i\in\mathcal{P}_j , C_{i}=1)$$
 
@@
 E(D_j\vert D_j > 0) & = \left (1- \pi_j\right)^{-1} E(D_j)   ,   \nonumber \\
-& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_{ij} Z_{ij}),\nonumber \\
-& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_{ij}) \frac{S(x_{ij})}{1-p_{ij}}, \nonumber \\
-& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} S(x_{ij}).
+& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}),\nonumber \\
+& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_{i}) \frac{S(x_{i})}{1-p_{i}}, \nonumber \\
+& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} S(x_{i}).
 \end{align*}
 
-### Variance of a positive declaration
+**Variance of a positive declaration**
 
@@
 
-And then,
+Then,
 
@@
 
-As the $(Y_{ij})_{i\in \mathcal{P}_j}$ are independent, $Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_{ij}) = \sum_{i \in\mathcal{P}_j} Var(C_{ij}.Z_{ij})$.
+As the $(Y_{i})_{i\in \mathcal{P}_j}$ are independent, $Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_{i}) = \sum_{i \in\mathcal{P}_j} Var(C_{i}.Z_{i})$.
 
 \begin{align*}
-Var(C_{ij} Z_{ij}) & = E( C_{ij}^2 Z_{ij}^2 ) - E(C_{ij} Z_{ij})^2,\\
-& =  E( C_{ij}^2) E(Z_{ij}^2 ) -E(C_{ij})^2 E(Z_{ij})^2,\\
-& = (1-p_{ij}) E(Z_{ij}^2 ) - (1-p_{ij})^2 E(Z_{ij})^2,\\
-& = (1-p_{ij})(Var(Z_{ij})+E(Z_{ij})^2)-(1-p_{ij})^2E(Z_{ij})^2,\\
-& = \frac{S(x_{ij})^2}{1-p_{ij}}(e^{\sigma^2}-1)+\frac{S(x_{ij})^2}{1-p_{ij}}-S(x_{ij})^2,\\
-& = \frac{S(x_{ij})^2}{1-p_{ij}}(e^{\sigma^2}-(1-p_{ij}))
+Var(C_{i} Z_{i}) & = E( C_{i}^2 Z_{i}^2 ) - E(C_{i} Z_{i})^2,\\
+& =  E( C_{i}^2) E(Z_{i}^2 ) -E(C_{i})^2 E(Z_{i})^2,\\
+& = (1-p_{i}) E(Z_{i}^2 ) - (1-p_{i})^2 E(Z_{i})^2,\\
+& = (1-p_{i})(Var(Z_{i})+E(Z_{i})^2)-(1-p_{i})^2E(Z_{i})^2,\\
+& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-1)+\frac{S(x_{i})^2}{1-p_{i}}-S(x_{i})^2,\\
+& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))
 \end{align*}
 
-### Conclusion
+**Sum up of the main formulas**
 
-$$P(D_j = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{ij})\right \rbrace} = \pi_j$$
+$$P(D_j = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j$$
 
-$$D_j \vert D_j > 0 \sim \mathcal{LN}(E(D_j \vert D_j > 0); ln(\frac{Var(D_j \vert D_j > 0)}{E(D_j \vert D_j > 0)^2} + 1))$$
+$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
 
-$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{ij})}{1-\pi_j}$$
+$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
 
-$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{ij})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
+$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{j}}(e^{\sigma^2}-(1-p_{i}))$$
 
-$$Var(Y_{ij})=\frac{S(x_{ij})^2}{1-p_{ij}}(e^{\sigma^2}-(1-p_{ij}))$$
+Assuming $D_j \vert D_j > 0$ also follows a lognormal distribution we can write:
+
+$$D_j \vert D_j > 0 \sim \mathcal{LN}( \mu_j = E(D_j \vert D_j > 0), \sigma_j^2= ln(\frac{Var(D_j \vert D_j > 0)}{E(D_j \vert D_j > 0)^2} + 1))$$
 

### 2022-03-30 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3

@@
 
-floatsintext      : no
+floatsintext     : yes
 figurelist       : no
@@
 library(kableExtra)
+library(knitr)
 library(latex2exp)
@@
 Observations are supposed to follow some distribution $\mathcal{L}_Y$.
-In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.
 
@@
 
+In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.
 $\mathcal{L}_Y$ is decomposed it 2 parts:
@@
 
-- if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM...)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$.
+- if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a positive continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$.
 
@@
 
-Fishermen must declare what they catch at the level of the statistical rectangle.
-Declarations are daily recorded in logbooks data.
-The declaration (denoted $D_j$) is a summation of all the $Y_i$ realized on the fishing positions belonging to the catch declaration $D_j$ and they can be expressed as:
+Fishermen must declare what they catch at the level of the statistical rectangle at a daily time step.
+These data are recorded in logbooks data.
+A declaration (denoted $D_j$) is a summation of all the $Y_i$ realized on the fishing positions belonging to the catch declaration $D_j$.
+They can be expressed as:
 
@@
 
-$\mathcal{P}_j$ is the vector of the fishing observations belonging to declaration $D_j$.
+$\mathcal{P}_j$ is the vector of the fishing observations belonging to the declaration $D_j$.
 The punctual observations $Y_i$ are indexed through $i \in [\![1,m_j]\!]$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.
@@
 
-In standard processing, $D_j$ (from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$.
+In standard processing, $D_j$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$.
 This is what we call uniform (or proportional) reallocation.
@@
 This model will be referred as the $D_j$ model.
-In contrast, the model fitted on punctual observations $Y_i$ will be referred as the $Y^*_i$ model.
+In contrast, the model fitted on reallocated observations $Y^*_i$ will be referred as the $Y^*_i$ model.
 
-In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | S_{\mathcal{P}_j},\mathcal{P}_j$.
+In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | \mathcal{P}_j,S_{\mathcal{P}_j}$.
 
@@
 
-If we consider $\mathcal{L}_{D_j|D_j>0}$ is lognormal (with parameterization described in SM...) then, $\mu_j = E(D|D_j>0)$ and $\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$.
+If we consider $\mathcal{L}_{D_j|D_j>0}$ is lognormal (with parameterization described in SM) then, $\mu_j = E(D|D_j>0)$ and $\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$.
+Such way, by fitting the model on declaration $D_j$ rather than reallocated catch $Y_i$ and linking $\mathcal{L}_D$ to $\mathcal{L}_Y$, we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation.
 All calculations to obtain these formulas are available in SM.
-By fitting the model on declaration $D_j$ rather than reallocated catch $Y_i$ and linking $\mathcal{L}_D$ to $\mathcal{L}_Y$, we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation.
 
@@
 
-![(\#fig:SchemeRealloc) Reallocation process.](images/realloc.png)
+![(\#fig:SchemeRealloc) Schematic representation of the reallocation process.](images/realloc.png)
 
@@
 
-To evaluate what bring the performance of the new approach, we assess models performance through simulation-estimation.
+To evaluate the impact of reallocation on model outputs and how our alternative approach can improve model accuracy, we conduct simulation-estimation.
 
-First, we conduct basic simulations in order to explore the base properties of the framework and the effect of some key variables on model performance.
+First, we conduct basic simulations in order to explore the effect of some key variables on model performance and also the base properties of the alternative models.
 We assume the study domain is a single statistical rectangle, only commercial data feed the model and the expression of the latent field in both simulation and estimation is simplified (the variability only depends on one covariate, not on a spatial random term).
-These simulations will be referred as single-square simulations.
+These simulations will be referred as "single-square simulations".
 
@@
 Regarding commercial data, the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for a single declaration in real data.
-All parameterizations are detailed in the table ....
+All parameterizations are detailed in the Table 1.
 
@@
 - the parameter of the species habitat relationship $\beta_S$ compared with its estimate $\hat \beta_S$.
+
+```{r}
+
+df <- data.frame(v1=c("$\\mu$","$\\beta_S$","$\\text{Range}$","$\\text{Marginal variance}$","$\\xi_{com}$","$\\sigma_{com}$","$k_{com}$","$\\xi_{sci}$","$\\sigma_{sci}$"),
+                 v2=c("2","2"," -- "," -- ","-1","1"," -- "," -- "," -- "),
+                 v3=c("2","2","1.5","0.5","-1","1","1","0","0.8"))
+
+# italic column headers
+colnames(df) <- c("\\text{Parameters}","\\text{Single-square simulations}","\\text{Multiple-square simulations}")
+
+kable(df, "latex", align=c("l","c","c"), booktabs=TRUE, escape = F, caption = 'Parameter values for the simulations')
+
+table_param <- t(data.frame(
+  c("$\\mu$","2","2"),
+  c("$\\beta_S$","2","2"),
+  c("$\\text{Range}$"," -- ","1.5"),
+  c("$\\text{Marginal variance}$"," -- ","0.5"),
+  c("$\\xi_{com}$","-1","-1"),
+  c("$\\sigma_{com}$","1","1"),
+  c("$k_{com}$"," -- ","1"),
+  c("$\\xi_{sci}$"," -- ","0"),
+  c("$\\sigma_{sci}$"," -- ","0.8")
+  ))
+
+```
 
@@
 
-- a golden standard configuration: observations are known exactly and the $Y^*_i$ model is fitted to the exact data.
+- a golden standard configuration: punctual observations ($Y_i$) are known exactly and the punctual-data model is fitted to the exact data.
 
 - a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations.
-We then fit the $Y^*_i$ model to the reallocated data ($Y_i^*$).
+We then fit the $Y_i^*$ model (the model fitted to punctual reallocated data).
 
-- a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the $D_j$ model to the data (i.e. the one working at the level of $D_j$).
+- a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the $D_j$ model to the data (i.e.
+the one fitted at the level of the declarations data).
 
@@
 
-The study area is based on the case study; it includes the whole coast of the Bay of Biscay (Figure \@ref(fig:MapSeveral)A).
+The study area is based on the case study; it includes the whole coast of the Bay of Biscay and cover several statistical rectangles (Figure \@ref(fig:MapSeveral)A).
 To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration).
@@
 (2) The centroid of the fishing zone is randomly sampled within this statistical rectangle.
-(3) The 10 fishing locations are randomly sampled within the fishing zone.
+(3) The 10 fishing punctual observations are randomly sampled within the fishing zone.
 The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km.
@@
 100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A).
-Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table ...).
+Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table 1).
 
@@
 
-- to assess what bring our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.
+- to assess what brings our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.
 
@@
 In addition to the 2 metrics introduced at the beginning of the section ($MSPE$ and species-habitat parameter $\beta_S$), we also compare the simulated/estimated values of the range parameter.
+
+Note that when including 2 datasets in the model, a difference in catchability should be accounted for and one of the two dataset will be considered as reference (see @alglave_combining_2022 for more details on parameterization).
+Here the reference level is set with scientific data and the parameter $k_{com}$ accounts for the relative ratio between scientific and commercial observation values.
+In simulations, $k_{com}$ is set to 1 so that both datasources are assumed to have the same catchability.
 
@@
 
-The $D_j$ model faced congvergence issues (some of the parameters were hardly estimated e.g.
+The $D_j$ model faced convergence issues (some of the parameters were hardly estimated e.g.
 the range parameter).
@@
 They are considered as exact commercial data of catches.
-Then, integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
+Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
 observation variance and zero-inflation parameter of commercial data).
@@
 
-## Single square analysis
+## Single-square simulations
 
@@
            reallocation = ifelse(reallocation == 1,"Yes","No"),
-           perc_convergence = round((1 - Convergence.sum / one.sum)*100,digits = 3))
+           perc_convergence = round((1 - Convergence.sum / one.sum)*100,digits = 3)) %>%
+  mutate(aggreg_obs = ifelse((aggreg_obs == "Yi" & reallocation == "Yes"),"Yi*",aggreg_obs))
 
@@
   dplyr::select(-Convergence.sum,-one.sum)
-
-knitr::kable(Results_conv_2,booktabs = T,align = "c",
-                  caption = "Percentage of convergence per simulation/model configuration at the level of a single rectangle")
 
@@
 Results <- Results %>%
-  mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dj","Yi")))
+  mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dj","Yi"))) %>%
+  mutate(relloc_aggreg = ifelse(aggreg_obs == F & reallocation == 1,"Reallocation: Yes -  Likelihood: Yi*",relloc_aggreg))
 
@@
                                 levels = c("Reallocation: No -  Likelihood: Yi",
-                                           "Reallocation: Yes -  Likelihood: Yi",
+                                           "Reallocation: Yes -  Likelihood: Yi*",
                                            "Reallocation: Yes -  Likelihood: Dj"))
@@
   scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
-                                "Reallocation: Yes -  Likelihood: Yi",
+                                "Reallocation: Yes -  Likelihood: Yi*",
                                 "Reallocation: Yes -  Likelihood: Dj"),
@@
   scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
-                               "Reallocation: Yes -  Likelihood: Yi",
+                               "Reallocation: Yes -  Likelihood: Yi*",
                                "Reallocation: Yes -  Likelihood: Dj"),
@@
 
-![(\#fig:PerfMetricSingle) Performance metric for single-square simulations.
-columns: number of commercial pings.
-x-axis: number of zones visited within each declaration.
-'Reallocation: ', data are or are not reallocated in simulations.
-'Likelihood: ', the likelihood is computed on reallocated observations $Y_i$ or on catch declarations $D_j$.
-Gold: golden standard.
-Gold: golden standard.
-Red: uniform reallocation situation.
-Green: model-based reallocation ($D_j$ model).
-Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the $D_j$ model.](images/Perf.metric_single_square.png)
-
 Reallocation has a major effect on predictions and estimates accuracy (Figure \@ref(fig:PerfMetricSingle)).
@@
 Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \@ref(fig:ParBiasSingle)).
-This alternative model have some convergence difficulties (Table 1) as 8% of the model did not converged when sample size is medium (100 pings) and only 3% did not when sample size is large (1000 pings).
+This alternative model have some convergence difficulties (Table 2) as 8% of the model did not converged when sample size is medium (100 pings) and only 3% did not when sample size is large (1000 pings).
+
+![(\#fig:PerfMetricSingle) Performance metric for single-square simulations.
+Columns: number of commercial fishing positions.
+x-axis: number of zones visited within each declaration.
+'Reallocation: ', data are or are not reallocated in simulations.
+'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $Y_i^*$ or on catch declarations $D_j$.
+Gold: golden standard.
+Red: uniform reallocation ($Y_i^*$ model).
+Green: model-based reallocation ($D_j$ model).
+Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the $D_j$ model (see Table 2).](images/Perf.metric_single_square.png){width=75%}
 
@@
 'Reallocation: ', data are or are not reallocated in simulations.
-'Likelihood: ', the likelihood is computed on reallocated observations $Y_i$ or on catch declarations $D_j$.
+'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $Y_i^*$ or on catch declarations $D_j$.
 Gold: golden standard.
-Red: uniform reallocation situation.
+Red: uniform reallocation ($Y_i^*$ model).
 Green: model-based reallocation ($D_j$ model).
@@
 
-## Multiple square analysis
+```{r}
 
-```{r}
+knitr::kable(Results_conv_2,booktabs = T,align = "c",
+                  caption = "Single-square simulations - Percentage of convergence per simulation-estimation configuration.")
+
+```
+
+\newpage
+
+## Multiple-square simulations
+
+```{r,warning=F}
 
@@
 
+Results_conv_2$aggreg_obs[which(Results_conv_2$aggreg_obs == "Yi")] <- "Yi*"
+
 Results_conv_2 <- Results_conv_2 %>%
@@
 Results_conv_2$`Likelihood level`[which(Results_conv_2$Model == "Scientific model" &
-                       Results_conv_2$`Likelihood level` == "Yi")] <- " "
-knitr::kable(Results_conv_2,booktabs = T,align = "c",
-                  caption = "Percentage of convergence per simulation/model configuration at the level of several ICES rectangles")
+                       Results_conv_2$`Likelihood level` == "Yi*")] <- " "
 
@@
 Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
-Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi"
+Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi*"
 Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific data"
-Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi","Dj","Scientific data"))
+Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi*","Dj","Scientific data"))
 
@@
   scale_color_manual(breaks = c("Scientific model",
-                                "Yi",
+                                "Yi*",
                                 "Dj"),
@@
 
-load("res/lkl_prof_beta.RData")
+# load("res/lkl_prof_beta.RData")
 
@@
 Whether the model is fitted at the punctual level or at the declaration level, the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models.
-This can be well evidenced when looking at Figure \@ref(fig:MapSeveral).
+This can be well illustrated from Figure \@ref(fig:MapSeveral).
 Integrating scientific and commercial data allows to (1) capture the hotspot missed by commercial data through scientific data and (2) better capture the local correlation structures through the dense commercial data.
@@
 
-Interestingly, uniform reallocation does not affect only the species-habitat relationship but also the spatial autocorrelation terms such as the range parameter.
+Interestingly, in addition to the species-habitat relationship, uniform reallocation also affects the spatial autocorrelation terms such as the range parameter.
 The model fitted on reallocated data ($Y_i^*$) provides biased range estimates while the $D_j$ model provides unbiased estimates.
@@
 The $D_j$ model allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
-This is evidenced in Figure \@ref(fig:MapSeveral) where the model fitted to reallocated catch $Y_i^*$ provides smoothed maps and does not captures the relatively small scale patterns that are shaped by the covariate.
-On the other hand the $D_j$ model (as the scientific-based model) better capture and disentangle the covariate effect and the spatial random effect and then provides predictions that better fit to the small-scale patterns of the species distribution.
-However, this goes with some difficulty in convergence as only 75% of the model built on catch declarations converge (Table 2).
+This is evidenced in Figure \@ref(fig:MapSeveral) where the model fitted to reallocated catch $Y_i^*$ provides smoothed maps and does not capture the relatively small scale patterns that are shaped by the covariate.
+On the other hand, the $D_j$ model (as the scientific-based model) better captures and disentangles the covariate effect and the spatial random effect and then provides predictions that better fit to the small-scale patterns of the species distribution.
+However, this goes with some difficulty in convergence as only 75% of the model built on catch declarations converge (Table 3).
 
-![(\#fig:PerfMetricSeveral) Performance metric for several rectangles simulations.
-columns: commercial data coverage.
-x-axis: likelihood level.
-$1^{st}$ row, red line: true value of $\beta_S$.](images/Perf.metric_multiple_square.png)
+![(\#fig:PerfMetricSeveral) Performance metric for the multiple-square simulations.
+Red line: true value for the range and the species-habitat parameter ($\beta_S$).
+Red: uniform reallocation ($Y_i^*$ model).
+Green: model-based reallocation ($D_j$ model).
+Blue: scientific-based model.](images/Perf.metric_multiple_square.png)
 
 ![(\#fig:MapSeveral) Relative distribution of simulated/estimated biomass field.
-A: Simulated biomass field with scientific samples (red) and rectangles that have not been sampled by commercial data (transparent rectangles).
+A: Simulated biomass field with scientific samples (red) and statistical rectangles.
+The rectangles that have not been sampled by commercial data are the transparent rectangles.
+They represent 1/3 of the full area.
 B: simulated biomass field.
 C: biomass field from the scientific-based model.
-$Y_i$: model fitted at the punctual observation level (D, E).
-$D_j$: $D_j$ model (F, G).](images/Map_multi_square.png)
+$Y_i^*$: $Y_i^*$ model (D, E).
+$D_j$: $D_j$ model (F, G).
+Scientific model: model fitted to scientific data only.
+Commercial model: model fitted to commercial data only.
+Integratedd model: model fitted to both data sources.](images/Map_multi_square.png)
+
+```{r}
+
+knitr::kable(Results_conv_2,booktabs = T,align = "c",
+                  caption = "Multiple-square simulations - Percentage of convergence per simulation-estimation configuration.")
+
+```
 
@@
   scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-  ggtitle("Integrated model",subtitle = "Commercial likelihood on Yi")+
+  ggtitle("Integrated model",subtitle = "Commercial likelihood on Yi*")+
   theme_bw()+
@@
 ## Parameter estimates
-est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi","Scientific model"))
+est_par_df_full_3$lkl[which(est_par_df_full_3$lkl == "Yi")] <- "Yi*"
+est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi*","Scientific model"))
 est_par_df_full_3 <- est_par_df_full_3 %>%
@@
 
-![(\#fig:CaseStudyPar) Parameters obtained with scientific-based model, the integrated model fitted on reallocated catch $Y_i^*$ and the integrated model fitted on catch declarations $D_j$.](images/par_plot.png)
-
-![(\#fig:CaseStudyMap) Maps obtained from (left) the scientific-based model, (center) the integrated model fitted on punctual observations $Y_i^*$, (right) the integrated model fitted on catch declarations $D_j$.](images/case_study_plot.png)
-
 Consistently with simulations, the $D_j$ model modifies some of the parameters estimates and revises the spatial pattern of the species distribution compared with the $Y^*_i$ model.
-In particular, the substrate effect is recovered in the $D_j$ model and fall in the same range of values of the scientific-based estimate (Figure \@ref(fig:CaseStudyMap)).
+In particular, the substrate effect is recovered in the $D_j$ model and fall in the same range as the scientific-based estimate (Figure \@ref(fig:CaseStudyMap)).
 The zero-inflation parameter $\xi$ is revised downwards (i.e.
 there are actually more zero-values than in the reallocated data) while the observation variance of commercial data is revised upwards (i.e.
-the commercial data are more noisy than expected when building likelihood on $Y_i$).
+the commercial data are noisier than expected when fitting the model on $Y_i^*$).
 
@@
 This emphasizes that uncertainty is probably underestimated in the $Y^*_i$ model compared with the $D_j$ model.
-Now, when comparing the scientific-based model and the integrated $D_j$ model, some parameters are better estimated than when there is only scientific data feeding the model.
+Now, when comparing the scientific-based model and the integrated $D_j$ model, some parameters are better estimated than when only scientific data feed the model.
 For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on $D_j$ substrate is significant and the confidence interval is smaller.
@@
 Regarding the maps of the species distribution, fitting the model at the $D_j$ level strongly modifies the model biomass field compared with the $Y_i^*$ model.
-In particular, the substrate covariate have a sharper effect on species distribution.
-Furthermore, the intensity of the hotspots are revised when fitting the model on $D_j$.
+In particular, the substrate covariate have a sharper effect on species distribution and the intensity of the hotspots are revised when fitting the model on $D_j$.
 
-Finally, the $D_j$ model fitted only on commercial data does not converge (while the one built on $Y_i^*$ does) emphasizing the model fitted on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
+Finally, the $D_j$ model fitted only on commercial data does not converge (while the one fitted on $Y_i^*$ does) emphasizing the model fitted on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
+
+![(\#fig:CaseStudyPar) Parameters obtained with scientific-based model, the integrated model fitted on reallocated catch $Y_i^*$ and the integrated model fitted on catch declarations $D_j$.](images/par_plot.png)
+
+![(\#fig:CaseStudyMap) Maps obtained from (left) the scientific-based model, (center) the integrated model fitted on punctual reallocated catch $Y_i^*$, (right) the integrated model fitted on catch declarations $D_j$.](images/case_study_plot.png)
 
@@
 \endgroup
+
+# Supplementary material
+
+**Notations**
+
+We model catch declarations $D_j$ (available at coarse resolution through logbooks data) as a sum of $Y_{i}$ punctual observations (which are unknown i.e.
+latent variables) each one realised at one fishing position $x_{i}$ (known through VMS data).
+
+We note:
+
+- $\mathcal{P}_j$ : the vector of all fishing positions related to the $j^{th}$ declaration.
+
+- $j \in \{1,...,l\}$ with $l$ the number of declarations.
+
+- $i \in \{1,...,m_j\}$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.
+
+$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$$
+
+**Reparameterization of the lognormal distribution**
+
+The lognormal distribution is usually written as $Z \sim \operatorname{L}(\rho;\sigma^2)$ with $Z=e^{\rho+\sigma N}$ and $N \sim \mathcal{N}(0,1)$.
+In this case, $E(Z)=e^{\rho + \frac{\sigma^2}{2}}$ and $Var(Z)=(e^{\sigma^2}-1)e^{2 \rho + \sigma^2}$.
+
+We choose to slightly reparameterize the lognormal distribution. Let's define $\rho = ln(\mu) - \frac{sigma^2}{2}$, then:
+
+- $Z=\mu e^{\sigma N - \frac{sigma^2}{2}}$
+
+- $E(Z)=\mu$
+
+- $Var(Z)=\mu^2(e^{\sigma^2} - 1) \Leftrightarrow	\sigma^2=ln(\frac{Var(Z)}{E(Z)^2} + 1)$
+
+**$D_j$ probability distribution and moments**
+
+We have to express the probability distribution of $D_j$ and its moments as a function of $Y_{i}$ and its related moments.
+Let's assume $Y_{i} = C_{i}.Z_{i}$ is a zero-inflated lognormal distribution with $C_{i}$ a binary random variable and $Z_{i}$ a lognormal random variable.
+
+$$C_{i} \sim \mathcal{B}(1-p_{i})$$ 
+with $p_{i}=exp(-e^\xi .S(x_{i}))$ the probability to obtain a zero value.
+
+$$Z_{i} \sim \operatorname{L}(\frac{S(x_{i})}{1-p_{i}},\sigma^2)$$
+
+Here, $Y_{i}$, $C_{i}$ and $Z_{i}$ are observations of a latent field $S(x_{i})$ at a sampled point $x_{i}$.
+
+**Probability of obtaining a zero declaration**
+
+\begin{align*}
+P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0 \vert S, X),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
+\end{align*}
+
+**Expectancy of a positive declaration**
+
+Following calculations are supposed to be conditionnal on $S$ and $X$.
+
+$$E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}\vert \exists i\in\mathcal{P}_j , C_{i}=1)$$
+
+\begin{align*}
+E(D_j\vert D_j > 0) & = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / P\left ( D_j > 0\right ),   \\
+& = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / \left (1-\pi_j\right).   \\
+\end{align*}
+
+As $E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } ) = E(D_j ),$
+
+\begin{align*}
+E(D_j\vert D_j > 0) & = \left (1- \pi_j\right)^{-1} E(D_j)   ,   \nonumber \\
+& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}),\nonumber \\
+& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_{i}) \frac{S(x_{i})}{1-p_{i}}, \nonumber \\
+& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} S(x_{i}).
+\end{align*}
+
+**Variance of a positive declaration**
+
+$$Var(D_j \vert D_j >0)  = E(D_j^2 \vert D_j >0)- E(D_j \vert D_j >0)^2.$$
+
+$$E(D_j^2 \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 1_{\left \lbrace D_j >0\right\rbrace}) = (1-\pi_j)^{-1} E(D_j^2 )$$
+
+$$E(D_j \vert D_j >0)^2 = ((1-\pi_j)^{-1} E(D_j 1_{\left \lbrace D_j >0\right\rbrace}))^2 =  (1-\pi_j)^{-2} E(D_j)^2$$
+
+Then,
+
+$$Var(D_j \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 ) - (1-\pi_j)^{-2} E(D_j)^2 = (1-\pi_j)^{-1} Var(D_j) - \frac{\pi_j}{(1-\pi_j)^2} E(D_j)^2.$$
+
+As the $(Y_{i})_{i\in \mathcal{P}_j}$ are independent, $Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_{i}) = \sum_{i \in\mathcal{P}_j} Var(C_{i}.Z_{i})$.
+
+\begin{align*}
+Var(C_{i} Z_{i}) & = E( C_{i}^2 Z_{i}^2 ) - E(C_{i} Z_{i})^2,\\
+& =  E( C_{i}^2) E(Z_{i}^2 ) -E(C_{i})^2 E(Z_{i})^2,\\
+& = (1-p_{i}) E(Z_{i}^2 ) - (1-p_{i})^2 E(Z_{i})^2,\\
+& = (1-p_{i})(Var(Z_{i})+E(Z_{i})^2)-(1-p_{i})^2E(Z_{i})^2,\\
+& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-1)+\frac{S(x_{i})^2}{1-p_{i}}-S(x_{i})^2,\\
+& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))
+\end{align*}
+
+**Sum up of the main formulas**
+
+$$P(D_j = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j$$
+
+$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
+
+$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
+
+$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{j}}(e^{\sigma^2}-(1-p_{i}))$$
+
+Assuming $D_j \vert D_j > 0$ also follows a lognormal distribution we can write:
+
+$$D_j \vert D_j > 0 \sim \operatorname{L}( \mu_j = E(D_j \vert D_j > 0), \sigma_j^2= ln(\frac{Var(D_j \vert D_j > 0)}{E(D_j \vert D_j > 0)^2} + 1))$$

### 2022-03-30 Version_M2_plusieurszones/draft/paper_3/paper_3.tex
paper_3

--- 2022-01-07 Version_M2_plusieurszones/draft/paper_3/paper_3.tex
@@
   english,
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{lmodern}
@@
 \keywords{keywords\newline\indent Word count: X}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
 \usepackage{lineno}
@@
 
-\hypertarget{model}{%
-\subsection{Model}\label{model}}
+We base our approach on the model developped by Alglave et al.
+(2022).
+In this framework, all observations were supposed to be realised at the same scale.
+Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and potentially combine these data with other (more accurate) datasets such as scientific data.
 
-See \texttt{3\_formulas.rmd}.
+\hypertarget{defining-the-problem-from-a-modelling-point-of-view}{%
+\subsection{Defining the problem from a modelling point of view}\label{defining-the-problem-from-a-modelling-point-of-view}}
+
+\textbf{Definition of the variables}
+
+Let's define the latent field \(S\) and the punctual observations \(Y\).
+\(S\) is a spatial log-Gaussian Field (GF) defined as \(\log(S(x))=\mu+\beta .
+\Gamma(x) + \delta(x)\) (Figure \ref{fig:SchemeRealloc}).
+We consider a discrete study domain where \(x \in [\![1,n]\!]\) stands for the discrete locations.
+\(\beta .
+\Gamma(x)\) is the covariate term with \(\Gamma\) the matrix of design of the covariates and \(\beta\) the effect of the covariates.
+\(\delta(x)\) is a GF capturing spatial correlation (\(\delta \sim \mathcal{N}(0,\Sigma)\)).
+
+All fishing locations \(x_i\) are supposed to be known through VMS data.
+At each fishing location \(x_i\), fishermen realize a catch \(Y_i\) conditionnally on the latent field values (referred hereafter as a punctual catch).
+Observations are supposed to follow some distribution \(\mathcal{L}_Y\).
+
+\[Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma^2)\]
+
+In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by Thorson (2018) and already used in Alglave et al.
+(2022).
+\(\mathcal{L}_Y\) is decomposed it 2 parts:
+
+\begin{itemize}
+\item
+  the probability to obtain a zero-catch which is modelled as a Bernoulli variable with probability \(p_{i}=\exp(-e^\xi .S(x_{i}))\).
+\(\xi\) is a parameter controlling zero-inflation.
+When \(\xi\) increases, the amount of zero in the data decreases.
+\item
+  if the catch is positive, the probability to obtain some observation value \(y_i\) is modelled through a positive continous distribution \(\operatorname{L}\) (here a lognormal distribution - its parameterization is described in SM)) with mean component \(\frac{S(x_{i})}{1-p_{i}}\) and variance term \(\sigma^2\).
+\end{itemize}
+
+\[
+\operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i)\right) =
+\left\{
+    \begin{array}{ll}
+        p_i & \text { if } y_{i}=0 \\
+        \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0
+    \end{array}
+\right.
+\]
+
+Fishermen must declare what they catch at the level of the statistical rectangle at a daily time step.
+These data are recorded in logbooks data.
+A declaration (denoted \(D_j\)) is a summation of all the \(Y_i\) realized on the fishing positions belonging to the catch declaration \(D_j\).
+They can be expressed as:
+
+\[D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}\]
+
+\(\mathcal{P}_j\) is the vector of the fishing observations belonging to the declaration \(D_j\).
+The punctual observations \(Y_i\) are indexed through \(i \in [\![1,m_j]\!]\) with \(m_j\) the number of fishing positions belonging to the \(j^{th}\) declaration.
+
+\textbf{Catch reallocation: uniform reallocation or model-based reallocation}
+
+In standard processing, \(D_j\) (available from logbooks) are reallocated uniformly on related \(x_i\) (available from VMS) so that derived punctual observations \(Y^*_i\) are computed as \(Y^*_i=D_j/m_j\).
+This is what we call uniform (or proportional) reallocation.
+In this case, inference of species distribution is directly computed through reallocated \(Y^*_i\) assuming these are the exact punctual observations.
+This strongly simplifies the actual process of observation and have most likely repercussion on model performance.
+
+To overcome such limitation, an alternative is to consider that only the catch declarations \(D_j\) are observed while \(Y_i\) are not (they are latent variables).
+Such way, we define some distribution \(\mathcal{L}_D\) as the distribution of the catch declarations:
+
+\[D_j | \mathcal{P}_j,S_{\mathcal{P}_j} \sim \mathcal{L}_D( S_{\mathcal{P}_j},\xi,\sigma^2)\]
+
+with \(S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))\), \(\xi\) the zero-inflation parmeter and \(\sigma^2\) the variance parameter.
+As \(D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}\), it is possible to relate \(D_j\), \(Y_i\) and \(S_{\mathcal{P}_j}\) through the distribution \(\mathcal{L}_Y\) and \(\mathcal{L}_D\).
+Our approach is to match the moments of \(D_j\) (obtained from the moments of \(\mathcal{L}_Y\)) and \(\mathcal{L}_D\) and then to infer \(S(x)\) and \(Y_i\) from \(D_j\).
+This is what we call model-based reallocation.
+This model will be referred as the \(D_j\) model.
+In contrast, the model fitted on reallocated observations \(Y^*_i\) will be referred as the \(Y^*_i\) model.
+
+In the following, \(Y\) and \(D\) will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as \(Y_i\) or \(D_j\) instead of \(Y_i \vert x_i, S(x_i)\) and \(D_j | \mathcal{P}_j,S_{\mathcal{P}_j}\).
+
+\textbf{Matching \(\mathcal{L}_Y\) and \(\mathcal{L}_D\)}
+
+First, as at the punctual level (\(Y\)), the declarations can be decomposed in 2 components: (1) the probability to obtain a zero-value declaration and, if the declaration is positive, (2) the probability to obtain some declaration value \(d_j\).
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\tightlist
+\item
+  We derive the probability to obtain a zero-declaration \(P(D_j = 0)\) by simply multiplying the probability to obtain a zero-punctual observation \(P(Y_i=0 )\) to all fishing points \(x_i \in \mathcal{P}_j\).
+\end{enumerate}
+
+\begin{align*}
+P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
+\end{align*}
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  If the declaration is positive, we assume the probability to obtain catch \(d_j\) follows a continous distribution \(\mathcal{L}_{D_j|D_j>0}\) with mean component \(\mu_j\) and variance component \(\sigma_j\).
+From \(\mathcal{L}_Y\) we can derive \(E(D_j \vert D_j > 0)\) and \(Var(D_j \vert D_j > 0)\) (given below) and re-express \(\mu_j\) and \(\sigma_j\) as a function of these two moments.
+\end{enumerate}
+
+\[E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}\]
+\[Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2\]
+\[Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))\]
+
+If we consider \(\mathcal{L}_{D_j|D_j>0}\) is lognormal (with parameterization described in SM) then, \(\mu_j = E(D|D_j>0)\) and \(\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)\).
+Such way, by fitting the model on declaration \(D_j\) rather than reallocated catch \(Y_i\) and linking \(\mathcal{L}_D\) to \(\mathcal{L}_Y\), we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation.
+All calculations to obtain these formulas are available in SM.
+
+The inference is realized through Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation (Kristensen, Nielsen, Berg, Skaug, \& Bell, 2016).
+
+\begin{figure}
+\centering
+\includegraphics{images/realloc.png}
+\caption{\label{fig:SchemeRealloc} Schematic representation of the reallocation process.}
+\end{figure}
 
@@
 
-\hypertarget{single-square-simulations}{%
-\subsubsection{Single-square simulations}\label{single-square-simulations}}
+To evaluate the impact of reallocation on model outputs and how our alternative approach can improve model accuracy, we conduct simulation-estimation.
 
-How many simulations? 100 or more?
+First, we conduct basic simulations in order to explore the effect of some key variables on model performance and also the base properties of the alternative models.
+We assume the study domain is a single statistical rectangle, only commercial data feed the model and the expression of the latent field in both simulation and estimation is simplified (the variability only depends on one covariate, not on a spatial random term).
+These simulations will be referred as \enquote{single-square simulations}.
 
-How do we simulate the covariate?
+Then, we extend the simulation-estimation study to several rectangles to get closer to a real case study configuration.
+We add a spatial random effect in the latent field (for both simulation and estimation) and we also simulate scientific data (in addition to commercial data) to explore the contribution of both datasets in inference.
+These simulations will be referred as multiple-square simulations.
 
-Still some lack of convergence.
+In these two sets of simulation-estimation, the covariate is modelled as a continous GRF that we suppose known at each point of the grid.
+The covariate effect is fixed to \(\beta_S=2\) and the intercept is also fixed to \(\mu=2\).
+Regarding commercial data, the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for a single declaration in real data.
+All parameterizations are detailed in the Table 1.
 
-Number of pings \textbar{} nu:
-1\textbar10
-10\textbar100
-100\textbar1000
+Each time we compare several model configurations and evaluate these in regards to 2 metrics:
 
-Size within a fishing sequence: radius = 3
+\begin{itemize}
+\tightlist
+\item
+  the MSPE which quantifies the accuracy of the spatial predictions of the latent field over the spatial domain (\(n\) is the number of locations over the grid).
+\end{itemize}
 
-Number of fishing zones: 1\textbar3\textbar5
+\[MSPE=\frac{\sum_x (S(x) - \hat{S}(x) ) ^2}{n}\]
 
-\hypertarget{multiple-squares-simulations}{%
-\subsubsection{Multiple-squares simulations}\label{multiple-squares-simulations}}
+\begin{itemize}
+\tightlist
+\item
+  the parameter of the species habitat relationship \(\beta_S\) compared with its estimate \(\hat \beta_S\).
+\end{itemize}
 
-How do we simulate the covariate?
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-2}Parameter values for the simulations}
+\centering
+\begin{tabular}[t]{lcc}
+\toprule
+\text{Parameters} & \text{Single-square simulations} & \text{Multiple-square simulations}\\
+\midrule
+$\mu$ & 2 & 2\\
+$\beta_S$ & 2 & 2\\
+$\text{Range}$ & -- & 1.5\\
+$\text{Marginal variance}$ & -- & 0.5\\
+$\xi_{com}$ & -1 & -1\\
+\addlinespace
+$\sigma_{com}$ & 1 & 1\\
+$k_{com}$ & -- & 1\\
+$\xi_{sci}$ & -- & 0\\
+$\sigma_{sci}$ & -- & 0.8\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\textbf{Single-square simulations}
+
+Two important variables may affect the accuracy of model outputs: the amount of commercial data and the number of fishing zones explored within one single declaration.
+In these simulations-estimations, we intend to explore the effect of these 2 variables.
+
+First, the amount of commercial data; it is expected that when increasing the amount of data the accuracy of the predictions will be improved.
+In simulations, we progressively increase the number of fishing positions (10, 100 and 1000 fishing pings).
+As there are 10 pings per declaration, this means that the number of declarations increases respectively (1, 10 and 100 declarations).
+
+Furthermore, the samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle).
+Note that we make a distinction between fishing zones and statistical rectangles: fishing zones are included in a statistical rectangle and there can be several fishing zones within a statistical rectangles.
+Consequently, a declaration can mix catches that have been realized in zones of high biomass and in zones of low biomass.
+For instance, a declaration can aggregate data from several fishing operations that occured the same day within the same statistical rectangle, but that were realized on 2 distinct types of fishing grounds (and then on two distinct habitats).
+Reallocating uniformly the declaration will strongly homogeneise the actual catch, will mix up information from to different grounds and then may lead to a strong loss of accuracy in model outputs.
+It is expected that when the number of fishing zones within a declaration increases, the accuracy of the model outputs decreases.
+To assess the effect of such process, we simulated the pings of a fishing declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones.
+
+Concretely, for each fishing declaration the fishing pings are sampled in 2 steps: first the centroid of the zones are randomly sampled over the statistical rectangle (here the simulation domain) and then the fishing positions are randomly sampled within the radius of the centroid of the zones.
+The zone size was chosen to be a square of side 7.
+This is a realistic order of magnitude if we consider each zone is one fishing operation and that for each fishing operation the distance of the operation will not exceed 30km.
+
+At each fishing position, a catch is realized conditionnally on the value of the latent field and follow the probability distribution described above (\(\mathcal{L}_y\)).
+
+We compare three simulation/estimation configurations:
+
+\begin{itemize}
+\item
+  a golden standard configuration: punctual observations (\(Y_i\)) are known exactly and the punctual-data model is fitted to the exact data.
+\item
+  a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations.
+We then fit the \(Y_i^*\) model (the model fitted to punctual reallocated data).
+\item
+  a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the \(D_j\) model to the data (i.e.~the one fitted at the level of the declarations data).
+\end{itemize}
+
+In addition to the metrics we introduced to compare model configuration (\(MSPE\) and species-habitat parameter \(\beta_S\)), we also assess the simulated/estimated values for the intercept \(\mu\), the observation variance parameter \(\sigma^2\) and the zero-inflation parameter \(\xi\).
+
+\textbf{Multiple-square simulations}
+
+In this case, we extend the study to more than one single rectangle, we simulate scientific data and include these in inference.
+Finally, in addition to the covariate, we also simulate/estimate a spatial random effect in the latent field as a Matérn GRF.
+
+The study area is based on the case study; it includes the whole coast of the Bay of Biscay and cover several statistical rectangles (Figure \ref{fig:MapSeveral}A).
+To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration).
+Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area.
+Similarly to the single-square simulations, the sampling of the commercial points of a declaration are realized in 3 steps.
+(1) The declaration is randomly affected to one of the ICES rectangle.
+(2) The centroid of the fishing zone is randomly sampled within this statistical rectangle.
+(3) The 10 fishing punctual observations are randomly sampled within the fishing zone.
+The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km.
+Note that we do not explore the effect of exploring several zones within the same declaration as it is already done in the single-square simulations.
+
+100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \ref{fig:MapSeveral}A).
+Scientific observations are simulated following the observation equations of \(\mathcal{L}_Y\) (with specific parameters for scientific data - Table 1).
+
+We compare several model configurations:
+
+\begin{itemize}
+\item
+  to assess what brings our alternative approach, we compare the \(Y^*_i\) model and the \(D_j\) model.
+\item
+  to assess the information brought by each data source, we compare models built on scientific data only (scientific-based models), models built on commercial data only (commercial-based models) and models combining both data sources (integrated models).
+\end{itemize}
+
+In addition to the 2 metrics introduced at the beginning of the section (\(MSPE\) and species-habitat parameter \(\beta_S\)), we also compare the simulated/estimated values of the range parameter.
+
+Note that when including 2 datasets in the model, a difference in catchability should be accounted for and one of the two dataset will be considered as reference (see Alglave et al.
+(2022) for more details on parameterization).
+Here the reference level is set with scientific data and the parameter \(k_{com}\) accounts for the relative ratio between scientific and commercial observation values.
+In simulations, \(k_{com}\) is set to 1 so that both datasources are assumed to have the same catchability.
 
@@
 
+To illustrate our method on a real case study, we applied the approach to the common sole of the Bay of Biscay.
+VMS-logbooks data were extracted for the bottom trawlers fleet (OTB).
+The methods to cross VMS-logbooks data and to filter the fleet is already extensively described in the previous papers (Alglave et al., 2022) and is not developped further here.
+Scientific data were extracted from the DATRAS database for the Orhago beam trawl survey (Gérard, 2003; ICES, 2018b).
+To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - ICES (2018a)).
+To illustrate the method, we compare the outputs of (1) the scientific-based model, (2) the integrated model fitted to scientific data and reallocated commercial data (\(Y^*_i\) model) and (3) the integrated model fitted to scientific and declarations data (\(D_j\) model).
+
+The \(D_j\) model faced convergence issues (some of the parameters were hardly estimated e.g.~the range parameter).
+To ease convergence, we integrated in the analysis onboard observer data for the same fleet.
+They are considered as exact commercial data of catches.
+Integrating these data allow to have direct information on \(Y_i\) and to better estimate the observation equations parameters (i.e.~observation variance and zero-inflation parameter of commercial data).
+
+Furthermore, we adopt a step estimation procedure to initialize the optimization algorithm for the \(D_j\) model.
+We first fit the \(Y^*_i\) model and use the estimates of the model to initialize the optimization of the \(D_j\) model.
+We eventually fix the parameters that are hard to estimate in first steps of optimization (intercept \(\mu\), covariate effect \(\beta_S\), range and marginal variance) and let them free iteratively.
+
 \hypertarget{results}{%
@@
 
-\hypertarget{single-square-analysis}{%
-\subsection{Single square analysis}\label{single-square-analysis}}
+\hypertarget{single-square-simulations}{%
+\subsection{Single-square simulations}\label{single-square-simulations}}
+
+Reallocation has a major effect on predictions and estimates accuracy (Figure \ref{fig:PerfMetricSingle}).
+Reallocating data conduct to a 10 to 200 times decrease in accuracy for spatial predictions when fitting the model to reallocated catch (MSPE gold compared to red boxplots).
+Accuracy decreases as the number of visited zones within a declaration increases.
+Besides, \(\beta_S\) estimates are biased and reallocation leads to the loss of the species-habitat relationship as the number of fishing zones (within a declaration) increases (\(\beta_S\) estimates get closer to 0).
+Increasing the number of samples does not improve inference.
+Regarding other parameters estimates, the zero-inflation parameter (\(\xi\)) is over-estimated (i.e.~when uniformly reallocating commercial data, the quantity of positive observations is under-estimated), the observation variance (\(\sigma\)) is underestimated (i.e.~the data is estimated to be less noisy than they actually are) and the intercept of the latent field (\(\mu\)) is slightly over-estimated (Figure \ref{fig:ParBiasSingle}).
+
+The \(D_j\) model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \ref{fig:PerfMetricSingle}).
+Still, the \(D_j\) model outputs are not as accurate as the ones of the golden-standard.
+Furthermore, the zero-inflation parameter is unbiased when the model is fitted to catch declarations \(D_j\).
+Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \ref{fig:ParBiasSingle}).
+This alternative model have some convergence difficulties (Table 2) as 8\% of the model did not converged when sample size is medium (100 pings) and only 3\% did not when sample size is large (1000 pings).
+
+\begin{figure}
+\centering
+\includegraphics[width=0.75\textwidth,height=\textheight]{images/Perf.metric_single_square.png}
+\caption{\label{fig:PerfMetricSingle} Performance metric for single-square simulations.
+Columns: number of commercial fishing positions.
+x-axis: number of zones visited within each declaration.
+\enquote{Reallocation:}, data are or are not reallocated in simulations.
+\enquote{Likelihood:}, the likelihood is computed on exact punctual observations \(Y_i\), on reallocated observations \(Y_i^*\) or on catch declarations \(D_j\).
+Gold: golden standard.
+Red: uniform reallocation (\(Y_i^*\) model).
+Green: model-based reallocation (\(D_j\) model).
+Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the \(D_j\) model (see Table 2).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{images/par_plot_single_square.png}
+\caption{\label{fig:ParBiasSingle} Parameters relative bias for single-square simulations.
+\enquote{Reallocation:}, data are or are not reallocated in simulations.
+\enquote{Likelihood:}, the likelihood is computed on exact punctual observations \(Y_i\), on reallocated observations \(Y_i^*\) or on catch declarations \(D_j\).
+Gold: golden standard.
+Red: uniform reallocation (\(Y_i^*\) model).
+Green: model-based reallocation (\(D_j\) model).
+Only the simulations with 1000 fishing positionss are represented.
+Black line: zero value.
+Red line: parameter true value.}
+\end{figure}
 
@@
 
-\caption{\label{tab:unnamed-chunk-1}Percentage of convergence for each simulation/model configuration}
+\caption{\label{tab:unnamed-chunk-5}Single-square simulations - Percentage of convergence per simulation-estimation configuration.}
 \centering
@@
 \toprule
-Nb samples & Fishing sequence & Reallocation & Likelihood level & Convergence (\%)\\
+Fishing positions & Declarations & Reallocation & Likelihood level & Convergence (\%)\\
 \midrule
 10 & 1 & No & Yi & 99.668\\
-10 & 1 & Yes & Yi & 0.333\\
+10 & 1 & Yes & Yi* & 0.333\\
 10 & 1 & Yes & Dj & 0.000\\
 100 & 10 & No & Yi & 100.000\\
-100 & 10 & Yes & Yi & 100.000\\
+100 & 10 & Yes & Yi* & 100.000\\
 \addlinespace
@@
 1000 & 100 & No & Yi & 100.000\\
-1000 & 100 & Yes & Yi & 100.000\\
+1000 & 100 & Yes & Yi* & 100.000\\
 1000 & 100 & Yes & Dj & 97.333\\
@@
 \end{table}
+
+\newpage
+
+\hypertarget{multiple-square-simulations}{%
+\subsection{Multiple-square simulations}\label{multiple-square-simulations}}
+
+Whether the model is fitted at the punctual level or at the declaration level, the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models.
+This can be well illustrated from Figure \ref{fig:MapSeveral}.
+Integrating scientific and commercial data allows to (1) capture the hotspot missed by commercial data through scientific data and (2) better capture the local correlation structures through the dense commercial data.
+
+Furthermore, consistently with single-square simulations, uniform reallocation conducts to a loss in both the predictions accuracy and the species-habitat relationship (Figure \ref{fig:PerfMetricSeveral}) compared to the model built on commercial declarations (\(D_j\)).
+
+Interestingly, in addition to the species-habitat relationship, uniform reallocation also affects the spatial autocorrelation terms such as the range parameter.
+The model fitted on reallocated data (\(Y_i^*\)) provides biased range estimates while the \(D_j\) model provides unbiased estimates.
+This is a consequence of the loss of the species-habitat relationship: when uniformly reallocating declarations, part of the variability related to the covariate effect is captured by the random effect.
+Consequently, the range parameter captures both the autocorrelation related to the actual random effect and to the covariate.
+The \(D_j\) model allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
+This is evidenced in Figure \ref{fig:MapSeveral} where the model fitted to reallocated catch \(Y_i^*\) provides smoothed maps and does not capture the relatively small scale patterns that are shaped by the covariate.
+On the other hand, the \(D_j\) model (as the scientific-based model) better captures and disentangles the covariate effect and the spatial random effect and then provides predictions that better fit to the small-scale patterns of the species distribution.
+However, this goes with some difficulty in convergence as only 75\% of the model built on catch declarations converge (Table 3).
+
+\begin{figure}
+\centering
+\includegraphics{images/Perf.metric_multiple_square.png}
+\caption{\label{fig:PerfMetricSeveral} Performance metric for the multiple-square simulations.
+Red line: true value for the range and the species-habitat parameter (\(\beta_S\)).
+Red: uniform reallocation (\(Y_i^*\) model).
+Green: model-based reallocation (\(D_j\) model).
+Blue: scientific-based model.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{images/Map_multi_square.png}
+\caption{\label{fig:MapSeveral} Relative distribution of simulated/estimated biomass field.
+A: Simulated biomass field with scientific samples (red) and statistical rectangles.
+The rectangles that have not been sampled by commercial data are the transparent rectangles.
+They represent 1/3 of the full area.
+B: simulated biomass field.
+C: biomass field from the scientific-based model.
+\(Y_i^*\): \(Y_i^*\) model (D, E).
+\(D_j\): \(D_j\) model (F, G).
+Scientific model: model fitted to scientific data only.
+Commercial model: model fitted to commercial data only.
+Integratedd model: model fitted to both data sources.}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-8}Multiple-square simulations - Percentage of convergence per simulation-estimation configuration.}
+\centering
+\begin{tabular}[t]{ccc}
+\toprule
+Model & Likelihood level & Convergence (\%)\\
+\midrule
+Commercial model & Yi* & 100.000\\
+Commercial model & Dj & 75.377\\
+Integrated model & Yi* & 100.000\\
+Integrated model & Dj & 76.382\\
+Scientific model &  & 100.000\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{real-case-study}{%
+\subsection{Real case study}\label{real-case-study}}
+
+Consistently with simulations, the \(D_j\) model modifies some of the parameters estimates and revises the spatial pattern of the species distribution compared with the \(Y^*_i\) model.
+In particular, the substrate effect is recovered in the \(D_j\) model and fall in the same range as the scientific-based estimate (Figure \ref{fig:CaseStudyMap}).
+The zero-inflation parameter \(\xi\) is revised downwards (i.e.~there are actually more zero-values than in the reallocated data) while the observation variance of commercial data is revised upwards (i.e.~the commercial data are noisier than expected when fitting the model on \(Y_i^*\)).
+
+In addition, uncertainty is also revised when fitting the model at the declaration level.
+For instance, when comparing the \(Y^*_i\) model to the \(D_j\) model, the confidence intervals of \(\beta_S\), the marginal variance, the range, \(\xi_{com}\), \(\sigma_{com}\) are much wider.
+This emphasizes that uncertainty is probably underestimated in the \(Y^*_i\) model compared with the \(D_j\) model.
+Now, when comparing the scientific-based model and the integrated \(D_j\) model, some parameters are better estimated than when only scientific data feed the model.
+For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on \(D_j\) substrate is significant and the confidence interval is smaller.
+
+On the contrary, other parameters do not seems be well estimated in either the \(Y_i^*\) or the \(D_j\) models.
+For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on \(Y_i\) and revised downards when working on \(D_j\).
+This is consistent with simulations results, see Figure \ref{fig:ParBiasSingle}.
+
+Regarding the maps of the species distribution, fitting the model at the \(D_j\) level strongly modifies the model biomass field compared with the \(Y_i^*\) model.
+In particular, the substrate covariate have a sharper effect on species distribution and the intensity of the hotspots are revised when fitting the model on \(D_j\).
+
+Finally, the \(D_j\) model fitted only on commercial data does not converge (while the one fitted on \(Y_i^*\) does) emphasizing the model fitted on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
+
+\begin{figure}
+\centering
+\includegraphics{images/par_plot.png}
+\caption{\label{fig:CaseStudyPar} Parameters obtained with scientific-based model, the integrated model fitted on reallocated catch \(Y_i^*\) and the integrated model fitted on catch declarations \(D_j\).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{images/case_study_plot.png}
+\caption{\label{fig:CaseStudyMap} Maps obtained from (left) the scientific-based model, (center) the integrated model fitted on punctual reallocated catch \(Y_i^*\), (right) the integrated model fitted on catch declarations \(D_j\).}
+\end{figure}
 
@@
 
+\hypertarget{supplementary-material}{%
+\section{Supplementary material}\label{supplementary-material}}
+
+\textbf{Notations}
+
+We model catch declarations \(D_j\) (available at coarse resolution through logbooks data) as a sum of \(Y_{i}\) punctual observations (which are unknown i.e.~latent variables) each one realised at one fishing position \(x_{i}\) (known through VMS data).
+
+We note:
+
+\begin{itemize}
+\item
+  \(\mathcal{P}_j\) : the vector of all fishing positions related to the \(j^{th}\) declaration.
+\item
+  \(j \in \{1,...,l\}\) with \(l\) the number of declarations.
+\item
+  \(i \in \{1,...,m_j\}\) with \(m_j\) the number of fishing positions belonging to the \(j^{th}\) declaration.
+\end{itemize}
+
+\[D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}\]
+
+\textbf{Reparameterization of the lognormal distribution}
+
+The lognormal distribution is usually written as \(Z \sim \operatorname{L}(\rho;\sigma^2)\) with \(Z=e^{\rho+\sigma N}\) and \(N \sim \mathcal{N}(0,1)\).
+In this case, \(E(Z)=e^{\rho + \frac{\sigma^2}{2}}\) and \(Var(Z)=(e^{\sigma^2}-1)e^{2 \rho + \sigma^2}\).
+
+We choose to slightly reparameterize the lognormal distribution. Let's define \(\rho = ln(\mu) - \frac{sigma^2}{2}\), then:
+
+\begin{itemize}
+\item
+  \(Z=\mu e^{\sigma N - \frac{sigma^2}{2}}\)
+\item
+  \(E(Z)=\mu\)
+\item
+  \(Var(Z)=\mu^2(e^{\sigma^2} - 1) \Leftrightarrow \sigma^2=ln(\frac{Var(Z)}{E(Z)^2} + 1)\)
+\end{itemize}
+
+\textbf{\(D_j\) probability distribution and moments}
+
+We have to express the probability distribution of \(D_j\) and its moments as a function of \(Y_{i}\) and its related moments.
+Let's assume \(Y_{i} = C_{i}.Z_{i}\) is a zero-inflated lognormal distribution with \(C_{i}\) a binary random variable and \(Z_{i}\) a lognormal random variable.
+
+\[C_{i} \sim \mathcal{B}(1-p_{i})\]
+with \(p_{i}=exp(-e^\xi .S(x_{i}))\) the probability to obtain a zero value.
+
+\[Z_{i} \sim \operatorname{L}(\frac{S(x_{i})}{1-p_{i}},\sigma^2)\]
+
+Here, \(Y_{i}\), \(C_{i}\) and \(Z_{i}\) are observations of a latent field \(S(x_{i})\) at a sampled point \(x_{i}\).
+
+\textbf{Probability of obtaining a zero declaration}
+
+\begin{align*}
+P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0 \vert S, X),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
+\end{align*}
+
+\textbf{Expectancy of a positive declaration}
+
+Following calculations are supposed to be conditionnal on \(S\) and \(X\).
+
+\[E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}\vert \exists i\in\mathcal{P}_j , C_{i}=1)\]
+
+\begin{align*}
+E(D_j\vert D_j > 0) & = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / P\left ( D_j > 0\right ),   \\
+& = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / \left (1-\pi_j\right).   \\
+\end{align*}
+
+As \(E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } ) = E(D_j ),\)
+
+\begin{align*}
+E(D_j\vert D_j > 0) & = \left (1- \pi_j\right)^{-1} E(D_j)   ,   \nonumber \\
+& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}),\nonumber \\
+& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_{i}) \frac{S(x_{i})}{1-p_{i}}, \nonumber \\
+& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} S(x_{i}).
+\end{align*}
+
+\textbf{Variance of a positive declaration}
+
+\[Var(D_j \vert D_j >0)  = E(D_j^2 \vert D_j >0)- E(D_j \vert D_j >0)^2.\]
+
+\[E(D_j^2 \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 1_{\left \lbrace D_j >0\right\rbrace}) = (1-\pi_j)^{-1} E(D_j^2 )\]
+
+\[E(D_j \vert D_j >0)^2 = ((1-\pi_j)^{-1} E(D_j 1_{\left \lbrace D_j >0\right\rbrace}))^2 =  (1-\pi_j)^{-2} E(D_j)^2\]
+
+Then,
+
+\[Var(D_j \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 ) - (1-\pi_j)^{-2} E(D_j)^2 = (1-\pi_j)^{-1} Var(D_j) - \frac{\pi_j}{(1-\pi_j)^2} E(D_j)^2.\]
+
+As the \((Y_{i})_{i\in \mathcal{P}_j}\) are independent, \(Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_{i}) = \sum_{i \in\mathcal{P}_j} Var(C_{i}.Z_{i})\).
+
+\begin{align*}
+Var(C_{i} Z_{i}) & = E( C_{i}^2 Z_{i}^2 ) - E(C_{i} Z_{i})^2,\\
+& =  E( C_{i}^2) E(Z_{i}^2 ) -E(C_{i})^2 E(Z_{i})^2,\\
+& = (1-p_{i}) E(Z_{i}^2 ) - (1-p_{i})^2 E(Z_{i})^2,\\
+& = (1-p_{i})(Var(Z_{i})+E(Z_{i})^2)-(1-p_{i})^2E(Z_{i})^2,\\
+& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-1)+\frac{S(x_{i})^2}{1-p_{i}}-S(x_{i})^2,\\
+& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))
+\end{align*}
+
+\textbf{Sum up of the main formulas}
+
+\[P(D_j = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j\]
+
+\[E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}\]
+
+\[Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2\]
+
+\[Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{j}}(e^{\sigma^2}-(1-p_{i}))\]
+
+Assuming \(D_j \vert D_j > 0\) also follows a lognormal distribution we can write:
+
+\[D_j \vert D_j > 0 \sim \operatorname{L}( \mu_j = E(D_j \vert D_j > 0), \sigma_j^2= ln(\frac{Var(D_j \vert D_j > 0)}{E(D_j \vert D_j > 0)^2} + 1))\]
+
 \end{document}

### 2022-03-30 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_2

@@
 We base our approach on the model developped by @alglave_combining_2022.
-In this framework, all observations were supposed to be realised at the same scale.
-Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and potentially combine these data with other (more accurate) datasets such as scientific data.
+In this framework, all observations were supposed to be realised at a fine scale.
+Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and combine these data with other fine scale datasets such as scientific data.
 
@@
 
-![(\#fig:SchemeRealloc) Schematic representation of the reallocation process.](images/realloc.png)
+![(\#fig:SchemeRealloc) Schematic representation of the reallocation process.
+The biomass field (the background field) depends on a covariate and a spatial random effect.
+The covariate is the $x$ axis.
+It has a positive effect on biomass values (i.e.
+biomass is higher on the right of the grid than on the left).
+The spatial random effect conduct to a hotspot on the bottom-right of the latent field.
+The study domain is considered as a statistical rectangle (grey square).
+Fishermen sample catches in areas of poor biomass where the covariate is relatively low (blue points) and in areas of higher biomass where the covariate is higher and eventually in the hotspot of biomass (orange and red points).
+These catches belong to the same declaration $j$ and are summed to constitute the declaration $D_j=50$.
+The declaration is declared at the level of the statistical rectangle.
+From VMS data, we know the fishing positions $x_i$.
+In standard processing, $D_j$ are then uniformly reallocated over the fishing positions $x_i$.
+This strongly homogeneizes the catch.
+In particular, the effect of the habitat is no more evidenced in the reallocated catch $Y_i^*$.](images/realloc.png)
 
@@
 We add a spatial random effect in the latent field (for both simulation and estimation) and we also simulate scientific data (in addition to commercial data) to explore the contribution of both datasets in inference.
-These simulations will be referred as multiple-square simulations.
+These simulations will be referred as "multiple-square simulations".
 

### 2022-04-02 Version_M2_plusieurszones/pres/GDREcostat_April2022/GDREcostat_2022.Rmd
gdr

@@
 subtitle : "Going from coarse landings data to fine scale species distribution"
-author : "Baptiste Alglave,  Marie-Pierre Etienne, Youen Vermard,  Mathieu Woillez, Etienne Rivot"
+author : "Baptiste Alglave, Marie-Pierre Etienne, Kasper Kristensen, Youen Vermard,  Mathieu Woillez, Etienne Rivot"
 
@@
 Access to more data \\
-Exact locations usually available
+Exact locations available
 \end{column}
@@
 \normalsize
-Low amount of samples
+Small sample size
 \end{column}
@@
 \onslide<2>{
-\footnotesize
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Some specific proposition in the context of fishery science: \\
-\tiny
-Alglave Baptiste, Rivot Etienne, Etienne Marie-Pierre, Woillez Mathieu, Thorson James T, Vermard Youen (2022).
+\normalsize
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Some specific application in the context of fishery science: \\
+\scriptsize
+\textit{Alglave Baptiste, Rivot Etienne, Etienne Marie-Pierre, Woillez Mathieu, Thorson James T, Vermard Youen (2022).
 \textbf{Combining scientific survey and commercial catch data to map fish distribution.} ICES Journal of Marine Science IN PRESS.
-\url{https://doi.org/10.1093/icesjms/fsac032}
+\url{https://doi.org/10.1093/icesjms/fsac032}}
 \vspace{\baselineskip}
@@
 
-\vspace{\baselineskip}
-
 \begin{columns}
@@
 \begin{column}{0.1\textwidth}
-\center \tiny {\bf Resolution}
+\center \scriptsize {\bf Spatial \\ resolution}
 \end{column}
 \begin{column}{0.45\textwidth}
-\center \tiny Catch are daily declared at the resolution of ICES rectangles
+\center \scriptsize Catch are daily declared at the resolution of ICES rectangles
 \end{column}
 \begin{column}{0.45\textwidth}
-\center \tiny VMS pings are vessels GPS locations emitted each hour
+\center \scriptsize VMS pings are vessels GPS locations emitted each hour
 \end{column}
@@
 
-### Modelling change of support
+### Two alternative procedures to reallocate catches
 
@@
   \begin{tabular}{ M{2cm} M{4cm} M{4cm} }
-  \textcolor{BaptisteRed}{\textbf{Actual situation}} & 
-  $Y_i | \textcolor{BaptisteBlue}{S}(x_i), x_i \sim \mathcal{L}_Y(\textcolor{BaptisteBlue}{S}(x_i),\xi,\sigma)$          &
-  $Y_i=\textcolor{BaptisteRed}{Y_i^*}$ \\
+  \textcolor{BaptisteRed}{\textbf{Current situation}} & 
+  $Y_i | \textcolor{BaptisteBlue}{S}(x_i), x_i \sim \mathcal{L}_Y(\textcolor{BaptisteBlue}{S}(x_i),\xi,\sigma^2)$          &
+  $Y_i=\frac{\textcolor{BaptisteLightGreen}{D_j}}{ \operatorname{n}(\textcolor{BaptisteLightGreen}{\mathcal{P}_j})}=\textcolor{BaptisteRed}{Y_i^*}$ \\
   \end{tabular}
@@
   $\textcolor{BaptisteLightGreen}{D_j}=\sum_{\textcolor{BaptisteLightGreen}{i \in \mathcal{P}_j}}{Y_{i}}$          &
-  $\textcolor{BaptisteLightGreen}{D_j} | \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\textcolor{BaptisteLightGreen}{\mathcal{P}_j} \sim \mathcal{L}_D( \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\xi,\sigma)$
-  & \textcolor{BaptisteBlue}{\ding{224}}  Match $\mathcal{L}_D$ and $\mathcal{L}_Y$ moments \\
+  $\textcolor{BaptisteLightGreen}{D_j} | \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\textcolor{BaptisteLightGreen}{\mathcal{P}_j} \sim \mathcal{L}_D( \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\xi,\sigma^2)$
+  & \textcolor{BaptisteBlue}{\ding{224}}  Match $\mathcal{L}_D$ and $\mathcal{L}_Y$ moments \tiny (appendix) \\ 
   \end{tabular}
@@
 \center
-\textbf{Simulation/estimation}
+\textbf{Simulation-estimation}
 
@@
 \begin{column}{0.55\textwidth}
-\scriptsize
+\footnotesize
 \textbf{Simulation}
@@
 \begin{itemize}
-\item \textcolor{BaptisteBlue}{\textbf{Latent field}} (\textcolor{BaptisteOrange}{covariate} + \textcolor{BaptisteBrown}{spatial random effect})
-\item \textbf{Commercial data} (3000 samples over 2/3 of the area)
-\item \textcolor{BaptisteLightGreen}{\textbf{Reallocation process}} (10 locations per declaration)
-\item \textcolor{SurveyBlue}{\textbf{Scientific data}} (100 samples over the whole the area)
+\item \scriptsize \textcolor{BaptisteBlue}{\textbf{Latent field}} \tiny (\textcolor{BaptisteOrange}{covariate} + \textcolor{BaptisteBrown}{spatial random effect})
+\item \scriptsize \textbf{Commercial data} \tiny (3000 samples over 2/3 of the area)
+\item \scriptsize \textcolor{BaptisteLightGreen}{\textbf{Reallocation process}} \tiny (10 locations per declaration)
+\item \scriptsize \textcolor{SurveyBlue}{\textbf{Scientific data}} \tiny (100 samples over the whole the area)
 \end{itemize}
@@
 \onslide<2-4>{
-
-\textbf{Estimation} \\
+\footnotesize
+\textbf{Estimation} \scriptsize \\
 Comparison of 3 model configurations: \\
 \tiny \vspace{\baselineskip} \scriptsize
-1/ Model fitted to \textcolor{SurveyBlue}{scientific data} (reference data) \\
+1/ Model fitted to \textbf{\textcolor{SurveyBlue}{scientific data}} only \\
 \tiny \vspace{\baselineskip} \scriptsize
-2/ Integrated model with \textcolor{BaptisteRed}{commercial likelihood built on $Y_i^*$} \\
+2/ \textcolor{BaptisteDarkGrey}{\textbf{Integrated model}} (= \textbf{\textcolor{SurveyBlue}{scientific}} + \textbf{commercial data}) with \textcolor{BaptisteRed}{commercial likelihood built on $\bf{Y_i^*}$} \\
 \tiny \vspace{\baselineskip} \scriptsize
-3/ Integrated model with \textcolor{BaptisteLightGreen}{commercial likelihood built on $D_j$} \\
+3/ \textcolor{BaptisteDarkGrey}{\textbf{Integrated model}} with \textcolor{BaptisteLightGreen}{commercial likelihood built on $\bf{D_j}$} \\
 \tiny \vspace{\baselineskip}
@@
 
-\scriptsize
-\textbf{Model evaluation} \\
+\footnotesize
+\textbf{Model evaluation} \scriptsize \\
 1/ Mean square prediction error \\
-$$MSPE=\frac{\sum_x (\textcolor{BaptisteBlue}{S(x)} - \textcolor{BaptisteBlue}{\hat{S}(x)} ) ^2}{n}$$
+$$MSPE=\frac{\sum_{x=1}^n (\textcolor{BaptisteBlue}{S(x)} - \textcolor{BaptisteBlue}{\hat{S}(x)} ) ^2}{n}$$
 
@@
 $$\textcolor{BaptisteOrange}{\beta_S} = 2 \, \, \, \text{versus} \, \, \, \textcolor{BaptisteOrange}{\hat{\beta}_S}$$
+\vspace{\baselineskip}
 
@@
 \onslide<4>{
-
-\vspace{\baselineskip}
 
@@
 \includegraphics[width=2cm]{images/solea_solea.png}
+\vspace{\baselineskip}
+\vspace{\baselineskip}
 \end{column}
@@
 \tiny \textcolor{SurveyBlue}{\textbf{Survey data:}} Orhago\\
-\textbf{Commercial data:} OTB-DEF traulers (to ease convergence onboard observer data were integrated in fitting)\\
-\textbf{Fitted models:} same as simulation\\
+\textbf{Commercial data:} OTB-DEF trawlers (to ease convergence onboard observer data were integrated in the fit)\\
+\textbf{Fitted models:} same as simulations\\
 \textcolor{BaptisteOrange}{\textbf{Covariate:}} substrate
+
+\normalsize \vspace{\baselineskip}
 
@@
   Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
-  Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi"
+  Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi*"
   Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific model"
-  Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi","Dj","Scientific model"))
+  Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi*","Dj","Scientific model"))
 
@@
     scale_color_manual(breaks = c("Scientific model",
-                                  "Yi",
+                                  "Yi*",
                                   "Dj"),
@@
     scale_color_distiller(palette = "Spectral")+
-    ggtitle("Commercial likelihood on Yi")+
+    ggtitle("Commercial likelihood on Yi*")+
     theme_bw()+
@@
 
-### Case study: sole in the Bay of Biscay
+### Case study: Sole in the Bay of Biscay
 
@@
     scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-    ggtitle("Commercial likelihood on Yi")+
+    ggtitle("Commercial likelihood on Yi*")+
     theme_bw()+
@@
   ## Parameter estimates
-  est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi","Scientific model"))
+  est_par_df_full_3$lkl[which(est_par_df_full_3$lkl == "Yi")] <- "Yi*"
+  est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi*","Scientific model"))
   est_par_df_full_3 <- est_par_df_full_3 %>%
@@
 
-  ggsave("images/par_plot.png",width = 6,height = 7.5)
+  ggsave("images/par_plot.png",width = 6/1.75,height = 7.5/1.75)
 
@@
 \begin{center}
-\includegraphics[width=5cm]{images/par_plot.png}
+\includegraphics[width=4.5cm]{images/par_plot.png}
 \end{center}
-\tiny *n.b.: punctual commercial data (i.e. observer data) needed to be integrated in inference in order to make the model converge and some parmaters needed to be step estimated (e.g. range)
+\scriptsize
+\onslide<2>{
+
+The \textcolor{BaptisteDarkGrey}{\textbf{integrated model}} fitted to \textcolor{BaptisteLightGreen}{$\bf{D_j}$}:
+
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Recovers the species-habitat relationship ($\bf{\textcolor{BaptisteOrange}{\beta_S}}$)
+
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Modifies the \textbf{contrasts of the map} \tiny (shape and intensity of the hotspots/coldspots) \scriptsize
+
+}
 
@@
 
-\item Framework that integrates \textbf{catch declarations data} (rough resolution of catch) and \textbf{scientific/onboard observer data} (exact locations of catch) \\
+\item \textcolor{BaptisteDarkGrey}{\textbf{Integrated framework}} that combines \textcolor{BaptisteLightGreen}{\textbf{catch declarations data}} (rough resolution) and \textcolor{SurveyBlue}{\textbf{scientific data}} (exact locations) \\
+\textcolor{BaptisteBlue}{\ding{224}} Allows to estimate the \textcolor{BaptisteOrange}{\textbf{habitat effect}} through commercial data \\
+\textcolor{BaptisteBlue}{\ding{224}} Modifies the \textbf{contrasts of the map} (hotspots vs. coldspots)
+
+\tiny \vspace{\baselineskip} \normalsize
 
@@
 
-\textcolor{BaptisteBlue}{\ding{224}} \textbf{Still question:} how to ease convergence ? \\
+\item \textbf{Some limits:}
+
+\textcolor{BaptisteBlue}{\ding{224}} How to ease convergence ? \\
+
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Need to make the hypothesis that fishing locations ($\textcolor{BaptisteLightGreen}{\mathcal{P}_j}$) are known
 
@@
 
-\vspace{\baselineskip}
+\tiny \vspace{\baselineskip} \normalsize
 
@@
 
-\item \textbf{ Change of support issues } \\
+\item \textbf{Is it a generic framework ?} \\
 
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} not so frequent in spatial ecology, \\
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} The overall approach is, \scriptsize \\ \textit{(i.e. modelling observed aggregated observations as a sum of latent punctual observations)}
 
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} but increasing amount of data available that may face such issues \\
+\normalsize \textbf{\textcolor{BaptisteBlue}{\ding{224}}} But need to adapt the observation model to the data \scriptsize \\ \textit{(here zeroinflated positive continuous data)}
 
@@
 
-\item \textbf{Is it a generic framework ?} \\
+\item \textbf{ Change of support applications} \\
 
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} the overall approach is, \tiny \\ (i.e.
-modelling observed aggregagetd catches as a convolution of latent exact catches) \\ \vspace{\baselineskip} \normalsize
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Increasing amount of available data that may face such issues \\
 
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} but need to adapt the observation model to the data \tiny \\ (here zeroinflated continuous data)   \\ \vspace{\baselineskip} \normalsize
-
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} and make the hypothesis that fishing locations ($\textcolor{BaptisteLightGreen}{\mathcal{P}_j}$) are known \scriptsize \\ (in reality $\textcolor{BaptisteLightGreen}{\mathcal{P}_j}$ is a trajectory and the set of fishing points are a realization of this trajectory) \\ \vspace{\baselineskip} \normalsize
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Here is one proposition to integrate aggregated data in SDM \\
 
@@
 
+\tiny \vspace{\baselineskip}
+
 \centering \Large \textcolor{BaptisteGrey}{\textbf{Thank you for your attention!}}
@@
 \begin{figure}
-    \begin{tabular}{c c}
-    \includegraphics[width=4.5cm]{images/peche_afrique_artisanale.jpg} &
-    \includegraphics[width=4.5cm]{images/peche_artisanale.jpg} \\
-    \includegraphics[width=4.5cm]{images/peche_hauturiere.jpg} &
-    \includegraphics[width=4.5cm]{images/thon_rouge.jpg}
-    \end{tabular}
+    \includegraphics[width=11cm]{images/front_page.png}
 \end{figure}
+
+#
+
+\footnotesize
+
+\textbf{Punctual observation model} ($Y_i$):
+
+\scriptsize $\operatorname{L}(y,\mu,\sigma^2)$ is the lognormal likelihood for observation $y$, mean $\mu$ and variance $\sigma^2$
+
+$Y$ and $D$ are supposed conditional on S and x \footnotesize
+
+$$
+\operatorname{P}\left(Y_i=y_{i} \right) =
+\left\{
+    \begin{array}{ll}
+        p_i & \text { if } y_{i}=0 \\
+        \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\mu_i=\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0
+    \end{array}
+\right.
+$$
+
+$$p_{i}=\exp(-e^\xi .S(x_{i}))$$
+
+\tiny \vspace{\baselineskip} \footnotesize
+
+\textbf{Declaration model} ($D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$):
+
+\begin{align*}
+P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0)
+                      = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
+\end{align*}
+
+$$\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right) =  \operatorname{L}\left(d_{j},\mu_j = E(D_j|d_j>0),\sigma_j^2 = ln(\frac{Var(D_j|d_j>0)}{E(D_j|d_j>0)^2} + 1) \right)$$
+
+$$E(D_j \vert d_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
+
+$$Var(D_j \vert d_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
+
+$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$

### 2022-04-02 Version_M2_plusieurszones/Scripts/In_progress/consistency_test.R

@@ -1,0 +1,66 @@
+
+ggplot(loc_x)+
+  geom_histogram(aes(x=bathy,fill=factor(substr_Mud_sediment)))
+
+
+fit_IM_res_no_realloc$Report$jnll_comp
+fit_IM_res_realloc$Report$jnll_comp
+fit_IM_res_sci$Report$jnll_comp
+
+
+## Integrated model
+obj_int <- fit_IM_res_realloc$Obj
+SD_int <- fit_IM_res_realloc$SD
+obj_int$fn()
+pl_int <- as.list(SD_int,"Est")
+
+## Scientific model
+obj_sci <- fit_IM_res_sci$Obj
+SD_sci <- fit_IM_res_sci$SD
+obj_sci$fn()
+pl_sci <- as.list(SD_sci,"Est")
+par_sci <- SD_sci$par.fixed
+f_sci <- as.numeric(obj_sci$fn(par_sci))
+
+## Fixed
+par_int <- SD_int$par.fixed # extractBaseParameters(obj_sci, pl_sci, pl_int)
+par_int <- par_int[which(names(par_int) %in% names(par_sci))]
+
+if(F %in% ("logtau" %in% names(par_int))){
+  
+  par_int <- c(par_int,-2.424969)
+  names(par_int)[length(par_int)] <- "logtau"
+  par_order <- which(names(par_sci) == "logtau")
+  par_int_2 <- c(par_int[1:(par_order-1)],par_int[length(par_int)],par_int[(par_order):(length(par_int)-1)])
+  par_int <- par_int_2
+  
+}
+
+par_int[1] <- par_sci[1]
+
+f_int <- as.numeric(obj_sci$fn(par_int))
+def_free <- sum(par_sci != 0)
+fixed <- 1 - pchisq( 2 * (f_int-f_sci), df=def_free )
+
+## Random
+# par_int.all <- extractBaseParameters(obj_sci, pl_sci, pl_int, all=TRUE)
+par_sci.all <- obj_sci$env$last.par.best
+par_int.all <- obj_int$env$last.par.best
+par_int.all <- par_int.all[which(names(par_int.all) %in% names(par_sci.all))]
+
+if(F %in% ("logtau" %in% names(par_int.all))){
+  
+  par_int.all <- c(par_int.all,-2.424969)
+  names(par_int.all)[length(par_int.all)] <- "logtau"
+  par_order <- which(names(par_sci.all) == "logtau")
+  par_int.all_2 <- c(par_int.all[1:(par_order-1)],
[…]

### 2022-04-20 Version_M2_plusieurszones/draft/paper_3/paper_3.tex

@@
 (2022).
-In this framework, all observations were supposed to be realised at the same scale.
-Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and potentially combine these data with other (more accurate) datasets such as scientific data.
+In this framework, all observations were supposed to be realised at a fine scale.
+Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and combine these data with other fine scale datasets such as scientific data.
 
@@
 \includegraphics{images/realloc.png}
-\caption{\label{fig:SchemeRealloc} Schematic representation of the reallocation process.}
+\caption{\label{fig:SchemeRealloc} Schematic representation of the reallocation process.
+The biomass field (the background field) depends on a covariate and a spatial random effect.
+The covariate is the \(x\) axis.
+It has a positive effect on biomass values (i.e.~biomass is higher on the right of the grid than on the left).
+The spatial random effect conduct to a hotspot on the bottom-right of the latent field.
+The study domain is considered as a statistical rectangle (grey square).
+Fishermen sample catches in areas of poor biomass where the covariate is relatively low (blue points) and in areas of higher biomass where the covariate is higher and eventually in the hotspot of biomass (orange and red points).
+These catches belong to the same declaration \(j\) and are summed to constitute the declaration \(D_j=50\).
+The declaration is declared at the level of the statistical rectangle.
+From VMS data, we know the fishing positions \(x_i\).
+In standard processing, \(D_j\) are then uniformly reallocated over the fishing positions \(x_i\).
+This strongly homogeneizes the catch.
+In particular, the effect of the habitat is no more evidenced in the reallocated catch \(Y_i^*\).}
 \end{figure}
@@
 We add a spatial random effect in the latent field (for both simulation and estimation) and we also simulate scientific data (in addition to commercial data) to explore the contribution of both datasets in inference.
-These simulations will be referred as multiple-square simulations.
+These simulations will be referred as \enquote{multiple-square simulations}.
 

### 2022-04-20 Version_M2_plusieurszones/pres/GDREcostat_April2022/old.Rmd

---
title: "Going from coarse landings data to fine scale species distribution"
subtitle : "Infering species distribution from aggregated spatial data"
author : "Baptiste Alglave,  Marie-Pierre Etienne, Youen Vermard,  Mathieu Woillez, Etienne Rivot"

fontsize: 10pt
bibliography : ["ref.bib"]
csl : ices-journal-of-marine-science.csl

output:
  beamer_presentation:
    theme: "Darmstadt"
    colortheme: "default"
    slide_level: 3
    # template: mytemplate.tex
    includes:
      in_header: header-simple.tex
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(rnaturalearth)
library(stringr)
library(sf)
library(maps)
library(ggplot2)
library(ggspatial)
library(cowplot)

col_scale <- function(palette = "Spectral", name = "", limits = NULL) {
  scale_colour_distiller(palette = palette,   # spectral colour scale
                         guide = "colourbar", # continuous colour bar
                         name = name,
                         limits = limits)
}

```

# Context

## Essential Fish Habitats

\begin{columns}
\begin{column}{0.5\textwidth}
Fishery ressources renewal highly rely on some key areas called\\
\textbf{Essential Fish Habitats (EFH)} :
\tiny \vspace{\baselineskip} \normalsize
\begin{itemize}
\item Spawning and reproduction grounds
\item Nursery grounds
\item Migration routes
\end{itemize}
\onslide<2>{
\center \Large \textbf{\textcolor{BaptisteBlue}{$\Downarrow$}}

\center \normalsize \textbf{\textcolor{BaptisteBlue}{Identify and caracterise these areas ?}}
}

\end{column}
\begin{column}{0.4\textwidth}

\begin{figure}[htb]
  \centering
  \includegraphics[width = 5cm]{"images/Brown_et_al_2018.png"}\quad
  \caption{\tiny Conceptual diagram of common life-history stages of fish in coastal habitats: S = mature adults during spawning, J = immature juveniles, and F = feeding adults not in spawning. Arrows represent migrations (M).}
\end{figure}
\end{column}
\end{columns}

\hfill \tiny [@brownConflictsCoastalZone2018]

<!-- \begin{columns} -->
<!-- \column{0.5\textwidth} -->
<!-- <!-- Context: -->
<!-- <!-- - -->
<!-- - truc -->
<!-- \column{0.5\textwidth} -->
<!-- \end{columns} -->

## Data

### Scientific data

```{r,include=F,warning=FALSE}

load("res/ptsurvey_wgs84_full.RData")

load("map/strata_ZN.RData")
load("map/strata_ZCC.RData")
load("map/strata_ZCL.RData")
load("map/strata_ZS.RData")

mapBase <- ne_countries(scale = "medium", returnclass = "sf")

plot_1 <- ggplot()+
    geom_sf(data = strata_ZN,
            alpha = 0.35,
            size = 0.75,
            color="#7FC97F",
            fill ="#7FC97F")+
    geom_sf(data = strata_ZS,
            alpha = 0.35,
            size = 0.75,
            color="#FFED6F",
            fill="#FFED6F")+
    geom_sf(data = strata_ZCC,
            alpha = 0.35,
            size = 0.75,
            color="white",
            fill="white")+
    geom_sf(data = strata_ZCL,
            alpha = 0.35,
            size = 0.75,
            color="#EF8A62",
            fill="#EF8A62")+
    geom_sf(data = mapBase)+
    geom_point(data = ptsurvey_wgs84_full,
               aes(x=long,y=lati,col = Sci.obs_spp))+
    coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
      theme(aspect.ratio = 1,legend.title = element_blank(),
          panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "skyblue"),
          legend.position = "none")+
    scale_color_gradient(low = "white", high = "red")+
  xlab("")+ylab("")+
  annotation_scale(location = "bl", line_width = .5) +
  annotation_north_arrow(location = "tl", height = unit(0.7, "cm"), width = unit(0.7, "cm"))

```

\begin{columns}
\begin{column}{0.4\textwidth}
\center \huge \textbf{\textcolor{BaptisteBlue}{\ding{58}}} \\
\normalsize
\begin{itemize}
\item Standardized sampling plan
\end{itemize}
\center \Huge \fontsize{45}{15}{\textbf{\textcolor{BaptisteBlue}{-}}} \\
\normalsize
\begin{itemize}
\item Expensive
\item Sparse and limited spatio-temporal coverage
\end{itemize}
\end{column}
\begin{column}{0.6\textwidth}

```{r, echo=FALSE, fig.align='center',out.width="120%",fig.cap="Orhago survey data - 2018"}

plot(plot_1)

```

\end{column}
\end{columns}

### Commercial data (VMS x logbooks)

```{r,include=F,warning=FALSE}

load("map/ICES_rect.RData")
load("res/points_sf.RData")

ICES_rect_2 <- ICES_rect %>%
  filter(str_detect(ICESNAME,"24E|23E|22E|21E|20E|19E|18E|17E|16E|15E"))

vms_plot <- points_sf %>%
  filter(str_detect(SI_DATE,"11/2018")) %>%
  # geom_sf_label(data = ICES_rect,aes(label = ICESNAME))+
  ggplot()+
  geom_sf(data = ICES_rect_2,fill="skyblue")+
  geom_sf(data = mapBase)+
  theme(panel.background = element_rect(fill = "skyblue"))+
  geom_sf(col="black")+
  coord_sf(xlim = c(-6,1), ylim = c(43,48), expand = FALSE)+
  theme(legend.position = "none")+
  annotation_scale(location = "bl", line_width = .5) +
  annotation_north_arrow(location = "tl", height = unit(0.7, "cm"), width = unit(0.7, "cm"))

```

\begin{columns}
\begin{column}{0.4\textwidth}
\center \huge \textbf{\textcolor{BaptisteBlue}{\ding{58}}} \\
\footnotesize
\begin{itemize}
\item Available data on the full year and dense coverage
\item Costless (for scientists)
\end{itemize}
\vspace{\baselineskip}
\center \fontsize{45}{15}{ \textbf{\textcolor{BaptisteBlue}{-}} } \\
\footnotesize
\begin{itemize}
\item Preferential behavior towards ressource
\item Non-homogeneous fleets (varying catchability and targeting behavior)
\item Reallocation of landings data on VMS pings
\end{itemize}
\end{column}
\begin{column}{0.6\textwidth}

```{r, echo=FALSE, fig.align='center',out.width="120%",fig.cap="VMS data for otter trauls (OTB-DEF) - November 2018"}

plot(vms_plot)

```

\end{column}
\end{columns}

## Scientific question and objectives

\centering \large Development of a hierarchical spatio-temporal model combining scientific and commercial data to infer fish spatial distribution

\vspace{\baselineskip}

\textbf{\textcolor{BaptisteBlue}{$\Rightarrow$}} Identification of Essential Fish Habitats (with a focus on spawning grounds)

\onslide<2>{

\vspace{\baselineskip}
\vspace{\baselineskip}

\raggedright \textbf{4 main dimensions :}

\vspace{\baselineskip}

\centering \textbf{\textcolor{BaptisteBlue}{Methodology} \quad \textcolor{BaptisteLightGreen}{Ecology} \quad \textcolor{orange}{Fishery} \quad \textcolor{yellow}{Management}}

}

# Spatio-temporal model

![\footnotesize Basic ideas of the spatio-temporal model](images/model_rep.png){width=105%}

## Sampling process

---

![\footnotesize Uniform sampling](images/preferential_sampling_map_b_0_without_sci.png){width=70%}

---

![\footnotesize Moderate preferential sampling](images/preferential_sampling_map_b_1_without_sci.png){width=70%}

---

![\footnotesize Strong preferential sampling](images/preferential_sampling_map_b_3_without_sci.png){width=70%}

---

::: {.block}
###
$$
\textcolor{BaptisteOrange}{\mathrm{X}_{com \,f }} \sim
\mathcal{IPP}(\textcolor{BaptisteOrange}{\lambda_f(x,t)})
$$
$$
log(\textcolor{BaptisteOrange}{\lambda_f(x,t)})=
\textcolor{BaptisteOrange}{\alpha_{X\,f}(t)} +
\textcolor{BaptisteOrange}{b_f(t)} . \log(\textcolor{BaptisteBlue}{S(x,t)})+
\textcolor{BaptisteOrange}{\eta_f(x,t)}
$$
:::

\vspace{\baselineskip}

\begin{columns}

\column{0.5\textwidth}

\scriptsize
\begin{itemize}

\item $\textcolor{BaptisteOrange}{\mathrm{X}_{com \,f }}$ : inhomogeneous point process

\item $\textcolor{BaptisteOrange}{\lambda_f(x,t)}$ : commercial sampling intensity

\item $\textcolor{BaptisteOrange}{\alpha_{X\,f}(t)}$ : intercept (seasonnal and yearly variation)

\item $\textcolor{BaptisteOrange}{b_f(t)}$ : targeting parameter

\item $\textcolor{BaptisteBlue}{S(x,t)}$ : biomass field

\item $\textcolor{BaptisteOrange}{\eta(x,t)} \sim N(0, Q_S^{-1})$

\item $f$ : fleet

\end{itemize}

\column{0.5\textwidth}
```{r, echo=FALSE, fig.align='center',out.width="75%"}
knitr::include_graphics("images/preferential_sampling_map_b_3_without_sci.png")
```
\end{columns}

# Fitting the model and analysing outputs

## Case study

<!-- \item Yearly survey  \tiny (November) \scriptsize -->

\scriptsize
\begin{itemize}
\item Bay of Biscay common sole \tiny (catch = landings, well-known ecology, commercial interest) \scriptsize
\item Commercial data from otter trawls targeting demersal species \tiny (fishing time as proxy of effort, good coverage of the area, relatively homogeneous catchability and targeting behavior) \scriptsize
\item Filter mature fraction of landings
\item Period : 2008 - 2018, monthly time step \tiny (132 time steps)
\end{itemize}

```{r,echo=F}

if(!file.exists("images/HF_OTB_DEF_plot.RData") & !file.exists("images/HF_OTB_CEP_plot.RData") & !file.exists("images/HF_OTB_DEF_plot.RData")){

  load("res/OTB_mature_commercial_data.RData")

  HF_OTB_DEF_plot <- ptvms_wgs84 %>%
    filter(str_detect(LE_MET_level6,"OTB_DEF_>=70_0")) %>%
    group_by(long,lati) %>%
    dplyr::summarise(HF = sum(HF)) %>%
    # filter(HF > 4) %>%
    ggplot()

  save(file="images/HF_OTB_DEF_plot.RData",data=HF_OTB_DEF_plot)

  HF_OTB_CEP_plot <- ptvms_wgs84 %>%
    filter(str_detect(LE_MET_level6,"OTB_CEP_>=70_0")) %>%
    group_by(long,lati) %>%
    dplyr::summarise(HF = sum(HF)) %>%
    # filter(HF > 4) %>%
    ggplot()

  save(file="images/HF_OTB_CEP_plot.RData",data=HF_OTB_CEP_plot)

  load("res/OTT_mature_commercial_data.RData")

  HF_OTT_DEF_plot <- ptvms_wgs84 %>%
    filter(str_detect(LE_MET_level6,"OTT_DEF_>=70_0")) %>%
    group_by(long,lati) %>%
    dplyr::summarise(HF = sum(HF)) %>%
    ggplot()
    # filter(HF > 4) %>%

  save(file="images/HF_OTT_DEF_plot.RData",data=HF_OTT_DEF_plot)

}else{

  load("images/HF_OTB_DEF_plot.RData")
  load("images/HF_OTB_CEP_plot.RData")
  load("images/HF_OTT_DEF_plot.RData")

}

```

```{r,echo=F,fig.align='center',out.width="100%",fig.cap="Orhago catch weights for sole - All data from 2008 to 2018."}

# load("res/Survey.data_catch_wgt.RData")
# Survey.data_catch_wgt %>%
#   ggplot()+
#   # geom_sf(data = ICES_rect_2,fill="skyblue")+
#   theme(panel.background = element_rect(fill = "skyblue"),
#         legend.title = element_blank())+
#   geom_point(aes(x=long,y=lati,col=CatchWgt_Solea_solea))+
#   scale_color_gradient2(midpoint = 1,
#                         low = "mediumpurple1",
#                         mid = "white",
#                         high = "red",
#                         space = "Lab" )+
#   geom_sf(data = mapBase)+
#   coord_sf(xlim = c(-6,1), ylim = c(43,48), expand = FALSE)+
#   # theme(legend.position = "none")+
#   # annotation_scale(location = "bl", line_width = .5)+
#   # annotation_north_arrow(location = "tl", height = unit(0.7, "cm"), width = unit(0.7, "cm"))+
#   xlab("")+ylab("")

```

\begin{columns}
\begin{column}{0.6\textwidth}

```{r,echo=F,fig.align='center',out.width="80%",fig.cap="Map of effort in fishing hours - Otter trawls targeting demersal fish (OTB-DEF) - Agregated data from 2008 to 2018."}

HF_OTB_DEF_plot_2 <- HF_OTB_DEF_plot + 
    # geom_sf(data = ICES_rect_2,fill="skyblue")+
    theme(panel.background = element_rect(fill = "skyblue"),
          legend.title = element_blank())+
  geom_point(aes(x=long,y=lati,col=HF),size=2.5,shape=15)+
  scale_color_gradient2(midpoint = 1,
                          low = "mediumpurple1",
                          mid = "white",
                          high = "red",
                          space = "Lab" )+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,1), ylim = c(43,48), expand = FALSE)+
    # theme(legend.position = "none")+
    # annotation_scale(location = "bl", line_width = .5)+
    # annotation_north_arrow(location = "tl", height = unit(0.7, "cm"), width = unit(0.7, "cm"))+
    xlab("")+ylab("")

plot(HF_OTB_DEF_plot_2)

```

\end{column}
\begin{column}{0.4\textwidth}

\begin{figure}
  \centering
  \includegraphics[width = 3cm]{"images/solea_solea.png"}\quad
  \caption{\tiny \textit{Solea solea} (Linnaeus, 1758)}
\end{figure}

\end{column}
\end{columns}

## Predictions of the relative biomass random field

```{r,include=F,warning=F}

## Load model outputs
load("res/fit_IM_res_b.season.year.RData")
load("res/loc_x_pred.RData")
SD <- fit_IM_res$SD
Report <- fit_IM_res$Report

time_step_t <- 1:12
year_y <- 2008:2018
time.step_df <- expand.grid(time_step_t,year_y)
colnames(time.step_df) <- c("Month","Year")
time.step_df <- time.step_df %>%
  arrange(Year,Month) %>%
  mutate(Month = ifelse(Month < 10,paste0("0",Month),Month)) %>%
  mutate(Year_Month = paste0(Year,"_",Month)) %>%
  mutate(t = 1:nrow(time.step_df))

# Build data
S_x_df <- cbind(loc_x_pred[,c("cell","x","y")],S_x = Report$S_p[1:nrow(loc_x_pred),]) %>%
  pivot_longer(cols = starts_with("S_x."),names_to = "t", values_to = "S_x") %>%
  mutate(t = as.numeric(str_replace(t,"S_x.",""))) %>%
  inner_join(time.step_df)

Ab_df <- S_x_df %>%
  group_by(Year_Month) %>%
  dplyr::summarise(Ab = sum(S_x))

S_x_df <- inner_join(S_x_df,Ab_df) %>%
  mutate(S_x_rel = S_x/Ab)

## time step
time_step_df <- S_x_df %>%
  group_by(Year_Month) %>%
  dplyr::select(Year_Month) %>%
  slice(1)
time_step_df$time_step = c(1:nrow(time_step_df))

```

```{r,echo=F,fig.align="center",warning=F,out.width="275px",fig.cap="Monthly relative distribution of sole for the year 2018"}

plot_S_x <- ggplot() +
    geom_point(data = S_x_df[which(str_detect(S_x_df$Year_Month,as.character(2018))),],
               aes(x=x,y=y,col = S_x_rel))+
    geom_sf(data = mapBase)+
    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
    scale_color_gradient(low = "white", high = "red")+xlab("")+ylab("")+
    theme(aspect.ratio = 1,
          legend.title = element_blank(),
          panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "skyblue"),
          panel.spacing.x = unit(6, "mm"),
          legend.position = "none")+
    facet_wrap(.~Year_Month)

plot(plot_S_x)

```

## Spatial patterns (EOF)

How to summarize the information of 132 maps and extract the main distribution patterns ? 
\onslide<2-3>{
\tiny\newline
\center \large \textcolor{BaptisteBlue}{\textbf{$\Rightarrow$ Empirical Orthogonal Functions (EOF)}}
}
\onslide<3>{
$$\overrightarrow{\mathbf{S}}^*_{t}=\sum_{k=1}^{K} \alpha_{k}(t) \vec{p}^{k}+\vec{n}_{t}$$
\footnotesize
\raggedright
\begin{itemize}
\item $\overrightarrow{\mathbf{S}}^*_{t}$ : normalized biomass field.

\item $\vec{p}^{k}$ : maps (or spatial patterns) which capture best the variance of $\overrightarrow{\mathbf{S}}^*_{t}$.

\item $\alpha_{k}(t)$ : temporal index informing when $\overrightarrow{\mathbf{S}}^*_{t}$ is distributed following the spatial pattern $\vec{p}^{k}$.

\item $\vec{n}_{t}$ : noise \tiny (or none relevant variation of the variable of interest).
\end{itemize}
\tiny

}

\hfill \tiny @vonstorchSpatialPatternsEOFs1999; @wikleSpatiotemporalStatistics2019

---

```{r,include=F,warning=F}

nEOF <- 10
EOF_mat <- S_x_df %>%
  dplyr::select(cell,Year_Month,S_x_rel) %>%
  dplyr::rename(S_x = S_x_rel) %>%
  pivot_wider(names_from = Year_Month,values_from = S_x) %>%
  as.matrix() %>% t()
EOF_mat <- EOF_mat[-1,]

# Compute mean pattern
spat_mean <- apply(EOF_mat, 2, mean)
spat_mean_df <- cbind(loc_x_pred,spat_mean = spat_mean)
nT <- nrow(EOF_mat)

## Then subtract and standardize
Zspat_detrend <- EOF_mat - outer(rep(1, nT), spat_mean)
Zt <- 1/sqrt(nT - 1)*Zspat_detrend

## Singular value decomposition
E <- svd(Zt)
V <- E$v
colnames(E$v) <- paste0("EOF", 1:nrow(EOF_mat)) # label columns
EOFs <- cbind(loc_x_pred[,c("x","y","cell")], E$v)
TS <- data.frame(E$u) %>%        # convert U to data frame
  mutate(t = 1:nrow(E$u)) %>%    # add a time field
  gather(eof, PC, -t)            # put columns (except time) into long-table format with EOF-PC as key-value pair
TS$nPC <- TS$PC * sqrt(nT-1)

eigen_df <- data.frame(eigen.values = E$d,perc.var = E$d/sum(E$d),cum.perc.var = cumsum(E$d/sum(E$d)),EOF=1:length(E$d))

EOFs_2 <- EOFs %>%
  select_at(c("x","y","cell",paste0("EOF",1:nEOF))) %>%
  pivot_longer(cols = starts_with("EOF")) %>%
  dplyr::rename(EOF = name) %>%
  filter(EOF=="EOF1" | EOF=="EOF2")

EOF_map <- ggplot(EOFs_2) + 
  geom_point(aes(x = x, y = y, col = -value)) +
  col_scale(name = "S_x") + theme_bw() + # ,limits = c(-0.1,0.1)
  # geom_sf(data = limites_orhago,size=1.5,col="skyblue")+
  geom_sf(data=mapBase)+
  xlab("Longitude (deg)") + ylab("Latitude (deg)") +
  coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
  theme(legend.title = element_blank()) +
  facet_wrap(.~EOF)

## Time series
EOFset1 <- E$u[1:(nT), 1:nEOF] * sqrt(nT - 1)
EOFset1 <- EOFset1 %>% data.frame
colnames(EOFset1) <- paste0("PC",1:nEOF)
EOFset1_2 <- EOFset1 %>%
  mutate(time_step = 1:nrow(EOFset1)) %>%
  pivot_longer(cols = starts_with("PC")) %>%
  dplyr::rename(PC = name) %>%
  filter(PC == "PC1") %>%
  inner_join(time_step_df)

EOF_ts <- ggplot()+
  geom_point(data = EOFset1_2,aes(x=Year_Month,y=-value,col=PC,group=PC),size=2)+
  geom_line(data = EOFset1_2,aes(x=Year_Month,y=-value,col=PC,group=PC),size=1)+
  geom_vline(xintercept=EOFset1_2$Year_Month[which(str_detect(EOFset1_2$Year_Month,"_02"))],linetype="dashed", color = "skyblue", size = 1)+
  geom_hline(yintercept = 0,linetype="dashed", color = "grey", size = 1)+
  theme_bw() +
  xlab("")+ylab("")+
  # facet_wrap(.~PC,ncol = 1) +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        plot.margin = margin(2, 2, 2, 2, unit = "pt"))+
  geom_text(aes(x = -Inf, y = -Inf, label = "2008"), hjust = -.10, vjust = 1.5) +
  geom_text(aes(x = max(EOFset1_2$Year_Month), y = -Inf, label = "2018"), hjust = 1, vjust = 1.5) +
  coord_cartesian(clip = "off")

```

```{r,echo=F,fig.align="center",out.width="300px",fig.cap="First EOF map (up) and related principal component (down). Blue line : February (reproduction period)."}

plot_grid(EOF_map,NULL,EOF_ts,ncol = 1,rel_heights = c(2,0.1,2))

```

---

\begin{columns}
\begin{column}{0.65\textwidth}

```{r,echo=F,fig.align="center",warning=F,fig.cap="First EOF map - Main spatial pattern"}

plot(EOF_map)

```

\end{column}
\begin{column}{0.35\textwidth}

\begin{figure}[htb]
  \centering
  \includegraphics[width = 4cm]{"images/arbault_et_al_1986.png"}\quad
  \caption{\tiny Eggs and Larvae distribution}
\end{figure}
\end{column}
\end{columns}
\hfill \tiny [@arbaultEstimationStockSole1986]

## Targeting parameter

\begin{columns}
\begin{column}{0.4\textwidth}

\centering \textbf{Parameters \textcolor{BaptisteOrange}{$b_f(t)$}}

\tiny \vspace{\baselineskip}

\footnotesize \textcolor{BaptisteGrey}{[0.8 ; 1.2]}

\end{column}
\begin{column}{0.2\textwidth}

\centering \textbf{vs.}

\end{column}
\begin{column}{0.4\textwidth}

\centering \textbf{Proportion of sole in landings}

\tiny \vspace{\baselineskip}

Alternative targeting indicator

\vspace{\baselineskip}

\footnotesize \textcolor{BaptisteGrey}{[2,5 \% ; 7.5 \%]}

\end{column}
\end{columns}

```{r,include=F}

load("res/fit_IM_res_b.season.year.RData")
SD <- fit_IM_res$SD
Report <- fit_IM_res$Report
b_season <- data.frame(Month = 1:12,
b_season=c(0,SD$par.fixed[which(names(SD$par.fixed)=="par_bseason")])) %>%
  mutate(Month = ifelse(Month<10,paste0("0",Month),Month)) # %>%
b_year <- data.frame(Year = 2008:2018,
                     b_year=c(0,SD$par.fixed[which(names(SD$par.fixed)=="par_byear")]))
b.seasonal_df <- inner_join(time.step_df,b_season)
b.seasonal_df <- inner_join(b.seasonal_df,b_year)
b.seasonal_df <- b.seasonal_df %>%
  mutate(b = SD$par.fixed[which(names(SD$par.fixed)=="par_b")]) %>%
  mutate(b_seas.year = b + b_season + b_year)

load("res/fit_IM_res_b.norm.RData")
SD <- fit_IM_res$SD
Report <- fit_IM_res$Report
b.norm_df <- cbind(time.step_df,b_norm = SD$par.random[which(names(SD$par.random)=="par_b")])

load("res/catch_prof_full_df_3.RData")
catch_prof_full_df_4 <- catch_prof_full_df_3 %>%
  filter(spp == "Solea_solea") %>%
  mutate(LE_KG_rel.wgt = LE_KG * LE_KG_rel )

b_df <- inner_join(b.seasonal_df,b.norm_df)
b_df <- inner_join(b_df,catch_prof_full_df_4[,c("Year_Month","LE_KG_rel","LE_KG_rel.wgt")])

```

\vspace{\baselineskip}

```{r,echo=F,fig.align="center",fig.asp=1/3,warning=F,out.width="300px"}

b_seas.year_vec <- (b_df$b_seas.year) / (sd(b_df$b_seas.year))
LE_KG_rel_vec <- (b_df$LE_KG_rel)/(sd(b_df$LE_KG_rel))
b_norm_vec <- (b_df$b_norm)/(sd(b_df$b_norm))

# plot(b_seas.year_vec,LE_KG_rel_vec)
# cor.test(b_seas.year_vec,LE_KG_rel_vec,method = "pearson")
# plot(b_norm_vec,LE_KG_rel_vec)
# cor.test(b_norm_vec,LE_KG_rel_vec,method = "pearson")

b_plot <- ggplot(data=b_df,aes(x=Year_Month,group=1))+

  geom_point(aes(y= (b_seas.year - mean(b_seas.year)) / (sd(b_seas.year)),color="b parameters"),size = 2)+
  geom_line(aes(y= (b_seas.year - mean(b_seas.year)) / (sd(b_seas.year)),color="b parameters"),size = 1)+

  geom_point(aes(y=(LE_KG_rel - mean(LE_KG_rel))/(sd(LE_KG_rel)),color="Sole proportion in landings"),size = 2,alpha = 0.75)+
  geom_line(aes(y=(LE_KG_rel - mean(LE_KG_rel))/(sd(LE_KG_rel)),color="Sole proportion in landings"),size = 1,alpha = 0.75)+

  # geom_point(aes(y=(b_norm - mean(b_norm))/(sd(b_norm)),color="b_random"),size = 2,alpha = 0.75)+
  # geom_line(aes(y=(b_norm - mean(b_norm))/(sd(b_norm)),color="b_random"),size = 1,alpha = 0.75)+

  geom_vline(xintercept=b_df$Year_Month[which(str_detect(b_df$Year_Month,"_02"))],linetype="dashed", color = "skyblue", size = 1)+
  theme_bw()+xlab("")+
  ylab("")+

  theme(legend.text = element_text(size=14),
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        plot.margin = margin(0, 0, 0, 0, unit = "pt"),
        legend.position = "bottom",
        aspect.ratio = 1/4)+
  geom_text(aes(x = -Inf, y = -Inf, label = "2008"), hjust = -.10, vjust = 1.5) +
  geom_text(aes(x = max(b_df$Year_Month), y = -Inf, label = "2018"), hjust = 1, vjust = 1.5) +
  coord_cartesian(clip = "off")

plot(b_plot)

```

\tiny
Variables have been normalized

# Discussion

## Conlusions

Approach with the potential to :

\tiny \vspace{\baselineskip} \normalsize
\begin{itemize}

\item capture main spatial patterns of the exploited fraction of the population \onslide<2-4>{\textbf{\textcolor{BaptisteBlue}{$\Leftrightarrow$ reproduction ecology of sole}}}

\tiny \vspace{\baselineskip} \normalsize

\onslide<3-4>{\item estimate the strength of targeting towards the species of interest and its evolution in time}

\vspace{\baselineskip}

\onslide<4>{\centering \large \textcolor{BaptisteGrey}{\textbf{$\Rightarrow$ Need for expert validation and additional applications}}}

\end{itemize}

## Perspectives

\begin{itemize}

\item Application to other case studies / other geographic areas

\begin{figure}
    \begin{tabular}{c c c c}
    \includegraphics[width=2cm]{images/merluccius_merluccius.jpg} &
    \includegraphics[width=2cm]{images/dicentrarchus_labrax.jpg} &
    \includegraphics[width=1.75cm]{images/lophius_spp.jpg} &
    \includegraphics[width=2cm]{images/sardina_pilchardus.jpg} \\
    \tiny \textbf{Hake} &
    \tiny \textbf{Sea bass} &
    \tiny \textbf{Anglerfish} &
    \tiny \textbf{Pilchard}\\
    \end{tabular}
\end{figure}
\tiny \vspace{\baselineskip} \normalsize

\onslide<2-3>{\item Extension of the model to include a multispecific targeting component and other drivers of targeting (tradition, management, etc.)}

\vspace{\baselineskip}

\onslide<3>{\item Effect of landings reallocation on vessel GPS positions ? How does it affect model outputs ? How to account for reallocation in the modelling approach ?}

\end{itemize}

---

\centering \Large \textcolor{BaptisteGrey}{\textbf{Thank you for your attention!}}

\tiny \vspace{\baselineskip}

\begin{figure}
    \begin{tabular}{c c}
    \includegraphics[width=4.5cm]{images/peche_afrique_artisanale.jpg} &
    \includegraphics[width=4.5cm]{images/peche_artisanale.jpg} \\
    \includegraphics[width=4.5cm]{images/peche_hauturiere.jpg} &
    \includegraphics[width=4.5cm]{images/thon_rouge.jpg}
    \end{tabular}
\end{figure}

# Bibliography

\tiny

### 2022-04-20 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3

@@
 ```
+
+# Introduction
+
+Spatial ecological data are becoming more and more accessible every days thanks to the huge effort from the scientific community to provide information in the blind spots of ecology, evolution and conservation [@nathan2022big,@hampton2013big,@gremillet2022big].
+These data are crucial to face the current challenges related to large scales ecological issues (e.g.
+tracking global change - @maureaud_are_2020) as well as smaller scale questions (e.g.
+tracking animal movement - @nathan2022big).
+However, these new data are highly heterogeneous in size, type and sampling design making there combination a methodological challenge [@pacifici2017integrating,@fletcher_practical_2019,@miller2019recent,@renner2019combining,@isaac2020data].
+For instance, ones often compare scientific survey data - that benefit from a controlled sampling plan and standard protocol - to citizen science data - that provide more data but that is also caracterized by misreporting or opportunistic sampling and then by potential sampling bias [@botella2021jointly,@steen2021spatial].
+Another less frequently mentionned data are declarations data.
+We refer to declaration data as the mandatory data that must be reported by some agent as a legal requirement to proceed with his activity.
+As they are mandatory, declaration data are usually very large datasets such as commercial catch declarations data in fisheries science [@hintzenVMStoolsOpensourceSoftware2012] or harvest data in hunting [@bauder2021mismatched,@gilbert2021integrating]; such way, these may be very informative of species spatial ecology.
+However, they are often declared at the scale of coarse spatial units while scientific survey (and commonly citizen science data) are usually reported with their exact locations.
+Furthermore, these administrative units do not generaly have a resolution that is relevant for ecological analysis and conservation [@pacifici2019resolving].
+Developing methods to infer spatial ecological process at a fine scale from coarse declarations data and integrate these with higher resolution data is then a challenge that will give access to huge amount of information.
+
+In the statistical literature, these issues are classified within "change of support" (COS) issues. It is recognized as a   Broadly speaking, COS is
+
+Combining these may require extra-care and a critical look at the data as well as methodological development to correct for these bias [@fletcher_practical_2019,@miller2019recent].
+
+Typically, data may be recorded following a standardized sampling plan with a controlled protocol.
+These data are considered as high quality data; however, they are often characterized by a low amount of samples due to the technical support they require and to their sampling cost.
+These are typically referred to as 'scientific survey' data (see for instance, the DATRAS online database which gathers the data from the main European Trawl scientific surveys [@ices_report_2014]).
+Alternatively, data may have been recorded from opportunistic sampling The strong development of citizen science data have allowed to provide additionnal information relying on volunteers to record data
+
+Recently, methods have been developped to combine high quality data benefiting from a standardized protocol and controled sampling plan with opportunistically sampled data that are
+
+the increasing amount of citizen science data
+
+They provide standardized ecological data on demersal fish over the whole European shelf.
+Alternatively, the Ebird plateform provide citizen science records of birds and is today extensively used in spatial and population models for ecological and conservation purposes [@sullivan2014ebird].
+
+ and new methods have been developped to integrate all these datasources  .
 

### 2022-04-21 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3

@@
 Furthermore, these administrative units do not generaly have a resolution that is relevant for ecological analysis and conservation [@pacifici2019resolving].
-Developing methods to infer spatial ecological process at a fine scale from coarse declarations data and integrate these with higher resolution data is then a challenge that will give access to huge amount of information.
-
-In the statistical literature, these issues are classified within "change of support" (COS) issues. It is recognized as a   Broadly speaking, COS is
+Developing methods to infer spatial ecological process at a fine scale from coarse declarations data and integrate these with higher resolution data is then a challenge that can give access to huge amount of information.
+
+Reconciling spatial (and eventually temporal) scales properly when different set of observations do not have the same resolution is a well known issue in geography, ecology, agriculture, geology and statistics [@gotway2002combining].
+Inferring values of a variable at points or blocks different from those that have been observed is commonly refered as the *change of support problem* (COS) in the statistical literature.
+Issues related to COS can usually be classified within 3 categories:
+
+- (1) spatial and temporal misalignement (when for example some multivariate data, where not all variables have been observed over all locations - see for example @finley2014bayesian).
+
+- (2) modifiable areal unit problem (MAUP) when data are aggregated over large scales
+
+Several studies have tackled this issue; for instance, @gotway2002combining and @gelfand2010misaligned proposed extensive reviews on COS and how it can be handled in statistical modelling.
+@mugglin2000fully, @gelfand2001change, @gotway2007geostatistical, @wikle2005combining proposed generic approaches for adressing COS in spatial or spatio-temporal context.
+kim2016change extended the framework from @wikle2005combining to account for temporal modelling and time-varying parameters.
+In the context of health analysis, @young2007linking proposed to compare some standard (and rough) way to connect data that do not have the same spatial resolution with a method properly accounting for COS.
+berrocal2010bivariate and berrocal2010spatio proposed a spatio-temporal method for fusing several air pollution data: one from coarse resolution but with full spatial coverage and another recorded at point level, with sparse distribution but where records almost corresponds to the true value of the process.
+In climate science, downscaling is common practice and is usually realized by combining large scale with lower scale dynamic models; more recently, @reich2014spectral and @parker2015multiresolution proposed a spectral statistical approach to downscale information from large scale model to lower scales.
+The
+
+[@gotway2002combining,@gotway2007geostatistical,@young2007linking,@berrocal2010bivariate,@berrocal2010spatio,@ren2013hierarchical,@reich2014spectral,@parker2015multiresolution,@kim2016change,@pacifici2019resolving,@gilbert2021integrating].
+Typically,
+
+It is recognized as a Broadly speaking, COS is
 

### 2022-04-26 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
remove generated files from git

--- 2022-03-30 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
@@
 output            :
-  papaja::apa6_pdf
+  papaja::apa6_pdf:
+    includes:
+            in_header: header.sty
     # reference_docx: mystyles_number_test.docx
@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE)
 
@@
 
-# Material and methods
-
-We base our approach on the model developped by @alglave_combining_2022.
-In this framework, all observations were supposed to be realised at a fine scale.
-Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and combine these data with other fine scale datasets such as scientific data.
-
-## Defining the problem from a modelling point of view
-
-**Definition of the variables**
-
-Let's define the latent field $S$ and the punctual observations $Y$.
-$S$ is a spatial log-Gaussian Field (GF) defined as $\log(S(x))=\mu+\beta .
-\Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)).
+#  Introduction 
+* Commencer par la motivation peche
+--> knowledge of essential habitat --> Species distribution model --> link between biomass and habitat to identify essential habitat
+-->  Data from scientific survey, precisely geolocalized standard sampling protocol, expensive and time consuming
+--> On the other hand declaration data, spatialized at coarser level  but cheap and reglementaire
+Combining the two sources of data.
+
+This problem is not unique, other exampl in terms of apllication
+
+Change of support form fine to coarse sptial level
+
+A solution the reallocation process
+
+** Pour la discu, simulation conditionnelle ? **
+
+We proposed an alternative approach to account for the change of support and we compare it to the approach proposed in alglave but also to a golden standard to assess the consequences of these methods.
+
+Paper organized as follow
+
+Section 2. presentation of the reference model and detail the reallocation process
+
+Section 3. Presnetation of an alternaion way to deal with change of support
+
+Section 4. simulation study to asses the bias and/or the loss of precision of the two methodes compared with the golden standard
+
+Section 5. Appllication on sole of the Bay of Biscay
+
+Section 6. Discu/conclu
+
+## An explicitly spatialized catch model
+
+In  @alglave_combining_2022, the authors propose a hierarchical spatial model to combine scientific survey data obtained through a standardized sampling protocol and catch data harvested by fishermen.
+This section provides the key ingredient of this model and raises some concern on the spatialization of the commercial catch data.
+
+Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ the spatial repartition of the biomass for a species of interest.
+$\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta .
+\Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$.
+As the catch are   zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ by a mixture model between a Dirac mass at $0$ and Log Normal distribution
+
+\begin{equation}
+Y_i | S(x_i) \overset{ind}{\sim} p_i \delta_0+ (1-p_i) B_i, \quad B_i \sim \mathcal{L}\left ( \frac{S(x_i)}{1-p_i} ,\sigma^2\right),
+\end{equation}
+
+<!-- 
+Je pense qu'il ne faut pas parler de la grille discrete tout de suite car S est défini sur un espace continu et moralement l'intensité du processus de poisson aussi.
+le passage à la grille est la pour simplifier l'expression de la vraisemblance, on peut dire qu'on fait une approximation pour des raisons numériques
+
 We consider a discrete study domain where $x \in [\![1,n]\!]$ stands for the discrete locations.
@@
 $\delta(x)$ is a GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$).
+-->
 

### 2022-04-26 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
intro

--- 2022-04-26 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
@@
 ---
-title             : "Going from coarse landings data to fine scale species distribution"
+title             : "Downscaling coarse observations to predict continuous wild species spatial distribution"
 
@@
 
-References for change of support issues in integrated modelling:
-
-https://esajournals.onlinelibrary.wiley.com/doi/epdf/10.1002/ecy.2710
-
-https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12793
-
-https://www.sciencedirect.com/science/article/pii/S0169534719302551#b0010
-
-https://www.sciencedirect.com/science/article/pii/S0006320721001993
+# Introduction
+
+**Attention:** L'introduction n'est pas finie
+
+**Context**
+
+Spatial ecological data are becoming more and more accessible every days thanks to the huge effort from the scientific community to provide information for ecology, evolution and conservation [@nathan2022big;@hampton2013big;@gremillet2022big].
+These data are crucial to face the current challenges related to both large and small scales ecological issues (e.g.
+tracking climate change [@maureaud_are_2020], tracking animal movement [@nathan2022big], mapping species distribution [@isaac2020data]).
+In particular, a key point of these studies lies in understanding what are the drivers of the ecological processes.
+For instance, in species distributions modelling, one will try to find the main covariates (e.g.
+substrat, temperature, depth/altitude) that shape species distribution [@guisan_predictive_2000;@planqueUnderstandingWhatControls2011].
+Combining these new datasources and more standard datasources is expected to improve our understanding of these processes and to fill the gaps in the blind spots of ecology.
+However, these datasources are highly heterogeneous in size, type and sampling design making there combination a methodological challenge [@pacifici2017integrating;@fletcher_practical_2019;@miller2019recent;@renner2019combining;@isaac2020data].
+For instance, combining scientific standardized data and citizen science data has been an important question in the last decade as they prooved to be highly complementary.
+The first one benefit from standardized protocol, controlled sapling plan and is usually considered as high-quality data.
+The second one provide a larger amount of data with lower cost, but is also caracterized by misreporting or opportunistic sampling and then by potential sampling bias [@botella2021jointly;@steen2021spatial].
+
+Another less frequently mentionned data are declarations data (we refer to declaration data as the mandatory data that must be reported by some agent as a legal requirement to proceed with his activity).
+As they are mandatory, declarations data are usually very large datasets and thus they may be very informative of species spatial ecology.
+A common example are commercial catch declarations data in fisheries science [@hintzenVMStoolsOpensourceSoftware2012] or harvest data in hunting [@bauder2021mismatched;@gilbert2021integrating].
+These data are most often declared at the scale of coarse spatial units while scientific survey (and commonly citizen science data) are usually reported with their exact locations.
+Furthermore, these administrative units do not generaly have a resolution that is relevant for ecological analysis and conservation [@pacifici2019resolving].
+
+Consequently (in ecological applications but in other fields too), to bypass the difference of resolution it is standard to apply rough hypothesis through geoprocessing technics based on proportional allocation and centroid smoothing [@gotway2007geostatistical;@young2007linking;@hintzenVMStoolsOpensourceSoftware2012].
+However, these methods do not provide a measure of uncertainty related to the processing of the data and can lead to bias not only in estimates of the mean and the variance, but also in any statistic that would be derived from these estimates.
+Overall, consequences of such procedures are hard to predict and can result in strong and deleterious biases [@gotway2007geostatistical;@pacifici2019resolving].
+Developing statistical methods to properly infer spatial ecological process at a fine scale from coarse declarations data and integrate these with higher resolution data is then a challenge that could give access to huge amount of information.
+
+**The COS issue**
+
+Reconciling spatial (and eventually temporal) scales properly when different set of observations do not have the same resolution is a well known issue in geography, ecology, agriculture, geology and statistics [@gotway2002combining].
+Inferring values of a variable at points or blocks different from those that have been observed is commonly refered as the *change of support problem* (COS) in the statistical literature [@gotway2002combining and @gelfand2010misaligned].
+COS issues can be classified within 2 categories - that are not necessarily mutually exclusive - [@pacifici2019resolving]:
+
+- (1) 'spatial and temporal misalignement' (when for example some multivariate data, where not all variables have been observed over all locations - see for example @finley2014bayesian).
+
+- (2) 'modifiable areal unit' problem (MAUP) when data are aggregated over increasingly larger geographic scales (e.g.
+collected at point level but regrouped/declared at coarse level - see for instance @gilbert2021integrating).
+Our case of application clearly refers to MAUP.
+Note that within this category, another category exist and is commonly refered to the 'ecological fallacy'.
+This occures when the individual response to a covariate differs from the response estimated from grouping individuals.
+This means that conclusions from a fine-resolution analysis can differ from an analysis at a coarser scale based on the aggregation of the fine-resolution data.
+
+Since 2000, several studies have described how could be overcome the different COS issues; @mugglin2000fully, @gelfand2001change, @gotway2007geostatistical, @wikle2005combining proposed generic approaches (and extensions of these approaches - @kim2016change) for adressing COS in spatial or spatio-temporal context.
+In health analysis @young2007linking proposed to compare some standard (and rough) way to connect data that do not have the same spatial resolution with a method properly accounting for COS.
+@berrocal2010bivariate and @berrocal2010spatio proposed a spatio-temporal method for fusing several air pollution data: one from coarse resolution but with full spatial coverage and another recorded at point level, with sparse distribution but where records almost corresponds to the true value of the process.
+In climate science, @reich2014spectral and @parker2015multiresolution proposed a spectral statistical approach to downscale information from large scale model to lower scales.
+In the field of ecology, few recent studies have tackled such issues: @finley2014bayesian provided a framework for integrating spatially misaligned data, @hefley2017bias proposed a solution based on COS to account for location error in presence-only data, @pacifici2019resolving introduced a framework for integrating datasources of different resolution to map species distribution.
+Applying similar ideas, @gilbert2021integrating integrated harvest data (aggregated data) and camera trap (punctual data) to map several wildlife species in Wisconsin.
+
+**The point that we adress in this paper**
+
+Still, applications in ecology remains sparse in regards to the huge amount of data that is available.
+Besides, one of the limiting point consist in the kind of observational data that can be fitted to the existing COS framework.
+Indeed, the framework that were developped so far and their related applications mainly limited their scope to relatively simple observational data: count data were modelled through Poisson processes [@mugglin2000fully;@gotway2007geostatistical;@pacifici2019resolving;@gilbert2021integrating] and continous data were modelled through Gaussian or Gamma distributions [@gelfand2001change;@wikle2005combining;@berrocal2010spatio].
+However, ecological data do not always consist of observation that can be modelled with standard probability distributions.
+For instance, in frequent cases data may be zero-inflated and positive-continous data.
+Several studies have developed models to handle properly such punctual data in a computationaly efficient way [@lecomte2013compound;@thorson2018three].
+However, these may complicate a bit the way COS is tackled when dealing with an aggregation of such complex data as their convolution may not be as simple as Poisson or Gaussian ones.
+
+In this paper, we aim at illustrating how to deal with change of support in ecological applications when the observational data is complex (e.g.
+zero-inflated and positive continous).
+We develop our approach on an existing framework developped by @alglave_combining_2022 in the field of marine and fisheries ecology.
+The framework aims at predicting the spatial distribution of fish species based on 2 datasets: scientific survey data and commercial catch declarations data.
+Commercial catch declarations are declared at the level of ICES rectangles (resolution of 0.5 x 1) while scientific data benefit from exact location records.
+Usually in standard processing, declarations data are reallocated uniformly over there GPS fishing positions (available through Vessel Monitoring Satellites - VMS) in order to improve its spatial resolution.
+
+In the following, we first describe the original model integrating both datasources (part 1).
+We then propose a generic statistical solution that allow to properly tackle the change of support issue and adapt it to our specific case (part 2).
+
+Finally, we compare this method to the current geoprocessing method in 3 steps:
+
+- A first simulation-estimation step at the scale of a single statistical rectangle where a the biomass distribution is simplified in order to explore the base properties of our framework, a more realistic simulation-estimation step (part 3).
+
+- A second simulation-estimation step at the scale of several statistical rectangles.
+The latent field is complexified and the set of simulation is parameterized so as to get closer to a realistic case study (part 4).
+
+- A final step where we compare the 2 methods on a real case study (common sole in the Bay of Biscay) and illustrate the main results that are consitent with simulations-estimations (part 5).
 

### 2022-04-28 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
intro

@@
 ```{r analysis-preferences}
+
 # Seed for random number generation
@@
 
-Spatial ecological data are becoming more and more accessible every days thanks to the huge effort from the scientific community to provide information for ecology, evolution and conservation [@nathan2022big;@hampton2013big;@gremillet2022big].
-These data are crucial to face the current challenges related to both large and small scales ecological issues (e.g.
-tracking climate change [@maureaud_are_2020], tracking animal movement [@nathan2022big], mapping species distribution [@isaac2020data]).
-In particular, a key point of these studies lies in understanding what are the drivers of the ecological processes.
-For instance, in species distributions modelling, one will try to find the main covariates (e.g.
+Spatial ecological data are becoming more and more accessible every days through new technologies and thanks to the huge effort from the scientific community to provide intensive information for ecology, evolution and conservation [@nathan2022big;@hampton2013big;@gremillet2022big].
+These data are crucial to face the current challenges related to large and small scales ecological questions: for instance, tracking climate change [@maureaud_are_2020], following animal movement [@nathan2022big] or mapping species distribution [@isaac2020data].
+In particular, a key point of these studies lies in understanding the drivers of the ecological processes under study.
+For instance, when investigating species distribution one will try to find the main covariates (e.g.
 substrat, temperature, depth/altitude) that shape species distribution [@guisan_predictive_2000;@planqueUnderstandingWhatControls2011].
-Combining these new datasources and more standard datasources is expected to improve our understanding of these processes and to fill the gaps in the blind spots of ecology.
-However, these datasources are highly heterogeneous in size, type and sampling design making there combination a methodological challenge [@pacifici2017integrating;@fletcher_practical_2019;@miller2019recent;@renner2019combining;@isaac2020data].
-For instance, combining scientific standardized data and citizen science data has been an important question in the last decade as they prooved to be highly complementary.
-The first one benefit from standardized protocol, controlled sapling plan and is usually considered as high-quality data.
+
+These datasources are often highly heterogeneous in size, type and sampling design making there combination a methodological challenge [@pacifici2017integrating;@fletcher_practical_2019;@miller2019recent;@renner2019combining;@isaac2020data].
+For instance, in species distribution modelling, recent studies have investigated how to combine scientific standardized data with citizen science data as they prooved to be highly complementary.
+The first one benefit from standardized protocol, controlled sampling plan and is usually considered as high-quality data.
 The second one provide a larger amount of data with lower cost, but is also caracterized by misreporting or opportunistic sampling and then by potential sampling bias [@botella2021jointly;@steen2021spatial].
@@
 As they are mandatory, declarations data are usually very large datasets and thus they may be very informative of species spatial ecology.
-A common example are commercial catch declarations data in fisheries science [@hintzenVMStoolsOpensourceSoftware2012] or harvest data in hunting [@bauder2021mismatched;@gilbert2021integrating].
+A common example of such datasources are commercial catch declarations data in fisheries science [@hintzenVMStoolsOpensourceSoftware2012] or harvest data in hunting [@bauder2021mismatched;@gilbert2021integrating].
 These data are most often declared at the scale of coarse spatial units while scientific survey (and commonly citizen science data) are usually reported with their exact locations.
@@
 Consequently (in ecological applications but in other fields too), to bypass the difference of resolution it is standard to apply rough hypothesis through geoprocessing technics based on proportional allocation and centroid smoothing [@gotway2007geostatistical;@young2007linking;@hintzenVMStoolsOpensourceSoftware2012].
-However, these methods do not provide a measure of uncertainty related to the processing of the data and can lead to bias not only in estimates of the mean and the variance, but also in any statistic that would be derived from these estimates.
+However, these methods do not provide a measure of uncertainty related to the processing of the data.
 Overall, consequences of such procedures are hard to predict and can result in strong and deleterious biases [@gotway2007geostatistical;@pacifici2019resolving].
+These bias do not only affect the mean and the variance, but also any statistic that would be derived from these estimates.
 Developing statistical methods to properly infer spatial ecological process at a fine scale from coarse declarations data and integrate these with higher resolution data is then a challenge that could give access to huge amount of information.
@@
 
-- (1) 'spatial and temporal misalignement' (when for example some multivariate data, where not all variables have been observed over all locations - see for example @finley2014bayesian).
+- (1) 'spatial and temporal misalignement' (the case of some multivariate data where not all variables have been observed over all locations - see for example @finley2014bayesian).
 
@@
 
-Since 2000, several studies have described how could be overcome the different COS issues; @mugglin2000fully, @gelfand2001change, @gotway2007geostatistical, @wikle2005combining proposed generic approaches (and extensions of these approaches - @kim2016change) for adressing COS in spatial or spatio-temporal context.
-In health analysis @young2007linking proposed to compare some standard (and rough) way to connect data that do not have the same spatial resolution with a method properly accounting for COS.
+Since 2000, several studies have described how could be overcome COS issues; @mugglin2000fully, @gelfand2001change, @gotway2007geostatistical, @wikle2005combining proposed generic approaches (and extensions of these approaches - @kim2016change) for adressing COS in spatial or spatio-temporal context.
+In health analysis, @young2007linking proposed to compare some standard (and rough) way to connect data that do not have the same spatial resolution with a method properly accounting for COS.
 @berrocal2010bivariate and @berrocal2010spatio proposed a spatio-temporal method for fusing several air pollution data: one from coarse resolution but with full spatial coverage and another recorded at point level, with sparse distribution but where records almost corresponds to the true value of the process.
 In climate science, @reich2014spectral and @parker2015multiresolution proposed a spectral statistical approach to downscale information from large scale model to lower scales.
-In the field of ecology, few recent studies have tackled such issues: @finley2014bayesian provided a framework for integrating spatially misaligned data, @hefley2017bias proposed a solution based on COS to account for location error in presence-only data, @pacifici2019resolving introduced a framework for integrating datasources of different resolution to map species distribution.
+In the field of ecology, some recent studies have tackled such issues: @finley2014bayesian provided a framework for integrating spatially misaligned data, @hefley2017bias proposed a solution based on COS to account for location error in presence-only data, @pacifici2019resolving introduced a framework for integrating datasources of different resolution to map species distribution.
 Applying similar ideas, @gilbert2021integrating integrated harvest data (aggregated data) and camera trap (punctual data) to map several wildlife species in Wisconsin.
@@
 
-Still, applications in ecology remains sparse in regards to the huge amount of data that is available.
-Besides, one of the limiting point consist in the kind of observational data that can be fitted to the existing COS framework.
+Still, applications in ecology remains sparse in regards to the huge amount of data that is now available.
+One of the limiting point consist in the kind of observational data that can be fitted to the existing COS framework.
 Indeed, the framework that were developped so far and their related applications mainly limited their scope to relatively simple observational data: count data were modelled through Poisson processes [@mugglin2000fully;@gotway2007geostatistical;@pacifici2019resolving;@gilbert2021integrating] and continous data were modelled through Gaussian or Gamma distributions [@gelfand2001change;@wikle2005combining;@berrocal2010spatio].
-However, ecological data do not always consist of observation that can be modelled with standard probability distributions.
+However, ecological data do not always consist of observations that can be modelled with standard probability distributions.
 For instance, in frequent cases data may be zero-inflated and positive-continous data.
-Several studies have developed models to handle properly such punctual data in a computationaly efficient way [@lecomte2013compound;@thorson2018three].
+Several studies have developed models to handle properly such data in a computationaly efficient way [@lecomte2013compound;@thorson2018three].
 However, these may complicate a bit the way COS is tackled when dealing with an aggregation of such complex data as their convolution may not be as simple as Poisson or Gaussian ones.
@@
 The framework aims at predicting the spatial distribution of fish species based on 2 datasets: scientific survey data and commercial catch declarations data.
-Commercial catch declarations are declared at the level of ICES rectangles (resolution of 0.5 x 1) while scientific data benefit from exact location records.
-Usually in standard processing, declarations data are reallocated uniformly over there GPS fishing positions (available through Vessel Monitoring Satellites - VMS) in order to improve its spatial resolution.
+Commercial catch declarations are declared at the level of ICES rectangles (resolution of 0.5° x 1°) while scientific data benefit from exact location records.
+Usually in standard processing, declarations data are reallocated uniformly over their GPS fishing positions (available through Vessel Monitoring Satellites - VMS) in order to improve its spatial resolution.
 
@@
 
-- A first simulation-estimation step at the scale of a single statistical rectangle where a the biomass distribution is simplified in order to explore the base properties of our framework, a more realistic simulation-estimation step (part 3).
+- A first simulation-estimation step at the scale of a single statistical rectangle.
+Our aim is to explore the base properties of our framework and then the biomass distribution is simplified to the most basic version (part 3).
 
 - A second simulation-estimation step at the scale of several statistical rectangles.
-The latent field is complexified and the set of simulation is parameterized so as to get closer to a realistic case study (part 4).
+The latent field is complexified and the set of simulations is parameterized so as to get closer to a realistic case study (part 4).
 

### 2022-05-04 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3
WIP

@@
 
-**The point that we adress in this paper**
+**Focus of the paper**
 
@@
 In  @alglave_combining_2022, the authors propose a hierarchical spatial model to combine scientific survey data obtained through a standardized sampling protocol and catch data harvested by fishermen.
-This section provides the key ingredient of this model and raises some concern on the spatialization of the commercial catch data.
-
-Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ the spatial repartition of the biomass for a species of interest.
-$\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta .
-\Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$.
-As the catch are   zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ by a mixture model between a Dirac mass at $0$ and Log Normal distribution
-
-\begin{equation}
-Y_i | S(x_i) \overset{ind}{\sim} p_i \delta_0+ (1-p_i) B_i, \quad B_i \sim \mathcal{L}\left ( \frac{S(x_i)}{1-p_i} ,\sigma^2\right),
-\end{equation}
+This section provides a biref overview of the key ingredient of this model and raises some concern on the spatialization of the commercial catch data.
+
+Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ a spatial random field which represents the biomass for a species of interest.
+$\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta \, \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$.
+As the catch are   zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ with a mixture of a Dirac mass at $0$ and a Log Normal distribution, the proportion of the mixture being defined  conditionally on the random field $\bm{S}$.
+
+\begin{align}
+Y_i | S(x_i) & \overset{ind}{\sim} p_i \delta_0+ (1-p_i) B_i, \quad B_i \sim \mathcal{L}\left ( \frac{S(x_i)}{1-p_i} ,\sigma^2\right), \cr \nonumber
+\mbox{with } p_i &= \exp(-e^\xi S(x_{i}))
+\end{align}
 
@@
 
-All fishing locations $x_i$ are supposed to be known through VMS data.
-At each fishing location $x_i$, fishermen realize a catch $Y_i$ conditionnally on the latent field values (referred hereafter as a punctual catch).
+In this approach all fishing locations $x_i$ and the coressponding catch are supposed to be known.
+
+<!-- however those fishinf locations are actually derived from a different dataset containing the hourly localisation of the the commercial vessels, the VMS dataset.
+This dataset is processed to identify the fishing location among all the recorded locations.
+-->
+
+<!-- At each fishing location $x_i$, fishermen realize a catch $Y_i$ conditionnally on the latent field values (referred hereafter as a punctual catch).
 Observations are supposed to follow some distribution $\mathcal{L}_Y$.
-
-$$Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma^2)$$
-
-In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.
-$\mathcal{L}_Y$ is decomposed it 2 parts:
-
-- the probability to obtain a zero-catch which is modelled as a Bernoulli variable with probability $p_{i}=\exp(-e^\xi .S(x_{i}))$.
+-->
+
+<!-- $$Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma^2)$$ -->
+
+<!-- In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.
+$\mathcal{L}_Y$ is decomposed it 2 parts: -->
+
+<!-- - the probability to obtain a zero-catch which is modelled as a Bernoulli variable with probability $p_{i}=\exp(-e^\xi .S(x_{i}))$.
 $\xi$ is a parameter controlling zero-inflation.
 When $\xi$ increases, the amount of zero in the data decreases.
-
-- if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a positive continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$.
-
-$$
-\operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i)\right) =
-\left\{
-    \begin{array}{ll}
-        p_i & \text { if } y_{i}=0 \\
-        \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0
-    \end{array}
-\right.
-$$
-
-Fishermen must declare what they catch at the level of the statistical rectangle at a daily time step.
+-->
+
+<!-- - if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a positive continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$.
+-->
+
+<!-- $$ -->
+<!-- \operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i)\right) = -->
+<!-- \left\{ -->
+<!--     \begin{array}{ll} -->
+<!--         p_i & \text { if } y_{i}=0 \\ -->
+<!--         \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0 -->
+<!--     \end{array} -->
+<!-- \right. -->
+<!-- $$ -->
+
+Actually, fishermen do not have to declare this precise information, but it is mandatory, for every vessel to declare the total daily catch aggregated at a given administrative spatial unit named  statistical rectangle.
 These data are recorded in logbooks data.
-A declaration (denoted $D_j$) is a summation of all the $Y_i$ realized on the fishing positions belonging to the catch declaration $D_j$.
-They can be expressed as:
-
-$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$$
-
-$\mathcal{P}_j$ is the vector of the fishing observations belonging to the declaration $D_j$.
-The punctual observations $Y_i$ are indexed through $i \in [\![1,m_j]\!]$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.
-
-**Catch reallocation: uniform reallocation or model-based reallocation**
-
-In standard processing, $D_j$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$.
+For a given vessel, on a given day, a declaration (denoted $D_j$) is therefore the sum of all individual catch $Y_i$ realized in the unit associated with the catch  declaration $D_j$, more formally:
+
+$$D_j=\sum_{x_i \in \mathcal{A}_j}{Y(x_i)},$$
+
+where $\mathcal{A}_j$ is the administrative spatial unit associated to declaration $j$.
+
+The original model proposed in  @alglave_combining_2022 is defined at the individual catch scale, those catch are  derived from the catch declaration combined with the VMS data.
+This is classically done in fisheries science REFREFEREFREF by using a uniform reallocation process.
+This process consists in counting, for a given vessel, on a given day,  the number $m_j$ of fishing points in $\mathcal{A}_j$ and define for each $x_i\in\mathcal{A}_j,$, the associated individual catch $Y(x_i)=D_j/m_j$.
+As a direct consequence, the so reconstructed individual catch are similar to a Catch Per Unit of Effort (CPUE), effort being measure in hours.
+As noted by @alglave_combining_2022, this process has several drawbacks.
+First, as a consequence of the reallocation process, the so reconstructed individual catches tend to exhibit smoother pattern than the original catches.
+Second, the actual number of data is the total number of declaration while the number of data after the reallocation process is the number of fishing locations, which is approximately 10 times the number of declarations.
+From a statistical point of view, this overestismation of the number of informative data tends to produce excessively narrow confidence intervals.
+
+<!-- **Catch reallocation: uniform reallocation or model-based reallocation** -->
+
+<!-- In standard processing, $D_j$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$.
 This is what we call uniform (or proportional) reallocation.
@@
 This strongly simplifies the actual process of observation and have most likely repercussion on model performance.
-
-To overcome such limitation, an alternative is to consider that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables).
-Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations:
-
-$$D_j | \mathcal{P}_j,S_{\mathcal{P}_j} \sim \mathcal{L}_D( S_{\mathcal{P}_j},\xi,\sigma^2)$$
-
-with $S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))$, $\xi$ the zero-inflation parmeter and $\sigma^2$ the variance parameter.
-As $D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$, it is possible to relate $D_j$, $Y_i$ and $S_{\mathcal{P}_j}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$.
+-->
+
+<!-- To overcome such limitations, an alternative is to consider that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables).
+Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->
+
+To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_j$ instead of the individual catch $Y(x_i)$.
+As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_j$ which consists of a sum of $m_j$ random variables folowwing a mixture distribution.
+This distribution has no  known analytical form.
+However the catch data also exhibit some zero inflation and a skewed repartition of the values and a mixture model would be a good candidate.
+Theefore, we propose to model $D_j$ as a mixture between a Dirac mass in $0$ and a lognormal distribution.
+In order to compare the original approach with our proposition, we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original propostion:
+
+<!-- that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables).
+Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->
+
+$$D_j | \mathcal{P}_j,\bm{S} \sim \mathcal{L}_D( \mu_j,\xi,\sigma^2)$$
+
+with $\mathcal{P}_j=(x_1,...,x_i, ...,x_{m_j})$, the list of the fishing points associated to declaration $D_j$,  $\mu_j$ the expected positive biomass, $\xi$ the zero-inflation parameter and $\sigma^2$ the variance parameter.
+
+<!-- As $D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$, it is possible to relate $D_j$, $Y_i$ and $S_{\mathcal{P}_j}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$.
 Our approach is to match the moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_j$.
@@
 In contrast, the model fitted on reallocated observations $Y^*_i$ will be referred as the $Y^*_i$ model.
-
-In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | \mathcal{P}_j,S_{\mathcal{P}_j}$.
+-->
+
+<!-- In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | \mathcal{P}_j,S_{\mathcal{P}_j}$.
+-->
 

### 2022-05-05 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
WIP

@@
 
+\begin{equation}
+\label{eq:defY}
+Y_i | S(x_i)  \overset{ind}{\sim} \M{p_i}{\frac{S(x_i)}{1-p_i}}{\sigma^2}
+\end{equation}
+
 \begin{align}
@@
 This is classically done in fisheries science REFREFEREFREF by using a uniform reallocation process.
-This process consists in counting, for a given vessel, on a given day,  the number $m_j$ of fishing points in $\mathcal{A}_j$ and define for each $x_i\in\mathcal{A}_j,$, the associated individual catch $Y(x_i)=D_j/m_j$.
+This process consists in counting, for a given vessel, on a given day,  the number $m_j$ of fishing points in $\mathcal{A}_j$ and define for each $x_i\in\mathcal{A}_j,$ the associated individual catch $\Yr(x_i)\defeq D_j/m_j$.
 As a direct consequence, the so reconstructed individual catch are similar to a Catch Per Unit of Effort (CPUE), effort being measure in hours.
 As noted by @alglave_combining_2022, this process has several drawbacks.
-First, as a consequence of the reallocation process, the so reconstructed individual catches tend to exhibit smoother pattern than the original catches.
-Second, the actual number of data is the total number of declaration while the number of data after the reallocation process is the number of fishing locations, which is approximately 10 times the number of declarations.
+First, as a consequence of the reallocation process, the reconstructed individual catches tend to exhibit smoother pattern than the original catches.
+Second, the actual number of data is the total number of catch declaration while the number of data after the reallocation process is the number of fishing locations, which is approximately 10 times the number of declarations.
 From a statistical point of view, this overestismation of the number of informative data tends to produce excessively narrow confidence intervals.
@@
 
-To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_j$ instead of the individual catch $Y(x_i)$.
-As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_j$ which consists of a sum of $m_j$ random variables folowwing a mixture distribution.
+To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_j$ instead of the reconstructed individual catch $\Yr(x_i)$.
+As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_j$ which consists of a sum of $m_j$ random variables following a mixture distribution.
 This distribution has no  known analytical form.
-However the catch data also exhibit some zero inflation and a skewed repartition of the values and a mixture model would be a good candidate.
-Theefore, we propose to model $D_j$ as a mixture between a Dirac mass in $0$ and a lognormal distribution.
-In order to compare the original approach with our proposition, we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original propostion:
+However the catch declaration data also exhibit some zero inflation and a long tail repartition of the values: a mixture model would be a good candidate.
+Therefore, we propose to use the same family of distribution for  $D_j$ than the one use for $\Yr$: a mixture between a Dirac mass in $0$ and a lognormal distribution.
+This is not exact, as the distribution of $D_j$ should be defined as convolution of mixtures but would result in quite untractable model.
+We somehow approximate the sum of  the original mixture distributions  by a mixture distribution of the same kind and we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original propostion:
 

### 2022-05-05 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
WIP

@@
 
-## An explicitly spatialized catch model
+## A spatialized catch model for aggregated data
 
@@
 \label{eq:defY}
-Y_i | S(x_i)  \overset{ind}{\sim} \M{p_i}{\frac{S(x_i)}{1-p_i}}{\sigma^2}
+Y_i | S(x_i)  \overset{ind}{\sim} \M{p_i}{\mu_i}{\sigma^2}
 \end{equation}
+with $p_i \defeq \exp(-e^\xi S(x_{i})$ the proportion of the mixture, $\mu_i\defeq 
+\frac{S(x_i)}{1-p_i}$ the expected catch when positive and $\sigma^2$ its variance (this parametrization of a lognormal distribution is detailed in SMXXXX and corresponds to
 
 \begin{align}
-Y_i | S(x_i) & \overset{ind}{\sim} p_i \delta_0+ (1-p_i) B_i, \quad B_i \sim \mathcal{L}\left ( \frac{S(x_i)}{1-p_i} ,\sigma^2\right), \cr \nonumber
-\mbox{with } p_i &= \exp(-e^\xi S(x_{i}))
+\P\left (Y_i =0 | S(x_i) \right) & = p_i=  \exp(-e^\xi S(x_{i}) ) \\
+\E\left ( Y_i \vert Y_i >0 , S(x_i)\right)  & =  \frac{S(x_i)}{1-p_i},\\  \nonumber
+\V ar( \log{Y_i}\vert Y_i >0, S(x_i) )& = \mu_i^2 (e^{\sigma^2}-1), \cr \nonumber
 \end{align}
@@
 Therefore, we propose to use the same family of distribution for  $D_j$ than the one use for $\Yr$: a mixture between a Dirac mass in $0$ and a lognormal distribution.
-This is not exact, as the distribution of $D_j$ should be defined as convolution of mixtures but would result in quite untractable model.
+This is not exact, as the distribution of $D_j$ should be defined as convolution of mixtures which  would result in an untractable model.
 We somehow approximate the sum of  the original mixture distributions  by a mixture distribution of the same kind and we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original propostion:
@@
 
-$$D_j | \mathcal{P}_j,\bm{S} \sim \mathcal{L}_D( \mu_j,\xi,\sigma^2)$$
-
-with $\mathcal{P}_j=(x_1,...,x_i, ...,x_{m_j})$, the list of the fishing points associated to declaration $D_j$,  $\mu_j$ the expected positive biomass, $\xi$ the zero-inflation parameter and $\sigma^2$ the variance parameter.
+$$D_j | \mathcal{P}_j,\bm{S} \sim \mathcal{L}_D( p^D_j, \mu^D_j, \sigma^{2,D})$$
+
+with $\mathcal{P}_j=(x_1,...,x_i, ...,x_{m_j})$, the list of the fishing points associated to declaration $D_j$,  $\mu^D_j$ the expected positive biomass, $p_j^D$ the proportion of the mixture and $\sigma^{2,D}$ the variance parameter.
 
@@
 
-**Matching $\mathcal{L}_Y$ and $\mathcal{L}_D$**
-
-First, as at the punctual level ($Y$), the declarations can be decomposed in 2 components: (1) the probability to obtain a zero-value declaration and, if the declaration is positive, (2) the probability to obtain some declaration value $d_j$.
-
-(1) We derive the probability to obtain a zero-declaration $P(D_j = 0)$ by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.
+**Matching the individual catch and the declaration levels**
+
+In order to define compatible models for $Y$ the actual individual catches and $D$ the catch declarations, we choose to match the key quantities of the two distributions.
+
+(1) As the $Y_1, \ldots, Y_{m_j}$ are independent conditionally on $\bm{S}$,  the probability to obtain a zero-declaration $P(D_j = 0)$ is obtained by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.
 
 \begin{align*}
-P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0),\nonumber \\
+\P(D_j = 0\vert \bm{S}, x_1, \ldots x_{m_j})  & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0\vert \bm{S}, x_i),\nonumber \\
                       & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
@@
 
-(2) If the declaration is positive, we assume the probability to obtain catch $d_j$ follows a continous distribution $\mathcal{L}_{D_j|D_j>0}$ with mean component $\mu_j$ and variance component $\sigma_j$.
-From $\mathcal{L}_Y$ we can derive $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$ (given below) and re-express $\mu_j$ and $\sigma_j$ as a function of these two moments.
-
-$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
-$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
-$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$
-
-If we consider $\mathcal{L}_{D_j|D_j>0}$ is lognormal (with parameterization described in SM) then, $\mu_j = E(D|D_j>0)$ and $\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$.
-Such way, by fitting the model on declaration $D_j$ rather than reallocated catch $Y_i$ and linking $\mathcal{L}_D$ to $\mathcal{L}_Y$, we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation.
-All calculations to obtain these formulas are available in SM.
+(2) The continuous component of the mixture is defined by the expected mean of a positive declaration and a transformation of its variance. It is straightforward to prove (see  SMXXXXX) that
+
+$$\E(D_j \vert D_j > 0)= \frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
+$$\V ar(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} \V ar(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
+$$\V ar(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$
+
+By using this second model, we expect to represent more accurately the observation process and avoid some of the drawbacks of the estimationbased on reallocated indivdual catch data $Y^r$.
 
 The inference is realized through Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].
+
+XXXX  Il faut aussi présenter la partie commercial ou dire que c'est comme dans Alglave 2022
 
@@
 **Single-square simulations**
+XXX Infor sur le nombre de simus ???
 
@@
 - a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations.
-We then fit the $Y_i^*$ model (the model fitted to punctual reallocated data).
+We then fit the $Y_i^r$ model (the model fitted to individual reallocated catch data).
 
@@
 In addition to the metrics we introduced to compare model configuration ($MSPE$ and species-habitat parameter $\beta_S$), we also assess the simulated/estimated values for the intercept $\mu$, the observation variance parameter $\sigma^2$ and the zero-inflation parameter $\xi$.
-
-**Multiple-square simulations**
-
-In this case, we extend the study to more than one single rectangle, we simulate scientific data and include these in inference.
-Finally, in addition to the covariate, we also simulate/estimate a spatial random effect in the latent field as a Matérn GRF.
-
-The study area is based on the case study; it includes the whole coast of the Bay of Biscay and cover several statistical rectangles (Figure \@ref(fig:MapSeveral)A).
-To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration).
-Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area.
-Similarly to the single-square simulations, the sampling of the commercial points of a declaration are realized in 3 steps.
-(1) The declaration is randomly affected to one of the ICES rectangle.
-(2) The centroid of the fishing zone is randomly sampled within this statistical rectangle.
-(3) The 10 fishing punctual observations are randomly sampled within the fishing zone.
-The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km.
-Note that we do not explore the effect of exploring several zones within the same declaration as it is already done in the single-square simulations.
-
-100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A).
-Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table 1).
-
-We compare several model configurations:
-
-- to assess what brings our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.
-
-- to assess the information brought by each data source, we compare models built on scientific data only (scientific-based models), models built on commercial data only (commercial-based models) and models combining both data sources (integrated models).
-
-In addition to the 2 metrics introduced at the beginning of the section ($MSPE$ and species-habitat parameter $\beta_S$), we also compare the simulated/estimated values of the range parameter.
-
-Note that when including 2 datasets in the model, a difference in catchability should be accounted for and one of the two dataset will be considered as reference (see @alglave_combining_2022 for more details on parameterization).
-Here the reference level is set with scientific data and the parameter $k_{com}$ accounts for the relative ratio between scientific and commercial observation values.
-In simulations, $k_{com}$ is set to 1 so that both datasources are assumed to have the same catchability.
-
-## Case-study: sole of the Bay of Biscay
-
-To illustrate our method on a real case study, we applied the approach to the common sole of the Bay of Biscay.
-VMS-logbooks data were extracted for the bottom trawlers fleet (OTB).
-The methods to cross VMS-logbooks data and to filter the fleet is already extensively described in the previous papers [@alglave_combining_2022] and is not developped further here.
-Scientific data were extracted from the DATRAS database for the Orhago beam trawl survey [@biais_gerard_orhago_2003;@ices_report_2018].
-To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - @ices_report_2018-1).
-To illustrate the method, we compare the outputs of (1) the scientific-based model, (2) the integrated model fitted to scientific data and reallocated commercial data ($Y^*_i$ model) and (3) the integrated model fitted to scientific and declarations data ($D_j$ model).
-
-The $D_j$ model faced convergence issues (some of the parameters were hardly estimated e.g.
-the range parameter).
-To ease convergence, we integrated in the analysis onboard observer data for the same fleet.
-They are considered as exact commercial data of catches.
-Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
-observation variance and zero-inflation parameter of commercial data).
-
-Furthermore, we adopt a step estimation procedure to initialize the optimization algorithm for the $D_j$ model.
-We first fit the $Y^*_i$ model and use the estimates of the model to initialize the optimization of the $D_j$ model.
-We eventually fix the parameters that are hard to estimate in first steps of optimization (intercept $\mu$, covariate effect $\beta_S$, range and marginal variance) and let them free iteratively.
-
-# Results
 
@@
 
+**Multiple-square simulations**
+
+In this case, we extend the study to more than one single rectangle, we simulate scientific data and include these in inference.
+Finally, in addition to the covariate, we also simulate/estimate a spatial random effect in the latent field as a Matérn GRF.
+
+The study area is based on the case study; it includes the whole coast of the Bay of Biscay and cover several statistical rectangles (Figure \@ref(fig:MapSeveral)A).
+To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration).
+Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area.
+Similarly to the single-square simulations, the sampling of the commercial points of a declaration are realized in 3 steps.
+(1) The declaration is randomly affected to one of the ICES rectangle.
+(2) The centroid of the fishing zone is randomly sampled within this statistical rectangle.
+(3) The 10 fishing punctual observations are randomly sampled within the fishing zone.
+The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km.
+Note that we do not explore the effect of exploring several zones within the same declaration as it is already done in the single-square simulations.
+
+100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A).
+Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table 1).
+
+We compare several model configurations:
+
+- to assess what brings our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.
+
+- to assess the information brought by each data source, we compare models built on scientific data only (scientific-based models), models built on commercial data only (commercial-based models) and models combining both data sources (integrated models).
+
+In addition to the 2 metrics introduced at the beginning of the section ($MSPE$ and species-habitat parameter $\beta_S$), we also compare the simulated/estimated values of the range parameter.
+
+Note that when including 2 datasets in the model, a difference in catchability should be accounted for and one of the two dataset will be considered as reference (see @alglave_combining_2022 for more details on parameterization).
+Here the reference level is set with scientific data and the parameter $k_{com}$ accounts for the relative ratio between scientific and commercial observation values.
+In simulations, $k_{com}$ is set to 1 so that both datasources are assumed to have the same catchability.
+
+## Case-study: sole of the Bay of Biscay
+
+To illustrate our method on a real case study, we applied the approach to the common sole of the Bay of Biscay.
+VMS-logbooks data were extracted for the bottom trawlers fleet (OTB).
+The methods to cross VMS-logbooks data and to filter the fleet is already extensively described in the previous papers [@alglave_combining_2022] and is not developped further here.
+Scientific data were extracted from the DATRAS database for the Orhago beam trawl survey [@biais_gerard_orhago_2003;@ices_report_2018].
+To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - @ices_report_2018-1).
+To illustrate the method, we compare the outputs of (1) the scientific-based model, (2) the integrated model fitted to scientific data and reallocated commercial data ($Y^*_i$ model) and (3) the integrated model fitted to scientific and declarations data ($D_j$ model).
+
+The $D_j$ model faced convergence issues (some of the parameters were hardly estimated e.g.
+the range parameter).
+To ease convergence, we integrated in the analysis onboard observer data for the same fleet.
+They are considered as exact commercial data of catches.
+Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
+observation variance and zero-inflation parameter of commercial data).
+
+Furthermore, we adopt a step estimation procedure to initialize the optimization algorithm for the $D_j$ model.
+We first fit the $Y^*_i$ model and use the estimates of the model to initialize the optimization of the $D_j$ model.
+We eventually fix the parameters that are hard to estimate in first steps of optimization (intercept $\mu$, covariate effect $\beta_S$, range and marginal variance) and let them free iteratively.
+
+# Results
+
 \newpage
@@
 
-**Expectancy of a positive declaration**
+**Expectation of a positive declaration**
 

### 2022-05-06 Version_M2_plusieurszones/pres/RESSTE_2022/RESSTE_2022.Rmd
RESSTE

--- 2022-04-02 Version_M2_plusieurszones/pres/GDREcostat_April2022/GDREcostat_2022.Rmd
@@
 ---
-title: "Infering species distribution from aggregated spatial data"
-subtitle : "Going from coarse landings data to fine scale species distribution"
+title: "Downscaling coarse observations to predict continuous species spatio-temporal distribution"
+subtitle : "Going from coarse landings data to fine scale fish distribution"
 author : "Baptiste Alglave, Marie-Pierre Etienne, Kasper Kristensen, Youen Vermard, Mathieu Woillez, Etienne Rivot"
@@
 }
+
+### Two alternative procedures to reallocate catches
+
+\textbf{Punctual observation model} ($Y_i$)
+
+\scriptsize $\operatorname{L}(y,\mu,\sigma^2)$ is the lognormal likelihood for observation $y$, mean $\mu$ and variance $\sigma^2$
+
+$Y$ and $D$ are supposed conditional on S and x \footnotesize
+
+$$
+\operatorname{P}\left(Y_i=y_{i} \right) =
+\left\{
+    \begin{array}{ll}
+        p_i & \text { if } y_{i}=0 \\
+        \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\mu_i=\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0
+    \end{array}
+\right.
+$$
+
+$$p_{i}=\exp(-e^\xi .S(x_{i}))$$
+
+\normalsize \vspace{\baselineskip} \vspace{\baselineskip}
+
+\textcolor{BaptisteLightGreen}{\textbf{Declaration model}} ($D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$)
+
+\footnotesize
+
+\begin{align*}
+P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0)
+                      = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j
+\end{align*}
+
+\normalsize \vspace{\baselineskip}
+
+$$\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right) = \quad \textbf{?}$$
+
+### Specifying $\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right)$
+
+\normalsize
+
+\textbf{Compute the moments}
+
+\footnotesize
+
+$$E(D_j \vert d_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
+
+$$Var(D_j \vert d_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
+
+$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$
+
+\vspace{\baselineskip}
+\vspace{\baselineskip}
+
+\normalsize
+
+\textbf{Consider $D_j$ is Lognormal too}
+
+\footnotesize
+
+\vspace{\baselineskip}
+
+$$
+\begin{aligned}
+&\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right) = \\ \\ & \quad \quad \operatorname{L}\left(d_{j},\mu_j = E(D_j|d_j>0),\sigma_j^2 = ln(\frac{Var(D_j|d_j>0)}{E(D_j|d_j>0)^2} + 1) \right)
+\end{aligned}
+$$
 
@@
 \end{figure}
-
-#
-
-\footnotesize
-
-\textbf{Punctual observation model} ($Y_i$):
-
-\scriptsize $\operatorname{L}(y,\mu,\sigma^2)$ is the lognormal likelihood for observation $y$, mean $\mu$ and variance $\sigma^2$
-
-$Y$ and $D$ are supposed conditional on S and x \footnotesize
-
-$$
-\operatorname{P}\left(Y_i=y_{i} \right) =
-\left\{
-    \begin{array}{ll}
-        p_i & \text { if } y_{i}=0 \\
-        \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\mu_i=\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0
-    \end{array}
-\right.
-$$
-
-$$p_{i}=\exp(-e^\xi .S(x_{i}))$$
-
-\tiny \vspace{\baselineskip} \footnotesize
-
-\textbf{Declaration model} ($D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$):
-
-\begin{align*}
-P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0)
-                      = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
-\end{align*}
-
-$$\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right) =  \operatorname{L}\left(d_{j},\mu_j = E(D_j|d_j>0),\sigma_j^2 = ln(\frac{Var(D_j|d_j>0)}{E(D_j|d_j>0)^2} + 1) \right)$$
-
-$$E(D_j \vert d_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
-
-$$Var(D_j \vert d_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
-
-$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$

### 2022-05-06 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
WIP

@@
 \label{eq:defY}
-Y_i | S(x_i)  \overset{ind}{\sim} \M{p_i}{\mu_i}{\sigma^2}
+Y_i | S(x_i)  \overset{ind}{\sim} \M{p_i}{\mu_i}{\sigma^2},
 \end{equation}
@@
 
+where $p_i\defeq \exp(-e^\xi S(x_{i}))$ is the mixture weight, $\mu_i \defeq \frac{S(x_i)}{1-p_i}$ is the mean of the lognormal distribution and $\sigma^2$ its variance, so that  
 \begin{align}
 \P\left (Y_i =0 | S(x_i) \right) & = p_i=  \exp(-e^\xi S(x_{i}) ) \\
-\E\left ( Y_i \vert Y_i >0 , S(x_i)\right)  & =  \frac{S(x_i)}{1-p_i},\\  \nonumber
-\V ar( \log{Y_i}\vert Y_i >0, S(x_i) )& = \mu_i^2 (e^{\sigma^2}-1), \cr \nonumber
+\E\left ( Y_i \vert Y_i >0 , S(x_i)\right)  & =  \mu_i = \frac{S(x_i)}{1-p_i},\\  \nonumber
+\V ar( Y_i\vert Y_i >0, S(x_i) )& = \mu_i^2 (e^{\sigma^2}-1), \cr \nonumber
 \end{align}
@@
 The inference is realized through Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].
-
-XXXX  Il faut aussi présenter la partie commercial ou dire que c'est comme dans Alglave 2022
+We made two simplifications.
+We use the Stochastic Partial Differential Equations (SPDE) approach to represent the spatiel field and we use integrated Nested Lappalace approximation to compute le likelihood.
+The stochastic random field is also approximated by a piecewise constant process defined on a fine grid.
+
+XXXX  Il faut aussi présenter la partie scientific ou dire que c'est comme dans Alglave 2022
 
@@
 
-## Simulation-estimation
-
-To evaluate the impact of reallocation on model outputs and how our alternative approach can improve model accuracy, we conduct simulation-estimation.
+## Comparison of the two approaches based on a simulation study
+
+To evaluate the impact of reallocation on model outputs and how our alternative approach can improve model accuracy, we conduct a simulation study.
 
@@
 
-In these two sets of simulation-estimation, the covariate is modelled as a continous GRF that we suppose known at each point of the grid.
+In these two sets of simulation studies, the covariate is modelled as a continous GRF that we suppose known at each point of the grid (when present the spatial random component is also a GRF but its values are not observed).
 The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$.
@@
 Two important variables may affect the accuracy of model outputs: the amount of commercial data and the number of fishing zones explored within one single declaration.
-In these simulations-estimations, we intend to explore the effect of these 2 variables.
+In these simulations, we intend to explore the effect of these 2 variables.
 
@@
 Furthermore, the samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle).
-Note that we make a distinction between fishing zones and statistical rectangles: fishing zones are included in a statistical rectangle and there can be several fishing zones within a statistical rectangles.
+Note that we make a distinction between fishing zones and statistical rectangles: fishing zones are included in a statistical rectangle and there can be several fishing zones within the same statistical rectangles.
 Consequently, a declaration can mix catches that have been realized in zones of high biomass and in zones of low biomass.

### 2022-05-23 Version_M2_plusieurszones/pres/RESSTE_2022/RESSTE_2022.Rmd
resste

@@
 ---
-\btVFill
+<!-- \btVFill -->
 \begin{center} \Large
 \textbf{How to integrate all these datasources?}
-\vspace{\baselineskip}
+\end{center}
 \large
@@
 
-\btVFill
+\onslide<2-3>{
 
+\begin{block}{ \center \textbf{Change of support} }
+
+\footnotesize
+\textbf{Common issue in statistical literature} \vspace{\baselineskip}
+
+\textbf{"Modifiable areal unit" problem (MAUP):} aggregation of data over increasingly larger geographic scales (e.g.
+data collected at point level but regrouped/declared at coarse level) \vspace{\baselineskip}
+
+\textbf{Several fields of application:} climatology, health science, ecology \vspace{\baselineskip}
+
+But mainly \textbf{standard observational data (Poisson, Gaussian)}, while data may be more complex in ecological applications (e.g. zero-inflated lognormal data)
+\begin{center}
+\textbf{\textcolor{BaptisteBlue}{\ding{224}} Objective of our work:} provide an approach that suits for complex data
 \end{center}
-\onslide<2>{
-\normalsize
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Some specific application in the context of fishery science: \\
-\scriptsize
+
+\end{block}
+
+}
+
+\onslide<3>{
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Base our model on an existing framework in the context of fishery science: \\
+\tiny
 \textit{Alglave Baptiste, Rivot Etienne, Etienne Marie-Pierre, Woillez Mathieu, Thorson James T, Vermard Youen (2022).
@@
   $\textcolor{BaptisteLightGreen}{D_j} | \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\textcolor{BaptisteLightGreen}{\mathcal{P}_j} \sim \mathcal{L}_D( \textcolor{BaptisteBlue}{S}_{\textcolor{BaptisteLightGreen}{\mathcal{P}_j}},\xi,\sigma^2)$
-  & \textcolor{BaptisteBlue}{\ding{224}}  Match $\mathcal{L}_D$ and $\mathcal{L}_Y$ moments \tiny (appendix) \\ 
+  & \textcolor{BaptisteBlue}{\ding{224}}  Match $\mathcal{L}_D$ and $\mathcal{L}_Y$ moments \\ 
   \end{tabular}
@@
 
+\onslide<2>{
+
 \textcolor{BaptisteLightGreen}{\textbf{Declaration model}} ($D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$)
@@
 
-$$\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right) = \quad \textbf{?}$$
+$$\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right) = \quad \textcolor{red}{\textbf{?}}$$
+
+}
 
@@
 
-\textbf{Compute the moments}
+\textbf{Compute the moments of $\mathbf{D_j\vert d_j > 0}$}
 
@@
 
-\textbf{Consider $D_j$ is Lognormal too}
+\onslide<2>{
+
+\textbf{Consider $\mathbf{D_j\vert d_j>0}$ is Lognormal too}
 
@@
 $$
+
+}
 
@@
 
-\item \textbf{ Change of support applications} \\
+\item \textbf{Moving to space-time ?} \\
 
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Increasing amount of available data that may face such issues \\
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Extending the observation model to account for \textbf{temporal misalignment} \\
 
-\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Here is one proposition to integrate aggregated data in SDM \\
+\textbf{\textcolor{BaptisteBlue}{\ding{224}}} Including spatio-\textbf{temporal covariance} in the latent field \\
 

### 2022-05-23 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd

@@
 In  @alglave_combining_2022, the authors propose a hierarchical spatial model to combine scientific survey data obtained through a standardized sampling protocol and catch data harvested by fishermen.
-This section provides a biref overview of the key ingredient of this model and raises some concern on the spatialization of the commercial catch data.
+This section provides a brief overview of the key ingredient of this model and raises some concerns on the spatialization of the commercial catch data.
 
 Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ a spatial random field which represents the biomass for a species of interest.
-$\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta \, \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$.
+$\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta \, \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$ and $\Gamma$ a field of covariate.
 As the catch are   zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ with a mixture of a Dirac mass at $0$ and a Log Normal distribution, the proportion of the mixture being defined  conditionally on the random field $\bm{S}$.
@@
 \label{eq:defY}
-Y_i | S(x_i)  \overset{ind}{\sim} \M{p_i}{\mu_i}{\sigma^2},
+Y_i | S(x_i)  \overset{ind}{\sim} \My{p_i}{\mu_i}{\sigma^2},
 \end{equation}
-with $p_i \defeq \exp(-e^\xi S(x_{i})$ the proportion of the mixture, $\mu_i\defeq 
-\frac{S(x_i)}{1-p_i}$ the expected catch when positive and $\sigma^2$ its variance (this parametrization of a lognormal distribution is detailed in SMXXXX and corresponds to
-
-where $p_i\defeq \exp(-e^\xi S(x_{i}))$ is the mixture weight, $\mu_i \defeq \frac{S(x_i)}{1-p_i}$ is the mean of the lognormal distribution and $\sigma^2$ its variance, so that  
+with $p_i \defeq \exp(-e^\xi S(x_{i}))$ the proportion of the mixture, $\mu_i\defeq 
+\frac{S(x_i)}{1-p_i}$ the expected catch when positive and $\sigma^2$ a transformation of its variance variance. This parametrization (see SMXXXX) of the mixture model corresponds to:
 \begin{align}
-\P\left (Y_i =0 | S(x_i) \right) & = p_i=  \exp(-e^\xi S(x_{i}) ) \\
+\P\left (Y_i =0 | S(x_i) \right) & = p_i=  \exp(-e^\xi S(x_{i}) ) \\  \nonumber
 \E\left ( Y_i \vert Y_i >0 , S(x_i)\right)  & =  \mu_i = \frac{S(x_i)}{1-p_i},\\  \nonumber
@@
 In this approach all fishing locations $x_i$ and the coressponding catch are supposed to be known.
+This is classically the case with scientific survey for which every catch is recorded as well as its geolocalization.
 
@@
 
-Actually, fishermen do not have to declare this precise information, but it is mandatory, for every vessel to declare the total daily catch aggregated at a given administrative spatial unit named  statistical rectangle.
+However, fishermen do not have to declare this precise information, but it is mandatory, for every vessel to declare the total daily catch aggregated at a given administrative spatial unit named  statistical rectangle.
 These data are recorded in logbooks data.
-For a given vessel, on a given day, a declaration (denoted $D_j$) is therefore the sum of all individual catch $Y_i$ realized in the unit associated with the catch  declaration $D_j$, more formally:
-
-$$D_j=\sum_{x_i \in \mathcal{A}_j}{Y(x_i)},$$
-
-where $\mathcal{A}_j$ is the administrative spatial unit associated to declaration $j$.
-
-The original model proposed in  @alglave_combining_2022 is defined at the individual catch scale, those catch are  derived from the catch declaration combined with the VMS data.
+For a given vessel, on a given day, a declaration (denoted $D$) is therefore the sum of all individual catch $Y_i$ realized in the administrative unit $\mathcal{A}_D$  associated with the catch  declaration $D$, more formally:
+
+$$D=\sum_{x_i \in \mathcal{A}_D}{Y(x_i)}.$$
+
+Since the original model proposed in @alglave_combining_2022 is defined at the individual catch scale, those catch are  derived from the logbooks data combined with the VMS data.
 This is classically done in fisheries science REFREFEREFREF by using a uniform reallocation process.
-This process consists in counting, for a given vessel, on a given day,  the number $m_j$ of fishing points in $\mathcal{A}_j$ and define for each $x_i\in\mathcal{A}_j,$ the associated individual catch $\Yr(x_i)\defeq D_j/m_j$.
-As a direct consequence, the so reconstructed individual catch are similar to a Catch Per Unit of Effort (CPUE), effort being measure in hours.
+This process consists in counting, for a given vessel, on a given day,  the number $m_j$ of fishing points in $\mathcal{A}_{D_j}$ associated with declaration $D_j$ and define for each $x_i\in\mathcal{A}_{D_j},$ the associated reallocated individual catch $\Yr(x_i)\defeq D_j/m_j$.
+As a direct consequence, the so reconstructed individual catch are similar to quantity widely used in Fishereis Sceince named  Catch Per Unit of Effort (CPUE), effort being measure in hours.
 As noted by @alglave_combining_2022, this process has several drawbacks.
 First, as a consequence of the reallocation process, the reconstructed individual catches tend to exhibit smoother pattern than the original catches. 
-Second, the actual number of data is the total number of catch declaration while the number of data after the reallocation process is the number of fishing locations, which is approximately 10 times the number of declarations.
+Second, the actual number of data is the total number of catch declarations while the number of data after the reallocation process is the number of fishing locations, which is approximately 10 times the number of declarations.
 From a statistical point of view, this overestismation of the number of informative data tends to produce excessively narrow confidence intervals.
@@
 However the catch declaration data also exhibit some zero inflation and a long tail repartition of the values: a mixture model would be a good candidate.
-Therefore, we propose to use the same family of distribution for  $D_j$ than the one use for $\Yr$: a mixture between a Dirac mass in $0$ and a lognormal distribution.
+Therefore, we propose to use the same family of distribution for  $D_j$ than the one use for the individual catch $Y$: a mixture between a Dirac mass in $0$ and a lognormal distribution.
 This is not exact, as the distribution of $D_j$ should be defined as convolution of mixtures which  would result in an untractable model.
-We somehow approximate the sum of  the original mixture distributions  by a mixture distribution of the same kind and we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original propostion:
+We somehow approximate the sum of  the original mixture distributions  by a mixture distribution of the same kind and we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original proposition:
 
@@
 
-$$D_j | \mathcal{P}_j,\bm{S} \sim \mathcal{L}_D( p^D_j, \mu^D_j, \sigma^{2,D})$$
+$$D_j | \mathcal{P}_j,\bm{S} \sim \Md{p^D_j}{\mu^D_j}{\sigma^{2,D}}$$
 
@@
 This is what we call model-based reallocation.
-This model will be referred as the $D_j$ model.
+This model will be referred as the Declaration Model.
 In contrast, the model fitted on reallocated observations $Y^*_i$ will be referred as the $Y^*_i$ model.
@@
 
-(1) As the $Y_1, \ldots, Y_{m_j}$ are independent conditionally on $\bm{S}$,  the probability to obtain a zero-declaration $P(D_j = 0)$ is obtained by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.
+(1) As the $Y_1, \ldots, Y_{m_j}$ are independent conditionally on $\bm{S}$,  the probability to obtain a zero-declaration $P(D_j = 0\vert \bm{S}, \mathcal{P}_j )$ is obtained by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.
 
 \begin{align*}
-\P(D_j = 0\vert \bm{S}, x_1, \ldots x_{m_j})  & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0\vert \bm{S}, x_i),\nonumber \\
-                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
+\P(D_j = 0\vert \bm{S}, \mathcal{P}_j)  & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0\vert \bm{S}, x_i),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = p_j^D
 \end{align*}
@@
 
-$$\E(D_j \vert D_j > 0)= \frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
-$$\V ar(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} \V ar(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
-$$\V ar(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$
-
-By using this second model, we expect to represent more accurately the observation process and avoid some of the drawbacks of the estimationbased on reallocated indivdual catch data $Y^r$.
-
-The inference is realized through Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].
-We made two simplifications.
-We use the Stochastic Partial Differential Equations (SPDE) approach to represent the spatiel field and we use integrated Nested Lappalace approximation to compute le likelihood.
+$$\E(D_j \vert D_j > 0, \bm{S}, \mathcal{P}_j)= \frac{\sum_{x_i \in \mathcal{P}_j} \E(Y_i \vert S(x_i)) }{1-p_j^D} = \frac{\sum_{x_i \in \mathcal{P}_j} S(x_{i})}{1-p_j^D},$$
+$$\V ar(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} \V ar(Y_{i})}{1-p_j^D} - \frac{\pi_j}{(1-p_j^D)^2}E(D_j)^2,$$
+$$\mbox{with } \V ar(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i})).$$
+
+By defining the observation process at the declaration level,  we expect to  avoid some of the drawbacks of the estimation based on reallocated individual catch data $Y^r$.
+
+ The inference is based on  maximum likelihood approach with two simplifications.
+We use the Stochastic Partial Differential Equations (SPDE) approach to represent the spatial field and we use integrated Nested Laplace approximation to compute le likelihood.
 The stochastic random field is also approximated by a piecewise constant process defined on a fine grid.
-
-XXXX  Il faut aussi présenter la partie scientific ou dire que c'est comme dans Alglave 2022
+The optimization of the likelihood relies Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].
+
+<!-- XXXX  Il faut aussi présenter la partie scientific ou dire que c'est comme dans Alglave 2022 -->
 
@@
 
-## Comparison of the two approaches based on a simulation study
-
-To evaluate the impact of reallocation on model outputs and how our alternative approach can improve model accuracy, we conduct a simulation study.
-
-First, we conduct basic simulations in order to explore the effect of some key variables on model performance and also the base properties of the alternative models.
-We assume the study domain is a single statistical rectangle, only commercial data feed the model and the expression of the latent field in both simulation and estimation is simplified (the variability only depends on one covariate, not on a spatial random term).
+We have finally three alternatives to estimate the spatial field of biomass from catch data:
+
+* the original approach from  @alglave_combining_2022 when the locations of  commercial catch (and eventually the scientific catch) are precisely geolocalized.
+This ideal situation, with no actual application,  will be named 'Spatial Model' and used as a reference for the comparison between the two alternatives.
+* the original model fitted with  commercial reallocated individual catch (and potentially few precisely geolocalized scientific data  as done in  @alglave_combining_2022.
+This approach will be refereed as "Reallocated Model".
+* the alternative approach detailed before where the biomass model is fitted  using commercial catch declaration at a corase spatial level and potentially few precisely geolocalized scientific data.
+This approach is named "Declaration Model".
+
+## Simulation studies
+
+To asses the drawbacks and the advantages of the different approaches, we conduct two different simulation studies. 
+<!-- To evaluate the impact of reallocation on model outputs and how our alternative approach can improve model accuracy, we conduct a simulation study. We have  -->
+First, we explore the base properties of the different models by a study at single statistical rectangle, based on commercial data and with a very simple spatial latent field which only depends on one covariate.
 These simulations will be referred as "single-square simulations".
 
-Then, we extend the simulation-estimation study to several rectangles to get closer to a real case study configuration.
-We add a spatial random effect in the latent field (for both simulation and estimation) and we also simulate scientific data (in addition to commercial data) to explore the contribution of both datasets in inference.
+Then, we extend the simulation study to several rectangles to get closer to a real case study configuration.
+We add a spatial random effect in the latent field and we also simulate precisely geolocalized scientific data (in addition to commercial data) to explore the contribution of both datasets in inference.
 These simulations will be referred as "multiple-square simulations".
 
-In these two sets of simulation studies, the covariate is modelled as a continous GRF that we suppose known at each point of the grid (when present the spatial random component is also a GRF but its values are not observed).
+In these two sets of simulation studies, there is a unique covariate,  modeled as a continuous GRF that we suppose known at each point of the grid (when present the spatial random component is also a GRF but its values are not observed).
+XXXX AJouter les paramètres du chmaps spatial 
 The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$.
@@
 
-Each time we compare several model configurations and evaluate these in regards to 2 metrics:
+The locations of the individual commercial catch are generally organized in spatial clusters, named fishing zones in the following.
+The simulation process mimics this property by sampling the fishing points using a Neymann Scott process: the centers of the fishing zones are sampled according to a Poisson process and the fishing points are then uniformely sampled within a squared area of side 7.
+At each fishing position, a catch is  sampled conditionally on the value of the latent field according to the model $\mathcal{M}_Y$.
+
+We compare the performance of the Spatial Model (the gold standard), the Reallocated Model and the Declaration Model  model configurations in regards to 2 metrics:
 
@@
 
-- the parameter of the species habitat relationship $\beta_S$ compared with its estimate $\hat \beta_S$.
+- the quality of the estimation of the parameter  $\beta_S$  which quantifies the species habitat relationship.
 
@@
 
-Two important variables may affect the accuracy of model outputs: the amount of commercial data and the number of fishing zones explored within one single declaration.
-In these simulations, we intend to explore the effect of these 2 variables.
-
-First, the amount of commercial data; it is expected that when increasing the amount of data the accuracy of the predictions will be improved.
-In simulations, we progressively increase the number of fishing positions (10, 100 and 1000 fishing pings).
-As there are 10 pings per declaration, this means that the number of declarations increases respectively (1, 10 and 100 declarations).
-
-Furthermore, the samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle).
+Two important variables may affect the accuracy of model outputs: the amount of commercial data and the number of fishing zones explored and aggregated within a catch declaration.
+The single-squares simulations intend to explore the effect of these two variables.
+
+First, increasing the amount of data the accuracy is expected to improve the estimations and the spatial predictions.
+We explore the  potential improvement of the spatial predictions brought by an increasing amount of fishing points (10, 100 and 1000) which correspond respectively to 1, 10 and 100 declarations, the number of fishing locations within a declaration being fixed to 10.
+
+Furthermore, the number of fishing zones within the statistical rectangle associated with a declaration might also affect the performance of the different approaches.
+We expect that the reallocation process will be less problematic when all the individual catch are spatially close.
+This situation corresponds to a declaration associated with only one fishing zone.
+<!-- closed    samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle).
 Note that we make a distinction between fishing zones and statistical rectangles: fishing zones are included in a statistical rectangle and there can be several fishing zones within the same statistical rectangles.
@@
 Reallocating uniformly the declaration will strongly homogeneise the actual catch, will mix up information from to different grounds and then may lead to a strong loss of accuracy in model outputs.
-It is expected that when the number of fishing zones within a declaration increases, the accuracy of the model outputs decreases.
-To assess the effect of such process, we simulated the pings of a fishing declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones.
-
+-->
+The accuracy of the Reallocated Model outputs is expected to decrease when the number of fishing zones increases.
+To assess the effect of such process, we simulated the fishing locations associated with a declaration declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones.
+
+XXXX Un petit schéma pour expliquer le  processus de similuation et mettre des images sur les mots de fishing zone, fishing points
+
+<!--
 Concretely, for each fishing declaration the fishing pings are sampled in 2 steps: first the centroid of the zones are randomly sampled over the statistical rectangle (here the simulation domain) and then the fishing positions are randomly sampled within the radius of the centroid of the zones.
@@
 This is a realistic order of magnitude if we consider each zone is one fishing operation and that for each fishing operation the distance of the operation will not exceed 30km.
+XXXX MPE  le coté raisonnable dépend de l'échelle de la carte. je vourdrais essayer de limiter le nombre de termes et donc éviter la notion de fishing operation-->
 
@@
 
-At each fishing position, a catch is realized conditionnally on the value of the latent field and follow the probability distribution described above ($\mathcal{L}_y$).
-
-We compare three simulation/estimation configurations:
-
-- a golden standard configuration: punctual observations ($Y_i$) are known exactly and the punctual-data model is fitted to the exact data.
-
-- a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations.
+<!-- We compare three simulation/estimation configurations: -->
+
+<!-- - a golden standard configuration: punctual observations ($Y_i$) are known exactly and the punctual-data model is fitted to the exact data. -->
+
+<!-- - a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations.
 We then fit the $Y_i^r$ model (the model fitted to individual reallocated catch data).
-
-- a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the $D_j$ model to the data (i.e.
+-->
+
+<!-- - a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the Declaration Model to the data (i.e.
 the one fitted at the level of the declarations data).
-
-In addition to the metrics we introduced to compare model configuration ($MSPE$ and species-habitat parameter $\beta_S$), we also assess the simulated/estimated values for the intercept $\mu$, the observation variance parameter $\sigma^2$ and the zero-inflation parameter $\xi$.
-
-## Single-square simulations
+-->
+
+For the single square simulation, in addition to the metrics introduced before ($MSPE$ and species-habitat parameter $\beta_S$), we also assess quality of the estitmation for  $\mu$, the observation variance parameter $\sigma^2$ and the zero-inflation parameter $\xi$.
 
@@
 ```
-
-Reallocation has a major effect on predictions and estimates accuracy (Figure \@ref(fig:PerfMetricSingle)).
-Reallocating data conduct to a 10 to 200 times decrease in accuracy for spatial predictions when fitting the model to reallocated catch (MSPE gold compared to red boxplots).
-Accuracy decreases as the number of visited zones within a declaration increases.
-Besides, $\beta_S$ estimates are biased and reallocation leads to the loss of the species-habitat relationship as the number of fishing zones (within a declaration) increases ($\beta_S$ estimates get closer to 0).
-Increasing the number of samples does not improve inference.
-Regarding other parameters estimates, the zero-inflation parameter ($\xi$) is over-estimated (i.e.
-when uniformly reallocating commercial data, the quantity of positive observations is under-estimated), the observation variance ($\sigma$) is underestimated (i.e.
-the data is estimated to be less noisy than they actually are) and the intercept of the latent field ($\mu$) is slightly over-estimated (Figure \@ref(fig:ParBiasSingle)).
-
-The $D_j$ model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \@ref(fig:PerfMetricSingle)).
-Still, the $D_j$ model outputs are not as accurate as the ones of the golden-standard.
-Furthermore, the zero-inflation parameter is unbiased when the model is fitted to catch declarations $D_j$.
-Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \@ref(fig:ParBiasSingle)).
-This alternative model have some convergence difficulties (Table 2) as 8% of the model did not converged when sample size is medium (100 pings) and only 3% did not when sample size is large (1000 pings).
-
-![(\#fig:PerfMetricSingle) Performance metric for single-square simulations.
-Columns: number of commercial fishing positions.
-x-axis: number of zones visited within each declaration.
-'Reallocation: ', data are or are not reallocated in simulations.
-'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $Y_i^*$ or on catch declarations $D_j$.
-Gold: golden standard.
-Red: uniform reallocation ($Y_i^*$ model).
-Green: model-based reallocation ($D_j$ model).
-Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the $D_j$ model (see Table 2).](images/Perf.metric_single_square.png){width=75%}
+The resultats are presented in Figures \@ref(fig:ParBiasSingle) and \@ref(fig:ParBiasSingle).
+
+![(\#fig:PerfMetricSingle) Performance metric for single-square simulations with a total of 100 or 1000 fishing positions in columns.
+The number of fishing zones visited within each declaration is represented on the x-axis.
+The result of the Spatial model is in yellow, in red the results of the Reallocated model and in green the Declaration model.
+Simulations conducted with 10 fishing positions are not represented as they encouter convergence issues as stated in Table 2.](images/Perf.metric_single_square.png){width=75%}
 
@@
 
+ Figure \@ref(fig:PerfMetricSingle) highlights that the reallocation process has a major effect on predictions and estimates accuracy.
+As expected, the reallocation process conducts to a 10 to 200 times decrease in accuracy for spatial predictions when fitting the Reallocated Model (MSPE gold compared to red boxplots).
+Accuracy decreases as the number of visited zones within a declaration increases.
+Besides, the estimation of $\beta_S$ is biased and reallocation leads to the loss of the species-habitat relationship as the number of fishing zones (within a declaration) increases ($\beta_S$ estimates get closer to 0).
+Increasing the number of samples does not improve inference.
+Figure \@ref(fig:ParBiasSingle) shows the over-estimation of  the zero-inflation parameter ($\xi$) when using the Reallocated Model.
+This is not surpising since the uniform reallocation of commercial data tends to decrease the proportion of zero since a realocated catch $Y^r$ is non-zer as soon as at least one of the individual catch associated with the same declaration is non -zero.
+The observation variance ($\sigma$) is underestimated i.e.
+the data is estimated to be less noisy than they actually are.
+The intercept of the latent field ($\mu$) is slightly over-estimated (Figure \@ref(fig:ParBiasSingle)).
+
+The Declaration model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \@ref(fig:PerfMetricSingle)) even so the model outputs are not as accurate as the ones of the golden-standard.
+Furthermore, the zero-inflation parameter is unbiased when the model is fitted to catch declarations.
+Other parameters (observation variance, intercept) are also better estimated than with the Reallocated Model even though they remain slightly biased (Figure \@ref(fig:ParBiasSingle)).
+This alternative model have some convergence issues (Table 2) as 8% of the model did not converged when sample size is medium (100 pings) and only 3% did not when sample size is large (1000 pings).
+
 ```{r}
@@
 **Multiple-square simulations**
-
-In this case, we extend the study to more than one single rectangle, we simulate scientific data and include these in inference.
-Finally, in addition to the covariate, we also simulate/estimate a spatial random effect in the latent field as a Matérn GRF.
+As mentioned before, we propose another set of simulations designed to be closer from the case study.
+The latent biomass process is modeled as the sum of the covariate effect and a random spatial field which represents the spatial structure not captured  by the covariate.
+This is possible as we extend tthe simulation to several statistical rectangles.
+We also simulate precisely localized scientific data as another source of information used to infer the spatial hidden biomass field.
 
 The study area is based on the case study; it includes the whole coast of the Bay of Biscay and cover several statistical rectangles (Figure \@ref(fig:MapSeveral)A).
-To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration).
-Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area.
-Similarly to the single-square simulations, the sampling of the commercial points of a declaration are realized in 3 steps.
+To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 individual catches per declaration).
+Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 of the area.
+Similarly to the single-square simulations, the sampling of the commercial fishing points associated with a declaration is realized in three steps.
 (1) The declaration is randomly affected to one of the ICES rectangle.
-(2) The centroid of the fishing zone is randomly sampled within this statistical rectangle.
+(2) The centroid of the fishing zone is uniformely  sampled within this statistical rectangle.
 (3) The 10 fishing punctual observations are randomly sampled within the fishing zone.
-The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km.
+The side of the squared fishing zone is set so as the extent of a fishing operation does not exceed 30 km.
 Note that we do not explore the effect of exploring several zones within the same declaration as it is already done in the single-square simulations.
 
-100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A).
-Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table 1).
+100 scientific precisely locazed scientific fishing points  are simulated following a random stratified plan; contrary to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A).
+Scientific observations are simulated following the observation equations of $\mathcal{M}_Y$ (with specific parameters for scientific data - Table 1).
 
@@
 
-- to assess what brings our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.
+- to assess what brings our alternative approach, we compare the Reallocated Model to the Declaration Model.
 
@@
 
-In addition to the 2 metrics introduced at the beginning of the section ($MSPE$ and species-habitat parameter $\beta_S$), we also compare the simulated/estimated values of the range parameter.
-
-Note that when including 2 datasets in the model, a difference in catchability should be accounted for and one of the two dataset will be considered as reference (see @alglave_combining_2022 for more details on parameterization).
+In addition to the 2 metrics introduced at the beginning of the section ($MSPE$ and species-habitat parameter $\beta_S$), we also compare the precision of the estimate for the range parameter.
+
+Note that when including 2 datasets in the model, a difference in efficiency (named catchability in the Fisheries Science vocabulary)  should be accounted for and one of the two dataset will be considered as reference (see @alglave_combining_2022 for more details on parameterization).
 Here the reference level is set with scientific data and the parameter $k_{com}$ accounts for the relative ratio between scientific and commercial observation values.
-In simulations, $k_{com}$ is set to 1 so that both datasources are assumed to have the same catchability.
-
-## Case-study: sole of the Bay of Biscay
-
-To illustrate our method on a real case study, we applied the approach to the common sole of the Bay of Biscay.
-VMS-logbooks data were extracted for the bottom trawlers fleet (OTB).
-The methods to cross VMS-logbooks data and to filter the fleet is already extensively described in the previous papers [@alglave_combining_2022] and is not developped further here.
-Scientific data were extracted from the DATRAS database for the Orhago beam trawl survey [@biais_gerard_orhago_2003;@ices_report_2018].
-To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - @ices_report_2018-1).
-To illustrate the method, we compare the outputs of (1) the scientific-based model, (2) the integrated model fitted to scientific data and reallocated commercial data ($Y^*_i$ model) and (3) the integrated model fitted to scientific and declarations data ($D_j$ model).
-
-The $D_j$ model faced convergence issues (some of the parameters were hardly estimated e.g.
-the range parameter).
-To ease convergence, we integrated in the analysis onboard observer data for the same fleet.
-They are considered as exact commercial data of catches.
-Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
-observation variance and zero-inflation parameter of commercial data).
-
-Furthermore, we adopt a step estimation procedure to initialize the optimization algorithm for the $D_j$ model.
-We first fit the $Y^*_i$ model and use the estimates of the model to initialize the optimization of the $D_j$ model.
-We eventually fix the parameters that are hard to estimate in first steps of optimization (intercept $\mu$, covariate effect $\beta_S$, range and marginal variance) and let them free iteratively.
-
-# Results
-
-\newpage
-
-## Multiple-square simulations
+In simulations, $k_{com}$ is set to 1 so that both data sources are assumed to have the same catchability.
+XXXX du coup est ce qu'on a intérêt à en parler ici ??
+XXXXX
 
@@
 
-Whether the model is fitted at the punctual level or at the declaration level, the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models.
+The contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model at the declaration level or at the individual reallocated catch level  are always smaller than their single-data counterparts.
 This can be well illustrated from Figure \@ref(fig:MapSeveral).
@@
 
-Furthermore, consistently with single-square simulations, uniform reallocation conducts to a loss in both the predictions accuracy and the species-habitat relationship (Figure \@ref(fig:PerfMetricSeveral)) compared to the model built on commercial declarations ($D_j$).
+Furthermore, consistently with single-square simulations, the Reallocated Model conducts to a loss in both the predictions accuracy and the species-habitat relationship (Figure \@ref(fig:PerfMetricSeveral)) compared to the Declaration Model.
 
 Interestingly, in addition to the species-habitat relationship, uniform reallocation also affects the spatial autocorrelation terms such as the range parameter.
-The model fitted on reallocated data ($Y_i^*$) provides biased range estimates while the $D_j$ model provides unbiased estimates.
+The Reallocated model provides biased range estimates while the Declaration model provides unbiased estimates.
 This is a consequence of the loss of the species-habitat relationship: when uniformly reallocating declarations, part of the variability related to the covariate effect is captured by the random effect.
 Consequently, the range parameter captures both the autocorrelation related to the actual random effect and to the covariate.
-The $D_j$ model allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
-This is evidenced in Figure \@ref(fig:MapSeveral) where the model fitted to reallocated catch $Y_i^*$ provides smoothed maps and does not capture the relatively small scale patterns that are shaped by the covariate.
-On the other hand, the $D_j$ model (as the scientific-based model) better captures and disentangles the covariate effect and the spatial random effect and then provides predictions that better fit to the small-scale patterns of the species distribution.
+The Declaration Model allows to recover and disentangle the effect of the species-habitat relationship and of the random effect.
+This is evidenced in Figure \@ref(fig:MapSeveral) where the Reallocated Model produces smoothed maps and does not capture the relatively small scale patterns that are shaped by the covariate.
+On the other hand, the Declaration model (as the scientific-based model) better captures and disentangles the covariate effect and the spatial random effect and then provides predictions that better fit to the small-scale patterns of the species distribution.
 However, this goes with some difficulty in convergence as only 75% of the model built on catch declarations converge (Table 3).
@@
 $Y_i^*$: $Y_i^*$ model (D, E).
-$D_j$: $D_j$ model (F, G).
+$D_j$:Declaration Model (F, G).
 Scientific model: model fitted to scientific data only.
@@
 
-<!-- see the file 'Scripts/In_progress/plot_multiple_square_2.R' -->
-
-## Real case study
+## Case-study: sole of the Bay of Biscay
+
+To illustrate our method on a real case study, we applied the approach to the common sole of the Bay of Biscay.
+VMS-logbooks data were extracted for the bottom trawlers fleet (OTB).
+The methods to cross VMS-logbooks data and to filter the fleet is already extensively described in the previous papers [@alglave_combining_2022] and is not developed further here.
+Scientific data were extracted from the DATRAS database for the Orhago beam trawl survey [@biais_gerard_orhago_2003;@ices_report_2018].
+To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - @ices_report_2018-1).
+To illustrate the method, we compare the outputs of (1) the Spatial model fitted with scientific data, (2) the integrated model Realocated model fitted to scientific data and reallocated individual catch data and (3) the integrated Declaration Model fitted to scientific and declarations data.
+
+The integrated Declaration Modelfaced convergence issues (some of the parameters were hardly estimated e.g.
+the range parameter).
+To ease convergence, we integrated in the analysis onboard observer data for the same fleet.
+They can be considered as individual commercial catch data.
+There are few vessels with onboard observers and therefore few precisely geolocalized commercial data XXXX Dire combien ??
+XXX Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
+observation variance and zero-inflation parameter of commercial data).
+
+Furthermore, as commonly done in complex fisheries model using automatic differentiation method [@fournier2012ad], we adopt a phase optimization procedure to initialize the optimization algorithm for the Declaration model.
+We first fit the Reallocated model and use the estimates of this model as starting point of the optimization algorithm used for the Declaration model estimation..
+We eventually fix the parameters that are hard to estimate in the initial optimization phases (intercept $\mu$, covariate effect $\beta_S$, range and marginal variance) to finally let them be  freely estimated in the following phases.
 
@@
 
-Consistently with simulations, the $D_j$ model modifies some of the parameters estimates and revises the spatial pattern of the species distribution compared with the $Y^*_i$ model.
-In particular, the substrate effect is recovered in the $D_j$ model and fall in the same range as the scientific-based estimate (Figure \@ref(fig:CaseStudyMap)).
+Consistently with simulations, the Declaration model modifies some of the parameters estimates and revises the spatial pattern of the species distribution compared with the Reallocated Model.
+In particular, the substrate effect is recovered in the Declaration Model and fall in the same range as the scientific-based estimate (Figure \@ref(fig:CaseStudyMap)).
 The zero-inflation parameter $\xi$ is revised downwards (i.e.
 there are actually more zero-values than in the reallocated data) while the observation variance of commercial data is revised upwards (i.e.
-the commercial data are noisier than expected when fitting the model on $Y_i^*$).
+the commercial data are noisier than estimated with the Reallocated Model).
 
 In addition, uncertainty is also revised when fitting the model at the declaration level.
-For instance, when comparing the $Y^*_i$ model to the $D_j$ model, the confidence intervals of $\beta_S$, the marginal variance, the range, $\xi_{com}$, $\sigma_{com}$ are much wider.
-This emphasizes that uncertainty is probably underestimated in the $Y^*_i$ model compared with the $D_j$ model.
-Now, when comparing the scientific-based model and the integrated $D_j$ model, some parameters are better estimated than when only scientific data feed the model.
-For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on $D_j$ substrate is significant and the confidence interval is smaller.
-
-On the contrary, other parameters do not seems be well estimated in either the $Y_i^*$ or the $D_j$ models.
-For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on $Y_i$ and revised downards when working on $D_j$.
-This is consistent with simulations results, see Figure \@ref(fig:ParBiasSingle).
-
-Regarding the maps of the species distribution, fitting the model at the $D_j$ level strongly modifies the model biomass field compared with the $Y_i^*$ model.
-In particular, the substrate covariate have a sharper effect on species distribution and the intensity of the hotspots are revised when fitting the model on $D_j$.
-
-Finally, the $D_j$ model fitted only on commercial data does not converge (while the one fitted on $Y_i^*$ does) emphasizing the model fitted on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
+For instance, when comparing the Reallocated Model to the Declaration Model, the confidence intervals of $\beta_S$, the marginal variance, the range, $\xi_{com}$, $\sigma_{com}$ are much wider.
+This emphasizes that uncertainty is probably underestimated in the Reallocated Model compared with the Declaration Model.
+Now, when comparing the scientific-based model and the integrated Declaration Model, some parameters are better estimated than when only scientific data feed the model.
+For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on the declarations, substrate is significant and the confidence interval is smaller.
+
+On the contrary, other parameters do not seem  well estimated in either the Reallocated or the Declaration Models.
+For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on the individual precisely geolocalized catch  and revised downwards when estimatedwith the Reallocated Model.
+This is consistent with the simulations results, see Figure \@ref(fig:ParBiasSingle).
+
+Regarding the maps of the species distribution, fitting the model at the declaration level strongly modifies the model biomass field compared with the Reallocated model.
+In particular, the substrate covariate have a sharper effect on species distribution and the intensity of the hotspots are revised when fitting the Declaration model.
+
+Finally, the Declaration Model fitted only on commercial data does not converge (while the one fitted on reallocated data does) emphasizing the model fitted on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
+This might be explained by the relative loss of spatail information due to the data agreagation.
 

### 2022-06-02 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3

@@
 
-**Attention:** L'introduction n'est pas finie
-
 **Context**
@@
 Reconciling spatial (and eventually temporal) scales properly when different set of observations do not have the same resolution is a well known issue in geography, ecology, agriculture, geology and statistics [@gotway2002combining].
-Inferring values of a variable at points or blocks different from those that have been observed is commonly refered as the *change of support problem* (COS) in the statistical literature [@gotway2002combining and @gelfand2010misaligned].
+Inferring values of a variable at points or blocks different from those that have been observed is commonly refered as the *change of support problem* (COS) in the statistical literature [@gotway2002combining,@gelfand2010misaligned].
 COS issues can be classified within 2 categories - that are not necessarily mutually exclusive - [@pacifici2019resolving]:
@@
 Our case of application clearly refers to MAUP.
-Note that within this category, another category exist and is commonly refered to the 'ecological fallacy'.
+Note that within this category, another category exist and is commonly refered as the 'ecological fallacy'.
 This occures when the individual response to a covariate differs from the response estimated from grouping individuals.
@@
 
-#  Introduction 
-* Commencer par la motivation peche
---> knowledge of essential habitat --> Species distribution model --> link between biomass and habitat to identify essential habitat
--->  Data from scientific survey, precisely geolocalized standard sampling protocol, expensive and time consuming
---> On the other hand declaration data, spatialized at coarser level  but cheap and reglementaire
-Combining the two sources of data.
-
-This problem is not unique, other exampl in terms of apllication
-
-Change of support form fine to coarse sptial level
-
-A solution the reallocation process
-
-** Pour la discu, simulation conditionnelle ? **
-
-We proposed an alternative approach to account for the change of support and we compare it to the approach proposed in alglave but also to a golden standard to assess the consequences of these methods.
-
-Paper organized as follow
-
-Section 2. presentation of the reference model and detail the reallocation process
-
-Section 3. Presnetation of an alternaion way to deal with change of support
-
-Section 4. simulation study to asses the bias and/or the loss of precision of the two methodes compared with the golden standard
-
-Section 5. Appllication on sole of the Bay of Biscay
-
-Section 6. Discu/conclu
-
 ## A spatialized catch model for aggregated data
@@
 In  @alglave_combining_2022, the authors propose a hierarchical spatial model to combine scientific survey data obtained through a standardized sampling protocol and catch data harvested by fishermen.
-This section provides a brief overview of the key ingredient of this model and raises some concerns on the spatialization of the commercial catch data.
+This section provides a brief overview of the key ingredient of this model and raises the main concerns on the spatialization of the commercial catch data.
 
 Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ a spatial random field which represents the biomass for a species of interest.
-$\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta \, \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$ and $\Gamma$ a field of covariate.
+$\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta \, \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$ and $\bm{\Gamma}=(\Gamma(x),x\in\mathcal{D})$ a field of covariate.
 As the catch are zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ with a mixture of a Dirac mass at $0$ and a Log Normal distribution, the proportion of the mixture being defined conditionally on the random field $\bm{S}$.
@@
 \begin{equation}
-\label{eq:defY}
 Y_i | S(x_i)  \overset{ind}{\sim} \My{p_i}{\mu_i}{\sigma^2},
 \end{equation}
-with $p_i \defeq \exp(-e^\xi S(x_{i}))$ the proportion of the mixture, $\mu_i\defeq 
-\frac{S(x_i)}{1-p_i}$ the expected catch when positive and $\sigma^2$ a transformation of its variance variance. This parametrization (see SMXXXX) of the mixture model corresponds to:
+
+with $p_i \defeq \exp(-e^\xi S(x_{i}))$ the proportion of the mixture, $\mu_i\defeq \frac{S(x_i)}{1-p_i}$ the expected catch when positive and $\sigma^2$ a transformation of its variance.
+This parametrization (see SMXXXX) of the mixture model corresponds to:
 \begin{align}
@@
 
-<!-- 
-Je pense qu'il ne faut pas parler de la grille discrete tout de suite car S est défini sur un espace continu et moralement l'intensité du processus de poisson aussi.
-le passage à la grille est la pour simplifier l'expression de la vraisemblance, on peut dire qu'on fait une approximation pour des raisons numériques
-
-We consider a discrete study domain where $x \in [\![1,n]\!]$ stands for the discrete locations.
-$\beta .
-\Gamma(x)$ is the covariate term with $\Gamma$ the matrix of design of the covariates and $\beta$ the effect of the covariates.
-$\delta(x)$ is a GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$).
--->
-
 In this approach all fishing locations $x_i$ and the coressponding catch are supposed to be known.
 This is classically the case with scientific survey for which every catch is recorded as well as its geolocalization.
-
-<!-- however those fishinf locations are actually derived from a different dataset containing the hourly localisation of the the commercial vessels, the VMS dataset.
-This dataset is processed to identify the fishing location among all the recorded locations.
--->
-
-<!-- At each fishing location $x_i$, fishermen realize a catch $Y_i$ conditionnally on the latent field values (referred hereafter as a punctual catch).
-Observations are supposed to follow some distribution $\mathcal{L}_Y$.
--->
-
-<!-- $$Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma^2)$$ -->
-
-<!-- In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.
-$\mathcal{L}_Y$ is decomposed it 2 parts: -->
-
-<!-- - the probability to obtain a zero-catch which is modelled as a Bernoulli variable with probability $p_{i}=\exp(-e^\xi .S(x_{i}))$.
-$\xi$ is a parameter controlling zero-inflation.
-When $\xi$ increases, the amount of zero in the data decreases.
--->
-
-<!-- - if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a positive continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$.
--->
-
-<!-- $$ -->
-<!-- \operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i)\right) = -->
-<!-- \left\{ -->
-<!--     \begin{array}{ll} -->
-<!--         p_i & \text { if } y_{i}=0 \\ -->
-<!--         \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0 -->
-<!--     \end{array} -->
-<!-- \right. -->
-<!-- $$ -->
 
@@
 
-$$D=\sum_{x_i \in \mathcal{A}_D}{Y(x_i)}.$$
+$$D=\sum_{i|x_i \in \mathcal{A}_D}{Y_i}$$
 
 Since the original model proposed in @alglave_combining_2022 is defined at the individual catch scale, those catch are derived from the logbooks data combined with the VMS data.
-This is classically done in fisheries science REFREFEREFREF by using a uniform reallocation process.
-This process consists in counting, for a given vessel, on a given day,  the number $m_j$ of fishing points in $\mathcal{A}_{D_j}$ associated with declaration $D_j$ and define for each $x_i\in\mathcal{A}_{D_j},$ the associated reallocated individual catch $\Yr(x_i)\defeq D_j/m_j$.
-As a direct consequence, the so reconstructed individual catch are similar to quantity widely used in Fishereis Sceince named  Catch Per Unit of Effort (CPUE), effort being measure in hours.
+This is classically done in fisheries science by using a uniform reallocation process [@hintzen_vmstools_2012,murrayEffectivenessUsingCPUE2013].
+This process consists in counting, for a given vessel, on a given day, the number $m_k$ of fishing points in $\mathcal{A}_{D_k}$ associated with declaration $D_k$ and define for each $x_i\in\mathcal{A}_{D_k},$ the associated reallocated individual catch $\Yr_i\defeq D_k/m_k$.
+As a direct consequence, the so reconstructed individual catch are similar to quantity widely used in fisheries science named Catch Per Unit of Effort (CPUE), effort being measured in hours.
 As noted by @alglave_combining_2022, this process has several drawbacks.
@@
 
-<!-- In standard processing, $D_j$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$.
+<!-- In standard processing, $D_k$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_k/m_k$.
 This is what we call uniform (or proportional) reallocation.
@@
 
-<!-- To overcome such limitations, an alternative is to consider that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables).
+<!-- To overcome such limitations, an alternative is to consider that only the catch declarations $D_k$ are observed while $Y_i$ are not (they are latent variables).
 Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->
 
-To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_j$ instead of the reconstructed individual catch $\Yr(x_i)$.
-As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_j$ which consists of a sum of $m_j$ random variables following a mixture distribution.
+To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_k$ instead of the reconstructed individual catch $\Yr_i$.
+As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_k$ which consists of a sum of $m_k$ random variables following a mixture distribution.
 This distribution has no known analytical form.
-However the catch declaration data also exhibit some zero inflation and a long tail repartition of the values: a mixture model would be a good candidate.
-Therefore, we propose to use the same family of distribution for  $D_j$ than the one use for the individual catch $Y$: a mixture between a Dirac mass in $0$ and a lognormal distribution.
-This is not exact, as the distribution of $D_j$ should be defined as convolution of mixtures which  would result in an untractable model.
-We somehow approximate the sum of  the original mixture distributions  by a mixture distribution of the same kind and we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original proposition:
-
-<!-- that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables).
+However, the catch declaration data also exhibit some zero inflation and a long tail repartition of the values, thus as for $Y_i$ a mixture model is a good candidate to model $D_k$.
+We define then the distribution of $D_k$ through the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original proposition:
+
+We somehow approximate the sum of the original mixture distributions by a mixture distribution of the same kind and we define the different key quantities, i.e.
+the mixture proportion, the expected positive catch declaration and its variance in order to match the original proposition:
+
+$$D_k | \mathcal{P}_k,\bm{S} \sim \Md{p^D_k}{\mu^D_k}{\sigma^{2,D}}$$
+
+with $\mathcal{P}_k=(1,...,i, ...,m_{k})$, the list of the fishing positions $(x_1,...,x_i, ...,x_{m_k})$ associated to declaration $D_k=\sum_{i \in \mathcal{P}_k}{Y_i}$, $\mu^D_k$ the expected positive biomass, $p_k^D$ the proportion of the mixture and $\sigma^{2,D}$ the variance parameter.
+
+<!-- that only the catch declarations $D_k$ are observed while $Y_i$ are not (they are latent variables).
 Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->
 
-$$D_j | \mathcal{P}_j,\bm{S} \sim \Md{p^D_j}{\mu^D_j}{\sigma^{2,D}}$$
-
-with $\mathcal{P}_j=(x_1,...,x_i, ...,x_{m_j})$, the list of the fishing points associated to declaration $D_j$,  $\mu^D_j$ the expected positive biomass, $p_j^D$ the proportion of the mixture and $\sigma^{2,D}$ the variance parameter.
-
-<!-- As $D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$, it is possible to relate $D_j$, $Y_i$ and $S_{\mathcal{P}_j}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$.
-Our approach is to match the moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_j$.
+<!-- As $D_k=\sum_{i \in \mathcal{P}_k}{Y_{i}}$, it is possible to relate $D_k$, $Y_i$ and $S_{\mathcal{P}_k}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$.
+Our approach is to match the moments of $D_k$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_k$.
 This is what we call model-based reallocation.
@@
 
-<!-- In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | \mathcal{P}_j,S_{\mathcal{P}_j}$.
+<!-- In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_k$ instead of $Y_i \vert x_i, S(x_i)$ and $D_k | \mathcal{P}_k,S_{\mathcal{P}_k}$.
 -->
@@
 
-In order to define compatible models for $Y$ the actual individual catches and $D$ the catch declarations, we choose to match the key quantities of the two distributions.
-
-(1) As the $Y_1, \ldots, Y_{m_j}$ are independent conditionally on $\bm{S}$,  the probability to obtain a zero-declaration $P(D_j = 0\vert \bm{S}, \mathcal{P}_j )$ is obtained by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.
+In order to define relate the individual observation level $Y$ and the catch declaration level $D$, we choose to match the key quantities of the two distributions.
+
+(1) As the $Y_1, \ldots, Y_{m_k}$ are independent conditionally on $\bm{S}$,  the probability to obtain a zero-declaration $\P(D_k = 0\vert \bm{S}, \mathcal{P}_k )$ is obtained by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $i \in \mathcal{P}_k$.
 
 \begin{align*}
-\P(D_j = 0\vert \bm{S}, \mathcal{P}_j)  & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0\vert \bm{S}, x_i),\nonumber \\
-                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = p_j^D
+\P(D_k = 0\vert \bm{S}, \mathcal{P}_k)  & = \P_{i \in \mathcal{P}_k} P(Y_{i} = 0\vert \bm{S}, x_i),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i \in \mathcal{P}_k} e^{\xi}. S(x_{i})\right \rbrace} = p_k^D
 \end{align*}
@@
 
-$$\E(D_j \vert D_j > 0, \bm{S}, \mathcal{P}_j)= \frac{\sum_{x_i \in \mathcal{P}_j} \E(Y_i \vert S(x_i)) }{1-p_j^D} = \frac{\sum_{x_i \in \mathcal{P}_j} S(x_{i})}{1-p_j^D},$$
-$$\V ar(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} \V ar(Y_{i})}{1-p_j^D} - \frac{\pi_j}{(1-p_j^D)^2}E(D_j)^2,$$
+$$\E(D_k \vert D_k > 0, \bm{S}, \mathcal{P}_k)= \frac{\sum_{i \in \mathcal{P}_k} \E(Y_i \vert S(x_i)) }{1-p_k^D} = \frac{\sum_{i \in \mathcal{P}_k} S(x_{i})}{1-p_k^D},$$
+$$\V ar(D_k \vert D_k > 0) = \frac{\sum_{i \in \mathcal{P}_k} \V ar(Y_{i})}{1-p_k^D} - \frac{\pi_k}{(1-p_k^D)^2}E(D_k)^2,$$
 $$\mbox{with } \V ar(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i})).$$
 
-By defining the observation process at the declaration level,  we expect to  avoid some of the drawbacks of the estimation based on reallocated individual catch data $Y^r$.
+Knowing the moment of $D_k$ (that only depends on the distribution of $Y_i$), it is required then to define a probability distribution for $D_k|D_k>0$ that is adapted to the underlying individual observations and express $\P(D_k|D_k>0)$ as a function of $\E(D_k \vert D_k > 0, \bm{S}, \mathcal{P}_k)$ and $\V ar(D_k \vert D_k > 0)$.
+We propose to use the same family of distribution for $D_k|D_k>0$ than the one used for the individual catch $Y_i|Y_i>0$ i.e.
+a lognormal distribution.
+This is an approximation that we discuss hereafter.
+
+By defining the observation process at the declaration level, we expect to avoid some of the drawbacks of the estimation based on reallocated individual catch data $\Yr$.
+
+Finally, note that the scientific data is considered known at each fishing locations and thus they can be simply modelled as standard individual data $Y_i$.
 
 The inference is based on  maximum likelihood approach with two simplifications.
-We use the Stochastic Partial Differential Equations (SPDE) approach to represent the spatial field and we use integrated Nested Laplace approximation to compute le likelihood.
+We use the Stochastic Partial Differential Equations (SPDE) approach to represent the spatial field [@lindgren_explicit_2011] and we use Laplace approximation to compute the likelihood.
 The stochastic random field is also approximated by a piecewise constant process defined on a fine grid.
 The optimization of the likelihood relies Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].
-
-<!-- XXXX  Il faut aussi présenter la partie scientific ou dire que c'est comme dans Alglave 2022 -->
 
@@
 Fishermen sample catches in areas of poor biomass where the covariate is relatively low (blue points) and in areas of higher biomass where the covariate is higher and eventually in the hotspot of biomass (orange and red points).
-These catches belong to the same declaration $j$ and are summed to constitute the declaration $D_j=50$.
+These catches belong to the same declaration $j$ and are summed to constitute the declaration $D_k=50$.
 The declaration is declared at the level of the statistical rectangle.
 From VMS data, we know the fishing positions $x_i$.
-In standard processing, $D_j$ are then uniformly reallocated over the fishing positions $x_i$.
+In standard processing, $D_k$ are then uniformly reallocated over the fishing positions $x_i$.
 This strongly homogeneizes the catch.
-In particular, the effect of the habitat is no more evidenced in the reallocated catch $Y_i^*$.](images/realloc.png)
-
-```{r}
-
-# save(data=gridpolygon_sf_2,file="res/simu_lf.RData")
-# 
-# plot_1 <- ggplot(gridpolygon_sf_2)+
-#   geom_sf(aes(fill=Strue_x),alpha=0.9)+
-#   scale_fill_distiller(palette = "Spectral")+
-#   theme_void()+theme(legend.position = "none")
-# 
-# ggsave("images/simu_lf.png",width = 7.5,height = 7.5,dpi = 500)
-
-```
+In particular, the effect of the habitat is no more evidenced in the reallocated catch $Y_i^r$.](images/realloc.png)
 
@@
 This ideal situation, with no actual application, will be named 'Spatial Model' and used as a reference for the comparison between the two alternatives.
-* the original model fitted with  commercial reallocated individual catch (and potentially few precisely geolocalized scientific data  as done in  @alglave_combining_2022.
-This approach will be refereed as "Reallocated Model".
-* the alternative approach detailed before where the biomass model is fitted  using commercial catch declaration at a corase spatial level and potentially few precisely geolocalized scientific data.
+* the original model fitted with commercial reallocated individual catch (and potentially few precisely geolocalized scientific data) as done in @alglave_combining_2022.
+This approach will be refered as "Reallocated Model".
+* the alternative approach detailed before where the biomass model is fitted  using commercial catch declaration at a coarse spatial level and potentially few precisely geolocalized scientific data.
 This approach is named "Declaration Model".
@@
 In these two sets of simulation studies, there is a unique covariate,  modeled as a continuous GRF that we suppose known at each point of the grid (when present the spatial random component is also a GRF but its values are not observed).
-XXXX AJouter les paramètres du chmaps spatial 
+
 The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$.
@@
 
-We compare the performance of the Spatial Model (the gold standard), the Reallocated Model and the Declaration Model  model configurations in regards to 2 metrics:
+We compare the performance of the Spatial Model (the gold standard), the Reallocated Model and the Declaration Model configurations in regards to 2 metrics:
 
@@
 
+To get enough replicates, we run the simulations 100 time.
+
 ```{r}
 
-df <- data.frame(v1=c("$\\mu$","$\\beta_S$","$\\text{Range}$","$\\text{Marginal variance}$","$\\xi_{com}$","$\\sigma_{com}$","$k_{com}$","$\\xi_{sci}$","$\\sigma_{sci}$"),
-                 v2=c("2","2"," -- "," -- ","-1","1"," -- "," -- "," -- "),
-                 v3=c("2","2","1.5","0.5","-1","1","1","0","0.8"))
+df <- data.frame(v1=c("$\\mu$","$\\beta_S$","$\\text{Range } (\\delta)$","$\\text{Marginal variance }  (\\delta)$","$\\text{Range } (\\Gamma)$","$\\text{Marginal variance }  (\\Gamma)$","$\\xi_{com}$","$\\sigma_{com}$","$k_{com}$","$\\xi_{sci}$","$\\sigma_{sci}$"),
+                 v2=c("2","2"," -- "," -- ","10 (cells)","0.5","-1","1"," -- "," -- "," -- "),
+                 v3=c("2","2","0.6 ($\\approx$ 50 km)","1","1.5 ($\\approx$ 120 km)","0.5","-1","1","1","0","0.8"))
 
@@
 **Single-square simulations**
-XXX Infor sur le nombre de simus ???
 
@@
 The accuracy of the Reallocated Model outputs is expected to decrease when the number of fishing zones increases.
-To assess the effect of such process, we simulated the fishing locations associated with a declaration declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones.
-
-XXXX Un petit schéma pour expliquer le  processus de similuation et mettre des images sur les mots de fishing zone, fishing points
+To assess the effect of such process, we simulated the fishing locations associated with a declaration declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones (Figure \@ref(fig:SimuSevZone)).
+
+```{r}
+
+## 1 zone
+load(file="res/fish_1zones.RData")
+
+centres <- fish_zones_res$centres
+peche_com_old <- fish_zones_res$peche_com_old
+gridpolygon_sf <- fish_zones_res$gridpolygon_sf
+
+centres_sf <- st_as_sf(centres,coords = c("x","y"))
+buffer_sf <- st_buffer(centres_sf,dist = 3,endCapStyle = 'SQUARE')
+
+gridpolygon_sf_2 <- gridpolygon_sf[st_intersects(gridpolygon_sf,buffer_sf) %>% lengths > 0,] %>%
+  mutate(centr = 1) %>%
+  dplyr::select(-layer)
+
+one_zones_plot <- ggplot()+
+  geom_point(data=centres,aes(x=x, y=y, col=as.factor(boats)),
+             shape = 4,size=4)+
+  geom_point(data=peche_com_old,aes(x=x, y=y, col=as.factor(boats)), size=4)+
+  geom_sf(data = gridpolygon_sf,alpha=0)+
+  geom_sf(data = gridpolygon_sf_2,alpha=0.2,fill="red")+
+  theme_void()+
+  theme(legend.position = "none",
+        plot.title = element_text(hjust = 0.5))+
+  ggtitle("One zone")
+
+## 3 zones
+load(file="res/fish_3zones.RData")
+
+centres <- fish_zones_res$centres
+peche_com_old <- fish_zones_res$peche_com_old
+gridpolygon_sf <- fish_zones_res$gridpolygon_sf
+
+centres_sf <- st_as_sf(centres,coords = c("x","y"))
+buffer_sf <- st_buffer(centres_sf,dist = 3,endCapStyle = 'SQUARE')
+
+gridpolygon_sf_2 <- gridpolygon_sf[st_intersects(gridpolygon_sf,buffer_sf) %>% lengths > 0,] %>%
+  mutate(centr = 1) %>%
+  dplyr::select(-layer)
+
+three_zones_plot <- ggplot()+
+  geom_point(data=centres,aes(x=x, y=y, col=as.factor(boats)),
+             shape = 4,size=4)+
+  geom_point(data=peche_com_old,aes(x=x, y=y, col=as.factor(boats)), size=4)+
+  geom_sf(data = gridpolygon_sf,alpha=0)+
+  geom_sf(data = gridpolygon_sf_2,alpha=0.2,fill="red")+
+  theme_void()+
+  theme(legend.position = "none",
+        plot.title = element_text(hjust = 0.5))+
+  ggtitle("Three zones")
+
+## 5 zones
+load(file="res/fish_5zones.RData")
+
+centres <- fish_zones_res$centres
+peche_com_old <- fish_zones_res$peche_com_old
+gridpolygon_sf <- fish_zones_res$gridpolygon_sf
+
+centres_sf <- st_as_sf(centres,coords = c("x","y"))
+buffer_sf <- st_buffer(centres_sf,dist = 3,endCapStyle = 'SQUARE')
+
+gridpolygon_sf_2 <- gridpolygon_sf[st_intersects(gridpolygon_sf,buffer_sf) %>% lengths > 0,] %>%
+  mutate(centr = 1) %>%
+  dplyr::select(-layer)
+
+five_zones_plot <- ggplot()+
+  geom_point(data=centres,aes(x=x, y=y, col=as.factor(boats)),
+             shape = 4,size=4)+
+  geom_point(data=peche_com_old,aes(x=x, y=y, col=as.factor(boats)), size=4)+
+  geom_sf(data = gridpolygon_sf,alpha=0)+
+  geom_sf(data = gridpolygon_sf_2,alpha=0.2,fill="red")+
+  theme_void()+
+  theme(legend.position = "none",
+        plot.title = element_text(hjust = 0.5))+
+  ggtitle("Five zones")
+
+sev_zones_plot <- plot_grid(one_zones_plot,
+                            three_zones_plot,
+                            five_zones_plot,nrow = 1)
+
+ggsave("images/sev_zones_plot.png",width = 9, height = 3)
+
+```
+
+![(\#fig:SimuSevZone) Simulations of 10 fishing points within 1, 3 and 5 fishing zones.
+The full grid correspond to a statistical rectangle.
+Cross are the centroid of the fishing zones.
+A declaration declared at the level of the statistical rectangle would be uniformly reallocated over these fishing points.](images/sev_zones_plot.png)
 
@@
 
-For the single square simulation, in addition to the metrics introduced before ($MSPE$ and species-habitat parameter $\beta_S$), we also assess quality of the estitmation for  $\mu$, the observation variance parameter $\sigma^2$ and the zero-inflation parameter $\xi$.
+For the single square simulation, in addition to the metrics introduced before ($MSPE$ and species-habitat parameter $\beta_S$), we also assess the quality of the estimation for  $\mu$, the observation variance parameter $\sigma^2$ and the zero-inflation parameter $\xi$.
 
@@
            perc_convergence = round((1 - Convergence.sum / one.sum)*100,digits = 3)) %>%
-  mutate(aggreg_obs = ifelse((aggreg_obs == "Yi" & reallocation == "Yes"),"Yi*",aggreg_obs))
+  mutate(aggreg_obs = ifelse((aggreg_obs == "Yi" & reallocation == "Yes"),"Yr",aggreg_obs))
 
@@
   mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dj","Yi"))) %>%
-  mutate(relloc_aggreg = ifelse(aggreg_obs == F & reallocation == 1,"Reallocation: Yes -  Likelihood: Yi*",relloc_aggreg))
+  mutate(relloc_aggreg = ifelse(aggreg_obs == F & reallocation == 1,"Reallocation: Yes -  Likelihood: Yr",relloc_aggreg))
 
@@
                                 levels = c("Reallocation: No -  Likelihood: Yi",
-                                           "Reallocation: Yes -  Likelihood: Yi*",
+                                           "Reallocation: Yes -  Likelihood: Yr",
                                            "Reallocation: Yes -  Likelihood: Dj"))
@@
   scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
-                                "Reallocation: Yes -  Likelihood: Yi*",
+                                "Reallocation: Yes -  Likelihood: Yr",
                                 "Reallocation: Yes -  Likelihood: Dj"),
@@
   scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
-                               "Reallocation: Yes -  Likelihood: Yi*",
+                               "Reallocation: Yes -  Likelihood: Yr",
                                "Reallocation: Yes -  Likelihood: Dj"),
@@
 ```
-The resultats are presented in Figures \@ref(fig:ParBiasSingle) and \@ref(fig:ParBiasSingle).
+
+The resultats are presented in Figures \@ref(fig:PerfMetricSingle) and \@ref(fig:ParBiasSingle).
 
@@
 The number of fishing zones visited within each declaration is represented on the x-axis.
-The result of the Spatial model is in yellow, in red the results of the Reallocated model and in green the Declaration model.
-Simulations conducted with 10 fishing positions are not represented as they encouter convergence issues as stated in Table 2.](images/Perf.metric_single_square.png){width=75%}
+The result of the Spatial model is in yellow, in red the results of the Reallocated model and in green the Declaration Model.
+Simulations conducted with 10 fishing positions are not represented as they encounter convergence issues as stated in Table 1.](images/Perf.metric_single_square.png){width=75%}
 
@@
 'Reallocation: ', data are or are not reallocated in simulations.
-'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $Y_i^*$ or on catch declarations $D_j$.
+'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $\Yr$ or on catch declarations $D_k$.
 Gold: golden standard.
-Red: uniform reallocation ($Y_i^*$ model).
-Green: model-based reallocation ($D_j$ model).
-Only the simulations with 1000 fishing positionss are represented.
+Red: uniform reallocation ($\Yr$ model).
+Green: model-based reallocation ($D_k$ model).
+Only the simulations with 1000 fishing positions are represented.
 Black line: zero value.
@@
 Figure \@ref(fig:ParBiasSingle) shows the over-estimation of  the zero-inflation parameter ($\xi$) when using the Reallocated Model.
-This is not surpising since the uniform reallocation of commercial data tends to decrease the proportion of zero since a realocated catch $Y^r$ is non-zer as soon as at least one of the individual catch associated with the same declaration is non -zero.
+This is not surpising since the uniform reallocation of commercial data tends to decrease the proportion of zero since a reallocated catch $Y^r$ is non-zero as soon as at least one of the individual catch associated with the same declaration is non-zero.
 The observation variance ($\sigma$) is underestimated i.e.
@@
 
-The Declaration model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \@ref(fig:PerfMetricSingle)) even so the model outputs are not as accurate as the ones of the golden-standard.
+The Declaration model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \@ref(fig:PerfMetricSingle)) even so the model outputs are not as accurate as the ones of the Spatial Model.
 Furthermore, the zero-inflation parameter is unbiased when the model is fitted to catch declarations.
@@
 **Multiple-square simulations**
+
 As mentioned before, we propose another set of simulations designed to be closer from the case study.
-The latent biomass process is modeled as the sum of the covariate effect and a random spatial field which represents the spatial structure not captured  by the covariate.
-This is possible as we extend tthe simulation to several statistical rectangles.
+The latent biomass process is modeled as the sum of a covariate effect and a random spatial field which represents the spatial structure not captured by the covariate.
+The covariate is simulated with wider autocorrelation than the random effect (Table 1).
+This is possible as we extend the simulation to several statistical rectangles.
 We also simulate precisely localized scientific data as another source of information used to infer the spatial hidden biomass field.
@@
 
-100 scientific precisely locazed scientific fishing points  are simulated following a random stratified plan; contrary to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A).
+100 scientific precisely localized scientific fishing points are simulated following a random stratified plan; contrary to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A).
 Scientific observations are simulated following the observation equations of $\mathcal{M}_Y$ (with specific parameters for scientific data - Table 1).
@@
 In addition to the 2 metrics introduced at the beginning of the section ($MSPE$ and species-habitat parameter $\beta_S$), we also compare the precision of the estimate for the range parameter.
-
-Note that when including 2 datasets in the model, a difference in efficiency (named catchability in the Fisheries Science vocabulary)  should be accounted for and one of the two dataset will be considered as reference (see @alglave_combining_2022 for more details on parameterization).
-Here the reference level is set with scientific data and the parameter $k_{com}$ accounts for the relative ratio between scientific and commercial observation values.
-In simulations, $k_{com}$ is set to 1 so that both data sources are assumed to have the same catchability.
-XXXX du coup est ce qu'on a intérêt à en parler ici ??
-XXXXX
 
@@
 
-Results_conv_2$aggreg_obs[which(Results_conv_2$aggreg_obs == "Yi")] <- "Yi*"
+Results_conv_2$aggreg_obs[which(Results_conv_2$aggreg_obs == "Yi")] <- "Yr"
 
@@
 Results_conv_2$`Likelihood level`[which(Results_conv_2$Model == "Scientific model" &
-                       Results_conv_2$`Likelihood level` == "Yi*")] <- " "
+                       Results_conv_2$`Likelihood level` == "Yr")] <- " "
 
@@
 Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
-Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi*"
+Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yr"
 Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific data"
-Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi*","Dj","Scientific data"))
+Results_plot$obs <- factor(Results_plot$obs,levels = c("Yr","Dj","Scientific data"))
 
@@
   scale_color_manual(breaks = c("Scientific model",
-                                "Yi*",
+                                "Yr",
                                 "Dj"),
@@
 Red line: true value for the range and the species-habitat parameter ($\beta_S$).
-Red: uniform reallocation ($Y_i^*$ model).
-Green: model-based reallocation ($D_j$ model).
+Red: uniform reallocation ($\Yr$ model).
+Green: model-based reallocation ($D_k$ model).
 Blue: scientific-based model.](images/Perf.metric_multiple_square.png)
@@
 C: biomass field from the scientific-based model.
-$Y_i^*$: $Y_i^*$ model (D, E).
-$D_j$:Declaration Model (F, G).
+$\Yr$: Reallocated Model (D, E).
+$D_k$: Declaration Model (F, G).
 Scientific model: model fitted to scientific data only.
 Commercial model: model fitted to commercial data only.
-Integratedd model: model fitted to both data sources.](images/Map_multi_square.png)
+Integrated model: model fitted to both data sources.](images/Map_multi_square.png)
 
@@
 To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - @ices_report_2018-1).
-To illustrate the method, we compare the outputs of (1) the Spatial model fitted with scientific data, (2) the integrated model Realocated model fitted to scientific data and reallocated individual catch data and (3) the integrated Declaration Model fitted to scientific and declarations data.
-
-The integrated Declaration Modelfaced convergence issues (some of the parameters were hardly estimated e.g.
+To illustrate the method, we compare the outputs of (1) the Spatial model fitted with scientific data, (2) the integrated Reallocated model fitted to both scientific data and reallocated individual catch data and (3) the integrated Declaration Model fitted to both scientific and declarations data.
+
+The integrated Declaration Model faced convergence issues (some of the parameters were hardly estimated e.g.
 the range parameter).
@@
 They can be considered as individual commercial catch data.
-There are few vessels with onboard observers and therefore few precisely geolocalized commercial data XXXX Dire combien ??
-XXX Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
+There are few vessels with onboard observers and therefore few precisely geolocalized commercial data (86 samples).
+Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e.
 observation variance and zero-inflation parameter of commercial data).
@@
 Furthermore, as commonly done in complex fisheries model using automatic differentiation method [@fournier2012ad], we adopt a phase optimization procedure to initialize the optimization algorithm for the Declaration model.
-We first fit the Reallocated model and use the estimates of this model as starting point of the optimization algorithm used for the Declaration model estimation..
+We first fit the Reallocated model and use the estimates of this model as starting point of the optimization algorithm used for the Declaration model estimation.
 We eventually fix the parameters that are hard to estimate in the initial optimization phases (intercept $\mu$, covariate effect $\beta_S$, range and marginal variance) to finally let them be freely estimated in the following phases.
@@
   scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-  ggtitle("Integrated model",subtitle = "Commercial likelihood on Yi*")+
+  ggtitle("Integrated model",subtitle = "Commercial likelihood on Yr")+
   theme_bw()+
@@
 ## Parameter estimates
-est_par_df_full_3$lkl[which(est_par_df_full_3$lkl == "Yi")] <- "Yi*"
-est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi*","Scientific model"))
+est_par_df_full_3$lkl[which(est_par_df_full_3$lkl == "Yi")] <- "Yr"
+est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yr","Scientific model"))
 est_par_df_full_3 <- est_par_df_full_3 %>%
@@
 On the contrary, other parameters do not seem  well estimated in either the Reallocated or the Declaration Models.
-For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on the individual precisely geolocalized catch  and revised downwards when estimatedwith the Reallocated Model.
+For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on the individual precisely geolocalized catch  and revised downwards when estimated with the Reallocated Model.
 This is consistent with the simulations results, see Figure \@ref(fig:ParBiasSingle).
@@
 Finally, the Declaration Model fitted only on commercial data does not converge (while the one fitted on reallocated data does) emphasizing the model fitted on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.
-This might be explained by the relative loss of spatail information due to the data agreagation.
-
-![(\#fig:CaseStudyPar) Parameters obtained with scientific-based model, the integrated model fitted on reallocated catch $Y_i^*$ and the integrated model fitted on catch declarations $D_j$.](images/par_plot.png)
-
-![(\#fig:CaseStudyMap) Maps obtained from (left) the scientific-based model, (center) the integrated model fitted on punctual reallocated catch $Y_i^*$, (right) the integrated model fitted on catch declarations $D_j$.](images/case_study_plot.png)
+This might be explained by the relative loss of spatial information due to the data agreagation.
+
+![(\#fig:CaseStudyPar) Parameters obtained with scientific-based model, the integrated model fitted on reallocated catch $\Yr$ and the integrated model fitted on catch declarations $D_k$.
+$k_com$ is the relative difference in catchability between scientific data and commercial data (i.e.
+it is a scaling parameter).
+This parameter captures the difference in catch efficiency between scientific and commercial gears.](images/par_plot.png)
+
+![(\#fig:CaseStudyMap) Maps obtained from (left) the scientific-based model, (center) the integrated model fitted on punctual reallocated catch $\Yr$, (right) the integrated model fitted on catch declarations $D_k$.
+](images/case_study_plot.png)
 
 # Discussion
+
+**Summary of the main findings**
+We built an integrated spatial framework:
+--> that allows to handle COS issues with complex data
+--> that can combine rough resolution data (catch declaration data) and high resolution data (scientific data) to provide fine resolution species predictions
+
+The Declaration Model allows to recover the effect of the habitat on species distribution while this was lost when reallocating uniformly the declarations of VMS pings.
+This modifies the resulting predictions and resulting maps better fit to the covariate shape.
+--> Consistent with litterature of COS (when dealing roughly with COS --> biased estimates)
+
+**Some simplifications / aproximations that may be investigated further**
+--> we assumed the fishing locations are known exactly and without error, however in the reality boats fish along a trajectory which is unknown and emit VMS pings along this trajectory.
+Potential combination with other frameworks to reconstruct the trajectory of vessels.
+
+--> We modelled the positive declarations as a lognormal distribution and assume this result from a set of other zero-inflated lognormal distribution.
+This should be seen as a first approximation to handle the issue of COS in our specific case.
+Our simulations and previous studies [@monnahan2021incorporating] higlhight that this gives good results in practice.
+Other observation modelled could be explored (a Gamma distribution instead of a Lognormal for instance).
+However, the relation between the parameters of a Gamma distribution (usually the shape $k$ and the scale $\theta$) and its moments ($k=\E[X]^2/\V[X]$, $\theta=\V[X]/\E[X]$) can be more complex than the parameterization of the Lognormal distribution we used here ($\mu=E(X)$, $\sigma^2=ln(\frac{Var(Z)}{E(Z)^2} + 1)$) which might cause problem during convergence.
+
+--> Individual observations improve the convergence of the model (similarly as @gilbert2021integrating).
+Need for individual observations (Obsmer, Survey).
+More and more punctual data that may be available through direct recording of the catch.
+Value of individual observations [@plet2020value].
+
+--> Covariate range may have effect on the bias.
+
+**Perspectives**
+--> moving to space-time: no major challenge regarding the observation process (same ideas that should be involved), more challenging regarding the latent field as this would require to specify a covariance function that is meaningfull for such processes.
 
@@
 
-We model catch declarations $D_j$ (available at coarse resolution through logbooks data) as a sum of $Y_{i}$ punctual observations (which are unknown i.e.
+We model catch declarations $D_k$ (available at coarse resolution through logbooks data) as a sum of $Y_{i}$ punctual observations (which are unknown i.e.
 latent variables) each one realised at one fishing position $x_{i}$ (known through VMS data).
@@
 
-- $\mathcal{P}_j$ : the vector of all fishing positions related to the $j^{th}$ declaration.
+- $\mathcal{P}_k$ : the vector of all fishing positions $x_i$ related to the $j^{th}$ declaration.
 
@@
 
-- $i \in \{1,...,m_j\}$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.
-
-$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$$
+- $i \in \{1,...,m_k\}$ with $m_k$ the number of fishing positions belonging to the $j^{th}$ declaration.
+
+$$D_k=\sum_{i \in \mathcal{P}_k}{Y_{i}}$$
 
@@
 
-**$D_j$ probability distribution and moments**
-
-We have to express the probability distribution of $D_j$ and its moments as a function of $Y_{i}$ and its related moments.
-Let's assume $Y_{i} = C_{i}.Z_{i}$ is a zero-inflated lognormal distribution with $C_{i}$ a binary random variable and $Z_{i}$ a lognormal random variable.
+**$D_k$ probability distribution and moments**
+
+We have to express the probability distribution of $D_k$ and its moments as a function of $Y_{i}$ and its related moments.
+Let's assume $Y_{i} = C_{i}.Z_{i}$ is a zero-inflated lognormal distribution with $C_{i}$ and $Z_i$ the two components of the mixture.
+$C_{i}$ is a binary random variable and $Z_{i}$ a lognormal random variable.
 
@@
 
+As mentionned in the core text, the probability to obtain a zero declaration is the probability that all individual observations within this declaration are null. This gives:
+
 \begin{align*}
-P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0 \vert S, X),\nonumber \\
-                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
+P(D_k = 0 \vert S, X) & = \prod_{i \in \mathcal{P}_k} P(Y_{i} = 0 \vert S, X),\nonumber \\
+                      & = \exp{ \left \lbrace- \sum_{i \in \mathcal{P}_k} e^{\xi}. S(x_{i})\right \rbrace} = \pi_k.
 \end{align*}
@@
 
-Following calculations are supposed to be conditionnal on $S$ and $X$.
-
-$$E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}\vert \exists i\in\mathcal{P}_j , C_{i}=1)$$
+Conditionnally on $\bm{S}$ and $\mathcal{P}_j$.
 
 \begin{align*}
-E(D_j\vert D_j > 0) & = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / P\left ( D_j > 0\right ),   \\
-& = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / \left (1-\pi_j\right).   \\
+E(D_k\vert D_k >0) & =  \sum_{i \in \mathcal{P}_k} E(C_{i} Z_{i}\vert \exists x_i\in\mathcal{P}_k , C_{i}=1)
+& = E(D_k 1_{ \left \lbrace D_k > 0\right\rbrace } )  / P\left ( D_k > 0\right ),   \\
+& = E(D_k 1_{ \left \lbrace D_k > 0\right\rbrace } )  / \left (1-\pi_k\right).   \\
 \end{align*}
 
-As $E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } ) = E(D_j ),$
+As $E(D_k 1_{ \left \lbrace D_k > 0\right\rbrace } ) = E(D_k ),$ we can write $E(D_k\vert D_k > 0)$ as:
 
 \begin{align*}
-E(D_j\vert D_j > 0) & = \left (1- \pi_j\right)^{-1} E(D_j)   ,   \nonumber \\
-& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}),\nonumber \\
-& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_{i}) \frac{S(x_{i})}{1-p_{i}}, \nonumber \\
-& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} S(x_{i}).
+E(D_k\vert D_k > 0) & = \left (1- \pi_k\right)^{-1} E(D_k)   ,   \nonumber \\
+& = \left (1- \pi_k\right)^{-1} \sum_{i \in \mathcal{P}_k} E(C_{i} Z_{i}),\nonumber \\
+& = \left (1- \pi_k\right)^{-1} \sum_{i \in \mathcal{P}_k} (1-p_{i}) \frac{S(x_{i})}{1-p_{i}}, \nonumber \\
+& = \left (1- \pi_k\right)^{-1}\sum_{i \in \mathcal{P}_k} S(x_{i}).
 \end{align*}
@@
 
-$$Var(D_j \vert D_j >0)  = E(D_j^2 \vert D_j >0)- E(D_j \vert D_j >0)^2.$$
-
-$$E(D_j^2 \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 1_{\left \lbrace D_j >0\right\rbrace}) = (1-\pi_j)^{-1} E(D_j^2 )$$
-
-$$E(D_j \vert D_j >0)^2 = ((1-\pi_j)^{-1} E(D_j 1_{\left \lbrace D_j >0\right\rbrace}))^2 =  (1-\pi_j)^{-2} E(D_j)^2$$
-
-Then,
-
-$$Var(D_j \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 ) - (1-\pi_j)^{-2} E(D_j)^2 = (1-\pi_j)^{-1} Var(D_j) - \frac{\pi_j}{(1-\pi_j)^2} E(D_j)^2.$$
-
-As the $(Y_{i})_{i\in \mathcal{P}_j}$ are independent, $Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_{i}) = \sum_{i \in\mathcal{P}_j} Var(C_{i}.Z_{i})$.
+The variance then can expressed as:
+
+$$Var(D_k \vert D_k >0)  = E(D_k^2 \vert D_k >0)- E(D_k \vert D_k >0)^2.$$
+with,
+
+\begin{align*}
+E(D_k^2 \vert D_k >0) & = (1-\pi_k)^{-1} E(D_k^2 1_{\left \lbrace D_k >0\right\rbrace}) 
+& = (1-\pi_k)^{-1} E(D_k^2 )
+\end{align*}
+
+and
+
+\begin{align*}
+E(D_k \vert D_k >0)^2 & = ((1-\pi_k)^{-1} E(D_k 1_{\left \lbrace D_k >0\right\rbrace}))^2 
+& =  (1-\pi_k)^{-2} E(D_k)^2
+\end{align*}
+
+Then, using these two expressions in the variance formula we obtain:
+
+\begin{align*}
+Var(D_k \vert D_k >0) & = (1-\pi_k)^{-1} E(D_k^2 ) - (1-\pi_k)^{-2} E(D_k)^2
+& = (1-\pi_k)^{-1} Var(D_k) - \frac{\pi_k}{(1-\pi_k)^2} E(D_k)^2.
+\end{align*}
+
+As the $(Y_{i})_{i \in \mathcal{P}_k}$ are independent, $Var(D_k) =\sum_{i \in\mathcal{P}_k} Var(Y_{i}) = \sum_{i \in\mathcal{P}_k} Var(C_{i}.Z_{i})$.
+
+Obtaining $Var(C_{i} Z_{i})$ is then straightforward due to conditionnal independence properties:
 
@@
 
-$$P(D_j = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j$$
-
-$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
-
-$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
+The main formulas of the model can be summarised as follow:
+
+- The probability to obtain a zero declaration
+
+$$P(D_k = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i \in \mathcal{P}_k} e^{\xi}. S(x_{i})\right \rbrace} = \pi_k$$
+
+- The expectancy of a positive declaration
+
+$$E(D_k \vert D_k > 0)=\frac{\sum_{i \in \mathcal{P}_k} S(x_{i})}{1-\pi_k}$$
+
+- The variance of a positive declaration
+
+$$Var(D_k \vert D_k > 0) = \frac{\sum_{i \in \mathcal{P}_k} Var(Y_{i})}{1-\pi_k} - \frac{\pi_k}{(1-\pi_k)^2}E(D_k)^2$$
+
+- The variance of an individual observation
 
@@
 
-Assuming $D_j \vert D_j > 0$ also follows a lognormal distribution we can write:
-
-$$D_j \vert D_j > 0 \sim \operatorname{L}( \mu_j = E(D_j \vert D_j > 0), \sigma_j^2= ln(\frac{Var(D_j \vert D_j > 0)}{E(D_j \vert D_j > 0)^2} + 1))$$
+Then, assuming $D_k \vert D_k > 0$ also follows a lognormal distribution we can write:
+
+$$D_k \vert D_k > 0 \sim \operatorname{L}( \mu_k = E(D_k \vert D_k > 0), \sigma_k^2= ln(\frac{Var(D_k \vert D_k > 0)}{E(D_k \vert D_k > 0)^2} + 1))$$

### 2022-06-02 Version_M2_plusieurszones/Scripts/function/commercial_scientific_14.R

@@ -304,38 +304,38 @@ simu_commercial_scientific <- function(Results,
   # Représentation graphique des donnees commerciales (4 graphes)
   
   # 1. Centres des zones de peche
-  # plot_centres = ggplot(centres) + geom_point(aes(x=x, y=y, col=as.factor(boats)), size=4) +
-  #   labs(color= "Bateau") +
-  #   theme_bw() +
-  #   theme(axis.title = element_blank(),legend.position='none')
-  
+  plot_centres = ggplot(centres) + geom_point(aes(x=x, y=y, col=as.factor(boats)), size=4) +
+    labs(color= "Bateau") +
+    theme_bw() +
+    theme(axis.title = element_blank(),legend.position='none')
+
   # 2. Points de peche
-  # plot_pointsdepechecomperboat = ggplot(peche_com_old) + geom_point(aes(x=x, y=y, col=as.factor(boats)), size=4) +
-  #   labs(color= "Bateau") +
-  #   theme_bw() +
-  #   theme(axis.title = element_blank(),legend.position='none')
+  plot_pointsdepechecomperboat = ggplot(peche_com_old) + geom_point(aes(x=x, y=y, col=as.factor(boats)), size=4) +
+    labs(color= "Bateau") +
+    theme_bw() +
+    theme(axis.title = element_blank(),legend.position='none')
   
   # 3. Points de peche dans les cellules
   pointsdepeche_com = as.data.frame(cbind(x = x_com, y = y_com, boats = boats_number, ncell = index_com_i, y_com_i = y_com_i))
-  # plot_pointsdepechecell = ggplot(pointsdepeche_com) +
-  #   geom_point(aes(x=x, y=y, col=as.factor(boats)), size=4) +
-  #   labs(color= "Bateau") +
-  #   theme_bw() +
-  #   theme(axis.title = element_blank(),legend.position='none')
+  plot_pointsdepechecell = ggplot(pointsdepeche_com) +
+    geom_point(aes(x=x, y=y, col=as.factor(boats)), size=4) +
+    labs(color= "Bateau") +
+    theme_bw() +
+    theme(axis.title = element_blank(),legend.position='none')
   
   # 4. Quantite pechee
-  # if (k==0){
-  #   plot_pointsdepecheqte = ggplot(pointsdepeche_com) + geom_point(aes(x=x, y=y, col=y_com_i), size=4) +
-  #     labs(color= "Quantite pechee") +
-  #     theme_bw() +
[…]

### 2022-06-02 Version_M2_plusieurszones/Scripts/reallocation_real_simu.R

@@ -54,7 +54,7 @@ if(species_to_plot == "Solea_solea"){
 }
 
 ## data file name
-sampling <- "simulate" # "simulate": simulate sampling, "from_tacsateflalo": take data points from tacsateflalo
+sampling <- "from_tacsateflalo" # "simulate": simulate sampling, "from_tacsateflalo": take data points from tacsateflalo
 simu_name <- "SimuFullArea"
 
 # file name for saving outputs

### 2022-06-02 Version_M2_plusieurszones/Scripts/simu_main_script.R

@@ -259,6 +259,7 @@ x_parallel1 <- clusterEvalQ(cl,{
   
   TMB::compile(paste0(TmbFile,"inst/executables/com_x_sci_data_14_scientific_commercial_simple.cpp"),"-O1 -g",DLLFLAGS="")
   
+  
   # If problems with the PATH
   # Sys.setenv(PATH = paste("C:/Rtools/bin", Sys.getenv("PATH"), sep=";")) # if "C:/Rtools/bin" is not in the PATH
   

### 2022-10-21 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
modif

@@
 
+numbersections : true
 floatsintext     : yes
@@
                 one.sum) %>%
-    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dj","Yi"),
+    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dk","Yi"),
            reallocation = ifelse(reallocation == 1,"Yes","No"),
@@
 Results <- Results %>%
-  mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dj","Yi"))) %>%
+  mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dk","Yi"))) %>%
   mutate(relloc_aggreg = ifelse(aggreg_obs == F & reallocation == 1,"Reallocation: Yes -  Likelihood: Yr",relloc_aggreg))
@@
                                            "Reallocation: Yes -  Likelihood: Yr",
-                                           "Reallocation: Yes -  Likelihood: Dj"))
+                                           "Reallocation: Yes -  Likelihood: Dk"))
 
@@
                                 "Reallocation: Yes -  Likelihood: Yr",
-                                "Reallocation: Yes -  Likelihood: Dj"),
+                                "Reallocation: Yes -  Likelihood: Dk"),
                      values=c("gold","#F8766D","#00BA38"))
@@
                                "Reallocation: Yes -  Likelihood: Yr",
-                               "Reallocation: Yes -  Likelihood: Dj"),
+                               "Reallocation: Yes -  Likelihood: Dk"),
                     values=c("gold","#F8766D","#00BA38"))+
@@
                 one.sum) %>%
-    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dj","Yi"),
+    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dk","Yi"),
            perc_convergence = round((1 - converge.sum / one.sum)*100,digits = 3))
@@
 Results_conv_2 <- Results_conv_2[-which(Results_conv_2$Model == "Scientific model" &
-                                          Results_conv_2$`Likelihood level` == "Dj"),]
+                                          Results_conv_2$`Likelihood level` == "Dk"),]
 Results_conv_2$`Likelihood level`[which(Results_conv_2$Model == "Scientific model" &
@@
 Results_plot$obs <- NA
-Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
+Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dk"
 Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yr"
 Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific data"
-Results_plot$obs <- factor(Results_plot$obs,levels = c("Yr","Dj","Scientific data"))
+Results_plot$obs <- factor(Results_plot$obs,levels = c("Yr","Dk","Scientific data"))
 
@@
                                 "Yr",
-                                "Dj"),
+                                "Dk"),
                      values=c("#619CFF","#F8766D","#00BA38"))
@@
                                 "Yr",
-                                "Dj"),
+                                "Dk"),
                      values=c("#619CFF","#F8766D","#00BA38"))+
@@
   scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-  ggtitle("Integrated model",subtitle = "Commercial likelihood on Dj")+
+  ggtitle("Integrated model",subtitle = "Commercial likelihood on Dk")+
   theme_bw()+
@@
 ## Parameter estimates
+est_par_df_full_3$lkl[which(est_par_df_full_3$lkl == "Dj")] <- "Dk"
+
 est_par_df_full_3$lkl[which(est_par_df_full_3$lkl == "Yi")] <- "Yr"
-est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yr","Scientific model"))
+est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dk","Yr","Scientific model"))
 est_par_df_full_3 <- est_par_df_full_3 %>%

### 2022-10-21 Version_M2_plusieurszones/pres/AFH_2022/AFH_2022.Rmd
modif

--- 2022-05-23 Version_M2_plusieurszones/pres/RESSTE_2022/RESSTE_2022.Rmd
@@
 ---
-title: "Downscaling coarse observations to predict continuous species spatio-temporal distribution"
-subtitle : "Going from coarse landings data to fine scale fish distribution"
-author : "Baptiste Alglave, Marie-Pierre Etienne, Kasper Kristensen, Youen Vermard, Mathieu Woillez, Etienne Rivot"
+title: "Inferring fine scale wild species distribution from spatially aggregated data"
+author : "Marie-Pierre Etienne, Baptiste Alglave, Kasper Kristensen, Youen Vermard, Mathieu Woillez, Etienne Rivot"
 

### 2022-10-21 Version_M2_plusieurszones/pres/AFH_2022/AFH_2022.tex
modif

--- 2022-10-21 Version_M2_plusieurszones/pres/AFH_2022/AFH_2022.Rmd
@@
----
-title: "Inferring fine scale wild species distribution from spatially aggregated data"
-author : "Marie-Pierre Etienne, Baptiste Alglave, Kasper Kristensen, Youen Vermard, Mathieu Woillez, Etienne Rivot"
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  10pt,
+  ignorenonframetext,
+]{beamer}
+\usepackage{pgfpages}
+\setbeamertemplate{caption}[numbered]
+\setbeamertemplate{caption label separator}{: }
+\setbeamercolor{caption name}{fg=normal text.fg}
+\beamertemplatenavigationsymbolsempty
+% Prevent slide breaks in the middle of a paragraph
+\widowpenalties 1 10000
+\raggedbottom
+\setbeamertemplate{part page}{
+  \centering
+  \begin{beamercolorbox}[sep=16pt,center]{part title}
+    \usebeamerfont{part title}\insertpart\par
+  \end{beamercolorbox}
+}
+\setbeamertemplate{section page}{
+  \centering
+  \begin{beamercolorbox}[sep=12pt,center]{part title}
+    \usebeamerfont{section title}\insertsection\par
+  \end{beamercolorbox}
+}
+\setbeamertemplate{subsection page}{
+  \centering
+  \begin{beamercolorbox}[sep=8pt,center]{part title}
+    \usebeamerfont{subsection title}\insertsubsection\par
+  \end{beamercolorbox}
+}
+\AtBeginPart{
+  \frame{\partpage}
+}
+\AtBeginSection{
+  \ifbibliography
+  \else
+    \frame{\sectionpage}
+  \fi
+}
+\AtBeginSubsection{
+  \frame{\subsectionpage}
+}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usetheme[]{Darmstadt}
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Inferring fine scale wild species distribution from spatially aggregated data},
+  pdfauthor={Marie-Pierre Etienne, Baptiste Alglave, Kasper Kristensen, Youen Vermard, Mathieu Woillez, Etienne Rivot},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\newif\ifbibliography
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% package -----------------------------------------
+\usepackage{graphicx}
+\usepackage{rotating}
+%\setbeamertemplate{caption}[numbered]
+\usepackage{hyperref}
+\usepackage{caption}
+\usepackage[normalem]{ulem}
+%\mode<presentation>
+\usepackage{wasysym}
+\usepackage{amsmath}
+\usepackage{caption}
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage{pifont}
+\usepackage{dashrule}
+\usepackage{setspace}
+\usepackage{array}
+\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
+\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
 
-fontsize: 10pt
-bibliography : ["ref.bib"]
-csl : ices-journal-of-marine-science.csl
+% ----------------------------------------------
+\usepackage[style=authoryear,sorting=ynt]{biblatex}
+\usepackage[font={small}]{caption}
 
-output:
-  beamer_presentation:
-    theme: "Darmstadt"
-    colortheme: "default"
-    slide_level: 3
-    # template: mytemplate.tex
-    includes:
-      in_header: header-simple.tex
----
+\captionsetup[figure]{labelformat=empty}% redefines the caption setup of the figures environment in the beamer class.
+\usepackage{caption}
+\captionsetup[figure]{font=tiny}
 
-```{r setup, include=FALSE}
+\setbeamertemplate{navigation symbols}{}
+\institute{06/2022 - AFH}
 
-knitr::opts_chunk$set(echo = TRUE)
+% logo -----------------------------------------
+\newsavebox{\logoA}
+\newsavebox{\logoB}
+\newsavebox{\logoC}
+\savebox{\logoA}{\includegraphics[width=3cm]{images/logo_ifremer.jpg}}
+\savebox{\logoB}{\includegraphics[width=3cm]{images/UMR_DECOD.jpg}}
+\savebox{\logoC}{\includegraphics[width=3cm]{images/logo_institut_agro.png}}
 
-library(dplyr)
-library(tidyr)
-library(rnaturalearth)
-library(stringr)
-library(sf)
-library(raster)
-library(maps)
-library(mapdata)
-library(ggplot2)
-library(ggspatial)
-library(cowplot)
-library(latex2exp)
-library(ggpubr)
+% \savebox{\logoD}{\includegraphics[width=1.5cm]{images/dtu_logo.jpg}}
+% \newsavebox{\logoD}
 
-col_scale <- function(palette = "Spectral", name = "", limits = NULL) {
-  scale_colour_distiller(palette = palette,   # spectral colour scale
-                         guide = "colourbar", # continuous colour bar
-                         name = name,
-                         limits = limits)
+\titlegraphic{%
+  \raisebox{.5\dimexpr\ht\logoB-\ht\logoA}{\usebox{\logoA}}% raise smaller logo into position
+  \hspace*{0.75cm}
+  \usebox{\logoC}
+  \hspace*{0.75cm}%
+  % \usebox{\logoC}
+  % \hspace*{1cm}%
+  \usebox{\logoB}
+
 }
 
-run_codes <- F
+% ----------------------------------------------
 
-grid_projection <- "+proj=longlat +datum=WGS84"
+\setbeamertemplate{title page}[empty]
 
-mapBase <- map("worldHires", fill = T, plot = F)
-mapBase <- st_as_sf(mapBase) %>% filter(ID %in% c("France","Spain"))
+\setbeamerfont{subtitle}{size=\small}
 
-```
+\setbeamercovered{transparent}
 
-# Context
+\definecolor{BaptisteBlue}{HTML}{6495ED}
+\definecolor{SurveyBlue}{HTML}{619CFF}
+\definecolor{BaptisteOrange}{HTML}{F39C12}
+\definecolor{BaptisteGreen}{HTML}{1E8449}
+\definecolor{BaptisteLightGreen}{HTML}{28B463}
+\definecolor{BaptisteGrey}{HTML}{A6ACAF}
+\definecolor{BaptisteRed}{HTML}{FF2A2A}
+\definecolor{BaptisteBrown}{HTML}{D45500}
+\definecolor{BaptisteDarkGrey}{HTML}{8C9193}
 
-### Spatial data in ecology
+% ---------------------------------------------------------
+% New command
 
-<!-- Standardized and high quality data, but low amount of samples -->
-<!-- more data, but no standardize sampling plan (e.g. problem related to preferential sampling) -->
-<!-- legal and massive data, but low quality, declared at the level of some administrative unit -->
+\newcommand{\btVFill}{\vskip0pt plus 1filll}
+
+% % make code-output smaller
+% \DefineVerbatimEnvironment{Highlighting}{Verbatim}{fontsize=\tiny,commandchars=\\\{\}}
+% 
+% % make console-output smaller:
+%   \makeatletter
+% \def\verbatim{\tiny\@verbatim \frenchspacing\@vobeyspaces \@xverbatim}
+% \makeatother
+%
+
+% \setlength{\parskip}{0pt}
+
+% \setlength{\OuterFrameSep}{-4pt}
+% \makeatletter
+% \preto{\@verbatim}{\topsep=-10pt \partopsep=-10pt }
+% \makeatother
+
+% ---------------------------------------------------------
+
+% \setbeamercolor{frametitle}{fg=clemsonpurple,bg=white}
+% \setbeamercolor{title}{fg=clemsonpurple,bg=white}
+% \setbeamercolor{local structure}{fg=clemsonpurple}
+% \setbeamercolor{section in toc}{fg=clemsonpurple,bg=white}
+% % \setbeamercolor{subsection in toc}{fg=clemsonorange,bg=white}
+% \setbeamercolor{item projected}{fg=clemsonpurple,bg=white}
+% \setbeamertemplate{itemize item}{\color{clemsonpurple}$\bullet$}
+% \setbeamertemplate{itemize subitem}{\color{clemsonpurple}\scriptsize{$\bullet$}}
+\let\Tiny=\tiny
+
+\AtBeginPart{}
+\AtBeginSection{}
+\AtBeginSubsection{}
+\AtBeginSubsubsection{}
+\setlength{\emergencystretch}{0em}
+\setlength{\parskip}{0pt}
+
+\title{Inferring fine scale wild species distribution from spatially aggregated
+data}
+\author{Marie-Pierre Etienne, Baptiste Alglave, Kasper Kristensen, Youen
+Vermard, Mathieu Woillez, Etienne Rivot}
+\date{}
+
+\begin{document}
+\frame{\titlepage}
+
+\hypertarget{context}{%
+\section{Context}\label{context}}
+
+\begin{frame}{Spatial data in ecology}
+\protect\hypertarget{spatial-data-in-ecology}{}
 
@@
 
----
-<!-- \btVFill -->
+\end{frame}
+
+\begin{frame}
+
 \begin{center} \Large
@@
 \large
+
 (especially when they do not have the same spatial resolution)
@@
 
-# Material and method
+\end{frame}
 
-### Commercial catch declarations data in fishery science
+\hypertarget{material-and-method}{%
+\section{Material and method}\label{material-and-method}}
 
-```{r,echo=F,fig.align='center',out.width="120%"}
-
-if(run_codes){
-
-  load("map/ICES_rect.RData")
-  load("res/points_sf.RData")
-
-  #-------------------
-  ## Logbook data plot
-  #-------------------
-  logbooks_df <- points_sf %>%
-    # filter(str_detect(SI_DATE,"11/2018")) %>%
-    group_by(ICESNAME_1) %>%
-    dplyr::summarise(Catch = sum(LE_KG_Solea_solea)/1000,
-                     Effort = sum(INTV)/60)
-
-  logbooks_sf <- st_join(ICES_rect,logbooks_df) %>%
-    filter(!is.na(Effort))
-
-  Logbooks_data <- ggplot()+
-    geom_sf(data = ICES_rect,fill="skyblue",alpha=0.25)+
-    geom_sf(data=logbooks_sf,aes(fill=Catch))+
-    geom_sf(data = mapBase)+
-    scale_fill_distiller(palette = "Spectral")+
-    theme(panel.background = element_rect(fill = "skyblue"),
-          plot.title = element_text(hjust=0.5),
-          legend.position = "none")+
-    geom_sf(col="black")+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    annotation_scale(location = "bl", line_width = .5)+
-    annotation_north_arrow(location = "tl", height = unit(0.7, "cm"),
-                           width = unit(0.7, "cm"))
-
-  ggsave(Logbooks_data,filename = "images/logbooks_data.png",width = 5,height = 5)
-
-  #---------------
-  ## VMS data plot
-  #---------------
-  ## Aggregate/discretize data
-  vms_df <- points_sf %>%
-    mutate(long = sf::st_coordinates(.)[,1],
-           lati = sf::st_coordinates(.)[,2]) %>%
-    as.data.frame %>%
-    dplyr::select(-geometry) %>%
-    mutate(long = round(long,digits = 2),
-           lati = round(lati,digits = 2)) %>%
-    group_by(long,lati) %>%
-    dplyr::summarise(Effort = sum(INTV)/60)
-
-  vms_sf <- st_as_sf(vms_df,coords = c("long","lati"),crs=grid_projection)
-
-  ## Build polygon
-  # Coordinates Bay of Biscay
-  grid_xmin <- -6
-  grid_xmax <- 0
-  grid_ymin <- 42
-  grid_ymax <- 48
-  grid_limit <- extent(c(grid_xmin,grid_xmax,grid_ymin,grid_ymax))
-
-  # Build grid
-  grid <- raster(grid_limit)
-  res(grid) <- 0.025
-  crs(grid) <- "+proj=longlat +datum=WGS84"
-  gridpolygon <- rasterToPolygons(grid)
-  gridpolygon$layer <- c(1:length(gridpolygon$layer))
-  gridpolygon_sf <- st_as_sf(gridpolygon)
-
-  vms_sf_2 <- st_join(gridpolygon_sf,points_sf)
-  vms_sf_3 <- vms_sf_2 %>%
-    group_by(layer) %>%
-    dplyr::summarise(Effort = sum(INTV)/60)
-
-  vms_sf_3 <- vms_sf_3 %>%
-    filter(!is.na(Effort))
-
-  # save(data=vms_sf_3,file="res/vms_sf_3.RData")
-
-  vms_data <- ggplot()+
-    geom_sf(data = ICES_rect,fill="skyblue",alpha=0.25)+
-    geom_sf(data=vms_sf_3,aes(col=Effort,fill=Effort))+
-    geom_sf(data = ICES_rect,alpha=0)+
-    geom_sf(data = mapBase)+
-    scale_color_distiller(direction = 1,palette = "OrRd")+ # palette = "OrRd"
-    scale_fill_distiller(direction = 1,palette = "OrRd")+ # ,palette = "Spectral"
-    theme(panel.background = element_rect(fill = "skyblue"),
-          plot.title = element_text(hjust=0.5),
-          legend.position = "none")+
-    geom_sf(col="black")+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)
-
-  ggsave(filename = "images/vms_data.png",width = 5,height = 5)
-
-}
-
-```
+\begin{frame}{Commercial catch declarations data in fishery science}
+\protect\hypertarget{commercial-catch-declarations-data-in-fishery-science}{}
 
@@
 
-### Two alternative procedures to reallocate catches
+\end{frame}
+
+\begin{frame}{Two alternative procedures to reallocate catches}
+\protect\hypertarget{two-alternative-procedures-to-reallocate-catches}{}
 
@@
 
-### Two alternative procedures to reallocate catches
+\end{frame}
 
-\textbf{Punctual observation model} ($Y_i$)
+\begin{frame}{Two alternative procedures to reallocate catches}
+\protect\hypertarget{two-alternative-procedures-to-reallocate-catches-1}{}
 
-\scriptsize $\operatorname{L}(y,\mu,\sigma^2)$ is the lognormal likelihood for observation $y$, mean $\mu$ and variance $\sigma^2$
+\textbf{Punctual observation model} (\(Y_i\))
 
-$Y$ and $D$ are supposed conditional on S and x \footnotesize
+\scriptsize \(\operatorname{L}(y,\mu,\sigma^2)\) is the lognormal
+likelihood for observation \(y\), mean \(\mu\) and variance \(\sigma^2\)
 
-$$
+\(Y\) and \(D\) are supposed conditional on S and x \footnotesize
+
+\[
 \operatorname{P}\left(Y_i=y_{i} \right) =
@@
 \right.
-$$
+\]
 
-$$p_{i}=\exp(-e^\xi .S(x_{i}))$$
+\[p_{i}=\exp(-e^\xi .S(x_{i}))\]
 
@@
 
-### Specifying $\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right)$
+\end{frame}
+
+\begin{frame}{Specifying
+\(\operatorname{P}\left(D_j=d_{j} \vert d_j > 0 \right)\)}
+\protect\hypertarget{specifying-operatornamepleftd_jd_j-vert-d_j-0-right}{}
 
@@
 
-$$E(D_j \vert d_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
+\[E(D_j \vert d_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}\]
 
-$$Var(D_j \vert d_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
+\[Var(D_j \vert d_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2\]
 
-$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$
+\[Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))\]
 
@@
 
-### Simulation-estimation and case study
+\end{frame}
 
-```{r,echo=F}
-
-if(run_codes){
-  ## Load simulated data
-  load("res/simu_df.RData")
-  load("res/list_input.RData")
-
-  simu_df_2 <- simu_df %>%
-    mutate(Ab = sum(S_x))
-
-  loc_x <- list_input$loc_x
-  sci.data_df <- data.frame(cell = list_input$index_sci_i,
-                            y_sci = list_input$y_sci_i) %>%
-    inner_join(loc_x[,c("x","y","cell","ICESNAME")])
-
-  com.data_df <- data.frame(cell = list_input$index_i,
-                            y_com = list_input$y_i2) %>%
-    inner_join(loc_x[,c("x","y","cell","ICESNAME")]) %>%
-    group_by(x,y,cell,ICESNAME) %>%
-    dplyr::summarise(n=n())
-
-  ## Grid data
-  grid <- raster(grid_limit)
-  res(grid) <- 0.05
-  crs(grid) <- "+proj=longlat +datum=WGS84"
-  gridpolygon <- rasterToPolygons(grid)
-  gridpolygon$layer <- c(1:length(gridpolygon$layer))
-  gridpolygon_sf <- st_as_sf(gridpolygon)
-
-  simu_sf <- st_as_sf(simu_df_2,coords = c("long","lati"),crs=grid_projection)
-
-  simu_sf_2 <- st_join(gridpolygon_sf,simu_sf) %>%
-    filter(!is.na(S_x))
-
-  ## Plot data
-  data_plot <- ggplot()+
-    geom_sf(data=simu_sf_2,aes(fill=log(S_x)),col=NA,alpha=0.5)+
-    geom_sf(data=simu_sf_2[which(simu_sf_2$ICESNAME %in% com.data_df$ICESNAME),],
-            aes(fill=log(S_x),col=log(S_x)))+
-    scale_fill_distiller(palette = "Spectral")+
-    scale_color_distiller(palette = "Spectral")+
-    geom_point(data=sci.data_df,aes(x=x,y=y),fill="#619CFF",shape=21,size=1)+
-    theme_bw()+
-    theme(plot.title = element_text(hjust = 0.5),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.title = element_blank(),
-          legend.position = "none")+
-    geom_sf(data=ICES_rect,alpha=0)+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  ggsave(filename = "images/data_plot.png",width = 5,height = 5)
-
-}
-
-```
+\begin{frame}{Simulation-estimation and case study}
+\protect\hypertarget{simulation-estimation-and-case-study}{}
 
@@
 
-# Results
+\end{frame}
 
-### Simulation-estimation
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-```{r,echo=F}
-
-if(run_codes){
-
-  ######################
-  ## Parameter estimates
-  ######################
-  load("res/Results_full_multi_square.RData")
-
-  Results_plot <- Results_2 %>%
-    filter(converge == 0) %>%
-    filter(simu_type == "Unsampled Rectangles")
-
-  Results_plot[,"RelBias_N"]=(Results_plot[,"N_est"]-Results_plot[,"N_true"])/Results_plot[,"N_true"]
-  Results_plot[,"RelBias_N.2"]=(Results_plot[,"N_est.2"]-Results_plot[,"N_true.2"])/Results_plot[,"N_true.2"]
-  Results_plot[,"RelBias_beta"]=(Results_plot[,"beta1_est"]-Results_plot[,"beta1_true"])/Results_plot[,"beta1_true"]
-
-  Results_plot$obs <- NA
-  Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
-  Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi*"
-  Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific model"
-  Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi*","Dj","Scientific model"))
-
-  Results_plot$Model[which(Results_plot$Model == "Scientific model")] <- "Scientific"
-  Results_plot$Model[which(Results_plot$Model == "Commercial model")] <- "Commercial"
-  Results_plot$Model[which(Results_plot$Model == "Integrated model")] <- "Integrated"
-
-  Results_plot <- Results_plot %>%
-    filter(Model != "Commercial")
-
-  # Results_plot_fullArea <- Results_plot
-  # Results_plot_UnsampArea <- Results_plot
-
-  beta_plot <- ggplot()+
-    geom_boxplot(data = Results_plot,
-                 aes(x = Model,
-                     y = beta1_est,
-                     fill = obs))+
-    geom_hline(yintercept = 0,linetype="dashed")+
-    theme_bw()+
-    theme(legend.title = element_blank(),
-          legend.position = "none",
-          aspect.ratio = 1)+
-    # facet_wrap(.~simu_type)+
-    ylab(TeX("$\\beta_S$")) +
-    xlab("")+
-    geom_hline(yintercept = 2,col="red",linetype="dashed")+
-    scale_color_manual(breaks = c("Scientific model",
-                                  "Yi*",
-                                  "Dj"),
-                       values=c("#619CFF","#F8766D","#00BA38"))
-
-  MSPE_S_plot <- ggplot()+
-    geom_boxplot(data = Results_plot,
-                 aes(x = Model,
-                     y = mspe,
-                     fill = obs))+
-    theme_bw()+
-    theme(legend.title = element_blank(),
-          legend.position = "none",
-          aspect.ratio = 1)+
-    # facet_wrap(.~simu_type)+
-    ylab("MSPE")+xlab("")+
-    scale_color_manual(breaks = c("Scientific model",
-                                  "Yi*",
-                                  "Dj"),
-                       values=c("#619CFF","#F8766D","#00BA38"))+
-    ylim(0,NA)
-
-  legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="bottom",legend.title = element_blank())))
-
-  pl_plot <- plot_grid(MSPE_S_plot,
-                       beta_plot,
-                       ncol = 1,
-                       legend,
-                       align="hv",
-                       rel_heights = c(1,1,0.15))
-
-  ggsave(filename = "images/res_simu_plot.png",width = 4,height = 7.5)
-
-  #######
-  ## Maps
-  #######
-  simu_plot <- ggplot(simu_df)+
-    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral")+
-    ggtitle("Simulated latent field")+
-    theme_bw()+
-    theme(plot.title = element_text(hjust = 0.5,face = "bold"),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.position = "none")+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  load("res/fit_IM_res_sci.RData")
-  sci_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
-  sci_plot <- ggplot(sci_df)+
-    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral")+
-    ggtitle("Scientific model")+
-    theme_bw()+
-    theme(plot.title = element_text(colour = "#619CFF",hjust = 0.5,face = "bold"),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.position = "none")+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  load("res/fit_IM_res_int_Yi.RData")
-  int_Yi_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
-  int_Yi_plot <- ggplot(int_Yi_df)+
-    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral")+
-    ggtitle("Commercial likelihood on Yi*")+
-    theme_bw()+
-    theme(plot.title = element_text(colour = "#F8766D",hjust = 0.5,face = "bold"),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.position = "none")+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  load("res/fit_IM_res_int_Dj.RData")
-  int_Dj_df <- data.frame(loc_x,S_x=fit_IM_res$Report$S_x)
-  int_Dj_plot <- ggplot(int_Dj_df)+
-    geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral")+
-    ggtitle("Commercial likelihood on Dj")+
-    theme_bw()+
-    theme(plot.title = element_text(colour = "#00BA38",hjust = 0.5,face = "bold"),
-          plot.subtitle = element_text(hjust = 0.5),
-          legend.position = "none")+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  map_simu <- plot_grid(simu_plot,sci_plot,int_Yi_plot,int_Dj_plot)
-
-  ggsave(filename = "images/map_simu_plot.png",width = 6,height = 7.5)
-
-}
-
-```
+\begin{frame}{Simulation-estimation}
+\protect\hypertarget{simulation-estimation}{}
 
@@
 
-### Case study: Sole in the Bay of Biscay
+\end{frame}
 
-```{r,echo=F}
-
-if(run_codes){
-
-  ## Load data
-  load("res/sci_df.RData")
-
-  load("res/realloc_rest_df.RData")
-
-  load("res/no.realloc_df.RData")
-
-  load("res/est_par_df_full_3.RData")
-
-  ## Plot maps
-  no.realloc_plot <- ggplot(no.realloc_df)+
-    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-    ggtitle("Commercial likelihood on Yi*")+
-    theme_bw()+
-    theme(legend.position = "none",
-          plot.title = element_text(colour = "#F8766D",hjust = 0.5,face = "bold"))+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  realloc_rest_plot <- ggplot(realloc_rest_df)+
-    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-    ggtitle("Commercial likelihood on Dj")+
-    theme_bw()+
-    theme(legend.position = "none",
-          plot.title = element_text(colour = "#00BA38",hjust = 0.5,face = "bold"))+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  sci_plot <- ggplot(sci_df)+
-    geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
-    scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
-    ggtitle("Scientific model")+
-    theme_bw()+
-    theme(legend.position = "none",
-          plot.title = element_text(colour = "#619CFF",hjust = 0.5,face = "bold"))+
-    geom_sf(data = mapBase)+
-    coord_sf(xlim = c(-6,0), ylim = c(43,48), expand = FALSE)+
-    xlab("")+ylab("")
-
-  # plot(log(no.realloc_df$S_x),log(sci_df$S_x))
-  # cor(log(no.realloc_df$S_x),log(sci_df$S_x),method = "spearman")
-  # plot(log(realloc_rest_df$S_x),log(sci_df$S_x))
-  # cor(log(realloc_rest_df$S_x),log(sci_df$S_x),method = "spearman")
-
-  case_study_plot <- plot_grid(NULL,sci_plot,no.realloc_plot,realloc_rest_plot,align="hv")
-
-  ggsave("images/case_study_plot.png",width = 6, height = 7.5)
-
-  ## Parameter estimates
-  est_par_df_full_3$lkl[which(est_par_df_full_3$lkl == "Yi")] <- "Yi*"
-  est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi*","Scientific model"))
-  est_par_df_full_3 <- est_par_df_full_3 %>%
-    filter(par_names != "substr_Sand_Coarse_substrate")
-  par_plot <- ggplot(est_par_df_full_3, aes(y=par_val, x=par_names))+
-    geom_point(
-      aes(color = lkl),
-      position = position_dodge(0.5),
-      size=2
-    )+
-    geom_errorbar(aes(ymin = CI.inf, ymax = CI.sup, color = lkl),
-                  position = position_dodge(0.5),width=0.4
-    )+
-    scale_color_manual(breaks = c("Scientific model",
-                                  "Yi*",
-                                  "Dj"),
-                       values=c("#619CFF","#F8766D","#00BA38"))+
-    geom_hline(yintercept=0, linetype="dashed", color = "red",alpha=0.4)+
-    xlab("")+ylab("")+
-    theme_bw()+
-    ggtitle("Parameters estimates")+
-    theme(plot.title = element_text(hjust = 0.5,face = "bold"),
-          legend.title = element_blank(),
-          legend.position = "bottom",
-          aspect.ratio = 1)+
-    coord_flip()+
-    scale_x_discrete(labels=c('k_com'=parse(text = TeX('$k_{com}$')),
-                              'Sigma_com'=parse(text = TeX('$\\sigma_{com}$')),
-                              'q1_com'=parse(text = TeX('$\\xi_{com}$')),
-                              'Sigma_sci'=parse(text = TeX('$\\sigma_{sci}$')),
-                              'q1_sci'=parse(text = TeX('$\\xi_{sci}$')),
-                              'Range'="Range",
-                              'MargSD'="M. var.",
-                              'substr_Mud_sediment'=parse(text = TeX('$\\beta_{S}$')),
-                              'intercept'=parse(text = TeX('$\\mu$'))))
-
-  ggsave("images/par_plot.png",width = 6/1.75,height = 7.5/1.75)
-
-}
-
-```
+\begin{frame}{Case study: Sole in the Bay of Biscay}
+\protect\hypertarget{case-study-sole-in-the-bay-of-biscay}{}
 
@@
 
-# Discussion
+\end{frame}
 
-### Discussion
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\begin{frame}{Discussion}
+\protect\hypertarget{discussion-1}{}
 
@@
 
-#
+\end{frame}
+
+\hypertarget{section}{%
+\section{}\label{section}}
+
+\begin{frame}{}
 
@@
 \end{figure}
+
+\end{frame}
+
+\end{document}

### 2022-10-21 Version_M2_plusieurszones/pres/RESSTE_2022/RESSTE_2022.tex
modif

--- 2022-10-21 Version_M2_plusieurszones/pres/AFH_2022/AFH_2022.tex
@@
 \hypersetup{
-  pdftitle={Inferring fine scale wild species distribution from spatially aggregated data},
-  pdfauthor={Marie-Pierre Etienne, Baptiste Alglave, Kasper Kristensen, Youen Vermard, Mathieu Woillez, Etienne Rivot},
+  pdftitle={Downscaling coarse observations to predict continuous species spatio-temporal distribution},
+  pdfauthor={Baptiste Alglave, Marie-Pierre Etienne, Kasper Kristensen, Youen Vermard, Mathieu Woillez, Etienne Rivot},
   hidelinks,
@@
 \setbeamertemplate{navigation symbols}{}
-\institute{06/2022 - AFH}
+\institute{04/2022 - RESSTE}
 
@@
 
-\title{Inferring fine scale wild species distribution from spatially aggregated
-data}
-\author{Marie-Pierre Etienne, Baptiste Alglave, Kasper Kristensen, Youen
+\title{Downscaling coarse observations to predict continuous species
+spatio-temporal distribution}
+\subtitle{Going from coarse landings data to fine scale fish distribution}
+\author{Baptiste Alglave, Marie-Pierre Etienne, Kasper Kristensen, Youen
 Vermard, Mathieu Woillez, Etienne Rivot}

### 2022-10-21 Version_M2_plusieurszones/Scripts/In_progress/figure_intro_manuscript.R

@@ -1,0 +1,108 @@
+library(sf)
+library(maps)
+library(mapdata)
+library(ggspatial)
+
+plot(st_geometry(points_sf_4))
+
+# world <- ne_countries(scale = "medium", returnclass = "sf")
+mapBase <- map("worldHires", fill = T, plot = F)
+mapBase <- st_as_sf(mapBase) %>% filter(ID %in% c("France","Spain"))
+
+VMS_data <- points_sf %>%
+  filter(month %in% c(10:12)) %>%
+  filter(LE_MET_level6 %in% c("OTB_DEF_>=70_0","OTB_CEP_>=70_0","OTT_DEF_>=70_0")) %>%
+  mutate(MET = str_sub(LE_MET_level6,1,3))
+
+
+plot_figure <- ggplot()+
+  theme_bw()+
+  geom_sf(data=VMS_data[which(VMS_data$MET == "OTT"),],
+          col="grey",alpha=0.05,shape=16,size=1)+
+  geom_sf(data=VMS_data[which(VMS_data$MET == "OTB"),],
+          col="bisque",alpha=0.1,shape=16,size=1)+
+  geom_sf(data=ObsMer_sf[which(ObsMer_sf$foCatEu6 == "OTT_DEF_>=70_0"),],
+          aes(size = CPUE),col = "darkgrey")+
+  geom_sf(data=ObsMer_sf[which(ObsMer_sf$foCatEu6 != "OTT_DEF_>=70_0"),],
+          aes(size = CPUE),col = "orange")+
+  geom_sf(data = st_as_sf(survey_layer_3),col="lightblue1",alpha=0,size=2)+
+  geom_point(data=ptsurvey_wgs84_sf_4,
+             aes(x = x, y = y,size = CatchWgt_spp),
+             col = "skyblue")+
+  theme(legend.position = "none")+
+  geom_sf(data = ICES_rect_2,alpha=0)+
+  geom_sf(data = mapBase)+
+  coord_sf(xlim = c(-5.5,0.25), ylim = c(43,48), expand = FALSE)+
+  annotation_scale(location = "bl", line_width = .5) +
+  annotation_north_arrow(location = "tl", height = unit(0.7, "cm"), width = unit(0.7, "cm"))+
+  xlab("")+ylab("")
+
+# x11()
+# plot_figure
+
+ggsave("C:/R_projects/phd_manuscript/Introduction/figures/several-datasources.png",
[…]

### 2022-10-21 Version_M2_plusieurszones/Scripts/In_progress/plot_multiple_square_2.R

@@ -3,7 +3,7 @@
 ##############################################################
 
 compute_profile <- F
-save_data <- T
+save_data <- F
 
 load("draft/paper_3/res/list_input.RData")
 
@@ -268,6 +268,8 @@ if(compute_profile){
 ############
 ## Plot maps
 ############
+
+
 max_val <- max(simu_df$S_x/sum(simu_df$S_x),
                sci_df$S_x/sum(sci_df$S_x),
                int_Yi_df$S_x/sum(int_Yi_df$S_x),
@@ -323,7 +325,7 @@ sci_plot <- ggplot(sci_df)+
 int_Yi_plot <- ggplot(int_Yi_df)+
   geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
   scale_color_distiller(palette = "Spectral",limits=c(0,max_val+0.01*max_val))+
-  ggtitle("Integrated model - Yi")+
+  ggtitle("Integrated model - Yr")+
   theme_bw()+
   theme(plot.title = element_text(hjust = 0.5),
         plot.subtitle = element_text(hjust = 0.5),
@@ -335,7 +337,7 @@ int_Yi_plot <- ggplot(int_Yi_df)+
 com_Yi_plot <- ggplot(com_Yi_df)+
   geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
   scale_color_distiller(palette = "Spectral",limits=c(0,max_val+0.01*max_val))+
-  ggtitle("Commercial model - Yi")+
+  ggtitle("Commercial model - Yr")+
   theme_bw()+
   theme(plot.title = element_text(hjust = 0.5),
         plot.subtitle = element_text(hjust = 0.5),
@@ -347,7 +349,7 @@ com_Yi_plot <- ggplot(com_Yi_df)+
 int_Dj_plot <- ggplot(int_Dj_df)+
   geom_point(aes(x=x,y=y,col=S_x/sum(S_x)),shape=15,size=2)+
   scale_color_distiller(palette = "Spectral",limits=c(0,max_val+0.01*max_val))+
-  ggtitle("Integrated model - Dj")+
+  ggtitle("Integrated model - Dk")+
   theme_bw()+
   theme(plot.title = element_text(hjust = 0.5),
         plot.subtitle = element_text(hjust = 0.5),
@@ -359,7 +361,7 @@ int_Dj_plot <- ggplot(int_Dj_df)+
[…]

### 2022-10-21 Version_M2_plusieurszones/Scripts/reallocation_real_simu.R

@@ -141,7 +141,13 @@ loc_x <- loc_x_sf %>%
          y=loc_x$y)
 
 ## Load tacsateflalo and obsmer
+foCatEu6_filter <- "OTT_DEF_>=70_0|OTB_CEP_>=70_0|OTB_DEF_>=70_0"
 source("Scripts/source/load_commercial_data.R")
+
+
+# ## Make figure manuscript
+# source("Scripts/In_progress/figure_intro_manuscript.R")
+
 
 ## Initialize loop
 #-----------------

### 2022-10-21 Version_M2_plusieurszones/Scripts/source/shape_ObsMer.R

@@ -13,8 +13,9 @@ HH_obsmer_2$quarter[which(HH_obsmer_2$month %in% c(4,5,6))] <- 2
 HH_obsmer_2$quarter[which(HH_obsmer_2$month %in% c(7,8,9))] <- 3
 HH_obsmer_2$quarter[which(HH_obsmer_2$month %in% c(10,11,12))] <- 4
 
+
 HH_obsmer_2 <- HH_obsmer_2 %>%
-  filter(str_detect(foCatEu6,"OTB_CEP_>=70_0|OTB_DEF_>=70_0")) %>%
+  filter(str_detect(foCatEu6,foCatEu6_filter)) %>%
   filter(year == year_y) %>%
   filter(y<48 & quarter == quarter_q) # month == month_m)
 
@@ -46,3 +47,20 @@ ObsMer_df <- inner_join(ObsMer_df,loc_x[,c("layer","cell","x","y","ICESNAME")])
 
 index_ObsM_i <- ObsMer_df$cell
 y_ObsM_i <- ObsMer_df$CPUE
+
+# vec_obs <- y_ObsM_i[which(y_ObsM_i != 0)]
+# lmfit <- lm(log(vec_obs) ~ 1)
+# 
+# descdist((vec_obs))
+# 
+# fit.gamma <- fitdist(vec_obs, "gamma")
+# plot(fit.gamma)
+# fit.gamma$aic
+# 
+# fit.norm <- fitdist(log(vec_obs), "norm")
+# plot(fit.norm)
+# fit.norm$aic
+# 
+# fit.lognorm <- fitdist(vec_obs, "lnorm")
+# plot(fit.lognorm)
+# fit.lognorm$aic
